### 2014-01-13 RoughWork/manuscript.tex
Added skeleton tex files for paper draft.

\documentclass[pdftex,12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{natbib}

\title{Storm Surges}

\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Abstract to be completed
\end{abstract}

\section{Introduction}\label{sec:intro}
\citep{masson2004modelling} %testing a citation

\section{Model Evaluation}\label{sec:model}
%A section about how we configured the model for SoG 
%To be inlcluded: NEMO overview, tides, rivers, bathymetry, atmospheric forcing, vertical/lateral mixing, boundary conditions, grid.

% Kate: A description of the tides and river forcing can be included here.

\section{Storm Surge Hindcasts}\label{sec:storm}

\section{Conclusions}\label{sec:conclusions}

\bibliographystyle{plain}
\bibliography{ref}

\end{document}

### 2014-01-21 RoughWork/manuscript.tex
Started adding rivers info to paper

@@
 \usepackage{natbib}
+\usepackage{amsmath}
 
@@
 % Kate: A description of the tides and river forcing can be included here. 
+\subsection{River forcing}
+River input provides a significant volume of freshwater to the Salish Sea and can influence stratification, circulation and primary productivity.
+However, most rivers in the domain are not gauged so parameterisations were required to represent river flow.
+\cite{morrison2011rivers} provides a method for estimating freshwater runoff in the Salish Sea region based on precipitation.
+Exact data were acquired from \cite{morrison2011rivers}, including the runoff volumes for each watershed for each year from 1970 to 2012, as well as monthly averages.
+
+Freshwater runoff from each watershed was divided between the rivers in that watershed, based on estimations of the area that each river drained in its catchment.
+The area drained by each river was estimated from Toporama maps by the Atlas of Canada and watershed maps available on the Washington State government website.
+The watersheds included in our model were Fraser (which represents approximately 44\% of the freshwater input into the region), Skagit (12\%), East Vancouver Island (North and South) (12\%), Howe (7\%), Bute (7\%), Puget (6\%), Juan de Fuca (5\%), Jervis (4\%) and Toba (3\%).
+
+The monthly flow from each river was input as a point source in the 3 grid points closest to the surface at the model point closest to the mouth of each river.
+Incoming water was assumed to be fresh, with a temperature of ??
+degrees C.
+A total of 150 rivers were parametrised by this method.
 
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{plainnat}
 \bibliography{ref}

### 2014-01-21 RoughWork/manuscript.tex
Added one sentence about tides :)

@@
 % Kate: A description of the tides and river forcing can be included here. 
-\subsection{River forcing}
+\subsection{Tidal forcing} %Kate
+The model was forced by tidal elevations and currents at the Juan de Fuca and Johnstone Strait boundaries. To be continued...
+
+\subsection{River forcing} %Kate
 River input provides a significant volume of freshwater to the Salish Sea and can influence stratification, circulation and primary productivity.
@@
 \cite{morrison2011rivers} provides a method for estimating freshwater runoff in the Salish Sea region based on precipitation.
-Exact data were acquired from \cite{morrison2011rivers}, including the runoff volumes for each watershed for each year from 1970 to 2012, as well as monthly averages.
+Monthly runoff volumes for each watershed for each year from 1970 to 2012 were acquired from \cite{morrison2011rivers}, as well as monthly averages.
 
@@
 
-The monthly flow from each river was input as a point source in the 3 grid points closest to the surface at the model point closest to the mouth of each river.
+The monthly flow from each river was input as a point source in the three grid points closest to the surface at the model point closest to the mouth of each river.
 Incoming water was assumed to be fresh, with a temperature of ??
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{plain}
 \bibliography{ref}

### 2014-01-24 RoughWork/manuscript.tex
Added some tide set up stuff

@@
 
-\section{Model Evaluation}\label{sec:model}
+\section{Model Configuration}
 %A section about how we configured the model for SoG 
-%To be inlcluded: NEMO overview, tides, rivers, bathymetry, atmospheric forcing, vertical/lateral mixing, boundary conditions, grid.
-
 % Kate: A description of the tides and river forcing can be included here. 
 \subsection{Tidal forcing} %Kate
-The model was forced by tidal elevations and currents at the Juan de Fuca and Johnstone Strait boundaries. To be continued...
+The model was forced by tidal elevations and currents at the Juan de Fuca and Johnstone Strait boundaries.
+Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on \citep{foreman00webtide}.
+The Johnstone Strait boundary was forced with...?
 
@@
 However, most rivers in the domain are not gauged so parameterisations were required to represent river flow.
-\cite{morrison2011rivers} provides a method for estimating freshwater runoff in the Salish Sea region based on precipitation.
-Monthly runoff volumes for each watershed for each year from 1970 to 2012 were acquired from \cite{morrison2011rivers}, as well as monthly averages.
+\citep{morrison2011rivers} provides a method for estimating freshwater runoff in the Salish Sea region based on precipitation.
+Monthly runoff volumes for each watershed for each year from 1970 to 2012 were acquired from \citep{morrison2011rivers}, as well as monthly averages.
 
@@
 degrees C.
-A total of 150 rivers were parametrised by this method.
+A total of 150 rivers were parameterised by this method.
+
+\section{Model Evaluation}\label{sec:model}
+%To be included: NEMO overview, tides, rivers, bathymetry, atmospheric forcing, vertical/lateral mixing, boundary conditions, grid. 
+\subsection{Tidal evaluation}
+The model was initially evaluated by comparing modelled harmonic constituents to measured harmonic constituents at tidal measuring stations throughout the domain.
+Comparisons were made using the complex difference, defined by \cite{foreman95tidal} as:
+
+To be continued!
 

### 2014-01-24 RoughWork/manuscript.tex
Added more details about tidal eval

@@
 \citep{masson2004modelling} %testing a citation
+%To be included: NEMO overview, tides, rivers, bathymetry, atmospheric forcing, vertical/lateral mixing, boundary conditions, grid.
 
@@
 %A section about how we configured the model for SoG 
-% Kate: A description of the tides and river forcing can be included here. 
-\subsection{Tidal forcing} %Kate
+\subsection{Model domain}
+
+\subsection{Boundary conditions}
+
+\subsubsection{Tidal forcing} %Kate
 The model was forced by tidal elevations and currents at the Juan de Fuca and Johnstone Strait boundaries.
-Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on \citep{foreman00webtide}.
+Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on \citep{foreman2000webtide}.
 The Johnstone Strait boundary was forced with...?
+(e.g.
+we might use tidal harmonics measured and calculated by \citep{thomson1980johnstone})
+
+\subsubsection{Temperature and salinity}
+Temperature and salinity at the Juan de Fuca boundary were taken from a climatology which was created from observations collected by (which Masson ref?) over the period of 19??-2004?.
+
+\subsection{Initial conditions}
 
@@
 
-Freshwater runoff from each watershed was divided between the rivers in that watershed, based on estimations of the area that each river drained in its catchment.
+Freshwater runoff from each watershed was divided between the rivers in that watershed.
 The area drained by each river was estimated from Toporama maps by the Atlas of Canada and watershed maps available on the Washington State government website.
-The watersheds included in our model were Fraser (which represents approximately 44\% of the freshwater input into the region), Skagit (12\%), East Vancouver Island (North and South) (12\%), Howe (7\%), Bute (7\%), Puget (6\%), Juan de Fuca (5\%), Jervis (4\%) and Toba (3\%).
+The watersheds included in our model were Fraser (which represents approximately 44\% of the freshwater input into our domain), Skagit (12\%), East Vancouver Island (North and South) (12\%), Howe (7\%), Bute (7\%), Puget (6\%), Juan de Fuca (5\%), Jervis (4\%) and Toba (3\%).
 
@@
 
+%we could show the location of all the rivers on the overall location map, or on a picture of our domain, since the rivers are on the edges of the domain so they won't block any bathymetry information
+
+\subsection{Atmospheric forcing}
+
 \section{Model Evaluation}\label{sec:model}
-%To be included: NEMO overview, tides, rivers, bathymetry, atmospheric forcing, vertical/lateral mixing, boundary conditions, grid. 
+
 \subsection{Tidal evaluation}
-The model was initially evaluated by comparing modelled harmonic constituents to measured harmonic constituents at tidal measuring stations throughout the domain.
-Comparisons were made using the complex difference, defined by \cite{foreman95tidal} as:
+The model was initially evaluated qualitatively by comparing patterns of tidal amplitude and phase to results from \citep{foreman1995tidal}.
+For example, the amphidromic dome around Victoria was produced in the $M_2$ results, as well as the monotonic decrease in $M_2$ amplitude moving northwards along the Strait of Georgia.
+Initially, the Johnstone Strait boundary was closed, but modelled $M_2$ amplitudes were too small compared to measured amplitudes...
+TBC
 
-To be continued!
+%perhaps a nice contour map of M2 and K1 amps and phases goes here?
+
+Once our model was reproducing observed tidal patterns, model results were quantitatively evaluated by comparing modelled harmonic constituents to measured harmonic constituents at tidal measuring stations throughout the domain.
+Comparisons were made using the complex difference (D), defined by \citep{foreman1995tidal} as:
+
+\begin{equation}
+D = [(A_0 \cos g_0 - A_m \cos g_m)^2 + (A_0 \sin g_0 - A_m \sin g_m)^2]^{1/2}
+\end{equation}
+where $A_0$, $A_m$, $g_0$ and $g_m$ are observed and modelled amplitudes and phases.
+
+%perhaps a table of complex differences at tidal stations similar to Table 1 of Foreman et al (1995)  (such as the one produced by tidetools.calc_diffs_meas_mod) here?
+%(would be cool to include the complex differences calculated at the VENUS nodes too)
+
+Complex differences were less than ??cm at all stations in our domain, which was assumed to be acceptable for our purposes. 
+%if it's favourable, we could compare our complex differences to Foreman et al (1995), who got an average of D=3cm for M2 and D=2.5cm for K1
 

### 2014-01-31 RoughWork/manuscript.tex
Added information about the open boundaries, values and conditions.

@@
 \subsubsection{Temperature and salinity}
-Temperature and salinity at the Juan de Fuca boundary were taken from a climatology which was created from observations collected by (which Masson ref?) over the period of 19??-2004?.
+Temperature and salinity at the Juan de Fuca boundary were taken from a weekly climatology which was created from results from a model covering the Salish Sea and the west coast of Vancouver Island \citep{massonfine2012}.
+Their results, originally on s-levels were interpolated onto z-levels and then onto the NEMO horizontal grid.
+To prepare the climatology all years (1995-2008) were averaged and results, approximately every 15 days, were interpolated to a weekly climatology.
 
-\subsection{Initial conditions}
+\subsubsection{Sea Surface Height}
+Sea surface height at the mouth of Juan de Fuca was set using values from the Tofino tide guage.
+A monthly climatology was produced using daily averages from 2000-2010, binning them by month, averaging and setting the yearly mean to zero.
+For the storm surge simulations, hourly variations in sea surface height were used.
+These values are the Tofino tide guage values, de-tided and with the zero reset as for the climatology.
+
+\subsubsection{Open Boundary Conditions}
+
+The model relaxed to the forced temperature and salinity over the 10 grid points (about 5km) closest to the open boundaries, using the NEMO FRS scheme \cite.
+The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \cite  The baroclinic velocities at the boundary were set to be equal to the values inside the boundary (zero-gradient boundary conditions).
+This scheme is not part of core NEMO.
+Zero gradient conditions were chose because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain.
 
@@
 \subsection{Atmospheric forcing}
+
+\subsection{Initial conditions}
+Initial conditions for temperature and salinity were taken from a CTD cast in the middle Strait of Georgia taken in Sept 2002 \citep{pawlowiczetal2007}.
+Conditions were initially uniform horizontally.
+Velocity was initialized at zero.
+
+\subsection{Spin-up}
+The model was spun up for a 15.5 months from the initial conditions above, starting Sep 16, 2002, using atmospheric forcing from 2002-2003, climatological temperature and salinity and sea surface height at the boundaries, with tides and climatological river output.
+All storm surge runs were started.....
 

### 2014-03-27 RoughWork/manuscript.tex
Added some references to bib file and some description of model domain. Bathy image too.

@@
 
-\section{Model Configuration}
+\section{Model Configuration}\label{sec:config}
 %A section about how we configured the model for SoG 
 \subsection{Model domain}
+The modelled domain extends from the Strait of Juan de Fuca to Puget Sound to Johnstone Strait as shown in Figure \ref{fig:domain}.
+Bathymetry from the Cascadia physiography dataset \cite{?} was smoothed to limit the difference in depth across grid cells.
+For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was performed to achieve constant depth across the first ten grid cells.
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.5]{Figures/bathy.png}
+\caption{Model domain including bathymetry, rivers (*), and storm surge locations (o) of interest.}\label{fig:domain}
+\end{figure}
 
 \subsection{Boundary conditions}
+%lateral and bottom friction, open boundaries, eta surface
 
-\subsubsection{Tidal forcing} %Kate
+\subsubsection{Tidal forcing} 
 The model was forced by tidal elevations and currents at the Juan de Fuca and Johnstone Strait boundaries.

### 2014-03-31 RoughWork/manuscript.tex
Added some references and description of ssh forcing.

@@
 The modelled domain extends from the Strait of Juan de Fuca to Puget Sound to Johnstone Strait as shown in Figure \ref{fig:domain}.
-Bathymetry from the Cascadia physiography dataset \cite{?} was smoothed to limit the difference in depth across grid cells.
-For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was performed to achieve constant depth across the first ten grid cells.
+Bathymetry from the Cascadia physiography dataset \citep{haugerud1999digital} was smoothed to limit the difference in depth across grid cells.
+For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells.
 
@@
 The model was forced by tidal elevations and currents at the Juan de Fuca and Johnstone Strait boundaries.
-Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on \citep{foreman2000webtide}.
-The Johnstone Strait boundary was forced with...?
-(e.g.
-we might use tidal harmonics measured and calculated by \citep{thomson1980johnstone})
+Tidal heights and currents for eight tidal constituents (K1 ,O1, P1, Q1, M2, K2, N2, S2) at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on work by \citet{foreman2000webtide}.
+The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major M2 and K1 constituents.
+Additionally, O1 and S2 elevation harmonics from their measuresments were employed.
+The remaining consituents were extrapolated from Webtide.
 
@@
 
-The model relaxed to the forced temperature and salinity over the 10 grid points (about 5km) closest to the open boundaries, using the NEMO FRS scheme \cite.
-The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \cite  The baroclinic velocities at the boundary were set to be equal to the values inside the boundary (zero-gradient boundary conditions).
+The model relaxed to the forced temperature and salinity over the 10 grid points (about 5km) closest to the open boundaries, using the NEMO FRS scheme \citep{madec2008nemo}. %other references>
+The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \citet{madec2008nemo}. %other references?
+The baroclinic velocities at the boundary were set to be equal to the values inside the boundary (zero-gradient boundary conditions).
 This scheme is not part of core NEMO.
-Zero gradient conditions were chose because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain.
+Zero gradient conditions were chosen because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain.
 
-\subsection{River forcing} %Kate
+\subsection{River forcing}
 River input provides a significant volume of freshwater to the Salish Sea and can influence stratification, circulation and primary productivity.
 However, most rivers in the domain are not gauged so parameterisations were required to represent river flow.
-\citep{morrison2011rivers} provides a method for estimating freshwater runoff in the Salish Sea region based on precipitation.
-Monthly runoff volumes for each watershed for each year from 1970 to 2012 were acquired from \citep{morrison2011rivers}, as well as monthly averages.
+\citet{morrison2011rivers} provides a method for estimating freshwater runoff in the Salish Sea region based on precipitation.
+Monthly runoff volumes for each watershed for each year from 1970 to 2012 were acquired from \citet{morrison2011rivers}, as well as monthly averages.
 
@@
 The monthly flow from each river was input as a point source in the three grid points closest to the surface at the model point closest to the mouth of each river.
-Incoming water was assumed to be fresh, with a temperature of ??
-degrees C.
+Incoming water was assumed to be fresh and at the temperature of the surface.
 A total of 150 rivers were parameterised by this method.
@@
 The model was spun up for a 15.5 months from the initial conditions above, starting Sep 16, 2002, using atmospheric forcing from 2002-2003, climatological temperature and salinity and sea surface height at the boundaries, with tides and climatological river output.
-All storm surge runs were started.....
+All storm surge runs were started three days prior to the event of interest with zero initial velocities and sea surface height and a stratification profile from model spin up.
+The modelled sea surface height adjusted to forcing in less than one day.
 
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{chicago}
 \bibliography{ref}

### 2014-04-03 RoughWork/manuscript.tex
Added description of grid and coastal/bottom bounday conditions.

@@
 For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells.
+As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computationally efficency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
+
+The rectangular numerical grid is divided into 900 by 400 by 40 grid cells, which results in a horizontal resolution of about 400 m by 500 m.
+The 40 vertical $z$-levels are stretched gradually in order to achieve higher resolution in the surface layer, with approximately 1 m vertical grid spacing down to 25 m (?) in depth.
+At the bottom boundary, partial $z$-levels are employed to limit large changes in bathymetry across grid cells \citep{madec2008nemo}.
 
@@
 %lateral and bottom friction, open boundaries, eta surface
+The model includes two open boundaries that connect to the Pacific Ocean, the western boundary of the Strait of Juan de Fucan as well as Johnstone Strait at the north, both of which are forced with eight tidal constituents, temperature and salinity climatologies, and the sea surface height anomaly.
+The details of these forcing conditions are described below.
+At coastal boundaries, the partial slip paramertrization implemented by NEMO \citep{madec2008nemo} is employed.
+At the ocean surface, atmospheric conditions and fresh water input from rivers are included and described in more detail below.
+Bottom friction is evaluated through a quadratic parametrization \citep{madec2008nemo} with a drag coefficient $C_D = 5\times 10^{-3}$.
+
+The free surface is evaluated .. (description of split-explicit and nonlinear free surface scheme here)
 

### 2014-04-23 RoughWork/manuscript.tex
Added description of NEMO, SSH equation, eddy viscosities.

--- 2014-03-31 RoughWork/manuscript.tex
@@
-\documentclass[pdftex,12pt]{article}
+\documentclass[pdftex,10pt]{article}
 \usepackage{fullpage}
@@
 %A section about how we configured the model for SoG 
+We have used the Nucleus for European Modelling of the Ocean (NEMO) framework in its regional configuration to develop an ocean model for the Strait of Georgia and Salish Sea.
+NEMO is a highly modularized tool used for studying ocean physics, ocean-ice interactions, and the biogeochemical properties of the ocean.
+NEMO's ocean core solves the three-dimensional hydrostatic equations of motion for an incompressible fluid under the Boussinesq approximation on a structured computational grid.
+Although not used in the present work, NEMO's options for grid nesting and biogeochemical coupling make it a useful tool for studying the complex physics and biogeochemical interactions within the Strait of Georgia.
+This work focuses on validating the physical set up of the Salish Sea model, in particular, determining appropriate forcing and boundary conditions for accurate reproduction of tidal amplitudes and phases as well as storm surge elevations.
+Future work will include biogeochemical coupling and data assimilation.
+
 \subsection{Model domain}
@@
 For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells.
+As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
+
+The curvilinear orthogonal numerical grid is divided into 400 by 900 by 40 grid cells, which results in an almost uniform horizontal resolution with grid spacing approximately 400 m by 500 m.
+The 40 vertical $z$-levels are stretched gradually in order to achieve higher resolution in the surface layer, with approximately 1 m vertical grid spacing down to about 20 m in depth.
+At the bottom boundary, partial $z$-levels are utilized in order to limit large changes in bathymetry across grid cells \citep{madec2008nemo}.
 
@@
 
-\subsection{Boundary conditions}
+\subsection{Boundary conditions and subgrid scale processes}
 %lateral and bottom friction, open boundaries, eta surface
+The model includes two open boundaries that connect to the Pacific Ocean, the western boundary of the Strait of Juan de Fuca as well as Johnstone Strait at the north, both of which are forced with eight tidal constituents, temperature and salinity climatologies, and the sea surface height anomaly.
+The details of these forcing conditions are described below.
+At coastal boundaries, the partial slip boundary condition, an approximation of the no slip boundary condition, is used.
+The partial slip boundary condition allows one to include the frictional effects of lateral boundaries without the restricitve resolution required to represent the lateral boundary layer under no slip conditions.
+A lateral eddy viscosity of 20 m $^2$ s $^{-1}$ parameterizes horizontal friction and the lateral eddy diffusivity is 20.5 m $^2$ s $^{-1}$ .
+At the ocean surface, meterological conditions and fresh water input from rivers are included and described in more detail below.
+Bottom friction is represented by a quadratic law for the bottom momentum flux with drag coefficient $C_D = 5\times 10^{-3}$.
+Vertical turbulence and mixing is calculated through the $k-\epsilon$ configuration of the generic length scale (GLS) turbulence closure \citep{umlauf2003generic} with background vertical eddy viscosity and diffusivity set to $1\times10^{-4}$ and $1\times10^{-5}$ m$^2$ s$^{-1}$ respectively.
+Details on the NEMO implementation of the partial slip lateral boundary condition, quadratic bottom friction law, and GLS turbulence closure scheme are provided by \citep{madec2008nemo}.
+
+In addition to the equations of motion, a prognostic equation for the sea surface height is solved at each time step.
+The inclusion the sea surface height equation requires a fairly restrictive time step due to the presence of high speed surface gravity waves.
+As such, the split-explicit time stepping algorithm is employed, where the free surface and barotropic equations are solved with a smaller time step than that used for the other variables.
+The model time step and barotropic time step are 10 s and 2 s respectively.
 
@@
 The monthly flow from each river was input as a point source in the three grid points closest to the surface at the model point closest to the mouth of each river.
-Incoming water was assumed to be fresh and at the temperature of the surface.
+Incoming water was assumed to be fresh and at surface temperatyre.
 A total of 150 rivers were parameterised by this method.
@@
 Velocity was initialized at zero.
+%Merge with below?
+%Perhaps image of spun-up stratifiction along Thalweg? Here or in model evaluation
 

### 2014-04-24 RoughWork/manuscript.tex
Updated descritpion of vertical grid spacing.

@@
 The curvilinear orthogonal numerical grid is divided into 400 by 900 by 40 grid cells, which results in an almost uniform horizontal resolution with grid spacing approximately 400 m by 500 m.
-The 40 vertical $z$-levels are stretched gradually in order to achieve higher resolution in the surface layer, with approximately 1 m vertical grid spacing down to about 20 m in depth.
+The 40 vertical $z$-levels are stretched gradually in order to achieve higher resolution in the surface layer, with 1 m vertical grid spacing down to about 10 m in depth.
+Below 10m the grid is stretched gradually with a maxmimum grid spacing of 27 m at the lowest layer.
 At the bottom boundary, partial $z$-levels are utilized in order to limit large changes in bathymetry across grid cells \citep{madec2008nemo}.

### 2014-04-24 RoughWork/manuscript.tex
Fixed typos

@@
 For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells.
-As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
+As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
 
@@
 The 40 vertical $z$-levels are stretched gradually in order to achieve higher resolution in the surface layer, with 1 m vertical grid spacing down to about 10 m in depth.
-Below 10m the grid is stretched gradually with a maxmimum grid spacing of 27 m at the lowest layer.
+Below 10m the grid is stretched gradually with a maximum grid spacing of 27 m at the lowest layer.
 At the bottom boundary, partial $z$-levels are utilized in order to limit large changes in bathymetry across grid cells \citep{madec2008nemo}.
@@
 
-\subsection{Boundary conditions and subgrid scale processes}
+\subsection{Boundary conditions and sub-grid scale processes}
 %lateral and bottom friction, open boundaries, eta surface
@@
 At coastal boundaries, the partial slip boundary condition, an approximation of the no slip boundary condition, is used.
-The partial slip boundary condition allows one to include the frictional effects of lateral boundaries without the restricitve resolution required to represent the lateral boundary layer under no slip conditions.
-A lateral eddy viscosity of 20 m $^2$ s $^{-1}$ parameterizes horizontal friction and the lateral eddy diffusivity is 20.5 m $^2$ s $^{-1}$ .
-At the ocean surface, meterological conditions and fresh water input from rivers are included and described in more detail below.
+The partial slip boundary condition allows one to include the frictional effects of lateral boundaries without the restrictive resolution required to represent the lateral boundary layer under no slip conditions.
+A lateral eddy viscosity of 20 m $^2$ s $^{-1}$ parametrizes horizontal friction and the lateral eddy diffusivity is 20.5 m $^2$ s $^{-1}$ .
+At the ocean surface, meteorological conditions and fresh water input from rivers are included and described in more detail below.
 Bottom friction is represented by a quadratic law for the bottom momentum flux with drag coefficient $C_D = 5\times 10^{-3}$.
@@
 The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major M2 and K1 constituents.
-Additionally, O1 and S2 elevation harmonics from their measuresments were employed.
-The remaining consituents were extrapolated from Webtide.
+Additionally, O1 and S2 elevation harmonics from their measurements were employed.
+The remaining constituents were extrapolated from Webtide.
 
@@
 \subsubsection{Sea Surface Height}
-Sea surface height at the mouth of Juan de Fuca was set using values from the Tofino tide guage.
+Sea surface height at the mouth of Juan de Fuca was set using values from the Tofino tide gauge.
 A monthly climatology was produced using daily averages from 2000-2010, binning them by month, averaging and setting the yearly mean to zero.
 For the storm surge simulations, hourly variations in sea surface height were used.
-These values are the Tofino tide guage values, de-tided and with the zero reset as for the climatology.
+These values are the Tofino tide gauge values, de-tided and with the zero reset as for the climatology.
 
@@
 
-The model relaxed to the forced temperature and salinity over the 10 grid points (about 5km) closest to the open boundaries, using the NEMO FRS scheme \citep{madec2008nemo}. %other references>
-The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \citet{madec2008nemo}. %other references?
+The model relaxed to the forced temperature and salinity over the 10 grid points (about 5km) closest to the open boundaries, using the NEMO flow relaxation scheme \citep{madec2008nemo}.
+%other references>
+The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \citep{madec2008nemo}. %other references?
 The baroclinic velocities at the boundary were set to be equal to the values inside the boundary (zero-gradient boundary conditions).
@@
 The monthly flow from each river was input as a point source in the three grid points closest to the surface at the model point closest to the mouth of each river.
-Incoming water was assumed to be fresh and at surface temperatyre.
+Incoming water was assumed to be fresh and at surface temperature.
 A total of 150 rivers were parameterised by this method.
@@
 %Merge with below?
-%Perhaps image of spun-up stratifiction along Thalweg? Here or in model evaluation
+%Perhaps image of spun-up stratification along Thalweg? Here or in model evaluation
 
@@
 
-\section{Storm Surge Hindcasts}\label{sec:storm}
+\section{Storm Surge Hind casts}\label{sec:storm}
 

### 2014-05-06 RoughWork/manuscript.tex
Added description of CGRF.

@@
 These values are the Tofino tide gauge values, de-tided and with the zero reset as for the climatology.
+The storm surge simulations also used the sea surface height anomaly from Port Hardy, forced at the northern boundary in Johnstone Strait.
 
@@
 \subsection{Atmospheric forcing}
+The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforcasting model suitable for use in ocean modelling \citep{smith2013new}.
+Forecasts from the period of 2002-2011 are available.
+Additionally, the inverse barometric effect of the atmospheric pressure is included, an important consideration in storm surge modelling.
 

### 2014-05-27 RoughWork/manuscript.tex
Outline of intro.

@@
 \section{Introduction}\label{sec:intro}
-\citep{masson2004modelling} %testing a citation
+
+%some general intro paragraph here
+
+%my attempt at physical oceanopgraphy SoG- Susan feel free to edit.
+The Strait of Georgia is a strongly stratified, semi-enclosed body of water located between Vancouver Island and the mainland of British Columbia.
+It is part of a larger system of waterways collectively known as the Salish Sea and is connected to the Pacific Ocean via the Strait of Juan de Fuca and Johnstone Strait.
+Several features related to the physical oceanopgraphy of this region render it difficut to model numerically.
+The outlfow from the Fraser River in the late spring is the primary source fresh water within the Strait of Georgia, and leads to large density gradients in the summer months.
+Additionally, strong tidal mixing through the San Juan and Gulf Islands produces horizontal density gradients that separate the saline waters of the Strait of Juan de Fuca and the fresher water of the Strait of Georgia.
+It is important to accurately represent the vertical mixing in this region as it sets the rate of export of fresh water through estuarine circulation.
+The strong and fast tidal currents through the narrow channels of the north, such as Discovery Passage and Johnstone Strait, pose a particular challenge for numerical models.
+Yet, several modelling efforts for this region exist, ranging in choices of grid structure, mixing parametrizations, and domain size and extent.
+
+%overview oF SoG/Salish Sea models
+
+%intro to storm surges
+Many coastal communities in the Strait of Georgia are at risk to flooding and property damaged caused by storm surges, a natural hazard arising from the combination of a strong storm and high tide.
+The low atmospheric pressure associated with a storm acts as an inverse barometer elevating the sea level.
+This effect in combination with strong winds pushing water up against the coast can cause flooding, particularly if the storm occurs during an unusually high tide.
+Additionally, there is a small contribution to increased sea level due to the thermal expansion of water during warmer years set by the El Nino Southern Oscillation (ENSO) \citep{abeys2011extreme}.
+%more on past events, significance
+
+%storm surge modelling
+
+%overview of paper: model configuration, model evaluation, storm surge hindcasts.
+
+\section{Model Configuration}\label{sec:config}
 %To be included: NEMO overview, tides, rivers, bathymetry, atmospheric forcing, vertical/lateral mixing, boundary conditions, grid.
 
-\section{Model Configuration}\label{sec:config}
 %A section about how we configured the model for SoG 
@@
 
-The curvilinear orthogonal numerical grid is divided into 400 by 900 by 40 grid cells, which results in an almost uniform horizontal resolution with grid spacing approximately 400 m by 500 m.
+The curvilinear orthogonal numerical grid is divided into 398 by 898 by 40 grid cells, which results in an almost uniform horizontal resolution with grid spacing approximately 440 m by 500 m.
 The 40 vertical $z$-levels are stretched gradually in order to achieve higher resolution in the surface layer, with 1 m vertical grid spacing down to about 10 m in depth.
@@
 \end{figure}
+%image needs some work: colour? labels for JdF, SoG, cities...
 
@@
 The details of these forcing conditions are described below.
-At coastal boundaries, the partial slip boundary condition, an approximation of the no slip boundary condition, is used.
+At coastal boundaries, the partial slip boundary condition, an approximation to no slip, is used.
 The partial slip boundary condition allows one to include the frictional effects of lateral boundaries without the restrictive resolution required to represent the lateral boundary layer under no slip conditions.
@@
 Vertical turbulence and mixing is calculated through the $k-\epsilon$ configuration of the generic length scale (GLS) turbulence closure \citep{umlauf2003generic} with background vertical eddy viscosity and diffusivity set to $1\times10^{-4}$ and $1\times10^{-5}$ m$^2$ s$^{-1}$ respectively.
-Details on the NEMO implementation of the partial slip lateral boundary condition, quadratic bottom friction law, and GLS turbulence closure scheme are provided by \citep{madec2008nemo}.
+Details on the NEMO implementation of the partial slip lateral boundary condition, quadratic bottom friction law, and GLS turbulence closure scheme are provided by \citet{madec2008nemo}.
 
@@
 \subsubsection{Tidal forcing} 
-The model was forced by tidal elevations and currents at the Juan de Fuca and Johnstone Strait boundaries.
-Tidal heights and currents for eight tidal constituents (K1 ,O1, P1, Q1, M2, K2, N2, S2) at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on work by \citet{foreman2000webtide}.
+The model was forced by eight tidal constituents (K1 ,O1, P1, Q1, M2, K2, N2, S2) at the Juan de Fuca and Johnstone Strait boundaries.
+Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on work by \citet{foreman2000webtide}.
 The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major M2 and K1 constituents.
@@
 
-The model relaxed to the forced temperature and salinity over the 10 grid points (about 5km) closest to the open boundaries, using the NEMO flow relaxation scheme \citep{madec2008nemo}.
-%other references>
-The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \citep{madec2008nemo}. %other references?
+The model is relaxed to the forced temperature and salinity over the 10 grid points (about 5km) closest to the open boundaries, using a flow relaxation scheme \citep{engedahl1995use}.
+%other references?
+The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \citep{flather1994storm, madec2008nemo}.
 The baroclinic velocities at the boundary were set to be equal to the values inside the boundary (zero-gradient boundary conditions).
@@
 The model was initially evaluated qualitatively by comparing patterns of tidal amplitude and phase to results from \citep{foreman1995tidal}.
-For example, the amphidromic dome around Victoria was produced in the $M_2$ results, as well as the monotonic decrease in $M_2$ amplitude moving northwards along the Strait of Georgia.
+For example, the amphidromic dome around Victoria was produced in the $M_2$ results, as well as the monotonic increase in $K_1$ amplitude moving northwards along the Strait of Georgia.
 Initially, the Johnstone Strait boundary was closed, but modelled $M_2$ amplitudes were too small compared to measured amplitudes...
@@
 
+\subsection{Stratification}
+
 \section{Storm Surge Hind casts}\label{sec:storm}

### 2014-06-27 RoughWork/manuscript.tex
Added acknowledgements and paper outline

@@
 It is part of a larger system of waterways collectively known as the Salish Sea and is connected to the Pacific Ocean via the Strait of Juan de Fuca and Johnstone Strait.
-Several features related to the physical oceanopgraphy of this region render it difficut to model numerically.
-The outlfow from the Fraser River in the late spring is the primary source fresh water within the Strait of Georgia, and leads to large density gradients in the summer months.
+Several features related to the physical oceanopgraphy of this region render it a challenge to model numerically.
+The outlfow from the Fraser River in the late spring is the primary source fresh water within the Strait of Georgia and leads to large vertical density gradients in the summer months.
 Additionally, strong tidal mixing through the San Juan and Gulf Islands produces horizontal density gradients that separate the saline waters of the Strait of Juan de Fuca and the fresher water of the Strait of Georgia.
@@
 %intro to storm surges
-Many coastal communities in the Strait of Georgia are at risk to flooding and property damaged caused by storm surges, a natural hazard arising from the combination of a strong storm and high tide.
+Many coastal communities in the Strait of Georgia are at risk to flooding and property damaged caused by storm surges, a natural hazard arising from the combination of a strong wind storm and high tide.
 The low atmospheric pressure associated with a storm acts as an inverse barometer elevating the sea level.
@@
 %overview of paper: model configuration, model evaluation, storm surge hindcasts.
+This paper is organized as follows: an overivew of the model configuration and domain including a description of open and surface boundary conditions is provided in section \ref{sec:config}.
+This section also describes the ocean model we have employed as well as choices in grid structure and mixing parameterizations.
+Next, section \ref{sec:model} presents an evaluation of model performance through comparisons of tidal amplitudes and phases with observations.
+The model's accuracy in producing storm surge hind casts is also asses in section \ref{sec:storm}.
+Finally, a summary and discussion of future research goals is provided in section \ref{sec:diss}.
 
@@
 The 40 vertical $z$-levels are stretched gradually in order to achieve higher resolution in the surface layer, with 1 m vertical grid spacing down to about 10 m in depth.
-Below 10m the grid is stretched gradually with a maximum grid spacing of 27 m at the lowest layer.
+Below 10m the grid is stretched gradually to a maximum grid spacing of 27 m at the lowest layer.
 At the bottom boundary, partial $z$-levels are utilized in order to limit large changes in bathymetry across grid cells \citep{madec2008nemo}.
@@
 The partial slip boundary condition allows one to include the frictional effects of lateral boundaries without the restrictive resolution required to represent the lateral boundary layer under no slip conditions.
-A lateral eddy viscosity of 20 m $^2$ s $^{-1}$ parametrizes horizontal friction and the lateral eddy diffusivity is 20.5 m $^2$ s $^{-1}$ .
+A lateral eddy viscosity of 20 m $^2$ s $^{-1}$ parametrizes horizontal friction and a lateral eddy diffusivity of 20.5 m $^2$ s $^{-1}$ is used.
 At the ocean surface, meteorological conditions and fresh water input from rivers are included and described in more detail below.
@@
 River input provides a significant volume of freshwater to the Salish Sea and can influence stratification, circulation and primary productivity.
-However, most rivers in the domain are not gauged so parameterisations were required to represent river flow.
+However, most rivers in the domain are not gauged so parameterizations were required to represent river flow.
 \citet{morrison2011rivers} provides a method for estimating freshwater runoff in the Salish Sea region based on precipitation.
@@
 \subsection{Stratification} 
+%outline: thalweg, plots of winter/summer stratty
 
 \section{Storm Surge Hind casts}\label{sec:storm}
+%outline: 
+%1. factors contributing to storm surge with focus on Feb 2006 event
+%2. Strong wind event but no surge (maybe Dec 2006 or Nov 2009)
+%3. High resolution wind (hopefully Dec 2012)
 
-\section{Conclusions}\label{sec:conclusions}
+\subsection{Factors Contributing to Storm Surges: Feb 4, 2006}
+
+\subsection{Strong wind event: }
+
+\subsection{High Resolution Atmospheric Forcing: Dec 17, 2012}
+
+\section{Discussion}\label{sec:diss}
+%future directions: running real time, biology and chemistry, model improvements?
+
+\section{Acknowledgements}\label{sec:ack}
+We would like to thank Michael Foreman, Diane Masson, and Luc Fillion for providing data used in model set up and evaluation.
+This project is funded by the Marine Environment Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada.
+%I'm likely missing a lot more (Keith, Youyu, Vasily, JP, Rich, Mark). Not sure about authorship but something to think about for later....
 

### 2014-07-03 RoughWork/manuscript.tex
Added template for table with tide comparisons
first storm surge figure

@@
 \usepackage{amsmath}
+\usepackage{epspdfconversion}
 
@@
 
+\begin{table}[h]
+\centering 
+\begin{tabular}{l c c c c c c} 
+\hline 
+& \multicolumn{3}{c}{$M_2$} & \multicolumn{3}{c}{$K_1$} \\ 
+\hline 
+Location   & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
+\hline 
+           &     &                      &          &     &                      &          \\
+\hline   
+Mean Error &     &                      &          &     &                      &          \\
+RMS Error  &     &                      &          &     &                      &          \\
+\hline  
+\end{tabular}
+\caption{Comparisons of modelled and observed amplitudes and phases.}
+\label{tab:comparison} 
+\end{table}
+%May have to include station numbers and locations on a map? Might be too busy with rivers too...
+%Should we use the new method of calculating tides?
+
 Complex differences were less than ??cm at all stations in our domain, which was assumed to be acceptable for our purposes. 
@@
 
+Forcing the model with only eight tidal constituents leads to an error in sea surface height predictions due to the omission of the next leading order tidal constituents.
+Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can be up to 40 cm (check) during times of high tidal range, and in particular, during the storm surge season.
+As such, storm surge model predictions were corrected by accounting for this error.
+
 \subsection{Factors Contributing to Storm Surges: Feb 4, 2006}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/feb2006.pdf}
+\caption{Comparison of observations and model output for the Feb.
+4, 2006 storm surge.
+Left: total water level observation (solid) and corrected model (dashed).
+Right: observed residuals (solid) and modelled residuals (dashed).}
+\label{fig:feb2006}
+\end{figure}
+%Note that colour may be an option here (no extra charge for A-O). But it will be worse for printing. Too much white space around image. Will fix later..
 

### 2014-07-03 RoughWork/manuscript.tex
Added description of feb2006 surge.

@@
 Next, section \ref{sec:model} presents an evaluation of model performance through comparisons of tidal amplitudes and phases with observations.
-The model's accuracy in producing storm surge hind casts is also asses in section \ref{sec:storm}.
+The model's accuracy in producing storm surge hindcasts is also assesed in section \ref{sec:storm}.
 Finally, a summary and discussion of future research goals is provided in section \ref{sec:diss}.
@@
 \subsubsection{Tidal forcing} 
-The model was forced by eight tidal constituents (K1 ,O1, P1, Q1, M2, K2, N2, S2) at the Juan de Fuca and Johnstone Strait boundaries.
+The model was forced by eight tidal constituents ($K_1$ ,$O_1$, $P_1$, $Q_1$, $M_2$, $K_2$, $N_2$, $S_2$) at the Juan de Fuca and Johnstone Strait boundaries.
 Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on work by \citet{foreman2000webtide}.
-The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major M2 and K1 constituents.
-Additionally, O1 and S2 elevation harmonics from their measurements were employed.
+The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major $M_2$ and $K_1$ constituents.
+Additionally, $O_1$ and $S_2$ elevation harmonics from their measurements were employed.
 The remaining constituents were extrapolated from Webtide.
@@
 
-\section{Storm Surge Hind casts}\label{sec:storm}
+\section{Storm Surge Hindcasts}\label{sec:storm}
 %outline: 
@@
 
-Forcing the model with only eight tidal constituents leads to an error in sea surface height predictions due to the omission of the next leading order tidal constituents.
-Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can be up to 40 cm (check) during times of high tidal range, and in particular, during the storm surge season.
+Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height predictions due to the omission of the next leading order tidal constituents such as $J_1$, $SA$, and $MM$.
+Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can be up to 40 cm (check) during times of high tidal range, and in particular, during the storm surge season Nov-Feb.
 As such, storm surge model predictions were corrected by accounting for this error.
+Additionally, since the sea surface height is calculated about a zero reverence level, the modelled water level is determined by adding a long term average of mean sea level at the desired location to the modelled sea surface height.
+
+Model results have been compared with observations by calculating both the total modelled water level and the modelled residual.
+.
+The modelled residual is calculated as the difference in model output between a simulation with all forcing conditions and a simulation forced only with tides and rivers.
+The observed water level is taken from observations by Fisheries and Oceans Canada (citation) and the observed residual is the observations minus tidal predictions.
+Tidal predictions are calculated using t\_tide \cite{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this cae, from the year prior surge event on interest.
+
+Three significiant surge or wind events have been considered.
+First, a study of a large surge event on Feb 4, 2006 is presented.
+This event caused siginificant damage and flooding in Boundary Bay, a community south of Richmond.
+(citation) and has been chosen as a case study to demonstrate the model's skill in reproducing a storm surge and to quantitatively assess forcing conditions that are important in storm surge hindcasting.
+%% Trying to decide Dec 15, 2005 or Nov 2009. I think it's more sense to do Nov 2009 - strong winds but no surge. 
+Next, an extreme wind event which resulted in damage to property in Vancouver and surrounding areas is discussed.
+This event occured on Dec 15, 2006 but did not result in significant flooding such as that which occured on Feb 4, 2006 because the storm did not occur at high tide.
+This case was choosen to idenitify the model's capability in handling large wind events that do not result in high water levels.
+Finally, a case with high resolution atmospheric forcing for a recent surge on Dec 17, 2012 is discussed.
 
 \subsection{Factors Contributing to Storm Surges: Feb 4, 2006}
+
+Simulation was run from Feb1-Feb7.
+The maximum water level observed at Point Atkinson on this day was 5.49 m, which is 0.8 m higher than the predicted tides, highlighting the significance of this event as it resulted in a large anomaly and also reached water levels close to the highest level ever recorded (5.61 m).
+
+The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations and corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River.
+The timing of maximum water level matches well between the observations and the model at all four locations, the most siginificant contribution to high water level being from the tides.
+There is also agood match in amplitude of the water level maximum at Point Atkinson and Cambell River, with a relative error less than
+2\%.
+However, the error at Victoria and Patricia Bay is more siginificant, at 12\% and 8\% respectively, likely due to error in $M_2$ tidal amplitudes as discussed above.
+The precise location and maginitude of the $M_2$ amphidromic dome around Victoria is very sensitive to model parameters such as bottom friction.
+
+The error in modelled residuals at Victoria and Campbell River is also apparent in these figures.
+Most noticebale is underestimate in modelled residual throughout most of the time period.
+Additionally, there is a delay of about 2-3 hours in the timing of maximum residual at most locations and the  relative errors in maximum residual range from 5\% at Point Atkinson to 29\% at Patricia Bay.
+%Question: is it better to present relative errors or just difference?
 
@@
 \includegraphics[scale=0.6]{Figures/feb2006.pdf}
-\caption{Comparison of observations and model output for the Feb.
-4, 2006 storm surge.
+\caption{Comparison of observations and model output for the Feb 4, 2006 storm surge.
 Left: total water level observation (solid) and corrected model (dashed).
@@
 \end{figure}
-%Note that colour may be an option here (no extra charge for A-O). But it will be worse for printing. Too much white space around image. Will fix later..
+%Note that colour may be an option here (no extra charge for A-O). But it will be worse for printing.
 
-\subsection{Strong wind event: }
+\subsection{Strong wind event: Dec 15, 2006}
 

### 2014-07-05 RoughWork/manuscript.tex
Added comment aout surge statistics

@@
 The observed water level is taken from observations by Fisheries and Oceans Canada (citation) and the observed residual is the observations minus tidal predictions.
-Tidal predictions are calculated using t\_tide \cite{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this cae, from the year prior surge event on interest.
+Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this cae, from the year prior surge event on interest.
 
@@
 %Question: is it better to present relative errors or just difference?
+%I think Hal Ritchie's talk with the storm surges had some stats presented. Look this up...
 

### 2014-07-08 RoughWork/manuscript.tex
Added notes to include spatial extent of surge

@@
 \subsubsection{Tidal forcing} 
-The model was forced by eight tidal constituents ($K_1$ ,$O_1$, $P_1$, $Q_1$, $M_2$, $K_2$, $N_2$, $S_2$) at the Juan de Fuca and Johnstone Strait boundaries.
+The model was forced by eight tidal constituents (K$_1$ ,O$_1$, P$_1$, Q$_1$, M$_2$, K$_2$, N$_2$, S$_2$) at the Juan de Fuca and Johnstone Strait boundaries.
 Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on work by \citet{foreman2000webtide}.
-The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major $M_2$ and $K_1$ constituents.
-Additionally, $O_1$ and $S_2$ elevation harmonics from their measurements were employed.
+The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major M$_2$ and K$_1$ constituents.
+Additionally, O$_1$ and S$_2$ elevation harmonics from their measurements were employed.
 The remaining constituents were extrapolated from Webtide.
@@
 The model was initially evaluated qualitatively by comparing patterns of tidal amplitude and phase to results from \citep{foreman1995tidal}.
-For example, the amphidromic dome around Victoria was produced in the $M_2$ results, as well as the monotonic increase in $K_1$ amplitude moving northwards along the Strait of Georgia.
-Initially, the Johnstone Strait boundary was closed, but modelled $M_2$ amplitudes were too small compared to measured amplitudes...
+For example, the amphidromic dome around Victoria was produced in the M$_2$ results, as well as the monotonic increase in $K_1$ amplitude moving northwards along the Strait of Georgia.
+Initially, the Johnstone Strait boundary was closed, but modelled M$_2$ amplitudes were too small compared to measured amplitudes...
 TBC
@@
 Once our model was reproducing observed tidal patterns, model results were quantitatively evaluated by comparing modelled harmonic constituents to measured harmonic constituents at tidal measuring stations throughout the domain.
-Comparisons were made using the complex difference (D), defined by \citep{foreman1995tidal} as:
+Comparisons were made using the complex difference ($D$), defined by \citep{foreman1995tidal} as:
 
@@
 \hline 
-& \multicolumn{3}{c}{$M_2$} & \multicolumn{3}{c}{$K_1$} \\ 
+& \multicolumn{3}{c}{M$_2$} & \multicolumn{3}{c}{K$_1$} \\ 
 \hline 
@@
 \end{tabular}
-\caption{Comparisons of modelled and observed amplitudes and phases.}
+\caption{Comparisons of modelled and observed amplitudes and phases.
+$R$ is the ration of modelled to observed amplitude.
+$\Delta \phi$ is the difference between modelled and observed phase.
+$D$ is the complex difference.
+}
 \label{tab:comparison} 
@@
 %% Trying to decide Dec 15, 2005 or Nov 2009. I think it's more sense to do Nov 2009 - strong winds but no surge. 
-Next, an extreme wind event which resulted in damage to property in Vancouver and surrounding areas is discussed.
+Next, an extreme wind event which resulted in property damage in Vancouver and surrounding areas is discussed.
 This event occured on Dec 15, 2006 but did not result in significant flooding such as that which occured on Feb 4, 2006 because the storm did not occur at high tide.
-This case was choosen to idenitify the model's capability in handling large wind events that do not result in high water levels.
+This case was choosen to identify the model's capability in handling large wind events that do not result in high water levels.
 Finally, a case with high resolution atmospheric forcing for a recent surge on Dec 17, 2012 is discussed.
@@
 \subsection{Factors Contributing to Storm Surges: Feb 4, 2006}
+%Include spatial extend of surge
 
@@
 The timing of maximum water level matches well between the observations and the model at all four locations, the most siginificant contribution to high water level being from the tides.
-There is also agood match in amplitude of the water level maximum at Point Atkinson and Cambell River, with a relative error less than
-2\%.
-However, the error at Victoria and Patricia Bay is more siginificant, at 12\% and 8\% respectively, likely due to error in $M_2$ tidal amplitudes as discussed above.
-The precise location and maginitude of the $M_2$ amphidromic dome around Victoria is very sensitive to model parameters such as bottom friction.
+There is also a good match in amplitude of the water level maximum at Point Atkinson and Cambell River, with a relative error less than 2\%.
+However, the error at Victoria and Patricia Bay is more siginificant, at 12\% and 8\% respectively, likely due to error in M$_2$ tidal amplitudes as discussed above.
+The precise location and magnitude of the M$_2$ amphidrome around Victoria is very sensitive to model parameters such as bottom friction.
 
@@
 Most noticebale is underestimate in modelled residual throughout most of the time period.
-Additionally, there is a delay of about 2-3 hours in the timing of maximum residual at most locations and the  relative errors in maximum residual range from 5\% at Point Atkinson to 29\% at Patricia Bay.
+Additionally, there is a delay of about 2-3 hours in the timing of maximum residual and the relative errors in maximum residual range from 5\% at Point Atkinson to 29\% at Patricia Bay.
 %Question: is it better to present relative errors or just difference?
@@
 
-\subsection{Strong wind event: Dec 15, 2006}
+\subsection{Strong wind event: Dec 15, 2006}% or Nov 2009
 
@@
 We would like to thank Michael Foreman, Diane Masson, and Luc Fillion for providing data used in model set up and evaluation.
-This project is funded by the Marine Environment Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada.
+This project is funded by the Marine Environmental Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada.
 %I'm likely missing a lot more (Keith, Youyu, Vasily, JP, Rich, Mark). Not sure about authorship but something to think about for later....

### 2014-07-08 RoughWork/manuscript.tex
Added a reference

@@
 
-Simulation was run from Feb1-Feb7.
+On Feb 4, 2006, a strong storm system with winds over 75 km/hr coinciding with high tide led to siginificant flooding in coastal comminuties of the Strait of Georgia \citep{romanowski2010storm}.
 The maximum water level observed at Point Atkinson on this day was 5.49 m, which is 0.8 m higher than the predicted tides, highlighting the significance of this event as it resulted in a large anomaly and also reached water levels close to the highest level ever recorded (5.61 m).
+The model was run from Feb 1 to 8, 2006 under varying forcing conditions in order to determine the siginicance of each factor contributing to the surge.
 
@@
 
-The error in modelled residuals at Victoria and Campbell River is also apparent in these figures.
-Most noticebale is underestimate in modelled residual throughout most of the time period.
-Additionally, there is a delay of about 2-3 hours in the timing of maximum residual and the relative errors in maximum residual range from 5\% at Point Atkinson to 29\% at Patricia Bay.
+The error in modelled residuals at Victoria and Patricia is also apparent in the panels on the right, most noticebale is an underestimate in modelled residual throughout most of the time period.
+Additionally, there is a delay of about 2-3 hours in the timing of maximum residual at all four locations.
+The relative errors in maximum residual range from 5\% at Point Atkinson to 29\% at Patricia Bay.
 %Question: is it better to present relative errors or just difference?

### 2014-07-10 RoughWork/manuscript.tex
Added discussion on factors contributing to storm surge.

--- 2014-07-08 RoughWork/manuscript.tex
@@
 Several features related to the physical oceanopgraphy of this region render it a challenge to model numerically.
-The outlfow from the Fraser River in the late spring is the primary source fresh water within the Strait of Georgia and leads to large vertical density gradients in the summer months.
+The outflow from the Fraser River in the late spring is the primary source fresh water within the Strait of Georgia and leads to large vertical density gradients in the summer months.
 Additionally, strong tidal mixing through the San Juan and Gulf Islands produces horizontal density gradients that separate the saline waters of the Strait of Juan de Fuca and the fresher water of the Strait of Georgia.
@@
 \centering 
-\begin{tabular}{l c c c c c c} 
+\begin{tabular}{l l c c c c c c} 
 \hline 
-& \multicolumn{3}{c}{M$_2$} & \multicolumn{3}{c}{K$_1$} \\ 
+& & \multicolumn{3}{c}{M$_2$} & \multicolumn{3}{c}{K$_1$} \\ 
 \hline 
-Location   & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
+Station number & Station name & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
 \hline 
-           &     &                      &          &     &                      &          \\
+               &              &     &                      &          &     &                      &          \\
 \hline   
-Mean Error &     &                      &          &     &                      &          \\
-RMS Error  &     &                      &          &     &                      &          \\
+Mean Error     &              &     &                      &          &     &                      &          \\
+RMS Error      &              &     &                      &          &     &                      &          \\
 \hline  
@@
 %outline: thalweg, plots of winter/summer stratty
+%Maybe note this
 
@@
 %3. High resolution wind (hopefully Dec 2012)
-
-Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height predictions due to the omission of the next leading order tidal constituents such as $J_1$, $SA$, and $MM$.
+A series of storm surge hindcasts is presented next.
+Storm surge simulations were started from rest at least three days prior to the event of interest to allow for adjustment of model currents.
+Stratification conditions from model spin-up at a time close the the event (within 5 days) were chosen to initialize the temperature and salinity fields.
+Boundary conditions include atmospheric forcing and fresh water input from rivers at the surface, sea surface height anomaly from observations at the open boundary, and tidal forcing with eight tidal constituents.
+Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height predictions due to the omission of the next leading order tidal constituents such as J$_1$.
 Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can be up to 40 cm (check) during times of high tidal range, and in particular, during the storm surge season Nov-Feb.
 As such, storm surge model predictions were corrected by accounting for this error.
-Additionally, since the sea surface height is calculated about a zero reverence level, the modelled water level is determined by adding a long term average of mean sea level at the desired location to the modelled sea surface height.
+Additionally, since the sea surface height is calculated about a zero reverence level, the modelled water level is determined by adding a long term average of mean sea level at the specified location to the modelled sea surface height.
 
 Model results have been compared with observations by calculating both the total modelled water level and the modelled residual.
-.
 The modelled residual is calculated as the difference in model output between a simulation with all forcing conditions and a simulation forced only with tides and rivers.
@@
 Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this cae, from the year prior surge event on interest.
+Additionally, several simulations with varying forcing conditions such as removing atmospsheric forcing or removing the sea surface height anomaly at the open boundaries have been completed.
 
@@
 
-Simulation was run from Feb1-Feb7.
+On Feb 4, 2006, a strong storm system with winds over 75 km/hr coinciding with an unsually high tide led to siginificant flooding in coastal comminuties of the Strait of Georgia \citep{romanowski2010storm}.
 The maximum water level observed at Point Atkinson on this day was 5.49 m, which is 0.8 m higher than the predicted tides, highlighting the significance of this event as it resulted in a large anomaly and also reached water levels close to the highest level ever recorded (5.61 m).
+A set of simulations from Feb 1 to 8, 2006 under varying forcing conditions has been completed in order to determine the siginicance of each factor contributing to the surge.
 
@@
 The timing of maximum water level matches well between the observations and the model at all four locations, the most siginificant contribution to high water level being from the tides.
-There is also a good match in amplitude of the water level maximum at Point Atkinson and Cambell River, with a relative error less than 2\%.
-However, the error at Victoria and Patricia Bay is more siginificant, at 12\% and 8\% respectively, likely due to error in M$_2$ tidal amplitudes as discussed above.
+There is also a good match in amplitude of the water level maximum at Point Atkinson and Campbell River, with a relative error less than 2\%.
+Indeed, there is only a 1.2 cm difference between observed and modelled water level elevation at Point Atkinson and a 9 cm difference at Campbell River.
+However, the error at Victoria and Patricia Bay is more siginificant, reaching 12\% and 8\% respectively (close to 40 cm in each case), partly due to error in tidal amplitudes as discussed above.
 The precise location and magnitude of the M$_2$ amphidrome around Victoria is very sensitive to model parameters such as bottom friction.
 
-The error in modelled residuals at Victoria and Campbell River is also apparent in these figures.
-Most noticebale is underestimate in modelled residual throughout most of the time period.
-Additionally, there is a delay of about 2-3 hours in the timing of maximum residual and the relative errors in maximum residual range from 5\% at Point Atkinson to 29\% at Patricia Bay.
+The error in modelled residuals at Victoria and Patricia Bay is also apparent in the panels on the right, most noticeable is an underestimate in modelled residual throughout most of the time period.
+Additionally, there is a delay of about 2-3 hours in the timing of maximum residual at all four locations.
+The errors in maximum residual range from 3.9 cm at Point Atkinson to 21 cm at Patricia Bay.
 %Question: is it better to present relative errors or just difference?
-%I think Hal Ritchie's talk with the storm surges had some stats presented. Look this up...
 
@@
 \end{figure}
-%Note that colour may be an option here (no extra charge for A-O). But it will be worse for printing.
+
+Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented in Figure \ref{fig:factors}.
+The modelled residual from three simulations have been plotted.
+First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly, is shown in blue.
+The modelled sea surface height has not been corrected as in the previous plot.
+Additionally, a simulation with all forcing except the atmospheric forcing is presented in green and a simulation with all forcing except the sea surface height anomaly is shown in red.
+It is clear that the sea surface height anomaly contributes most significantly to the surge at each of these locations.
+When the sea surface height anomaly is not included as a forcing condition, the surge drops from 78 cm to 28 cm at Point Atkinson.
+Additionally, the timing of the maximum surge occurs much later.
+A similar drop in the maximum surge and delay in the timing of the maximum surge is observed at the other locations.
+
+The contribution to the surge due to atmospheric forcing appears to be geographically dependent, its removal leading to a drop (about 15 cm) in the maximum surge at Point Atkinson and Campbell River but a slight increase (about 6 cm) at Victoria and Patricia Bay.
+Additionally, at Point Atkinson the timing of the maximum surge is earlier by about two hours when atmospheric effects are neglected.
+%Worried that this might be due to pressure/mountains at PA and CR
+%maybe look at winds to see why Vic/Pat Bay increased. Plan: map of wind vectors from CGRF (not for paper, just for me in aiding with explanation)
+
+Finally, an assessment of the importance in including tide-surge interactions is presented next.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
+\caption{Comparison of the modelled residual for three simulations: A simulation with all forcing (blue), a simulation without atmospheric forcing (green), and a simulation without the sea surface height anomaly.
+}
+\label{fig:factors}
+\end{figure}
+%not sure about the colours, but different line styles is really hard to make out. 
+%Maybe only need the surge plot?
 
@@
 \subsection{High Resolution Atmospheric Forcing: Dec 17, 2012}
+
+\subsection{Statisitics}
+%I think Hal Ritchie's talk with the storm surges had some stats presented. Look this up...
 

### 2014-07-11 RoughWork/manuscript.tex
Added disucssion on tide-surge interaction

@@
 %more on past events, significance
+
+Communities in the Strait of Georgia are partiuclarly susceptible to storm surges in the winter months when there is an increased propensity for storms and south easterly winds and geostrophic adjustment lead to higher coastal waters \citep{danard2003storm}.
+In the summer months, the wind direction changes to north westerly and mean sea level is lower.
+%work on this
 
@@
 
-Finally, an assessment of the importance in including tide-surge interactions is presented next.
+These comparisons point to the importance of including, as a forcing condition at the open bounday, the surge entering the domain from the Pacific Ocean.
+Storms travelling over the Pacific Ocean torwads the west coast of the northern United States and Canada can induce surges along the coastline which then enter the Salish Sea sytem through the Strait of Juan de Fuca.
+In a model domain that does not inlcude the Pacific Ocean, this effect must be inlcuded as a boundary condition.
+Previous simulations by \citet{murty1995storm} suggest that the inverese barometer effect and the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.
+A detailed assessment of the effect of the local atmospheric condtions and winds is included later where simulations have been performed with high resolution atmospheric forcing data.
 
@@
 %not sure about the colours, but different line styles is really hard to make out.
-%Maybe only need the surge plot?
+
+Finally, an assessment of the importance of including tide-surge interactions is presented next.
+A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}.
+Recall, the modelled residual is defined as the difference between the sea surface height in a simulation with all forcing and a simulaltion with tides amd rivers only.
+A comparison between the modelled residual and a surge only run can determine the significance of the tide-surge interaction.
+The most significant differences occur at Point Atkinson where the maximum modelled residual is 7 cm higher than the maximum of the surge only case.
+In contrast, the tide-surge interaction at Victoria is hardly noticeable.
+Indeed, a spatial map of the modelled residual and surge only differences (not shown here) indicate that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands.
+The passages between these islands are known for vigorous tidal mixing, which may have an effect on the surge propagation.
+However, the effect is a relatively small underestimate of the surge elevation and no changes to the timing of the surge.
+So, in operational settings, running a surge only simulation and adding the predicted tides a posteriori could be more computationally efficient with only a small decrease in the accuracy of the prediction.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/feb2006_tidesurge.pdf}
+\caption{Modelled residuals (solid) compared a simulation of the surge propagation without tidal forcing (dashed).}
+\label{fig:tidesurge}
+\end{figure}
 
@@
 %future directions: running real time, biology and chemistry, model improvements?
+%comments on wetting and drying (maybe most imporant in shallow regions)
 

### 2014-07-15 RoughWork/manuscript.tex
Tried making a nicer bathy image.
New bathymetry and domain image
Included no_pressure run in analysis, but didn't save the figure.

@@
 \usepackage{amsmath}
-\usepackage{epspdfconversion}
 
@@
 \centering
-\includegraphics[scale=0.5]{Figures/bathy.png}
-\caption{Model domain including bathymetry, rivers (*), and storm surge locations (o) of interest.}\label{fig:domain}
+\includegraphics[scale=0.5]{Figures/bathy.pdf}
+\caption{Model domain (highlighted area) including bathymetry, rivers (blue circles), and storm surge locations of interest  (red stars).}\label{fig:domain}
 \end{figure}
@@
 \includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
-\caption{Comparison of the modelled residual for three simulations: A simulation with all forcing (blue), a simulation without atmospheric forcing (green), and a simulation without the sea surface height anomaly.
+\caption{Comparison of the modelled residual for three simulations: A simulation with all forcing (blue), a simulation without atmospheric forcing (green), and a simulation without the sea surface height anomaly (red).
 }

### 2014-07-15 RoughWork/manuscript.tex
Modified feb2006 by removing a background state.

@@
 Storm surge simulations were started from rest at least three days prior to the event of interest to allow for adjustment of model currents.
-Stratification conditions from model spin-up at a time close the the event (within 5 days) were chosen to initialize the temperature and salinity fields.
+Stratification conditions from model spin-up at a time close to the event (within 5 days) were chosen to initialize the temperature and salinity fields.
 Boundary conditions include atmospheric forcing and fresh water input from rivers at the surface, sea surface height anomaly from observations at the open boundary, and tidal forcing with eight tidal constituents.
@@
 As such, storm surge model predictions were corrected by accounting for this error.
-Additionally, since the sea surface height is calculated about a zero reverence level, the modelled water level is determined by adding a long term average of mean sea level at the specified location to the modelled sea surface height.
+Additionally, since the sea surface height is calculated about a zero reverence level, the modelled water level is determined by adding a long term average of mean sea level at the specified location to the modelled sea surface height and removing a background anomaly caused by topographic pressure gradients since the atmospheric reforecasting model uses a terrain following coordinate system.
+(Nancy: I think we need a better way of adjusting come publication time).
 
@@
 
-Three significiant surge or wind events have been considered.
+Three significiant surge or wind events have been considered in detail, as well as an array of hindcasts to asses model ability in a statistical measure.
 First, a study of a large surge event on Feb 4, 2006 is presented.
@@
 %% Trying to decide Dec 15, 2005 or Nov 2009. I think it's more sense to do Nov 2009 - strong winds but no surge. 
-Next, an extreme wind event which resulted in property damage in Vancouver and surrounding areas is discussed.
+Second, an extreme wind event which resulted in property damage in Vancouver and surrounding areas is discussed.
 This event occured on Dec 15, 2006 but did not result in significant flooding such as that which occured on Feb 4, 2006 because the storm did not occur at high tide.
 This case was choosen to identify the model's capability in handling large wind events that do not result in high water levels.
-Finally, a case with high resolution atmospheric forcing for a recent surge on Dec 17, 2012 is discussed.
+Third, a case with high resolution atmospheric forcing for a recent surge on Dec 17, 2012 is discussed.
+Lastly, these cases presented in detail will be combined with $x$ additional runs for statistical evaluation of model performance.
 
@@
 The timing of maximum water level matches well between the observations and the model at all four locations, the most siginificant contribution to high water level being from the tides.
-There is also a good match in amplitude of the water level maximum at Point Atkinson and Campbell River, with a relative error less than 2\%.
-Indeed, there is only a 1.2 cm difference between observed and modelled water level elevation at Point Atkinson and a 9 cm difference at Campbell River.
-However, the error at Victoria and Patricia Bay is more siginificant, reaching 12\% and 8\% respectively (close to 40 cm in each case), partly due to error in tidal amplitudes as discussed above.
+There is also a good match in amplitude of the water level maximum at Point Atkinson and Campbell River, with a relative error less than 4\%.
+There is only a 7.2 cm difference between observed and modelled water level elevation at Point Atkinson and a 20.9 cm difference at Campbell River.
+Tthe error at Victoria and Patricia Bay is slightly more siginificant, reaching 9.6\% and 6.3\% respectively (close to 30 cm in each case), partly due to error in tidal amplitudes as discussed above.
 The precise location and magnitude of the M$_2$ amphidrome around Victoria is very sensitive to model parameters such as bottom friction.
 
-The error in modelled residuals at Victoria and Patricia Bay is also apparent in the panels on the right, most noticeable is an underestimate in modelled residual throughout most of the time period.
-Additionally, there is a delay of about 2-3 hours in the timing of maximum residual at all four locations.
-The errors in maximum residual range from 3.9 cm at Point Atkinson to 21 cm at Patricia Bay.
+The error in modelled residuals is also apparent in the panels on the right, most noticeable is an underestimate in maximum residual (or surge) at all locations.
+There is also a 2-3 hour delay in the timing of the maximum residual except at Victoria where the maximum residual in sync with observations.
+The errors in maximum residual range from 8.0 cm at Victoria to 28 cm at Campbell River.
+It is also apparent that Point Atkinson and Campbell River experienced the largest observed surges at 82 cm and 95 cm respectively.
+The surges at Victoria and Patricia Bay were observed to be 66 cm and 74 cm.
 %Question: is it better to present relative errors or just difference?
@@
 \centering
-\includegraphics[scale=0.6]{Figures/feb2006.pdf}
+\includegraphics[scale=0.6]{Figures/feb2006_bg.pdf}
 \caption{Comparison of observations and model output for the Feb 4, 2006 storm surge.

### 2014-07-15 RoughWork/manuscript.tex
Removed background in factors plots/discussion.

@@
 \end{figure}
-%image needs some work: colour? labels for JdF, SoG, cities...
+%image needs some work: labels for tide comparisons?
 
@@
 First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly, is shown in blue.
-The modelled sea surface height has not been corrected as in the previous plot.
 Additionally, a simulation with all forcing except the atmospheric forcing is presented in green and a simulation with all forcing except the sea surface height anomaly is shown in red.
 It is clear that the sea surface height anomaly contributes most significantly to the surge at each of these locations.
-When the sea surface height anomaly is not included as a forcing condition, the surge drops from 78 cm to 28 cm at Point Atkinson.
+When the sea surface height anomaly is not included as a forcing condition, the surge drops from 70 cm to 15.6 cm at Point Atkinson.
 Additionally, the timing of the maximum surge occurs much later.
 A similar drop in the maximum surge and delay in the timing of the maximum surge is observed at the other locations.
+Note that the press
 
-The contribution to the surge due to atmospheric forcing appears to be geographically dependent, its removal leading to a drop (about 15 cm) in the maximum surge at Point Atkinson and Campbell River but a slight increase (about 6 cm) at Victoria and Patricia Bay.
-Additionally, at Point Atkinson the timing of the maximum surge is earlier by about two hours when atmospheric effects are neglected.
-%Worried that this might be due to pressure/mountains at PA and CR
-%maybe look at winds to see why Vic/Pat Bay increased. Plan: map of wind vectors from CGRF (not for paper, just for me in aiding with explanation)
+The contribution to the surge due to atmospheric forcing appears to be geographically dependent, its removal leading to a drop of about 5 cm in the maximum surge at Point Atkinson and 10 cm at Campbell River.
+When atmospheric forces are neglected the timing of the maximum surge is earlier at Point Atkinson but later at Campbell River.
+Campbell River exhibits a strong response to atmopsheric conditions during midday on Feb 4, as demonstrated in a two sharp spikes in both the blue and red curves arounf 12:00 hr The green curve, which does not include any atmopsheric focring, does not exhibit this beahvious.
+The result is a peak surge earlier than in day than the no atmospheric forcing case.
+This response is also noted in the observations at Campbell River in Figure \ref{fig:feb2006} where the maxmium surge occurs earlier than at the other locations.
+At Victoria and Patricia Bay, the surge amplitude is mainly unaffected by the atmospheric forcing.
+%maybe look at winds/pressure to see why Vic/Pat Bay aren't affected?. Plan: map of wind vectors from CGRF (not for paper, just for me in aiding with explanation)
 
 These comparisons point to the importance of including, as a forcing condition at the open bounday, the surge entering the domain from the Pacific Ocean.
-Storms travelling over the Pacific Ocean torwads the west coast of the northern United States and Canada can induce surges along the coastline which then enter the Salish Sea sytem through the Strait of Juan de Fuca.
+Storms travelling over the Pacific Ocean torwads the west coast of the northern United States and Canada can induce elevated sea levels along the coastline which then enter the Salish Sea sytem through the Strait of Juan de Fuca.
 In a model domain that does not inlcude the Pacific Ocean, this effect must be inlcuded as a boundary condition.
@@
 \centering
-\includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
+\includegraphics[scale=0.6]{Figures/feb2006_factors_bg.pdf}
 \caption{Comparison of the modelled residual for three simulations: A simulation with all forcing (blue), a simulation without atmospheric forcing (green), and a simulation without the sea surface height anomaly (red).

### 2014-07-15 RoughWork/manuscript.tex
Removed a background state from tide-surge plots.

@@
 A comparison between the modelled residual and a surge only run can determine the significance of the tide-surge interaction.
+This figure shows there is almost no difference between the simulations at all locations, rendering the tide-sure interaction negligible.
+In operational settings, running a surge only simulation and adding the predicted tides a posteriori could be more computationally efficient with only a small decrease in the accuracy of the prediction.
+% I'm not sure if remocing a background state is really appropriate here.
+There could be some differences in the background due to the removal of the tides.
+This also suggests it is more appropriate to use the atmospheric pressure to remove the background...
+
+%Finally, an assessment of the importance of including tide-surge interactions is presented next.
+A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}.
+Recall, the modelled residual is defined as the difference between the sea surface height in a simulation with all forcing and a simulaltion with tides amd rivers only.
+A comparison between the modelled residual and a surge only run can determine the significance of the tide-surge interaction.
 The most significant differences occur at Point Atkinson where the maximum modelled residual is 7 cm higher than the maximum of the surge only case.
@@
 \centering
-\includegraphics[scale=0.6]{Figures/feb2006_tidesurge.pdf}
+\includegraphics[scale=0.6]{Figures/feb2006_tidesurge_bg.pdf}
 \caption{Modelled residuals (solid) compared a simulation of the surge propagation without tidal forcing (dashed).}

### 2014-07-17 RoughWork/manuscript.tex
Added new figure to text
Added description of surge propagation.

@@
 Below 10m the grid is stretched gradually to a maximum grid spacing of 27 m at the lowest layer.
-At the bottom boundary, partial $z$-levels are utilized in order to limit large changes in bathymetry across grid cells \citep{madec2008nemo}.
+At the bottom boundary, partial $z$-levels are utilized in order to limit large changes in bathymetry across grid cells \citep{madec2012nemo}.
 
@@
 Vertical turbulence and mixing is calculated through the $k-\epsilon$ configuration of the generic length scale (GLS) turbulence closure \citep{umlauf2003generic} with background vertical eddy viscosity and diffusivity set to $1\times10^{-4}$ and $1\times10^{-5}$ m$^2$ s$^{-1}$ respectively.
-Details on the NEMO implementation of the partial slip lateral boundary condition, quadratic bottom friction law, and GLS turbulence closure scheme are provided by \citet{madec2008nemo}.
+Details on the NEMO implementation of the partial slip lateral boundary condition, quadratic bottom friction law, and GLS turbulence closure scheme are provided by \citet{madec2012nemo}.
 
@@
 %other references?
-The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \citep{flather1994storm, madec2008nemo}.
+The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}.
 The baroclinic velocities at the boundary were set to be equal to the values inside the boundary (zero-gradient boundary conditions).
@@
 
-Three significiant surge or wind events have been considered in detail, as well as an array of hindcasts to asses model ability in a statistical measure.
-First, a study of a large surge event on Feb 4, 2006 is presented.
-This event caused siginificant damage and flooding in Boundary Bay, a community south of Richmond.
-(citation) and has been chosen as a case study to demonstrate the model's skill in reproducing a storm surge and to quantitatively assess forcing conditions that are important in storm surge hindcasting.
-%% Trying to decide Dec 15, 2005 or Nov 2009. I think it's more sense to do Nov 2009 - strong winds but no surge. 
-Second, an extreme wind event which resulted in property damage in Vancouver and surrounding areas is discussed.
-This event occured on Dec 15, 2006 but did not result in significant flooding such as that which occured on Feb 4, 2006 because the storm did not occur at high tide.
+Several significiant surge or wind events have been considered in detail, as well as an array of hindcasts to asses model ability in a statistical measure.
+These cases have been used to asses the model's skill in reproducing water level height and surge amplitude and timing during storm events.
+In addition, the contribution to the surge of several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed.
+We have also examined the spatial variabilty of the surge as it propagates through the domain and have considered a case with high-resolution atmospheric forcing.
+
+%First, a study of a large surge event on Feb 4, 2006 is presented.
+This event caused siginificant damage and flooding in Boundary Bay, a community south of Richmond, (citation) and has been chosen as a case study to demonstrate the model's skill in reproducing a storm surge and to quantitatively assess forcing conditions that are important in storm surge hindcasting.
+Second, an extreme wind event on Dec 15, 2006 is examined.
+This event produced winds over (x) (citation) but did not result in significant flooding because the storm did not occur at high tide, however, it did result in siginificant property damage due to high winds and waves.
+This event has been used to understand the spatial and temporal variability of the surge propagation.
+Next, a large wind event on Nov 18, 2009 that did not result in a siginificant surge or flooding is presented.
 This case was choosen to identify the model's capability in handling large wind events that do not result in high water levels.
-Third, a case with high resolution atmospheric forcing for a recent surge on Dec 17, 2012 is discussed.
+Next, a case with high resolution atmospheric forcing for a recent surge on Dec 17, 2012 is discussed.
 Lastly, these cases presented in detail will be combined with $x$ additional runs for statistical evaluation of model performance.
@@
 In operational settings, running a surge only simulation and adding the predicted tides a posteriori could be more computationally efficient with only a small decrease in the accuracy of the prediction.
-% I'm not sure if remocing a background state is really appropriate here.
-There could be some differences in the background due to the removal of the tides.
-This also suggests it is more appropriate to use the atmospheric pressure to remove the background...
+% I'm not sure if removing a background state is really appropriate here.
+There could be some differences in the background states because of the tides(ie where fresh water moves to, etc).
+I also think it is more appropriate to use the atmospheric pressure to remove the background state...
 
@@
 
-\subsection{Strong wind event: Dec 15, 2006}% or Nov 2009
+\subsection{Spatial Extent of Surge: Dec 15, 2006}
+On Dec 15, 2006, a major storm with wind gusts over 120 km/hr caused significant damage to many coastal communities on the Pacific Northwest.
+In British Columbia, over 250,000 residents were left without power due to downed trees and hydro lines.
+Although the storm was very strong and the observed residual at Point Atkinson reached 85 cm, this event did not coincide with high tide and so no siginificant flooding occurred in the Vancouver area\citep{forseth2006adaptation}.
+This event has been used to examine the propgation of the surge throughout the domain.
+First, an assessment of the accuracy of the hindcast is provided.
+
+The evolution and spatial distribution of the surge is highlighted in Figure \ref{fig:spatial} where the modelled residual is plotted every two hours beginning on Dec 14 at 18:30.
+The surge enters the system from the Strait of Juna de Fuca and Johnstone Strait, peaking between Dec 15 at 6:30 and Dec 15 at 10:30 and then begins to subside throughout the entire domain.
+Notably, the Strait of Georgia experiences the largest surge values, upwards of 70 cm, whereas the Juan de Fuca region experiences a maximum surge around 55 cm.
+Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges than the Vancouver Island side, likley due to the southerly wind in the atmosheric model pushing water against the mainland coast.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/dec2006_spatial.pdf}
+\caption{Propagation of the storm surge on Dec 14/15 2006.
+Modelled residual every 2 hours starting on Dec 14 at 18:30.
+Time advances from top left to bottom right.
+Note that the images are aligned with the numerical grid which is rotated $29^{\circ}$ counter-clockwise of North.
+}
+\label{fig:spatial}
+\end{figure}
+
+\subsection{Strong Wind Event: Nov 18, 2009}
 

### 2014-07-18 RoughWork/manuscript.tex
Included description of model winds for feb 2006

@@
 \subsection{Factors Contributing to Storm Surges: Feb 4, 2006}
-%Include spatial extend of surge
 
@@
 It is clear that the sea surface height anomaly contributes most significantly to the surge at each of these locations.
-When the sea surface height anomaly is not included as a forcing condition, the surge drops from 70 cm to 15.6 cm at Point Atkinson.
-Additionally, the timing of the maximum surge occurs much later.
-A similar drop in the maximum surge and delay in the timing of the maximum surge is observed at the other locations.
-Note that the press
+When the sea surface height anomaly is not included as a forcing condition, the surge drops from 70 cm to 15.6 cm at Point Atkinson and its maximum occurs eight hours later.
+A similar drop in the maximum surge and delay in timing of the maximum surge is observed at the other locations.
+The atmospheric model indicates southeasterly winds during the onset of the surge but after the peak surge has passed, the winds shift to westerly.
+A strong westerly wind would cause an elevated sea surface height at Point Atkinson as water piles against the coast, explaining the delay in peak surge when only atmospheric forcing is used since the winds to shift westerly after the peak surge has passed.
+A simulation with only wind forcing (no inverse barometer effect or anomaly at open boundary) produces a surge of 9.0 cm at Point Atkinson, a small fraction of the total observed surge.
+However, this wind-induced surge occurs well after the peak surge from observations and simulations with full forcing.
 
@@
 When atmospheric forces are neglected the timing of the maximum surge is earlier at Point Atkinson but later at Campbell River.
-Campbell River exhibits a strong response to atmopsheric conditions during midday on Feb 4, as demonstrated in a two sharp spikes in both the blue and red curves arounf 12:00 hr The green curve, which does not include any atmopsheric focring, does not exhibit this beahvious.
+Campbell River exhibits a strong response to atmopsheric conditions during midday on Feb 4, as demonstrated in a two sharp spikes in both the blue and red curves arounf 12:00 hr The green curve, which does not include any atmopsheric focring, does not exhibit this behaviour.
 The result is a peak surge earlier than in day than the no atmospheric forcing case.
@@
 Time advances from top left to bottom right.
-Note that the images are aligned with the numerical grid which is rotated $29^{\circ}$ counter-clockwise of North.
+Note that the images are oriented with the numerical grid which is rotated $29^{\circ}$ counter-clockwise of North.
 }

### 2014-07-21 RoughWork/manuscript.tex
Included surge timing/amplitude infor for Dec 2006

@@
 On Dec 15, 2006, a major storm with wind gusts over 120 km/hr caused significant damage to many coastal communities on the Pacific Northwest.
-In British Columbia, over 250,000 residents were left without power due to downed trees and hydro lines.
-Although the storm was very strong and the observed residual at Point Atkinson reached 85 cm, this event did not coincide with high tide and so no siginificant flooding occurred in the Vancouver area\citep{forseth2006adaptation}.
+Although the storm was very strong and the observed residual at Point Atkinson reached 84 cm, this event did not coincide with high tide and so no siginificant or reported flooding occurred in the Vancouver area\citep{forseth2006adaptation}.
 This event has been used to examine the propgation of the surge throughout the domain.
 First, an assessment of the accuracy of the hindcast is provided.
+
+Point Atkinson observed a maximum surge amplitude of 84 cm, which is approximately 20 cm higer than the model result, however the timing of the surge agrees almost exactly.
+The best agreement occurs at Victoria where the observed surge was 62 cm and the model returned 54 cm.
+At all of the locations, the model predicted the maximum surge to occur within 3.5 hours of the observed surge.
+Typically, the modelled surge was too early, in contrast to the later predictions in the Feb, 4 2006 case.
+Errors in the modelled water level at the timing of maximum surge were of the same order as the errors in the surge amplitude.
 

### 2014-07-21 RoughWork/manuscript.tex
Reorganize model configuration section so that it flows like a paper.

@@
 
-The curvilinear orthogonal numerical grid is divided into 398 by 898 by 40 grid cells, which results in an almost uniform horizontal resolution with grid spacing approximately 440 m by 500 m.
-The 40 vertical $z$-levels are stretched gradually in order to achieve higher resolution in the surface layer, with 1 m vertical grid spacing down to about 10 m in depth.
-Below 10m the grid is stretched gradually to a maximum grid spacing of 27 m at the lowest layer.
-At the bottom boundary, partial $z$-levels are utilized in order to limit large changes in bathymetry across grid cells \citep{madec2012nemo}.
+The curvilinear orthogonal numerical grid has been divided into 398 by 898 by 40 grid cells, which results in an almost uniform horizontal resolution with grid spacing approximately 440 m by 500 m.
+The 40 vertical $z$-levels were stretched gradually in order to achieve higher resolution in the surface layer, with 1 m vertical grid spacing down to about 10 m in depth.
+Below 10m the grid was stretched to a maximum grid spacing of 27 m at the lowest layer.
+At the bottom boundary, partial $z$-levels were utilized in order to limit large changes in bathymetry across grid cells \citep{madec2012nemo}.
+
+In addition to the equations of motion, a prognostic equation for the sea surface height is solved at each time step.
+The inclusion of the sea surface height equation requires a fairly restrictive time step due to the presence of high speed surface gravity waves.
+As such, the split-explicit time stepping algorithm is employed, where the free surface and barotropic equations are solved with a smaller time step than that used for the other variables.
+The model time step is and barotropic time step are 10 s and 2 s respectively.
+Small vertical grid spacing and large vertical velocities in the Boundary Pass and Haro Strait region impose a fairly restricitive step through the vertical CFL condition.
 
@@
 \subsection{Boundary conditions and sub-grid scale processes}
-%lateral and bottom friction, open boundaries, eta surface
-The model includes two open boundaries that connect to the Pacific Ocean, the western boundary of the Strait of Juan de Fuca as well as Johnstone Strait at the north, both of which are forced with eight tidal constituents, temperature and salinity climatologies, and the sea surface height anomaly.
-The details of these forcing conditions are described below.
-At coastal boundaries, the partial slip boundary condition, an approximation to no slip, is used.
-The partial slip boundary condition allows one to include the frictional effects of lateral boundaries without the restrictive resolution required to represent the lateral boundary layer under no slip conditions.
-A lateral eddy viscosity of 20 m $^2$ s $^{-1}$ parametrizes horizontal friction and a lateral eddy diffusivity of 20.5 m $^2$ s $^{-1}$ is used.
-At the ocean surface, meteorological conditions and fresh water input from rivers are included and described in more detail below.
-Bottom friction is represented by a quadratic law for the bottom momentum flux with drag coefficient $C_D = 5\times 10^{-3}$.
-Vertical turbulence and mixing is calculated through the $k-\epsilon$ configuration of the generic length scale (GLS) turbulence closure \citep{umlauf2003generic} with background vertical eddy viscosity and diffusivity set to $1\times10^{-4}$ and $1\times10^{-5}$ m$^2$ s$^{-1}$ respectively.
-Details on the NEMO implementation of the partial slip lateral boundary condition, quadratic bottom friction law, and GLS turbulence closure scheme are provided by \citet{madec2012nemo}.
-
-In addition to the equations of motion, a prognostic equation for the sea surface height is solved at each time step.
-The inclusion the sea surface height equation requires a fairly restrictive time step due to the presence of high speed surface gravity waves.
-As such, the split-explicit time stepping algorithm is employed, where the free surface and barotropic equations are solved with a smaller time step than that used for the other variables.
-The model time step and barotropic time step are 10 s and 2 s respectively.
-
-\subsubsection{Tidal forcing} 
-The model was forced by eight tidal constituents (K$_1$ ,O$_1$, P$_1$, Q$_1$, M$_2$, K$_2$, N$_2$, S$_2$) at the Juan de Fuca and Johnstone Strait boundaries.
+The model includes two open boundaries that connect to the Pacific Ocean, the western boundary of the Strait of Juan de Fuca as well as Johnstone Strait at the north, both of which are forced with eight tidal constituents (K$_1$ ,O$_1$, P$_1$, Q$_1$, M$_2$, K$_2$, N$_2$, S$_2$), temperature and salinity climatologies, and the sea surface height anomaly.
 Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on work by \citet{foreman2000webtide}.
@@
 The remaining constituents were extrapolated from Webtide.
+% Needs editing! Update when we are finished with tides.
 
-\subsubsection{Temperature and salinity}
 Temperature and salinity at the Juan de Fuca boundary were taken from a weekly climatology which was created from results from a model covering the Salish Sea and the west coast of Vancouver Island \citep{massonfine2012}.
@@
 To prepare the climatology all years (1995-2008) were averaged and results, approximately every 15 days, were interpolated to a weekly climatology.
+The model is relaxed to the forced temperature and salinity over the 10 grid points (about 5km) closest to the open boundaries, using a flow relaxation scheme \citep{engedahl1995use}.
+%other references?
 
-\subsubsection{Sea Surface Height}
 Sea surface height at the mouth of Juan de Fuca was set using values from the Tofino tide gauge.
@@
 The storm surge simulations also used the sea surface height anomaly from Port Hardy, forced at the northern boundary in Johnstone Strait.
-
-\subsubsection{Open Boundary Conditions}
-
-The model is relaxed to the forced temperature and salinity over the 10 grid points (about 5km) closest to the open boundaries, using a flow relaxation scheme \citep{engedahl1995use}.
-%other references?
 The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}.
@@
 
-\subsection{River forcing}
+At coastal boundaries, the partial slip boundary condition, an approximation to no slip, is used.
+Partial slip allows one to include the frictional effects of lateral boundaries without the restrictive resolution required to represent the lateral boundary layer under no slip conditions.
+A lateral eddy viscosity of 20 m $^2$ s $^{-1}$ parametrizes horizontal friction and a lateral eddy diffusivity of 20.5 m $^2$ s $^{-1}$ is used.
+Bottom friction is represented by a quadratic law for the bottom momentum flux with drag coefficient $C_D = 5\times 10^{-3}$.
+Vertical turbulence and mixing is calculated through the $k-\epsilon$ configuration of the generic length scale (GLS) turbulence closure \citep{umlauf2003generic} with background vertical eddy viscosity and diffusivity set to $1\times10^{-4}$ and $1\times10^{-5}$ m$^2$ s$^{-1}$ respectively.
+Details on the NEMO implementation of the partial slip lateral boundary condition, quadratic bottom friction law, and GLS turbulence closure scheme are provided by \citet{madec2012nemo}.
+
+At the ocean surface, meteorological conditions and fresh water input from rivers are included.
+The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforcasting model suitable for use in ocean modelling \citep{smith2013new}.
+Forecasts from the period of 2002-2011 are available.
+Additionally, the inverse barometric effect of the atmospheric pressure is included, an important consideration in storm surge modelling.
+
 River input provides a significant volume of freshwater to the Salish Sea and can influence stratification, circulation and primary productivity.
@@
 Incoming water was assumed to be fresh and at surface temperature.
-A total of 150 rivers were parameterised by this method.
+A total of 150 rivers were parameterised by this method and their locations are indicated by the blue dots in Figure \ref{fig:domain}.
 
-%we could show the location of all the rivers on the overall location map, or on a picture of our domain, since the rivers are on the edges of the domain so they won't block any bathymetry information
-
-\subsection{Atmospheric forcing}
-The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforcasting model suitable for use in ocean modelling \citep{smith2013new}.
-Forecasts from the period of 2002-2011 are available.
-Additionally, the inverse barometric effect of the atmospheric pressure is included, an important consideration in storm surge modelling.
-
-\subsection{Initial conditions}
 Initial conditions for temperature and salinity were taken from a CTD cast in the middle Strait of Georgia taken in Sept 2002 \citep{pawlowiczetal2007}.
-Conditions were initially uniform horizontally.
-Velocity was initialized at zero.
-%Merge with below?
-%Perhaps image of spun-up stratification along Thalweg? Here or in model evaluation
-
-\subsection{Spin-up}
+Conditions were initially uniform horizontally and velocity was initialized at zero.
 The model was spun up for a 15.5 months from the initial conditions above, starting Sep 16, 2002, using atmospheric forcing from 2002-2003, climatological temperature and salinity and sea surface height at the boundaries, with tides and climatological river output.
@@
 D = [(A_0 \cos g_0 - A_m \cos g_m)^2 + (A_0 \sin g_0 - A_m \sin g_m)^2]^{1/2}
-\end{equation}
+\end{equation}\label{eq:compdiff}
 where $A_0$, $A_m$, $g_0$ and $g_m$ are observed and modelled amplitudes and phases.
@@
 \caption{Comparisons of modelled and observed amplitudes and phases.
-$R$ is the ration of modelled to observed amplitude.
+$R$ is the ratio of modelled to observed amplitude.
 $\Delta \phi$ is the difference between modelled and observed phase.
@@
 \end{table}
-%May have to include station numbers and locations on a map? Might be too busy with rivers too...
+%May have to include station numbers and locations on a map? Maybe on the contour map, overlay with stations and green/red dots (hopefully mostly green)
 %Should we use the new method of calculating tides?
@@
 %outline: thalweg, plots of winter/summer stratty
-%Maybe note this
+%Maybe not this
 
@@
 Typically, the modelled surge was too early, in contrast to the later predictions in the Feb, 4 2006 case.
-Errors in the modelled water level at the timing of maximum surge were of the same order as the errors in the surge amplitude.
+Errors in the modelled water level at the time of maximum surge were of the same order as the errors in the surge amplitude.
 

### 2014-07-22 RoughWork/manuscript.tex
Added an abstract. Still needs work

@@
 
-\title{Storm Surges}
-
+\title{Storm surge modelling in the Strait of Georgia}
+\author{Nancy Soontiens, Susan Allen, Doug Latornell, Kate Le Sou\"{e}f}
 \date{\today}
@@
 \begin{abstract}
-Abstract to be completed
+The Strait of Georgia is a large, semi-enclosed body of water between Vancouver Island and the mainland of British Columbia connected to the Pacific Ocean via the Strait of Juan de Fuca at the south and Johnstone Strait at the north.
+During the winter months, coastal communities in the Strait of Georgia are particularly susceptible to flooding caused by storm surges, a natural hazard that can occur when a strong storm coincides with high tide.
+Storm surge hindcasts have been produced using a three-dimensional numerical ocean model for the Strait of Georgia and the surrounding bodies of water (Strait of Juan de Fuca, Puget Sound, and Johnstone Strait) collectively known as the Salish Sea.
+The numerical model employs the Nucleus for European Modelling of the Ocean (NEMO) in a regional configuration.
+The model has been evaluated through comparisons of tidal elevation harmonics, tidal currents, and storm surge water levels with observations.
+Important forcing factors contributing to storm surges have been assessed, finding that the surge entering the domain from the Pacific Ocean contributes most significantly.
+%needs a little more work.
 \end{abstract}

### 2014-07-22 RoughWork/manuscript.tex
Worked on introduction

@@
 During the winter months, coastal communities in the Strait of Georgia are particularly susceptible to flooding caused by storm surges, a natural hazard that can occur when a strong storm coincides with high tide.
-Storm surge hindcasts have been produced using a three-dimensional numerical ocean model for the Strait of Georgia and the surrounding bodies of water (Strait of Juan de Fuca, Puget Sound, and Johnstone Strait) collectively known as the Salish Sea.
+This investigation produces storm surge hindcasts using a three-dimensional numerical ocean model for the Strait of Georgia and the surrounding bodies of water (Strait of Juan de Fuca, Puget Sound, and Johnstone Strait) collectively known as the Salish Sea.
 The numerical model employs the Nucleus for European Modelling of the Ocean (NEMO) in a regional configuration.
-The model has been evaluated through comparisons of tidal elevation harmonics, tidal currents, and storm surge water levels with observations.
+The model is evaluated through comparisons of tidal elevation harmonics, tidal current ellipses, and storm surge water levels with observations.
 Important forcing factors contributing to storm surges have been assessed, finding that the surge entering the domain from the Pacific Ocean contributes most significantly.
@@
 
-%my attempt at physical oceanopgraphy SoG- Susan feel free to edit.
+%my attempt at physical oceanopgraphy in SoG- Susan feel free to edit.
 The Strait of Georgia is a strongly stratified, semi-enclosed body of water located between Vancouver Island and the mainland of British Columbia.
@@
 %overview oF SoG/Salish Sea models
+Early efforts by \citet{crean1988numerical, stronach1993update} implemented a series of models covering the Strait of Juan de Fuca, the Strait of Georgia and Puget Sound using a finite difference discretization of the equations of motion.
+%tide models (Foreman), baroclinic (Masson, PNWL, Sutherland and MacCready)
 
@@
 
-Communities in the Strait of Georgia are partiuclarly susceptible to storm surges in the winter months when there is an increased propensity for storms and south easterly winds and geostrophic adjustment lead to higher coastal waters \citep{danard2003storm}.
-In the summer months, the wind direction changes to north westerly and mean sea level is lower.
+Communities in the Strait of Georgia are partiuclarly susceptible to storm surges in the winter months when there is an increased propensity for storms and a higher mean sea level caused by southeasterly winds and geostrophic adjustment \citep{danard2003storm}.
+In the summer months, the wind direction changes to northwesterly and mean sea level is lower.
+In British Columbia waters, the largest tidal range is typically observed in the winter months during the perigean spring tide or ``King" tide, when the sun and moon are aligned with the Earth and the moon is at its perigee.
+As such, storm surges are typically observed between Novemeber and Feburary.
+Several historical examples have been documented by \citet{forseth2006adaptation}.
 %work on this

### 2014-07-24 RoughWork/manuscript.tex
Worked on past models of SoG in intro

@@
 \title{Storm surge modelling in the Strait of Georgia}
-\author{Nancy Soontiens, Susan Allen, Doug Latornell, Kate Le Sou\"{e}f}
+\author{Nancy Soontiens, Susan Allen, Doug Latornell, Kate Le Sou\"{e}f, Idalia Machuca}
 \date{\today}
@@
 This investigation produces storm surge hindcasts using a three-dimensional numerical ocean model for the Strait of Georgia and the surrounding bodies of water (Strait of Juan de Fuca, Puget Sound, and Johnstone Strait) collectively known as the Salish Sea.
-The numerical model employs the Nucleus for European Modelling of the Ocean (NEMO) in a regional configuration.
+The numerical model employs the Nucleus for European Modelling of the Ocean (NEMO) architecture in a regional configuration.
 The model is evaluated through comparisons of tidal elevation harmonics, tidal current ellipses, and storm surge water levels with observations.
 Important forcing factors contributing to storm surges have been assessed, finding that the surge entering the domain from the Pacific Ocean contributes most significantly.
+In addition, local wind patterns cause a slight increase in surge amplitude on the mainland side of the Strait of Georgia when compared to the Vancouver Island coastal areas.
+Surge amplitudes are found to be greater within the Strait of Georgia when compared to amplitudes in the Strait of Juan de Fuca.
 %needs a little more work.
@@
 %some general intro paragraph here
+%Need something about SoG size, geography, islands, sills, rivers....
 
@@
 %overview oF SoG/Salish Sea models
-Early efforts by \citet{crean1988numerical, stronach1993update} implemented a series of models covering the Strait of Juan de Fuca, the Strait of Georgia and Puget Sound using a finite difference discretization of the equations of motion.
+Early work by \citet{crean1988numerical, crean1988mathematical, stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia and Puget Sound using a finite difference discretization of the equations of motion.
+A barotropic version of their model was used to reproduce storm surge water level elevations with reasonable accuracy, however, this barotropic version did not include the effects of fresh water entering the system from rivers.
+Later, a baroclinc version of their model examined residual currents and more.
+%elaborate!
+ The baroclinic model of \citet{masson2004modelling} examined seasonal variability of deep water properties in the Strait of Georgia noting a strong sensitivity to vertical mixing over the sills.
+This model employed a sigma vertical coordinate system leading to errors in diapycnal mixing due to steep topographics slopes.
+%fix this up still.
 %tide models (Foreman), baroclinic (Masson, PNWL, Sutherland and MacCready)

### 2014-07-24 RoughWork/manuscript.tex
Excluded the background state from tide-surge comparison

@@
 
-\subsection{Model domain}
 The modelled domain extends from the Strait of Juan de Fuca to Puget Sound to Johnstone Strait as shown in Figure \ref{fig:domain}.
@@
 
-\subsection{Boundary conditions and sub-grid scale processes}
 The model includes two open boundaries that connect to the Pacific Ocean, the western boundary of the Strait of Juan de Fuca as well as Johnstone Strait at the north, both of which are forced with eight tidal constituents (K$_1$ ,O$_1$, P$_1$, Q$_1$, M$_2$, K$_2$, N$_2$, S$_2$), temperature and salinity climatologies, and the sea surface height anomaly.
@@
 These cases have been used to asses the model's skill in reproducing water level height and surge amplitude and timing during storm events.
-In addition, the contribution to the surge of several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed.
+In addition, the contribution to the surge due to several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed.
 We have also examined the spatial variabilty of the surge as it propagates through the domain and have considered a case with high-resolution atmospheric forcing.
@@
 
-Finally, an assessment of the importance of including tide-surge interactions is presented next.
+%Finally, an assessment of the importance of including tide-surge interactions is presented next.
 A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}.
@@
 
-%Finally, an assessment of the importance of including tide-surge interactions is presented next.
+Finally, an assessment of the importance of including the tide-surge interaction is presented next.
 A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}.
 Recall, the modelled residual is defined as the difference between the sea surface height in a simulation with all forcing and a simulaltion with tides amd rivers only.
-A comparison between the modelled residual and a surge only run can determine the significance of the tide-surge interaction.
-The most significant differences occur at Point Atkinson where the maximum modelled residual is 7 cm higher than the maximum of the surge only case.
+A comparison between the modelled residual and a surge-only run can determine the significance of the tide-surge interaction.
+The most significant differences occur at Point Atkinson where the maximum modelled residual is 7 cm lower than the maximum of the surge-only case.
 In contrast, the tide-surge interaction at Victoria is hardly noticeable.
-Indeed, a spatial map of the modelled residual and surge only differences (not shown here) indicate that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands.
+Indeed, a spatial map of the modelled residual and surge-only differences (not shown here) indicate that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands.
 The passages between these islands are known for vigorous tidal mixing, which may have an effect on the surge propagation.
 However, the effect is a relatively small underestimate of the surge elevation and no changes to the timing of the surge.
-So, in operational settings, running a surge only simulation and adding the predicted tides a posteriori could be more computationally efficient with only a small decrease in the accuracy of the prediction.
+So, in operational settings, performing a surge-only simulation and adding the predicted tides a posteriori could be more computationally efficient with only a small decrease in the accuracy of the prediction.
 
@@
 \centering
-\includegraphics[scale=0.6]{Figures/feb2006_tidesurge_bg.pdf}
-\caption{Modelled residuals (solid) compared a simulation of the surge propagation without tidal forcing (dashed).}
+\includegraphics[scale=0.6]{Figures/feb2006_tidesurge.pdf}
+\caption{Modelled residuals (solid) compared a simulation of the surge propagation without tidal forcing (dashed). A background state has not been removed from these plots. }
 \label{fig:tidesurge}

### 2014-07-24 RoughWork/manuscript.tex
Worked on modelling overview in intro

@@
 The Strait of Georgia is a large, semi-enclosed body of water between Vancouver Island and the mainland of British Columbia connected to the Pacific Ocean via the Strait of Juan de Fuca at the south and Johnstone Strait at the north.
-During the winter months, coastal communities in the Strait of Georgia are particularly susceptible to flooding caused by storm surges, a natural hazard that can occur when a strong storm coincides with high tide.
+During the winter months, coastal communities in the Strait of Georgia are at risk to flooding caused by storm surges, a natural hazard that can occur when a strong storm coincides with high tide.
 This investigation produces storm surge hindcasts using a three-dimensional numerical ocean model for the Strait of Georgia and the surrounding bodies of water (Strait of Juan de Fuca, Puget Sound, and Johnstone Strait) collectively known as the Salish Sea.
@@
 
-%some general intro paragraph here
-%Need something about SoG size, geography, islands, sills, rivers....
+%some general intro paragraph here. Need something about SoG size, geography, islands, sills, rivers, tides....
+The Strait of Georgia is a strongly stratified, semi-enclosed body of water located between Vancouver Island and the mainland of British Columbia.
+It is part of a larger system of waterways collectively known as the Salish Sea and is connected to the Pacific Ocean via the Strait of Juan de Fuca and Johnstone Strait.
 
 %my attempt at physical oceanopgraphy in SoG- Susan feel free to edit.
-The Strait of Georgia is a strongly stratified, semi-enclosed body of water located between Vancouver Island and the mainland of British Columbia.
-It is part of a larger system of waterways collectively known as the Salish Sea and is connected to the Pacific Ocean via the Strait of Juan de Fuca and Johnstone Strait.
 Several features related to the physical oceanopgraphy of this region render it a challenge to model numerically.
@@
 %overview oF SoG/Salish Sea models
-Early work by \citet{crean1988numerical, crean1988mathematical, stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia and Puget Sound using a finite difference discretization of the equations of motion.
+Early work by \citet{crean1988mathematical,crean1988numerical} and \citet{stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia and Puget Sound using a finite difference discretization of the equations of motion.
 A barotropic version of their model was used to reproduce storm surge water level elevations with reasonable accuracy, however, this barotropic version did not include the effects of fresh water entering the system from rivers.
-Later, a baroclinc version of their model examined residual currents and more.
-%elaborate!
- The baroclinic model of \citet{masson2004modelling} examined seasonal variability of deep water properties in the Strait of Georgia noting a strong sensitivity to vertical mixing over the sills.
-This model employed a sigma vertical coordinate system leading to errors in diapycnal mixing due to steep topographics slopes.
-%fix this up still.
-%tide models (Foreman), baroclinic (Masson, PNWL, Sutherland and MacCready)
+Later, a baroclinc version of their model examined residual circulation and fresh water flushing rates \citep{stronach1993update, marinone1996model}.
+This early work pointed to an important connection between the model's bottom friction parameterization and the location and amplitude of the M$_2$ amphidrome.
+They found that a local increase in the bottom friction coefficient, especially in the area of Haro Strait, was necessary in order to adequately reproduce the observed tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}.
+Other studies over a similar domain, such as the barotropic finite element model of \citet{foreman1995tidal}, have avoided local increases in bottom friction.
+\citet{foreman2004m} used data assimilation techinques to study M$_2$ tidal dissipation without local changes to the bottom frction in a two-dimensional finite element model.
+
+More recently, modelling studies have focused on baroclinic features.
+\citet{masson2004modelling} used the Princeton Ocean Model (POM) in a regional mode to examine the seasonal variability of deep water properties in the Strait of Georgia noting a strong sensitivity to vertical mixing over the sills.
+This model employed a sigma or terrain-following vertical coordinate system leading to spurious diapycnal mixing since, numerically, the lateral mixing is computed on sigma levels which may intersect isopycnals in the presence of steep topography.
+\citet{yang2010multi} developed a model of southern Strait of Georgia and Puget Sound using volume techniques on an unstructured grid.
+\citet{sutherland2011model} employed the Regional Ocean Modeling System (ROMS) to study estuarine circulation in the Salish Sea with a model domain that extended outside the Strait of Juan de Fuca into the Pacific Ocean and up to the northern Strait of Georgia.
+The passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.
+These studies have led to an improvement in understanding circulation properties in the Salish Sea and have highlighted important features that should be considered in future modelling work.
 
@@
 %overview of paper: model configuration, model evaluation, storm surge hindcasts.
-This paper is organized as follows: an overivew of the model configuration and domain including a description of open and surface boundary conditions is provided in section \ref{sec:config}.
-This section also describes the ocean model we have employed as well as choices in grid structure and mixing parameterizations.
+This paper is organized as follows: an overview of the model configuration and domain including a description of the ocean model we have employed  is provided in section \ref{sec:config}.
+This section also describes open and surface boundary conditions as well as choices in grid structure and mixing parameterizations.
 Next, section \ref{sec:model} presents an evaluation of model performance through comparisons of tidal amplitudes and phases with observations.
@@
 \end{figure}
-%image needs some work: labels for tide comparisons?
+%image may need more work
 
@@
 In contrast, the tide-surge interaction at Victoria is hardly noticeable.
-Indeed, a spatial map of the modelled residual and surge-only differences (not shown here) indicate that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands.
+Indeed, a spatial map of the modelled residual and surge-only difference (not shown here) indicates that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands.
 The passages between these islands are known for vigorous tidal mixing, which may have an effect on the surge propagation.
@@
 
-Point Atkinson observed a maximum surge amplitude of 84 cm, which is approximately 20 cm higer than the model result, however the timing of the surge agrees almost exactly.
+Point Atkinson observed a maximum surge amplitude of 84 cm, which is approximately 20 cm higher than the model result, however the timing of the surge agrees almost exactly.
 The best agreement occurs at Victoria where the observed surge was 62 cm and the model returned 54 cm.
 At all of the locations, the model predicted the maximum surge to occur within 3.5 hours of the observed surge.
-Typically, the modelled surge was too early, in contrast to the later predictions in the Feb, 4 2006 case.
+Typically, the modelled surge was too early, in contrast to the later predictions in the Feb 4, 2006 case.
 Errors in the modelled water level at the time of maximum surge were of the same order as the errors in the surge amplitude.

### 2014-07-24 RoughWork/manuscript.tex
A little bit more on intro paragraph

@@
 It is part of a larger system of waterways collectively known as the Salish Sea and is connected to the Pacific Ocean via the Strait of Juan de Fuca and Johnstone Strait.
+It is on average 222 km in length and 28 km wide and its maximum depth is 420 m \citep{thomson1981oceanography}.
+At its southern end, the Strait of Georgia is connected to the Strait of Juan de Fuca through several passages, the largest being Haro Strait.
+The San Juan and Gulf Islands are scattered throughout these passages where complex bathymetry and steep topographic slopes induce vigorous tidal mixing.
+Many rivers feed fresh water into the Salish Sea system, the most siginificant source being from the Fraser River.
+Mixed diurnal and semi-diurnal tides are found throughout most of the Salish Sea but a semi-dirunal amphidrome is present nearby Victoria, where the tides are mainly diurnal.
+This system is home to several large coastal communities, such as Vancouver and Victoria, serves as a waterway for commericial traffic and recreational use, and supports a large and active fishing industry.
 
@@
 %work on this
-
+%something about fetch
 %storm surge modelling

### 2014-07-26 RoughWork/manuscript.tex
Included description of nove 2009

@@
 
-\subsection{Strong Wind Event: Nov 18, 2009}
+\subsection{Strong Wind Event: Nov 19, 2009}
+Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered.
+As recorded in the Environment Canada Historical Climate Data archives, the maximum hourly wind speed at Popint Atkinson was 18.6 m/s and the largest gust observed was 25.8 m/s or 93 km/hr.
+In addtion, large wind speeds over 15 m/s persisted for over 9 hours.
+Yet, the surge associated with this event is rather small when compared the previous two cases as shown in Figure \ref{fig:nov2009} where we have plotted the wind speed at Point Atkinson, sea surface height anomaly forcing at the open boundaries, and observed and modelled residuals at Point Atkinson.
+First, the largest surge throughoutt this time frame did not coincide with the large wind event, reflected in both the model and observations.
+The observed surge reached 62.6 cm on Nov 20 at 10:00 hr, while the model recorded a maximum surge of 38.8 cm five hours later.
+The surge did not coincide with a high tide and so no high water level events were recorded.
+Although not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and a more siginificant surge entering the system at Tofino and Port Hardy.
+
+\citet{danard2003storm} discuss the role of wind fetch, the distance over which wind acts on the surface of a water basin, in setting up seiche motion in lakes.
+As wind blows over the surface of a lake, the water level becomes elevated on the downwind side and depressed on the upwind side.
+The longer the fetch, the more significant the tilting of the water surface.
+These changes in water level can also contribute to storm surges.
+During the high wind event on Nov 19, the winds at Point Atkinson were directed to the west, explaining the lack of surge associated with this high wind event since wind stresses would act to depress the water on the east side of the basin.
+Perhaps the small surge that occured after the high wind event has formed in response to the sea surface returning to equilibrium in a seiche-like motion after the winds have relaxed.
+The model appears to reproduce this surge with reasonable accuracy, however, there are some discrepencies between weather observations and the atmospheric forcing in use.
+A more detailed look at the senstivity of modelled surges to atmospheric forcing is considered in a case using a high-resolution weather product.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/nov2009_bg.pdf}
+\caption{Wind and storm surge conditions at Point Atkison from  strong wind event on Nov 19, 2009.
+Top: Wind speed from Environment Canada observations at Point Atkinson (solid) and and from the atmophseric model (CGRF) at a grid point near Point Atkinson (dashed).
+Middle: Sea surace height anomaly at Tofino (solid) and Port Hardy (dashed) used as forcing conditions at the western and northern open boundaries.
+Bottom: Observed (solid) and modelled (dashed) residuals at Point Atkinson.
+}
+\label{fig:nov2009}
+\end{figure}
 
@@
 
-\subsection{Statisitics}
+\subsection{Statistical evaluation}
 %I think Hal Ritchie's talk with the storm surges had some stats presented. Look this up...

### 2014-07-29 RoughWork/manuscript.tex
Added references for high tides.

--- 2014-07-24 RoughWork/manuscript.tex
@@
 During the winter months, coastal communities in the Strait of Georgia are at risk to flooding caused by storm surges, a natural hazard that can occur when a strong storm coincides with high tide.
-This investigation produces storm surge hindcasts using a three-dimensional numerical ocean model for the Strait of Georgia and the surrounding bodies of water (Strait of Juan de Fuca, Puget Sound, and Johnstone Strait) collectively known as the Salish Sea.
+This investigation produces storm surge hind casts using a three-dimensional numerical ocean model for the Strait of Georgia and the surrounding bodies of water (Strait of Juan de Fuca, Puget Sound, and Johnstone Strait) collectively known as the Salish Sea.
 The numerical model employs the Nucleus for European Modelling of the Ocean (NEMO) architecture in a regional configuration.
@@
 It is part of a larger system of waterways collectively known as the Salish Sea and is connected to the Pacific Ocean via the Strait of Juan de Fuca and Johnstone Strait.
+It is on average 222 km in length and 28 km wide and its maximum depth is 420 m \citep{thomson1981oceanography}.
+At its southern end, the Strait of Georgia is connected to the Strait of Juan de Fuca through several passages, the largest being Haro Strait.
+The San Juan and Gulf Islands are scattered throughout these passages where complex bathymetry and steep topographic slopes induce vigorous tidal mixing.
+Many rivers feed fresh water into the Salish Sea system, the most significant source being from the Fraser River.
+Mixed diurnal and semi-diurnal tides are found throughout most of the Salish Sea but a semi-diurnal amphidrome is present nearby Victoria, where the tides are mainly diurnal.
+This system is home to several large coastal communities, such as Vancouver and Victoria, serves as a waterway for commercial traffic and recreational use, and supports a large and active fishing industry.
 
-%my attempt at physical oceanopgraphy in SoG- Susan feel free to edit.
-Several features related to the physical oceanopgraphy of this region render it a challenge to model numerically.
+%my attempt at physical oceanography in SoG- Susan feel free to edit.
+Several features related to the physical oceanography of this region render it a challenge to model numerically.
 The outflow from the Fraser River in the late spring is the primary source fresh water within the Strait of Georgia and leads to large vertical density gradients in the summer months.
@@
 A barotropic version of their model was used to reproduce storm surge water level elevations with reasonable accuracy, however, this barotropic version did not include the effects of fresh water entering the system from rivers.
-Later, a baroclinc version of their model examined residual circulation and fresh water flushing rates \citep{stronach1993update, marinone1996model}.
-This early work pointed to an important connection between the model's bottom friction parameterization and the location and amplitude of the M$_2$ amphidrome.
+Later, a baroclinic version of their model examined residual circulation and fresh water flushing rates \citep{stronach1993update, marinone1996model}.
+This early work pointed to an important connection between the model's bottom friction parametrization and the location and amplitude of the M$_2$ amphidrome.
 They found that a local increase in the bottom friction coefficient, especially in the area of Haro Strait, was necessary in order to adequately reproduce the observed tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}.
 Other studies over a similar domain, such as the barotropic finite element model of \citet{foreman1995tidal}, have avoided local increases in bottom friction.
-\citet{foreman2004m} used data assimilation techinques to study M$_2$ tidal dissipation without local changes to the bottom frction in a two-dimensional finite element model.
+\citet{foreman2004m} used data assimilation techniques to study M$_2$ tidal dissipation without local changes to the bottom friction in a two-dimensional finite element model.
 
@@
 \citet{yang2010multi} developed a model of southern Strait of Georgia and Puget Sound using volume techniques on an unstructured grid.
-\citet{sutherland2011model} employed the Regional Ocean Modeling System (ROMS) to study estuarine circulation in the Salish Sea with a model domain that extended outside the Strait of Juan de Fuca into the Pacific Ocean and up to the northern Strait of Georgia.
+\citet{sutherland2011model} employed the Regional Ocean Modelling System (ROMS) to study estuarine circulation in the Salish Sea with a model domain that extended outside the Strait of Juan de Fuca into the Pacific Ocean and up to the northern Strait of Georgia.
 The passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.
@@
 
-Communities in the Strait of Georgia are partiuclarly susceptible to storm surges in the winter months when there is an increased propensity for storms and a higher mean sea level caused by southeasterly winds and geostrophic adjustment \citep{danard2003storm}.
+Communities in the Strait of Georgia are particularly susceptible to storm surges in the winter months when there is an increased propensity for storms and a higher mean sea level caused by southeasterly winds and geostrophic adjustment \citep{danard2003storm}.
 In the summer months, the wind direction changes to northwesterly and mean sea level is lower.
-In British Columbia waters, the largest tidal range is typically observed in the winter months during the perigean spring tide or ``King" tide, when the sun and moon are aligned with the Earth and the moon is at its perigee.
-As such, storm surges are typically observed between Novemeber and Feburary.
+In British Columbia waters, the largest tidal ranges occur in June and December at the summer and winter solstices \citep{thomson1981oceanography}.
+The highest water levels are typically observed in the winter around the perigean spring tide or ``King" tide, when the sun and moon are aligned with the Earth and the moon is at its perigee.
+As such, storm surges are typically observed between November and February.
 Several historical examples have been documented by \citet{forseth2006adaptation}.
 %work on this
+%something about fetch
 
@@
 
-%overview of paper: model configuration, model evaluation, storm surge hindcasts.
+%overview of paper: model configuration, model evaluation, storm surge hind casts.
 This paper is organized as follows: an overview of the model configuration and domain including a description of the ocean model we have employed  is provided in section \ref{sec:config}.
-This section also describes open and surface boundary conditions as well as choices in grid structure and mixing parameterizations.
+This section also describes open and surface boundary conditions as well as choices in grid structure and mixing parametrizations.
 Next, section \ref{sec:model} presents an evaluation of model performance through comparisons of tidal amplitudes and phases with observations.
-The model's accuracy in producing storm surge hindcasts is also assesed in section \ref{sec:storm}.
+The model's accuracy in producing storm surge hind casts is also assessed in section \ref{sec:storm}.
 Finally, a summary and discussion of future research goals is provided in section \ref{sec:diss}.
@@
 We have used the Nucleus for European Modelling of the Ocean (NEMO) framework in its regional configuration to develop an ocean model for the Strait of Georgia and Salish Sea.
-NEMO is a highly modularized tool used for studying ocean physics, ocean-ice interactions, and the biogeochemical properties of the ocean.
+NEMO is a highly modularized tool used for studying ocean physics, ocean-ice interactions, and the bio-geochemical properties of the ocean.
 NEMO's ocean core solves the three-dimensional hydrostatic equations of motion for an incompressible fluid under the Boussinesq approximation on a structured computational grid.
-Although not used in the present work, NEMO's options for grid nesting and biogeochemical coupling make it a useful tool for studying the complex physics and biogeochemical interactions within the Strait of Georgia.
+Although not used in the present work, NEMO's options for grid nesting and bio-geochemical coupling make it a useful tool for studying the complex physics and bio-geochemical interactions within the Strait of Georgia.
 This work focuses on validating the physical set up of the Salish Sea model, in particular, determining appropriate forcing and boundary conditions for accurate reproduction of tidal amplitudes and phases as well as storm surge elevations.
-Future work will include biogeochemical coupling and data assimilation.
+Future work will include bio-geochemical coupling and data assimilation.
 
 The modelled domain extends from the Strait of Juan de Fuca to Puget Sound to Johnstone Strait as shown in Figure \ref{fig:domain}.
-Bathymetry from the Cascadia physiography dataset \citep{haugerud1999digital} was smoothed to limit the difference in depth across grid cells.
+Bathymetry from the Cascadia physiography data set \citep{haugerud1999digital} was smoothed to limit the difference in depth across grid cells.
 For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells.
@@
 The model time step is and barotropic time step are 10 s and 2 s respectively.
-Small vertical grid spacing and large vertical velocities in the Boundary Pass and Haro Strait region impose a fairly restricitive step through the vertical CFL condition.
+Small vertical grid spacing and large vertical velocities in the Boundary Pass and Haro Strait region impose a fairly restrictive step through the vertical CFL condition.
 
@@
 At the ocean surface, meteorological conditions and fresh water input from rivers are included.
-The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforcasting model suitable for use in ocean modelling \citep{smith2013new}.
+The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2013new}.
 Forecasts from the period of 2002-2011 are available.
@@
 River input provides a significant volume of freshwater to the Salish Sea and can influence stratification, circulation and primary productivity.
-However, most rivers in the domain are not gauged so parameterizations were required to represent river flow.
+However, most rivers in the domain are not gauged so parametrizations were required to represent river flow.
 \citet{morrison2011rivers} provides a method for estimating freshwater runoff in the Salish Sea region based on precipitation.
@@
 Incoming water was assumed to be fresh and at surface temperature.
-A total of 150 rivers were parameterised by this method and their locations are indicated by the blue dots in Figure \ref{fig:domain}.
+A total of 150 rivers were parametrized by this method and their locations are indicated by the blue dots in Figure \ref{fig:domain}.
 
@@
 Conditions were initially uniform horizontally and velocity was initialized at zero.
-The model was spun up for a 15.5 months from the initial conditions above, starting Sep 16, 2002, using atmospheric forcing from 2002-2003, climatological temperature and salinity and sea surface height at the boundaries, with tides and climatological river output.
+The model was spun up for a 15.5 months from the initial conditions above, starting Sept 16, 2002, using atmospheric forcing from 2002-2003, climatological temperature and salinity and sea surface height at the boundaries, with tides and climatological river output.
 All storm surge runs were started three days prior to the event of interest with zero initial velocities and sea surface height and a stratification profile from model spin up.
@@
 The model was initially evaluated qualitatively by comparing patterns of tidal amplitude and phase to results from \citep{foreman1995tidal}.
-For example, the amphidromic dome around Victoria was produced in the M$_2$ results, as well as the monotonic increase in $K_1$ amplitude moving northwards along the Strait of Georgia.
+For example, the amphidrome around Victoria was produced in the M$_2$ results, as well as the monotonic increase in $K_1$ amplitude moving northwards along the Strait of Georgia.
 Initially, the Johnstone Strait boundary was closed, but modelled M$_2$ amplitudes were too small compared to measured amplitudes...
@@
 
-\section{Storm Surge Hindcasts}\label{sec:storm}
+\section{Storm Surge Hind casts}\label{sec:storm}
 %outline: 
@@
 %3. High resolution wind (hopefully Dec 2012)
-A series of storm surge hindcasts is presented next.
+A series of storm surge hind casts is presented next.
 Storm surge simulations were started from rest at least three days prior to the event of interest to allow for adjustment of model currents.
@@
 Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this cae, from the year prior surge event on interest.
-Additionally, several simulations with varying forcing conditions such as removing atmospsheric forcing or removing the sea surface height anomaly at the open boundaries have been completed.
+Additionally, several simulations with varying forcing conditions such as removing atmospheric forcing or removing the sea surface height anomaly at the open boundaries have been completed.
 
-Several significiant surge or wind events have been considered in detail, as well as an array of hindcasts to asses model ability in a statistical measure.
+Several significant surge or wind events have been considered in detail, as well as an array of hind casts to asses model ability in a statistical measure.
 These cases have been used to asses the model's skill in reproducing water level height and surge amplitude and timing during storm events.
 In addition, the contribution to the surge due to several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed.
-We have also examined the spatial variabilty of the surge as it propagates through the domain and have considered a case with high-resolution atmospheric forcing.
+We have also examined the spatial variability of the surge as it propagates through the domain and have considered a case with high-resolution atmospheric forcing.
 
 %First, a study of a large surge event on Feb 4, 2006 is presented.
-This event caused siginificant damage and flooding in Boundary Bay, a community south of Richmond, (citation) and has been chosen as a case study to demonstrate the model's skill in reproducing a storm surge and to quantitatively assess forcing conditions that are important in storm surge hindcasting.
+This event caused significant damage and flooding in Boundary Bay, a community south of Richmond, (citation) and has been chosen as a case study to demonstrate the model's skill in reproducing a storm surge and to quantitatively assess forcing conditions that are important in storm surge hind casting.
 Second, an extreme wind event on Dec 15, 2006 is examined.
-This event produced winds over (x) (citation) but did not result in significant flooding because the storm did not occur at high tide, however, it did result in siginificant property damage due to high winds and waves.
+This event produced winds over (x) (citation) but did not result in significant flooding because the storm did not occur at high tide, however, it did result in significant property damage due to high winds and waves.
 This event has been used to understand the spatial and temporal variability of the surge propagation.
-Next, a large wind event on Nov 18, 2009 that did not result in a siginificant surge or flooding is presented.
-This case was choosen to identify the model's capability in handling large wind events that do not result in high water levels.
+Next, a large wind event on Nov 18, 2009 that did not result in a significant surge or flooding is presented.
+This case was chosen to identify the model's capability in handling large wind events that do not result in high water levels.
 Next, a case with high resolution atmospheric forcing for a recent surge on Dec 17, 2012 is discussed.
@@
 
-On Feb 4, 2006, a strong storm system with winds over 75 km/hr coinciding with an unsually high tide led to siginificant flooding in coastal comminuties of the Strait of Georgia \citep{romanowski2010storm}.
+On Feb 4, 2006, a strong storm system with winds over 75 km/hr coinciding with an unusually high tide led to significant flooding in coastal communities of the Strait of Georgia \citep{romanowski2010storm}.
 The maximum water level observed at Point Atkinson on this day was 5.49 m, which is 0.8 m higher than the predicted tides, highlighting the significance of this event as it resulted in a large anomaly and also reached water levels close to the highest level ever recorded (5.61 m).
-A set of simulations from Feb 1 to 8, 2006 under varying forcing conditions has been completed in order to determine the siginicance of each factor contributing to the surge.
+A set of simulations from Feb 1 to 8, 2006 under varying forcing conditions has been completed in order to determine the significance of each factor contributing to the surge.
 
 The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations and corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River.
-The timing of maximum water level matches well between the observations and the model at all four locations, the most siginificant contribution to high water level being from the tides.
+The timing of maximum water level matches well between the observations and the model at all four locations, the most significant contribution to high water level being from the tides.
 There is also a good match in amplitude of the water level maximum at Point Atkinson and Campbell River, with a relative error less than 4\%.
 There is only a 7.2 cm difference between observed and modelled water level elevation at Point Atkinson and a 20.9 cm difference at Campbell River.
-Tthe error at Victoria and Patricia Bay is slightly more siginificant, reaching 9.6\% and 6.3\% respectively (close to 30 cm in each case), partly due to error in tidal amplitudes as discussed above.
+The error at Victoria and Patricia Bay is slightly more significant, reaching 9.6\% and 6.3\% respectively (close to 30 cm in each case), partly due to error in tidal amplitudes as discussed above.
 The precise location and magnitude of the M$_2$ amphidrome around Victoria is very sensitive to model parameters such as bottom friction.
@@
 When atmospheric forces are neglected the timing of the maximum surge is earlier at Point Atkinson but later at Campbell River.
-Campbell River exhibits a strong response to atmopsheric conditions during midday on Feb 4, as demonstrated in a two sharp spikes in both the blue and red curves arounf 12:00 hr The green curve, which does not include any atmopsheric focring, does not exhibit this behaviour.
+Campbell River exhibits a strong response to atmospheric conditions during midday on Feb 4, as demonstrated in a two sharp spikes in both the blue and red curves around 12:00 hr The green curve, which does not include any atmospheric coring, does not exhibit this behaviour.
 The result is a peak surge earlier than in day than the no atmospheric forcing case.
-This response is also noted in the observations at Campbell River in Figure \ref{fig:feb2006} where the maxmium surge occurs earlier than at the other locations.
+This response is also noted in the observations at Campbell River in Figure \ref{fig:feb2006} where the maximum surge occurs earlier than at the other locations.
 At Victoria and Patricia Bay, the surge amplitude is mainly unaffected by the atmospheric forcing.
@@
 
-These comparisons point to the importance of including, as a forcing condition at the open bounday, the surge entering the domain from the Pacific Ocean.
-Storms travelling over the Pacific Ocean torwads the west coast of the northern United States and Canada can induce elevated sea levels along the coastline which then enter the Salish Sea sytem through the Strait of Juan de Fuca.
-In a model domain that does not inlcude the Pacific Ocean, this effect must be inlcuded as a boundary condition.
-Previous simulations by \citet{murty1995storm} suggest that the inverese barometer effect and the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.
-A detailed assessment of the effect of the local atmospheric condtions and winds is included later where simulations have been performed with high resolution atmospheric forcing data.
+These comparisons point to the importance of including, as a forcing condition at the open boundary, the surge entering the domain from the Pacific Ocean.
+Storms travelling over the Pacific Ocean towards the west coast of the northern United States and Canada can induce elevated sea levels along the coastline which then enter the Salish Sea system through the Strait of Juan de Fuca.
+In a model domain that does not include the Pacific Ocean, this effect must be included as a boundary condition.
+Previous simulations by \citet{murty1995storm} suggest that the inverse barometer effect and the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.
+A detailed assessment of the effect of the local atmospheric conditions and winds is included later where simulations have been performed with high resolution atmospheric forcing data.
 
@@
 A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}.
-Recall, the modelled residual is defined as the difference between the sea surface height in a simulation with all forcing and a simulaltion with tides amd rivers only.
+Recall, the modelled residual is defined as the difference between the sea surface height in a simulation with all forcing and a simulation with tides and rivers only.
 A comparison between the modelled residual and a surge only run can determine the significance of the tide-surge interaction.
@@
 A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}.
-Recall, the modelled residual is defined as the difference between the sea surface height in a simulation with all forcing and a simulaltion with tides amd rivers only.
+Recall, the modelled residual is defined as the difference between the sea surface height in a simulation with all forcing and a simulation with tides and rivers only.
 A comparison between the modelled residual and a surge-only run can determine the significance of the tide-surge interaction.
@@
 On Dec 15, 2006, a major storm with wind gusts over 120 km/hr caused significant damage to many coastal communities on the Pacific Northwest.
-Although the storm was very strong and the observed residual at Point Atkinson reached 84 cm, this event did not coincide with high tide and so no siginificant or reported flooding occurred in the Vancouver area\citep{forseth2006adaptation}.
-This event has been used to examine the propgation of the surge throughout the domain.
-First, an assessment of the accuracy of the hindcast is provided.
+Although the storm was very strong and the observed residual at Point Atkinson reached 84 cm, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area\citep{forseth2006adaptation}.
+This event has been used to examine the propagation of the surge throughout the domain.
+First, an assessment of the accuracy of the hind cast is provided.
 
@@
 Notably, the Strait of Georgia experiences the largest surge values, upwards of 70 cm, whereas the Juan de Fuca region experiences a maximum surge around 55 cm.
-Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges than the Vancouver Island side, likley due to the southerly wind in the atmosheric model pushing water against the mainland coast.
+Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast.
 
@@
 
-\subsection{Strong Wind Event: Nov 18, 2009}
+\subsection{Strong Wind Event: Nov 19, 2009}
+Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered.
+As recorded in the Environment Canada Historical Climate Data archives, the maximum hourly wind speed at Point Atkinson was 18.6 m/s and the largest gust observed was 25.8 m/s or 93 km/hr.
+In addition, large wind speeds over 15 m/s persisted for over 9 hours.
+Yet, the surge associated with this event is rather small when compared the previous two cases as shown in Figure \ref{fig:nov2009} where we have plotted the wind speed at Point Atkinson, sea surface height anomaly forcing at the open boundaries, and observed and modelled residuals at Point Atkinson.
+Two significant surge events are noted, the first occurring on Nov 19 and a second, larger surge on Nov 20.
+The largest surge did not coincide with the strong wind event which peaked at Nov 19 01:00 hr, reflected in both the model and observations.
+The highest observed surge reached 62.6 cm on Nov 20 at 10:00 hr, while the model recorded a maximum surge of 38.8 cm five hours later.
+The surge did not occur during a high tide and so no high water level events were recorded.
+Although not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and a more significant surge entering the system at Tofino and Port Hardy.
+
+\citet{danard2003storm} discuss the role of wind fetch, the distance over which wind acts on the surface of a water basin, in setting up seiche motion in lakes and its contribution to storm surges.
+As wind blows over the surface of a lake, the water level becomes elevated on the downwind side and depressed on the upwind side in response to the wind-induced currents.
+The longer the fetch, the more significant the tilting of the water surface.
+These changes in water level can also contribute to storm surges.
+During the high wind event on Nov 19, the winds at Point Atkinson were directed to the west, explaining the lack of surge associated with this high wind event since wind stresses would act to depress the water on the east side of the basin.
+It is possible that the surge occurring on Nov 19 formed in response to the sea surface returning to equilibrium in a seiche-like motion after the winds have relaxed.
+The model appears to reproduce this surge with reasonable accuracy, however, there are some discrepancies in the wind speed and direction between weather observations and the atmospheric forcing employed.
+A more detailed look at the sensitivity of modelled surges to atmospheric forcing is considered in a case with a high-resolution weather product.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/nov2009_bg.pdf}
+\caption{Wind and storm surge conditions at Point Atkinson for a strong wind event on Nov 19, 2009.
+Top: Wind speed from Environment Canada observations at Point Atkinson (solid) and from the atmospheric model (CGRF) at a grid point near Point Atkinson (dashed).
+Middle: Sea surface height anomaly at Tofino (solid) and Port Hardy (dashed) used as forcing conditions at the western and northern open boundaries.
+Bottom: Observed (solid) and modelled (dashed) residuals at Point Atkinson.
+}
+\label{fig:nov2009}
+\end{figure}
 
@@
 %future directions: running real time, biology and chemistry, model improvements?
-%comments on wetting and drying (maybe most imporant in shallow regions)
+%comments on wetting and drying (maybe most important in shallow regions)
 

### 2014-07-29 RoughWork/manuscript.tex
Storm surge modelling in intro

@@
 This model employed a sigma or terrain-following vertical coordinate system leading to spurious diapycnal mixing since, numerically, the lateral mixing is computed on sigma levels which may intersect isopycnals in the presence of steep topography.
-\citet{yang2010multi} developed a model of southern Strait of Georgia and Puget Sound using volume techniques on an unstructured grid.
+\citet{yang2010multi} developed a model of southern Strait of Georgia and Puget Sound using finite volume techniques on an unstructured grid.
 \citet{sutherland2011model} employed the Regional Ocean Modelling System (ROMS) to study estuarine circulation in the Salish Sea with a model domain that extended outside the Strait of Juan de Fuca into the Pacific Ocean and up to the northern Strait of Georgia.
@@
 Additionally, there is a small contribution to increased sea level due to the thermal expansion of water during warmer years set by the El Nino Southern Oscillation (ENSO) \citep{abeys2011extreme}.
-%more on past events, significance
 
@@
 In British Columbia waters, the largest tidal ranges occur in June and December at the summer and winter solstices \citep{thomson1981oceanography}.
-The highest water levels are typically observed in the winter around the perigean spring tide or ``King" tide, when the sun and moon are aligned with the Earth and the moon is at its perigee.
-As such, storm surges are typically observed between November and February.
+The highest water levels are observed in the winter around the perigean spring tide or ``King" tide, when the sun and moon are aligned with the Earth and the moon is at its perigee.
+As such, storm surges in the Strait of Georgia are typically observed between November and February.
 Several historical examples have been documented by \citet{forseth2006adaptation}.
-%work on this
-%something about fetch
 
 %storm surge modelling
+%murty, danard, crean, non BC models (xu, flather)
+Coastal communities can prepare for and respond to storm surge hazards using predictions from storm surge models to determine if flooding conditions are probable.
+Like coastal ocean models, some of the complexity of a storm surge model depends on the properties of the geographic location of interest.
+In shallow regions, it is important to include the effects of wetting and drying which allows grid cells to alternate between land and water depending on the modelled sea surface height and depth \citep{hubbert1999storm, weisberg2006hurricane}.
+Other regions require modelling over a very large domain since non-local effects from the open sea may influence surge propagation \citep{weisberg2006hurricane, lane2009verification}.
+Such large domains typically use nested or unstructered grids to resolve the details of complicated coastlines.
+Appropriate treatment of rivers \citep{flather1994storm}, accurate bathymetry and inlcuding surface waves \citep{xu2010storm} have also been identified as important issues.
+Most of these models neglect the effects of stratification, either by treating the ocean as a constant temperature and salinity or by using a two-dimensional depth-averaged set of equations, an assumption that is assumed to be valid.
 
-%overview of paper: model configuration, model evaluation, storm surge hind casts.
+Storm surge modelling in the Strait of Georgia dates back to \citet{crean1988mathematical} in a barotropic finite difference model.
+Their work is highlighted in reviews of storm surges in Canadian waters by \citet{murty1995storm} and \citet{danard2003storm}.
+Compared to eastern Canada \citep{gray1984preliminary, bernier2006predicting,bobanovic2006forecasting}, relatively little has been published on storm surge modelling on the west coast of Canada.
+
+This article presents the development of a three-dimensional numerical ocean model for the Strait of Georgia and its accuracy in simulating storm surge hind casts.
+This baraoclinic model includes the effect of stratification and rivers, however, wetting and drying is not represented.
+Model evaluation has focused on regions where the near-coast slopes are steep.
 This paper is organized as follows: an overview of the model configuration and domain including a description of the ocean model we have employed is provided in section \ref{sec:config}.
@@
 %I think Hal Ritchie's talk with the storm surges had some stats presented. Look this up...
+\citep{bernier2006predicting}
 

### 2014-09-22 RoughWork/manuscript.tex
Added new figures to manuscript

@@
 During the winter months, coastal communities in the Strait of Georgia are at risk to flooding caused by storm surges, a natural hazard that can occur when a strong storm coincides with high tide.
-This investigation produces storm surge hind casts using a three-dimensional numerical ocean model for the Strait of Georgia and the surrounding bodies of water (Strait of Juan de Fuca, Puget Sound, and Johnstone Strait) collectively known as the Salish Sea.
+This investigation produces storm surge hindcasts using a three-dimensional numerical ocean model for the Strait of Georgia and the surrounding bodies of water (Strait of Juan de Fuca, Puget Sound, and Johnstone Strait) collectively known as the Salish Sea.
 The numerical model employs the Nucleus for European Modelling of the Ocean (NEMO) architecture in a regional configuration.
@@
 
-This article presents the development of a three-dimensional numerical ocean model for the Strait of Georgia and its accuracy in simulating storm surge hind casts.
+This article presents the development of a three-dimensional numerical ocean model for the Strait of Georgia and its accuracy in simulating storm surge hindcasts.
 This baraoclinic model includes the effect of stratification and rivers, however, wetting and drying is not represented.
@@
 Next, section \ref{sec:model} presents an evaluation of model performance through comparisons of tidal amplitudes and phases with observations.
-The model's accuracy in producing storm surge hind casts is also assessed in section \ref{sec:storm}.
+The model's accuracy in producing storm surge hindcasts is also assessed in section \ref{sec:storm}.
 Finally, a summary and discussion of future research goals is provided in section \ref{sec:diss}.
@@
 %3. High resolution wind (hopefully Dec 2012)
-A series of storm surge hind casts is presented next.
+A series of storm surge hindcasts is presented next.
 Storm surge simulations were started from rest at least three days prior to the event of interest to allow for adjustment of model currents.
@@
 
-Several significant surge or wind events have been considered in detail, as well as an array of hind casts to asses model ability in a statistical measure.
+Several significant surge or wind events have been considered in detail, as well as an array of hindcasts to asses model ability in a statistical measure.
 These cases have been used to asses the model's skill in reproducing water level height and surge amplitude and timing during storm events.
@@
 \centering
-\includegraphics[scale=0.6]{Figures/feb2006_bg.pdf}
+\includegraphics[scale=0.6]{Figures/feb2006.pdf}
 \caption{Comparison of observations and model output for the Feb 4, 2006 storm surge.
@@
 \centering
-\includegraphics[scale=0.6]{Figures/feb2006_factors_bg.pdf}
+\includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
 \caption{Comparison of the modelled residual for three simulations: A simulation with all forcing (blue), a simulation without atmospheric forcing (green), and a simulation without the sea surface height anomaly (red).
@@
 This event has been used to examine the propagation of the surge throughout the domain.
-First, an assessment of the accuracy of the hind cast is provided.
+First, an assessment of the accuracy of the hindcast is provided.
 

### 2014-10-02 RoughWork/manuscript.tex
Added images from new simulations. Began correcing text in Dec 2006 section

@@
 
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/tides.pdf}
+\caption{Evaluation of model tidal amplitude and phases for the M$_2$ and K$_1$ constituents. }
+\label{fig:tides}
+\end{figure}
+
 \subsection{Stratification} 
@@
 
-Point Atkinson observed a maximum surge amplitude of 84 cm, which is approximately 20 cm higher than the model result, however the timing of the surge agrees almost exactly.
-The best agreement occurs at Victoria where the observed surge was 62 cm and the model returned 54 cm.
+Point Atkinson observed a maximum surge amplitude of 84 cm, which is approximately 29 cm higher than the model result, however the timing of the surge agrees almost exactly.
+as displayed in Figure \ref{fig:dec2006} The best agreement occurs at Victoria where the observed surge was 62 cm and the model returned 44 cm.
 At all of the locations, the model predicted the maximum surge to occur within 3.5 hours of the observed surge.
@@
 
-The evolution and spatial distribution of the surge is highlighted in Figure \ref{fig:spatial} where the modelled residual is plotted every two hours beginning on Dec 14 at 18:30.
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/dec2006.pdf}
+\caption{Comparison of observed water level and surge amplitudes during the Dec 15, 2006 storm surge simulation. }
+\label{fig:dec2006}
+\end{figure}%table instead of figure?
+
+The evolution and spatial distribution of the surge is highlighted in Figure \ref{fig:spatial} where the modelled residual is plotted every hour beginning on Dec 15 at 05:30 UTC.
 The surge enters the system from the Strait of Juna de Fuca and Johnstone Strait, peaking between Dec 15 at 6:30 and Dec 15 at 10:30 and then begins to subside throughout the entire domain.
@@
 \includegraphics[scale=0.6]{Figures/dec2006_spatial.pdf}
-\caption{Propagation of the storm surge on Dec 14/15 2006.
-Modelled residual every 2 hours starting on Dec 14 at 18:30.
+\caption{Propagation of the storm surge on Dec 15 2006.
+Modelled residual every hour starting on Dec 15 at 05:30 UTC.
 Time advances from top left to bottom right.
-Note that the images are oriented with the numerical grid which is rotated $29^{\circ}$ counter-clockwise of North.
-}
+Wind vectors from the atmopsheric forcing are overlaid in black.}
 \label{fig:spatial}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/dec2006_spatial_sshonly.pdf}
+\caption{As in Figure \ref{fig:spatial} but for a simulation that does not include atmopsheric forcing. }
+\label{fig:spatial_sshonly}
 \end{figure}
@@
 \centering
-\includegraphics[scale=0.6]{Figures/nov2009_bg.pdf}
+\includegraphics[scale=0.6]{Figures/nov2009.pdf}
 \caption{Wind and storm surge conditions at Point Atkinson for a strong wind event on Nov 19, 2009.

### 2014-10-02 RoughWork/manuscript.tex
Edited captions

@@
 \includegraphics[scale=0.6]{Figures/tides.pdf}
-\caption{Evaluation of model tidal amplitude and phases for the M$_2$ and K$_1$ constituents. }
+\caption{Evaluation of model tidal amplitude and phases for the M$_2$ and K$_1$ constituents.
+On the left, a transect through the stations used for comparison.
+On the right, modelled M$_2$ (black) and K$_1$ (gray) amplitude (top) and phase (middle) compared with observations in the dashed curves of the same colour.
+The bottom plot displays the M$_2$ and K_$1$ complex differences.
+}
 \label{fig:tides}
@@
 As such, storm surge model predictions were corrected by accounting for this error.
-Additionally, since the sea surface height is calculated about a zero reverence level, the modelled water level is determined by adding a long term average of mean sea level at the specified location to the modelled sea surface height and removing a background anomaly caused by topographic pressure gradients since the atmospheric reforecasting model uses a terrain following coordinate system.
-(Nancy: I think we need a better way of adjusting come publication time).
+Additionally, since the sea surface height is calculated about a zero reverence level, the modelled water level is determined by adding a long term average of mean sea level at the specified location to the modelled sea surface height.
 
@@
 \includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
-\caption{Comparison of the modelled residual for three simulations: A simulation with all forcing (blue), a simulation without atmospheric forcing (green), and a simulation without the sea surface height anomaly (red).
+\caption{Comparison of the modelled residual for three simulations: A simulation with all forcing (solid black), a simulation without atmospheric forcing (solid grey), a simulation without the sea surface height anomaly (dashed black) and a simulation with only wind forcing, ie no sea surface height anomaly or pressure (dashed grey).
 }
@@
 \end{figure}
-%not sure about the colours, but different line styles is really hard to make out.
-
-%Finally, an assessment of the importance of including tide-surge interactions is presented next.
-A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}.
-Recall, the modelled residual is defined as the difference between the sea surface height in a simulation with all forcing and a simulation with tides and rivers only.
-A comparison between the modelled residual and a surge only run can determine the significance of the tide-surge interaction.
-This figure shows there is almost no difference between the simulations at all locations, rendering the tide-sure interaction negligible.
-In operational settings, running a surge only simulation and adding the predicted tides a posteriori could be more computationally efficient with only a small decrease in the accuracy of the prediction.
-% I'm not sure if removing a background state is really appropriate here.
-There could be some differences in the background states because of the tides(ie where fresh water moves to, etc).
-I also think it is more appropriate to use the atmospheric pressure to remove the background state...
 
@@
 \includegraphics[scale=0.6]{Figures/feb2006_tidesurge.pdf}
-\caption{Modelled residuals (solid) compared a simulation of the surge propagation without tidal forcing (dashed). A background state has not been removed from these plots. }
+\caption{Modelled residuals (solid) compared with a simulation of the surge propagation without tidal forcing (dashed). }
 \label{fig:tidesurge}

### 2014-10-03 RoughWork/manuscript.tex
Moved tide table skeleton to an appendix
Modified description of tide-surge interaction

@@
 
-\begin{table}[h]
-\centering 
-\begin{tabular}{l l c c c c c c} 
-\hline 
-& & \multicolumn{3}{c}{M$_2$} & \multicolumn{3}{c}{K$_1$} \\ 
-\hline 
-Station number & Station name & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
-\hline 
-               &              &     &                      &          &     &                      &          \\
-\hline   
-Mean Error     &              &     &                      &          &     &                      &          \\
-RMS Error      &              &     &                      &          &     &                      &          \\
-\hline  
-\end{tabular}
-\caption{Comparisons of modelled and observed amplitudes and phases.
-$R$ is the ratio of modelled to observed amplitude.
-$\Delta \phi$ is the difference between modelled and observed phase.
-$D$ is the complex difference.
-}
-\label{tab:comparison} 
-\end{table}
-%May have to include station numbers and locations on a map? Maybe on the contour map, overlay with stations and green/red dots (hopefully mostly green)
-%Should we use the new method of calculating tides?
-
 Complex differences were less than ??cm at all stations in our domain, which was assumed to be acceptable for our purposes. 
@@
 A comparison between the modelled residual and a surge-only run can determine the significance of the tide-surge interaction.
-The most significant differences occur at Point Atkinson where the maximum modelled residual is 7 cm lower than the maximum of the surge-only case.
+The most significant differences occur at Point Atkinson where the maximum of the surge-only case is 8.6 cm higher than maximum modelled residual.
 In contrast, the tide-surge interaction at Victoria is hardly noticeable.
-Indeed, a spatial map of the modelled residual and surge-only difference (not shown here) indicates that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands.
+Indeed, a spatial map of the difference between the modelled residual and surge-only difference (not shown here) indicates that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands.
 The passages between these islands are known for vigorous tidal mixing, which may have an effect on the surge propagation.
-However, the effect is a relatively small underestimate of the surge elevation and no changes to the timing of the surge.
+However, the effect is a relatively small overestimate of the surge elevation and no changes to the timing of the surge.
 So, in operational settings, performing a surge-only simulation and adding the predicted tides a posteriori could be more computationally efficient with only a small decrease in the accuracy of the prediction.
@@
 
+\section{Appendix}
+\begin{table}[h]
+\centering 
+\begin{tabular}{l l c c c c c c} 
+\hline 
+& & \multicolumn{3}{c}{M$_2$} & \multicolumn{3}{c}{K$_1$} \\ 
+\hline 
+Station number & Station name & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
+\hline 
+               &              &     &                      &          &     &                      &          \\
+\hline   
+Mean Error     &              &     &                      &          &     &                      &          \\
+RMS Error      &              &     &                      &          &     &                      &          \\
+\hline  
+\end{tabular}
+\caption{Comparisons of modelled and observed amplitudes and phases.
+$R$ is the ratio of modelled to observed amplitude.
+$\Delta \phi$ is the difference between modelled and observed phase.
+$D$ is the complex difference.
+}
+\label{tab:comparison} 
+\end{table}
+%May have to include station numbers and locations on a map? Maybe on the contour map, overlay with stations and green/red dots (hopefully mostly green)
+%Should we use the new method of calculating tides?
+
 \section{Acknowledgements}\label{sec:ack}

### 2014-10-08 RoughWork/manuscript.tex
Added description of statistics and table for Feb 2006

@@
 The numerical model employs the Nucleus for European Modelling of the Ocean (NEMO) architecture in a regional configuration.
-The model is evaluated through comparisons of tidal elevation harmonics, tidal current ellipses, and storm surge water levels with observations.
+The model is evaluated through comparisons of tidal harmonics, tidal current ellipses, and storm surge water levels with observations.
 Important forcing factors contributing to storm surges have been assessed, finding that the surge entering the domain from the Pacific Ocean contributes most significantly.
@@
 Such large domains typically use nested or unstructered grids to resolve the details of complicated coastlines.
-Appropriate treatment of rivers \citep{flather1994storm}, accurate bathymetry and inlcuding surface waves \citep{xu2010storm} have also been identified as important issues.
+Appropriate treatment of rivers \citep{flather1994storm}, accurate bathymetry and the inclusion surface waves \citep{xu2010storm} have also been identified as important issues.
 Most of these models neglect the effects of stratification, either by treating the ocean as a constant temperature and salinity or by using a two-dimensional depth-averaged set of equations, an assumption that is assumed to be valid.
@@
 
+%outline
 This article presents the development of a three-dimensional numerical ocean model for the Strait of Georgia and its accuracy in simulating storm surge hindcasts.
@@
 On the right, modelled M$_2$ (black) and K$_1$ (gray) amplitude (top) and phase (middle) compared with observations in the dashed curves of the same colour.
-The bottom plot displays the M$_2$ and K_$1$ complex differences.
+The bottom plot displays the M$_2$ and K$_1$ complex differences.
 }
@@
 \end{figure}
-
-\subsection{Stratification} 
-%outline: thalweg, plots of winter/summer stratty
-%Maybe not this
 
@@
 As such, storm surge model predictions were corrected by accounting for this error.
-Additionally, since the sea surface height is calculated about a zero reverence level, the modelled water level is determined by adding a long term average of mean sea level at the specified location to the modelled sea surface height.
+Additionally, since the sea surface height is calculated about a zero reference level, the modelled water level is determined by adding a long term average of mean sea level at the specified location to the modelled sea surface height.
 
@@
 The modelled residual is calculated as the difference in model output between a simulation with all forcing conditions and a simulation forced only with tides and rivers.
-The observed water level is taken from observations by Fisheries and Oceans Canada (citation) and the observed residual is the observations minus tidal predictions.
-Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this cae, from the year prior surge event on interest.
+Water level observations are taken from Fisheries and Oceans Canada (citation) and are used to calculate the osberved residual, defined as the observed water level minus tidal predictions.
+Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this case, from the year prior surge event of interest.
 Additionally, several simulations with varying forcing conditions such as removing atmospheric forcing or removing the sea surface height anomaly at the open boundaries have been completed.
@@
 We have also examined the spatial variability of the surge as it propagates through the domain and have considered a case with high-resolution atmospheric forcing.
+
+The model's skill is reported through a few statistical measures of the the total water level $\eta$ and the residual $\delta$. First, the mean absolute error ($MAE$), defined as
+\begin{equation}
+MAE = \text{mean}\left(\left| M - O \right|\right),
+\end{equation}
+where $M$ and $O$ are the modelled and observed quantities of interest, is a common measure that estimates the average error in the model.
+Next, a measure comparing the variance of the error to the variance of the observations defined as \citep{thompson2003prediction},
+\begin{equation}
+\gamma^2 = \frac{\text{var}\left(M-O\right)}{\text{var}\left(O\right)},
+\end{equation}
+is also calculated.
+Smaller values of $\gamma^2$ indicate better model skill.
+This quantity was used by \citet{bernier2006predicting} and \citet{bernier2010tide} in storm surge predictions in Eastern Canada and is caclulated here as a point of comparison.
+A third measure, called the Willmott skill score \citep{willmott1982some}, is also used to assess overall model skill in predicting the water level and residual.
+This quantity is defined as
+\begin{equation}
+WS = 1 - \frac{\sum_{i=1}^N \left(M_i - O_i\right)^2}{\sum_{i=1}^N \left(|M_i'| + |O_i'|\right)^2},
+\end{equation} 
+where $M_i' = M_i-\bar{O}$ and $O_i'=O_i-\bar{O}$ and $\bar{O}$ is the mean of the observations. This score is bounded by zero and one. Values close to one indicate good model skill.
 
@@
 On Feb 4, 2006, a strong storm system with winds over 75 km/hr coinciding with an unusually high tide led to significant flooding in coastal communities of the Strait of Georgia \citep{romanowski2010storm}.
-The maximum water level observed at Point Atkinson on this day was 5.49 m, which is 0.8 m higher than the predicted tides, highlighting the significance of this event as it resulted in a large anomaly and also reached water levels close to the highest level ever recorded (5.61 m).
-A set of simulations from Feb 1 to 8, 2006 under varying forcing conditions has been completed in order to determine the significance of each factor contributing to the surge.
+The maximum water level observed at Point Atkinson on this day was 5.46 m, which is 0.82 m higher than the predicted tides, highlighting the significance of this event as it resulted in a large anomaly and also reached water levels close to the highest level ever recorded (5.61 m).
+A set of simulations from Feb 1 to 8, 2006 under a variety of forcing conditions has been completed in order to determine the contribution of each forcing factor to the surge.
 
 The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations and corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River.
-The timing of maximum water level matches well between the observations and the model at all four locations, the most significant contribution to high water level being from the tides.
-There is also a good match in amplitude of the water level maximum at Point Atkinson and Campbell River, with a relative error less than 4\%.
-There is only a 7.2 cm difference between observed and modelled water level elevation at Point Atkinson and a 20.9 cm difference at Campbell River.
-The error at Victoria and Patricia Bay is slightly more significant, reaching 9.6\% and 6.3\% respectively (close to 30 cm in each case), partly due to error in tidal amplitudes as discussed above.
-The precise location and magnitude of the M$_2$ amphidrome around Victoria is very sensitive to model parameters such as bottom friction.
+The timing of maximum water level matches well between the observations and the model at all four locations, the most significant contribution to high water level being from the tides which are generally well-represented by the model.
+There is a good match in the water level maximum at Victoria and Patricia Bay, with a relative error less than 2\%.
+There is only a 4.2 cm difference between observed and modelled water level elevation at Victoria and a 6.9 cm difference at Patricia Bay, the model slightly underpredicting the maximum water level in each case.
+The error at Point Atkinson and Campbell River is slightly more significant, reaching 2.5\% and 2.7\% respectively, close to 14 cm overpredicted in each case.
+It is suprising that Victoria and Patricia Bay agree most closely with observations because the M$_2$ tidal amplitudes in the model are known to be too low in this area.
+In conjunction with the overpredicted water levels at Point Atkinson and Campbell River, it is likely that one of the forcing factors is overestimated in this example.
 
-The error in modelled residuals is also apparent in the panels on the right, most noticeable is an underestimate in maximum residual (or surge) at all locations.
-There is also a 2-3 hour delay in the timing of the maximum residual except at Victoria where the maximum residual in sync with observations.
-The errors in maximum residual range from 8.0 cm at Victoria to 28 cm at Campbell River.
-It is also apparent that Point Atkinson and Campbell River experienced the largest observed surges at 82 cm and 95 cm respectively.
+A comparison between observed and modelled residuals is also displayed in the panels on the right.
+In general, there is a very good agreement between the observed and modelled residules.
+The most noticeable difference is a 4-6 hour delay in the model's timing of the maximum residual except at Victoria where the maximum residual in sync with observations.
+The errors in maximum residual range from 6.0 cm to high at Victoria to 15.9 cm too low at Campbell River.
+Point Atkinson and Patricia Bay both experience a very good match with observed residuals with errors less than 2 cm.
+Note that Point Atkinson and Campbell River observed the largest observed surges at 82 cm and 95 cm respectively.
 The surges at Victoria and Patricia Bay were observed to be 66 cm and 74 cm.
-%Question: is it better to present relative errors or just difference?
 
@@
 \end{figure}
+
+The maximum water level and maximum residuals are important factors in storm surge hindcasts but a better indicator of overall model performance is provided by a group of statistical measures in Table \ref{tab:feb2006stat}.
+In general, $\gamma^2$ and $WS$ suggest good model agreement for the predicted water levels, most likely because the tidal signal dominates.
+The skill scores degrade slightly for the residuals but are still reasonable and indicate very good agreement.
+Note that the mean absolute error ($MAE$) is lower for the residuals, but this measure does not account for the larger variablity in water levels as opposed to residuals.
+
+\begin{table}[h]
+\centering 
+\begin{tabular}{l c c c c c c} 
+\hline 
+& \multicolumn{3}{c}{Water Level ($\eta$)}        & \multicolumn{3}{c}{Residual ($\delta$)} \\ 
+\hline 
+Location       & $MAE$ (cm) & $\gamma^2$ & $WS$   & $MAE$ (cm) & $\gamma^2$ & $WS$          \\
+\hline 
+Point Atkinson & 10.5       &   0.012    & 0.994  & 6.2        &  0.0646    & 0.979         \\
+Victoria       &  8.6       &   0.032    & 0.992  & 6.1        &  0.0851    & 0.970         \\
+Patricia Bay   &  9.1       &   0.024    & 0.992  & 6.9        &  0.0704    & 0.969         \\
+Campbell River & 13.1       &   0.041    & 0.990  & 5.8        &  0.0790    & 0.961         \\
+\hline 
+\end{tabular}
+\caption{Evaluation of model performance for the Feb 2006 simulation.
+Mean absoulte error (MAE), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and resdiuals ($\delta$) during the period displayed in Figure \ref{fig:feb2006}.}
+\label{tab:feb2006stat} 
+\end{table} 
+%anything else to include? statistics calculated in Feb 2006 -paper prodcution.ipynb and saved in statisitcs_feb2006.csv
 
@@
 
-\subsection{High Resolution Atmospheric Forcing: Dec 17, 2012}
-
-\subsection{Statistical evaluation}
-%I think Hal Ritchie's talk with the storm surges had some stats presented. Look this up...
-\citep{bernier2006predicting}
+%\subsection{High Resolution Atmospheric Forcing: Dec 17, 2012}
 

### 2014-10-08 RoughWork/manuscript.tex
Revised contributing factors analysis to reflect new simulations

@@
 
-The model's skill is reported through a few statistical measures of the the total water level $\eta$ and the residual $\delta$. First, the mean absolute error ($MAE$), defined as
+The model's skill is reported through a few statistical measures of the the total water level $\eta$ and the residual $\delta$. First, the mean absolute error ($\bar{e}$), defined as
 \begin{equation}
-MAE = \text{mean}\left(\left| M - O \right|\right),
+\bar{e} = \text{mean}\left(\left| M - O \right|\right),
 \end{equation}
@@
 The skill scores degrade slightly for the residuals but are still reasonable and indicate very good agreement.
-Note that the mean absolute error ($MAE$) is lower for the residuals, but this measure does not account for the larger variablity in water levels as opposed to residuals.
+Note that the mean absolute error ($\bar{e}$) is lower for the residuals, but this measure does not account for the larger variablity in water levels as opposed to residuals.
 
@@
 \centering 
-\begin{tabular}{l c c c c c c} 
+\begin{tabular}{|l |c c c | c c c|} 
 \hline 
-& \multicolumn{3}{c}{Water Level ($\eta$)}        & \multicolumn{3}{c}{Residual ($\delta$)} \\ 
+& \multicolumn{3}{|c|}{Water Level ($\eta$)}        & \multicolumn{3}{|c|}{Residual ($\delta$)} \\ 
 \hline 
-Location       & $MAE$ (cm) & $\gamma^2$ & $WS$   & $MAE$ (cm) & $\gamma^2$ & $WS$          \\
+Location       & $\bar{e}$ (cm) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\gamma^2$ & $WS$      \\
 \hline 
@@
 \caption{Evaluation of model performance for the Feb 2006 simulation.
-Mean absoulte error (MAE), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and resdiuals ($\delta$) during the period displayed in Figure \ref{fig:feb2006}.}
+Mean absoulte error ($\bar{e}$), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and resdiuals ($\delta$) during the period displayed in Figure \ref{fig:feb2006}.}
 \label{tab:feb2006stat} 
@@
 Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented in Figure \ref{fig:factors}.
-The modelled residual from three simulations have been plotted.
-First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly, is shown in blue.
-Additionally, a simulation with all forcing except the atmospheric forcing is presented in green and a simulation with all forcing except the sea surface height anomaly is shown in red.
+The modelled residual from four simulations have been plotted.
+First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly, is shown in the dashed black curve.
+Additionally, a simulation with all forcing except the atmospheric forcing is presented in solid grey and a simulation with all forcing except the sea surface height anomaly is shown solid black.
+Finally, a simulation with only wind forcing, that is no anomaly and no inverse barometer effect due to pressure, is shown in the dashed grey curve.
 It is clear that the sea surface height anomaly contributes most significantly to the surge at each of these locations.
-When the sea surface height anomaly is not included as a forcing condition, the surge drops from 70 cm to 15.6 cm at Point Atkinson and its maximum occurs eight hours later.
+When the sea surface height anomaly is not included as a forcing condition, the surge drops from a maximum 82 cm to 11 cm at Point Atkinson and its maximum occurs eight hours later.
 A similar drop in the maximum surge and delay in timing of the maximum surge is observed at the other locations.
-The atmospheric model indicates southeasterly winds during the onset of the surge but after the peak surge has passed, the winds shift to westerly.
-A strong westerly wind would cause an elevated sea surface height at Point Atkinson as water piles against the coast, explaining the delay in peak surge when only atmospheric forcing is used since the winds to shift westerly after the peak surge has passed.
-A simulation with only wind forcing (no inverse barometer effect or anomaly at open boundary) produces a surge of 9.0 cm at Point Atkinson, a small fraction of the total observed surge.
-However, this wind-induced surge occurs well after the peak surge from observations and simulations with full forcing.
+The atmospheric model indicates southeasterly winds near Point Atkinson during the onset of the surge but after the peak surge has passed, the winds shift to westerly.
+A strong westerly wind would cause an elevated sea surface height at Point Atkinson as water piles against the coast, explaining the delay in peak surge when only atmospheric forcing is used since the winds shift to westerly after the peak surge has passed.
+A simulation with only wind forcing produces a surge of 7.8 cm at Point Atkinson, a small fraction of the total observed surge, and only centimeters at the other locations.
+Note that this wind-induced surge occurs well after the peak surge from observations and simulations with full forcing.
 
-The contribution to the surge due to atmospheric forcing appears to be geographically dependent, its removal leading to a drop of about 5 cm in the maximum surge at Point Atkinson and 10 cm at Campbell River.
-When atmospheric forces are neglected the timing of the maximum surge is earlier at Point Atkinson but later at Campbell River.
-Campbell River exhibits a strong response to atmospheric conditions during midday on Feb 4, as demonstrated in a two sharp spikes in both the blue and red curves around 12:00 hr The green curve, which does not include any atmospheric coring, does not exhibit this behaviour.
-The result is a peak surge earlier than in day than the no atmospheric forcing case.
-This response is also noted in the observations at Campbell River in Figure \ref{fig:feb2006} where the maximum surge occurs earlier than at the other locations.
-At Victoria and Patricia Bay, the surge amplitude is mainly unaffected by the atmospheric forcing.
-%maybe look at winds/pressure to see why Vic/Pat Bay aren't affected?. Plan: map of wind vectors from CGRF (not for paper, just for me in aiding with explanation)
+The contribution to the surge due to atmospheric forcing appears to be geographically dependent, its removal leading to a drop of about 2.5 cm in the maximum surge at Point Atkinson and 2.3 cm at Campbell River.
+Interestingly, the surge amplitude increases by 4.8 cm at Victoria and 3.6 cm at Patricia Bay when atmospheric forces are negelcted.
+The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
+Additionally, the surge amplitude decreases by a few centimtres if the inverse barometer effect due to atmospheric pressure is not included.
 
@@
 Previous simulations by \citet{murty1995storm} suggest that the inverse barometer effect and the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.
-A detailed assessment of the effect of the local atmospheric conditions and winds is included later where simulations have been performed with high resolution atmospheric forcing data.
+A detailed assessment of the effect of the local atmospheric conditions and winds should be performed with high resolution atmospheric forcing data.
 
@@
 \includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
-\caption{Comparison of the modelled residual for three simulations: A simulation with all forcing (solid black), a simulation without atmospheric forcing (solid grey), a simulation without the sea surface height anomaly (dashed black) and a simulation with only wind forcing, ie no sea surface height anomaly or pressure (dashed grey).
+\caption{Comparison of the modelled residual for three simulations: A simulation with all forcing (solid black), a simulation without atmospheric forcing (solid grey), a simulation without the sea surface height anomaly (dashed black) and a simulation with only wind forcing, i.e.\ no sea surface height anomaly or pressure (dashed grey).
 }

### 2014-10-14 RoughWork/manuscript.tex
Updated comments on tide-surge interaction

@@
 \includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
-\caption{Comparison of the modelled residual for three simulations: A simulation with all forcing (solid black), a simulation without atmospheric forcing (solid grey), a simulation without the sea surface height anomaly (dashed black) and a simulation with only wind forcing, i.e.\ no sea surface height anomaly or pressure (dashed grey).
+\caption{Comparison of the modelled residual for four simulations: A simulation with all forcing (solid black), a simulation without atmospheric forcing (solid grey), a simulation without the sea surface height anomaly (dashed black) and a simulation with only wind forcing, i.e.\ no sea surface height anomaly or pressure (dashed grey).
 }
@@
 A comparison between the modelled residual and a surge-only run can determine the significance of the tide-surge interaction.
-The most significant differences occur at Point Atkinson where the maximum of the surge-only case is 8.6 cm higher than maximum modelled residual.
+The most significant differences occur at Point Atkinson where the maximum of the surge-only case is 8.7 cm higher than maximum modelled residual, about 10\% higher of the surge amplitude.
 In contrast, the tide-surge interaction at Victoria is hardly noticeable.
 Indeed, a spatial map of the difference between the modelled residual and surge-only difference (not shown here) indicates that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands.
-The passages between these islands are known for vigorous tidal mixing, which may have an effect on the surge propagation.
+The passages between these islands are known for vigorous tidal mixing, which may have a nonlinear effect on the surge propagation.
 However, the effect is a relatively small overestimate of the surge elevation and no changes to the timing of the surge.

### 2014-10-14 RoughWork/manuscript.tex
Included statistics for Dec 2006

@@
 The skill scores degrade slightly for the residuals but are still reasonable and indicate very good agreement.
-Note that the mean absolute error ($\bar{e}$) is lower for the residuals, but this measure does not account for the larger variablity in water levels as opposed to residuals.
+Note that the mean absolute error ($\bar{e}$) is lower for the residuals, but this measure does not account for the larger variablity in water levels due to the tides.
 
@@
 \hline 
-Point Atkinson & 10.5           &   0.012    & 0.994  & 6.2            &  0.0646    & 0.979     \\
-Victoria       &  8.6           &   0.032    & 0.992  & 6.1            &  0.0851    & 0.970     \\
-Patricia Bay   &  9.1           &   0.024    & 0.992  & 6.9            &  0.0704    & 0.969     \\
-Campbell River & 13.1           &   0.041    & 0.990  & 5.8            &  0.0790    & 0.961     \\
+Point Atkinson & 10.5           &   0.012    & 0.994  & 6.2            &  0.065     & 0.979     \\
+Victoria       &  8.6           &   0.032    & 0.992  & 6.1            &  0.085     & 0.970     \\
+Patricia Bay   &  9.1           &   0.024    & 0.992  & 6.9            &  0.070     & 0.969     \\
+Campbell River & 13.1           &   0.041    & 0.990  & 5.8            &  0.079     & 0.961     \\
 \hline 
@@
 %anything else to include? statistics calculated in Feb 2006 -paper prodcution.ipynb and saved in statisitcs_feb2006.csv 
+%mean of model and mean of observations?
 
@@
 On Dec 15, 2006, a major storm with wind gusts over 120 km/hr caused significant damage to many coastal communities on the Pacific Northwest.
-Although the storm was very strong and the observed residual at Point Atkinson reached 84 cm, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area\citep{forseth2006adaptation}.
-This event has been used to examine the propagation of the surge throughout the domain.
+Although the storm was very strong and the observed residual at Point Atkinson reached 84 cm, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area \citep{forseth2006adaptation}.
+This event has been used to examine the propagation of the surge throughout the domain and spatial variations in the surge amplitude.
 First, an assessment of the accuracy of the hindcast is provided.
 
-Point Atkinson observed a maximum surge amplitude of 84 cm, which is approximately 29 cm higher than the model result, however the timing of the surge agrees almost exactly.
-as displayed in Figure \ref{fig:dec2006} The best agreement occurs at Victoria where the observed surge was 62 cm and the model returned 44 cm.
+Point Atkinson observed a maximum surge amplitude of 84 cm, which is approximately 28 cm higher than the modelled residual, however the timing of the surge agrees almost exactly as displayed in Figure \ref{fig:dec2006}.
+The best agreement occurs at Victoria where the observed surge was 62 cm and the model returned 44 cm.
 At all of the locations, the model predicted the maximum surge to occur within 3.5 hours of the observed surge.
 Typically, the modelled surge was too early, in contrast to the later predictions in the Feb 4, 2006 case.
-Errors in the modelled water level at the time of maximum surge were of the same order as the errors in the surge amplitude.
+Error and skill scores for the modelled water level and the modelled residual are provided in Table \ref{dec2006stat}.
+The modelled residuals are on average too low and have a lower skill score than the Feb 2006 case.
+This could be due to a number of factors.
+First,the very large wind speeds observed during this storm are significantly underrepresented by the atmospheric model leading to lower wind stress and less piling of water against the coast.
+Second, the anomaly forcing at the open boundaries could be too low.
+The observed residual at Tofino, BC (a point near by but outside of the domain) is up to 10 cm lower than the observed residual at Neah Bay, WA (a point near by but inside of the domain).
+Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the case.
+%Skill score between CGRF and observed winds?
 
@@
 \includegraphics[scale=0.6]{Figures/dec2006.pdf}
-\caption{Comparison of observed water level and surge amplitudes during the Dec 15, 2006 storm surge simulation. }
+\caption{Comparison of observed and modelled water level and residuls during the Dec 15, 2006 storm surge simulation. }
 \label{fig:dec2006}
 \end{figure}%table instead of figure?
+
+\begin{table}[h]
+\centering 
+\begin{tabular}{|l |c c c | c c c|} 
+\hline 
+& \multicolumn{3}{|c|}{Water Level ($\eta$)}        & \multicolumn{3}{|c|}{Residual ($\delta$)} \\ 
+\hline 
+Location       & $\bar{e}$ (cm) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\gamma^2$ & $WS$      \\
+\hline 
+Point Atkinson & 10.5           &   0.025    & 0.994  & 11.1           &  0.275     & 0.891     \\
+Victoria       &  9.6           &   0.042    & 0.984  &  7.9           &  0.223     & 0.911     \\
+Patricia Bay   &  9.2           &   0.031    & 0.990  &  9.5           &  0.260     & 0.889     \\
+Campbell River & 10.6           &   0.030    & 0.992  & 10.6           &  0.247     & 0.905     \\
+\hline 
+\end{tabular}
+\caption{Evaluation of model performance for the Dec 2006 simulation.}
+\label{tab:dec2006stat} 
+\end{table}
 

### 2014-10-14 RoughWork/manuscript.tex
Modified description of spatial extent of surge

@@
 
-The evolution and spatial distribution of the surge is highlighted in Figure \ref{fig:spatial} where the modelled residual is plotted every hour beginning on Dec 15 at 05:30 UTC.
-The surge enters the system from the Strait of Juna de Fuca and Johnstone Strait, peaking between Dec 15 at 6:30 and Dec 15 at 10:30 and then begins to subside throughout the entire domain.
-Notably, the Strait of Georgia experiences the largest surge values, upwards of 70 cm, whereas the Juan de Fuca region experiences a maximum surge around 55 cm.
+The evolution and spatial distribution of the surge is highlighted in Figure \ref{fig:spatial} where the modelled residual is plotted every hour beginning on Dec 15 at 05:30 UTC through to Dec 15 at 14:30 UTC.
+The surge enters the system from the Strait of Juna de Fuca, peaking between Dec 15 at 9:30 and Dec 15 at 11:30 and then begins to subside throughout the entire domain.
+Notably, the Strait of Georgia experiences the largest surge values, upwards of 60 cm, whereas the Juan de Fuca region experiences a maximum surge around 45 cm.
 Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast.
+At about 11:30 UTC, the winds switch to northwesterly and then westerly, further amplifying the surge along the mainland coast.
 
@@
 \includegraphics[scale=0.6]{Figures/dec2006_spatial.pdf}
-\caption{Propagation of the storm surge on Dec 15 2006.
+\caption{Propagation of the storm surge on Dec 15, 2006.
 Modelled residual every hour starting on Dec 15 at 05:30 UTC.
@@
 \end{figure}
+
+An evaluation of the spatial distribution of the surge in a simulation without atmospheric forcing is provided in Figure \ref{fig:spatial_sshonly}.
+In this case, the surge amplitude across the Strait of Georgia, is uniform with no pronouced difference between the mainland side and Vancouver Island side.
+While the surge entering the domain from the Pacific Ocean is the leading order contribution, the winds may induce localized changes to the surge amplitude.
+Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland.
+Even without atmospheric forcing, there is still a siginificant difference in surge amplitude between the Strait of Georgria and the Strait of Juan de Fuca, perhaps because the Strait of Georgia is also affected by the surge entering the domain from Johnstone Strait at the northern boundary.
 

### 2014-10-14 RoughWork/manuscript.tex
Edited description of Nov 2009 storm

@@
 \centering 
-\begin{tabular}{|l |c c c | c c c|} 
+\begin{tabular}{|l |c c c c c | c c c c c|} 
 \hline 
-& \multicolumn{3}{|c|}{Water Level ($\eta$)}        & \multicolumn{3}{|c|}{Residual ($\delta$)} \\ 
+& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
 \hline 
-Location       & $\bar{e}$ (cm) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\gamma^2$ & $WS$      \\
+Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{m}}$ (o) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
 \hline 
-Point Atkinson & 10.5           &   0.025    & 0.994  & 11.1           &  0.275     & 0.891     \\
-Victoria       &  9.6           &   0.042    & 0.984  &  7.9           &  0.223     & 0.911     \\
-Patricia Bay   &  9.2           &   0.031    & 0.990  &  9.5           &  0.260     & 0.889     \\
-Campbell River & 10.6           &   0.030    & 0.992  & 10.6           &  0.247     & 0.905     \\
+Point Atkinson & 10.5           &  3.45                & 3.47                 &   0.025    & 0.994  & 11.1           &  0.212                 & 0.186                  &  0.275     & 0.891 \\
+Victoria       &  9.6           &  2.23                & 2.16                 &   0.042    & 0.984  &  7.9           &  0.190                 & 0.156                  &  0.223     & 0.911 \\
+Patricia Bay   &  9.2           &  2.63                & 2.58                 &   0.031    & 0.990  &  9.5           &  0.193                 & 0.149                  &  0.260     & 0.889 \\
+Campbell River & 10.6           &  3.23                & 3.28                 &   0.030    & 0.992  & 10.6           &  0.191                 & 0.163                  &  0.247     & 0.905 \\
 \hline 
@@
 Two significant surge events are noted, the first occurring on Nov 19 and a second, larger surge on Nov 20.
-The largest surge did not coincide with the strong wind event which peaked at Nov 19 01:00 hr, reflected in both the model and observations.
-The highest observed surge reached 62.6 cm on Nov 20 at 10:00 hr, while the model recorded a maximum surge of 38.8 cm five hours later.
-The surge did not occur during a high tide and so no high water level events were recorded.
-Although not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and a more significant surge entering the system at Tofino and Port Hardy.
+The largest surge did not coincide with the strong wind event which peaked at Nov 19 01:00 UTC, reflected in both the model and observations.
+The highest observed surge reached 62.5 cm on Nov 20 at 10:30 UTC, while the model recorded a maximum surge of 50.5 cm five hours later.
+On Nov 19, the observed surge reached upwards of 50 cm, which is low when compared to the other cases considered.
+Neither of the surges occured during a high tide and so no high water level events were recorded.
+Although not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and another surge entering the system at Tofino and Port Hardy.
+%citation for Environment Canada data?
 
@@
 These changes in water level can also contribute to storm surges.
-During the high wind event on Nov 19, the winds at Point Atkinson were directed to the west, explaining the lack of surge associated with this high wind event since wind stresses would act to depress the water on the east side of the basin.
-It is possible that the surge occurring on Nov 19 formed in response to the sea surface returning to equilibrium in a seiche-like motion after the winds have relaxed.
+During the high wind event on Nov 19, the winds at Point Atkinson were directed to the west, explaining the lack of strong surge observed during this event since wind stresses would act to depress the water on the east side of the basin.
 The model appears to reproduce this surge with reasonable accuracy, however, there are some discrepancies in the wind speed and direction between weather observations and the atmospheric forcing employed.
-A more detailed look at the sensitivity of modelled surges to atmospheric forcing is considered in a case with a high-resolution weather product.
 
@@
 
+Once again, an evaluation of the model skill in this hindcast is provided in Table \ref{tab:nov2009stats}.
+
 %\subsection{High Resolution Atmospheric Forcing: Dec 17, 2012}
+%not to be included
+
+%I have one other simulation for November 2006. Skill scores are comparable to the Feb 2006 case, so I'm not sure what this case adds to the overall article.
 

### 2014-10-14 RoughWork/manuscript.tex
Modified tables to include mean of obs/mod

@@
 \centering 
-\begin{tabular}{|l |c c c | c c c|} 
+\begin{tabular}{|l |c c c c c | c c c c c|} 
 \hline 
-& \multicolumn{3}{|c|}{Water Level ($\eta$)}        & \multicolumn{3}{|c|}{Residual ($\delta$)} \\ 
+& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
 \hline 
-Location       & $\bar{e}$ (cm) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\gamma^2$ & $WS$      \\
+Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
 \hline 
-Point Atkinson & 10.5           &   0.012    & 0.994  & 6.2            &  0.065     & 0.979     \\
-Victoria       &  8.6           &   0.032    & 0.992  & 6.1            &  0.085     & 0.970     \\
-Patricia Bay   &  9.1           &   0.024    & 0.992  & 6.9            &  0.070     & 0.969     \\
-Campbell River & 13.1           &   0.041    & 0.990  & 5.8            &  0.079     & 0.961     \\
+Point Atkinson & 10.5           & 3.46                 & 3.54                 &   0.012    & 0.994  & 6.2            &  0.273                 & 0.309                  & 0.065     & 0.979     \\
+Victoria       &  8.6           & 2.18                 & 2.19                 &   0.032    & 0.992  & 6.1            &  0.235                 & 0.278                  & 0.085     & 0.970     \\
+Patricia Bay   &  9.1           & 2.60                 & 2.64                 &   0.024    & 0.992  & 6.9            &  0.223                 & 0.278                  & 0.070     & 0.969     \\
+Campbell River & 13.1           & 3.20                 & 3.21                 &   0.041    & 0.990  & 5.8            &  0.303                 & 0.300                  & 0.079     & 0.961     \\
 \hline 
@@
 Typically, the modelled surge was too early, in contrast to the later predictions in the Feb 4, 2006 case.
-Error and skill scores for the modelled water level and the modelled residual are provided in Table \ref{dec2006stat}.
+Error and skill scores for the modelled water level and the modelled residual are provided in Table \ref{tab:dec2006stat}.
 The modelled residuals are on average too low and have a lower skill score than the Feb 2006 case.
@@
 \hline 
-Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{m}}$ (o) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
+Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
 \hline 
@@
 \end{tabular}
-\caption{Evaluation of model performance for the Dec 2006 simulation.}
+\caption{Evaluation of model performance for the Dec 2006 hindcast.}
 \label{tab:dec2006stat} 
@@
 
-Once again, an evaluation of the model skill in this hindcast is provided in Table \ref{tab:nov2009stats}.
+Once again, an evaluation of the model skill for this hindcast is provided in Table \ref{tab:nov2009stat}.
+
+\begin{table}[h]
+\centering 
+\begin{tabular}{|l |c c c c c | c c c c c|} 
+\hline 
+& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
+\hline 
+Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
+\hline 
+Point Atkinson & 14.0           &  3.45                & 3.55                 &   0.012    & 0.995  &  6.1           &  0.313                 & 0.331                  &  0.237     & 0.913 \\
+Victoria       &  9.9           &  2.25                & 2.24                 &   0.023    & 0.989  &  4.8           &  0.323                 & 0.293                  &  0.245     & 0.899 \\
+Patricia Bay   & 11.2           &  2.50                & 2.65                 &   0.020    & 0.997  &  6.0           &  0.277                 & 0.290                  &  0.272     & 0.901 \\
+Campbell River & 13.5           &  3.22                & 3.27                 &   0.019    & 0.991  &  4.9           &  0.326                 & 0.314                  &  0.217     & 0.924 \\
+\hline 
+\end{tabular}
+\caption{Evaluation of model performance for the Nov 2009 hindcast.}
+\label{tab:nov2009stat} 
+\end{table}
 

### 2014-10-14 RoughWork/manuscript.tex
Edited caption of table

@@
 \caption{Evaluation of model performance for the Feb 2006 simulation.
-Mean absoulte error ($\bar{e}$), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and resdiuals ($\delta$) during the period displayed in Figure \ref{fig:feb2006}.}
+Mean absolute error ($\bar{e}$), observed mean (subscript o), modelled mean (subscript m), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and resdiuals ($\delta$) during the period displayed in Figure \ref{fig:feb2006}.}
 \label{tab:feb2006stat} 

### 2014-10-14 RoughWork/manuscript.tex
Included description of Nov 2009 stats

@@
 Error and skill scores for the modelled water level and the modelled residual are provided in Table \ref{tab:dec2006stat}.
+All skill scores rank above 0.89.
 The modelled residuals are on average too low and have a lower skill score than the Feb 2006 case.
@@
 Notably, the Strait of Georgia experiences the largest surge values, upwards of 60 cm, whereas the Juan de Fuca region experiences a maximum surge around 45 cm.
-Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast.
+Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges (\~ 10 cm) than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast.
 At about 11:30 UTC, the winds switch to northwesterly and then westerly, further amplifying the surge along the mainland coast.
@@
 Once again, an evaluation of the model skill for this hindcast is provided in Table \ref{tab:nov2009stat}.
+The modelled water levels and residuals are too high, except at Victoria.
+Although Victoria has the smallest mean error in both the water level and residuals, its Willmott skill score and $\gamma^2$ values rank low when compared to the other stations.
+Generally, the skills scores are very good, scoring above 0.89 at all stations.
 

### 2014-10-16 RoughWork/manuscript.tex
Added comments for what to include in discussion
Started editing tides

@@
 The remaining constituents were extrapolated from Webtide.
+At the western boundary, the tidal elevation and currents forcing parameters were tuned for a good match wiht observations at locations within the Strait of Georgia, including Point Atkinson, Gibsons Landing, Winchelsea and Halfmoon Bay.
+At the nothern boundary, the tides were tuned for good matches at Yorke Island and Kelsey Bay.
+Further details on the tuning are provided in the appendix \ref{sec:appendix}.
 % Needs editing! Update when we are finished with tides.
@@
 %anything else to include? statistics calculated in Feb 2006 -paper prodcution.ipynb and saved in statisitcs_feb2006.csv 
-%mean of model and mean of observations?
+%I also have standard deviations, correlations, ...
 
@@
 In addition, large wind speeds over 15 m/s persisted for over 9 hours.
-Yet, the surge associated with this event is rather small when compared the previous two cases as shown in Figure \ref{fig:nov2009} where we have plotted the wind speed at Point Atkinson, sea surface height anomaly forcing at the open boundaries, and observed and modelled residuals at Point Atkinson.
+Yet, the surge associated with this event is rather small when compared to the previous two cases as shown in Figure \ref{fig:nov2009} where we have plotted the wind speed at Point Atkinson, sea surface height anomaly forcing at the open boundaries, and observed and modelled residuals at Point Atkinson.
 Two significant surge events are noted, the first occurring on Nov 19 and a second, larger surge on Nov 20.
 The largest surge did not coincide with the strong wind event which peaked at Nov 19 01:00 UTC, reflected in both the model and observations.
-The highest observed surge reached 62.5 cm on Nov 20 at 10:30 UTC, while the model recorded a maximum surge of 50.5 cm five hours later.
+At Point Atkinsons, the highest observed surge reached 62.5 cm on Nov 20 at 10:30 UTC, while the model recorded a maximum surge of 50.5 cm five hours later.
 On Nov 19, the observed surge reached upwards of 50 cm, which is low when compared to the other cases considered.
@@
 These changes in water level can also contribute to storm surges.
-During the high wind event on Nov 19, the winds at Point Atkinson were directed to the west, explaining the lack of strong surge observed during this event since wind stresses would act to depress the water on the east side of the basin.
-The model appears to reproduce this surge with reasonable accuracy, however, there are some discrepancies in the wind speed and direction between weather observations and the atmospheric forcing employed.
+During the high wind event on Nov 19 and Nov 20, the winds at Point Atkinson were directed to the west, explaining the lack of strong surge observed during this event since wind stresses would act to depress the water on the east side of the basin.
+In fact, the surge amplitude at Victoria on Nov 20 as 64.7 cm which is slightly higher than that observed at Point Atkinson (62.5 cm).
+Of the cases considered, this is the only example where the surge amplitude at Victoria was higher than that at Point Atkinson.
+In fact, this event observed a higher surge at Victoria than the Dec 2006 event, further highlighting the role that winds plays in setting up local variations in the surge.
+In this case, the model results do not reflect a higher surge at Victoria (46.1 cm versus 50.5 cm at Point Atkinson), perhaps due to an underestimate in the wind speed and some discrepancies in the wind direction of the atmospheric forcing.
 
@@
 %comments on wetting and drying (maybe most important in shallow regions)
+%summary of skill scores
 
-\section{Appendix}
+\section{Appendix}\label{sec:appendix}
 \begin{table}[h]
@@
 \hline 
-Station number & Station name & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
+            & Station name     & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
 \hline 
-               &              &     &                      &          &     &                      &          \\
+1           & Port Renfrew     &     &                      &          &     &                      &          \\
+2           & Sheringham Point &     &                      &          &     &                      &          \\
+3	    & Pedder Bay       &     &                      &          &     &                      &          \\
+4           & Esquimalt        &     &                      &          &     &                      &          \\
+5           & Victoria         &     &                      &          &     &                      &          \\
+6           & Clover Point     &     &                      &          &     &                      &          \\
+7           & Finnerty Cove    &     &                      &          &     &                      &          \\
+8           & Fulford Harbour  &     &                      &          &     &                      &          \\
+9           & Tumbo Channel    &     &                      &          &     &                      &          \\
+10          & Patos Island     &     &                      &          &     &                      &          \\
+11          & Whaler Bay       &     &                      &          &     &                      &          \\
+12          & Tsawwassen       &     &                      &          &     &                      &          \\
+13          & Sandheads        &     &                      &          &     &                      &          \\
+14          & Point Grey       &     &                      &          &     &                      &          \\
+15          & Point Atkinson   &     &                      &          &     &                      &          \\
+16          & Gibsons Landing  &     &                      &          &     &                      &          \\
+17          & Winchelsea       &     &                      &          &     &                      &          \\
+18          & Halfmoon Bay     &     &                      &          &     &                      &          \\
+19          & Irvines Landing  &     &                      &          &     &                      &          \\
+20          & Powell River     &     &                      &          &     &                      &          \\
+21          & Little River     &     &                      &          &     &                      &          \\
+22          & Lund             &     &                      &          &     &                      &          \\
+23          & Twin Islets      &     &                      &          &     &                      &          \\
+24          & Campbell River   &     &                      &          &     &                      &          \\
+25          & Maude Island E   &     &                      &          &     &                      &          \\
+26          & Nymphe Cove      &     &                      &          &     &                      &          \\
+27          & Seymour Narrows  &     &                      &          &     &                      &          \\
+28          & Brown Bay        &     &                      &          &     &                      &          \\
+29          & Chatham Point    &     &                      &          &     &                      &          \\
+30          & Kelsey Bay       &     &                      &          &     &                      &          \\
+31          & Yorke Island     &     &                      &          &     &                      &          \\
 \hline   
-Mean Error     &              &     &                      &          &     &                      &          \\
-RMS Error      &              &     &                      &          &     &                      &          \\
+            &  Mean Error      &     &                      &          &     &                      &          \\
+            &  RMS Error       &     &                      &          &     &                      &          \\
 \hline  

### 2014-10-16 RoughWork/manuscript.tex
added tide table to manuscript

@@
 \centering 
-\begin{tabular}{l l c c c c c c} 
+\begin{tabular}{l l l l c c c c c c} 
 \hline 
-& & \multicolumn{3}{c}{M$_2$} & \multicolumn{3}{c}{K$_1$} \\ 
+& & & & \multicolumn{3}{c}{M$_2$} & \multicolumn{3}{c}{K$_1$} \\ 
 \hline 
-            & Station name     & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
+            & Station name     & Longitude & Latitude & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) & $R$ & $ \Delta \phi$ (deg) & $D$ (cm) \\
 \hline 
-1           & Port Renfrew     &     &                      &          &     &                      &          \\
-2           & Sheringham Point &     &                      &          &     &                      &          \\
-3	    & Pedder Bay       &     &                      &          &     &                      &          \\
-4           & Esquimalt        &     &                      &          &     &                      &          \\
-5           & Victoria         &     &                      &          &     &                      &          \\
-6           & Clover Point     &     &                      &          &     &                      &          \\
-7           & Finnerty Cove    &     &                      &          &     &                      &          \\
-8           & Fulford Harbour  &     &                      &          &     &                      &          \\
-9           & Tumbo Channel    &     &                      &          &     &                      &          \\
-10          & Patos Island     &     &                      &          &     &                      &          \\
-11          & Whaler Bay       &     &                      &          &     &                      &          \\
-12          & Tsawwassen       &     &                      &          &     &                      &          \\
-13          & Sandheads        &     &                      &          &     &                      &          \\
-14          & Point Grey       &     &                      &          &     &                      &          \\
-15          & Point Atkinson   &     &                      &          &     &                      &          \\
-16          & Gibsons Landing  &     &                      &          &     &                      &          \\
-17          & Winchelsea       &     &                      &          &     &                      &          \\
-18          & Halfmoon Bay     &     &                      &          &     &                      &          \\
-19          & Irvines Landing  &     &                      &          &     &                      &          \\
-20          & Powell River     &     &                      &          &     &                      &          \\
-21          & Little River     &     &                      &          &     &                      &          \\
-22          & Lund             &     &                      &          &     &                      &          \\
-23          & Twin Islets      &     &                      &          &     &                      &          \\
-24          & Campbell River   &     &                      &          &     &                      &          \\
-25          & Maude Island E   &     &                      &          &     &                      &          \\
-26          & Nymphe Cove      &     &                      &          &     &                      &          \\
-27          & Seymour Narrows  &     &                      &          &     &                      &          \\
-28          & Brown Bay        &     &                      &          &     &                      &          \\
-29          & Chatham Point    &     &                      &          &     &                      &          \\
-30          & Kelsey Bay       &     &                      &          &     &                      &          \\
-31          & Yorke Island     &     &                      &          &     &                      &          \\
+1 & Port Renfrew & -124.43 & 48.55 & 1.16 & -9.4 & 16.63  & 1.03 & -4.94 & 4.17 \\
+2 & Sheringham Point & -123.92 & 48.38 & 1.19 & -16.47 & 17.95  & 1 & -4.44 & 4.24 \\
+3 & Pedder Bay & -123.55 & 48.33 & 0.96 & -27.39 & 15.91  & 0.99 & -2.41 & 2.66 \\
+4 & Esquimalt & -123.43 & 48.43 & 0.91 & -23.09 & 14.40  & 0.99 & -1.04 & 1.21 \\
+5 & Victoria & -123.37 & 48.42 & 0.89 & -17.77 & 11.62  & 1.04 & -1.83 & 3.22 \\
+6 & Clover Point & -123.35 & 48.41 & 0.83 & -17.95 & 13.39  & 1.03 & -2.23 & 3.14 \\
+7 & Finnerty Cove & -123.3 & 48.47 & 0.83 & -8.24 & 9.47  & 1.02 & -2.19 & 3.25 \\
+8 & Fulford Harbour & -123.45 & 48.77 & 0.82 & -3.08 & 10.78  & 1 & -0.63 & 0.83 \\
+9 & Tumbo Channel & -123.11 & 48.79 & 1.05 & 2.03 & 4.43  & 1.04 & 0.66 & 3.34 \\
+10 & Patos Island & -122.97 & 48.78 & 1.03 & 0.26 & 2.32  & 1.04 & -1.5 & 3.92 \\
+11 & Whaler Bay & -123.33 & 48.89 & 1 & -0.76 & 1.10  & 1.01 & -0.48 & 1.19 \\
+12 & Tsawwassen & -123.13 & 48.99 & 1 & 0.92 & 1.31  & 1.01 & 0.19 & 1.09 \\
+13 & Sandheads & -123.3 & 49.1 & 1.02 & -1.05 & 2.58  & 1.04 & -0.59 & 3.15 \\
+14 & Point Grey & -123.27 & 49.25 & 0.98 & -3.31 & 5.83  & 0.96 & -0.68 & 3.63 \\
+15 & Point Atkinson & -123.25 & 49.33 & 1.01 & -0.48 & 1.07  & 1 & -0.56 & 0.94 \\
+16 & Gibsons Landing & -123.5 & 49.4 & 0.98 & 0.75 & 2.34  & 1 & 0.96 & 1.47 \\
+17 & Winchelsea & -124.08 & 49.3 & 1.01 & -0.19 & 0.78  & 1.01 & 0.51 & 1.30 \\
+18 & Halfmoon Bay & -123.92 & 49.52 & 1.01 & -0.15 & 0.83  & 1.01 & 0.15 & 0.72\\ 
+19 & Irvines Landing & -124.05 & 49.63 & 1 & -0.32 & 0.59  & 1.01 & 0.24 & 1.12 \\
+20 & Powell River & -124.55 & 49.86 & 1 & -1.72 & 3.05  & 1 & -0.01 & 0.43 \\
+21 & Little River & -124.92 & 49.74 & 1.03 & 0.07 & 2.54  & 1.01 & 0.33 & 0.73\\ 
+22 & Lund & -124.76 & 49.98 & 1.02 & -2.52 & 4.98  & 1.02 & -0.95 & 2.45 \\
+23 & Twin Islets & -124.93 & 50.03 & 1.02 & -2.05 & 4.22  & 1.01 & -0.24 & 0.72\\ 
+24 & Campbell River & -125.24 & 50.04 & 0.99 & 0.51 & 1.12  & 1 & 0.94 & 1.39 \\
+25 & Maude Island E & -125.33 & 50.13 & 1.13 & -10.02 & 12.62  & 0.97 & -1.44 & 2.89\\ 
+26 & Nymphe Cove & -125.36 & 50.13 & 1.02 & -5.43 & 6.04  & 0.97 & -0.29 & 2.02 \\
+27 & Seymour Narrows & -125.34 & 50.13 & 0.63 & 31.31 & 53.40  & 1.14 & 10.91 & 17.28 \\
+28 & Brown Bay & -125.37 & 50.16 & 0.98 & -5.63 & 9.26  & 0.98 & -0.04 & 1.17 \\
+29 & Chatham Point & -125.44 & 50.33 & 1.09 & -5.49 & 12.18  & 0.98 & -3.18 & 3.85 \\
+30 & Kelsey Bay & -125.96 & 50.39 & 1 & -1.69 & 3.50  & 0.98 & -0.39 & 1.00 \\
+31 & Yorke Island & -125.97 & 50.44 & 0.99 & 1.59 & 3.32  & 1.01 & 1.23 & 1.37\\ 
 \hline 
-            &  Mean Error      &     &                      &          &     &                      &          \\
-            &  RMS Error       &     &                      &          &     &                      &          \\
+&   Mean Error (cm/deg) & & & 4.23  & 6.5 & 8.05   & 1.56 & 1.49 & 2.58 \\
+&   RMS Error (cm/deg)  & & & 7.67  & 10.73 & 12.73 & 2.41 & 2.57 & 3.91 \\
 \hline  

### 2014-10-17 RoughWork/manuscript.tex
Finished tides

@@
 Further details on the tuning are provided in the appendix \ref{sec:appendix}.
-% Needs editing! Update when we are finished with tides.
 
@@
 For example, the amphidrome around Victoria was produced in the M$_2$ results, as well as the monotonic increase in $K_1$ amplitude moving northwards along the Strait of Georgia.
-Initially, the Johnstone Strait boundary was closed, but modelled M$_2$ amplitudes were too small compared to measured amplitudes...
-TBC
-
-%perhaps a nice contour map of M2 and K1 amps and phases goes here?
-
-Once our model was reproducing observed tidal patterns, model results were quantitatively evaluated by comparing modelled harmonic constituents to measured harmonic constituents at tidal measuring stations throughout the domain.
+Once the model was reproducing observed tidal patterns, model results were quantitatively evaluated by comparing modelled harmonic constituents to measured harmonic constituents at tidal measuring stations throughout the domain.
 Comparisons were made using the complex difference ($D$), defined by \citep{foreman1995tidal} as:
@@
 where $A_0$, $A_m$, $g_0$ and $g_m$ are observed and modelled amplitudes and phases.
+Figure \ref{fig:tides} compares modelled and observed amplitude and phase for the $M_2$ and $K_1$ tidal constituents for stations along a transect from the Strait of Juan de Fuca through to the Strait of Georgia and Johnstone Strait.
+Complex differences for these two constituents are also provided.
+Comparisons have been made with data from \citet{foreman1995tidal}, \citet{foreman2004m} and \citet{foreman2012circulation}.
+Table  \ref{tab:comparison} lists station names and numbers in the Appendix \ref{sec:appendix}.
+
+In general, the K$_1$ tidal harmonics are well represented by the model with complex differences below 5 cm at all stations except Seymour Narrows (station 27).
+Seymour Narrows is a very constricted region in Johnstone Strait and is not well resolved by the model.
+As such, the M$_2$ amplitude and phase are also poorly reproduced here.
+In addition, M$_2$ amphidrome in the model, indicated by the drop in amplitude around station 5 (Victoria), is too strong resulting in lower than observed M$_2$ amplitudes in this region.
+Additionally, the M$_2$ phase errors in the Strait of Juan de Fuca region (stations 1-7) are quite large, resulting in large complex differences in this region.
+However, the match in the Strait of Georgia region (stations 10-24) is very good, with complex differences less than 6 cm, in part due to tidal tuning for a good match in the main area of interest.
+Details on the tidal tuning procedure are provided in the Appendix \ref{sec:appendix}.
+%still need to look up how other models of this region do with tides.
 
@@
 %(would be cool to include the complex differences calculated at the VENUS nodes too)
-
-Complex differences were less than ??cm at all stations in our domain, which was assumed to be acceptable for our purposes. 
+%perhaps a nice contour map of M2 and K1 amps and phases goes here?
 %if it's favourable, we could compare our complex differences to Foreman et al (1995), who got an average of D=3cm for M2 and D=2.5cm for K1
@@
 On the right, modelled M$_2$ (black) and K$_1$ (gray) amplitude (top) and phase (middle) compared with observations in the dashed curves of the same colour.
-The bottom plot displays the M$_2$ and K$_1$ complex differences.
-}
+The bottom plot displays the M$_2$ and K$_1$ complex differences.}
 \label{fig:tides}
@@
 Storm surge simulations were started from rest at least three days prior to the event of interest to allow for adjustment of model currents.
-Stratification conditions from model spin-up at a time close to the event (within 5 days) were chosen to initialize the temperature and salinity fields.
+Stratification conditions from model spin-up at a time close to the event (within five days) were chosen to initialize the temperature and salinity fields.
 Boundary conditions include atmospheric forcing and fresh water input from rivers at the surface, sea surface height anomaly from observations at the open boundary, and tidal forcing with eight tidal constituents.
-Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height predictions due to the omission of the next leading order tidal constituents such as J$_1$.
-Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can be up to 40 cm (check) during times of high tidal range, and in particular, during the storm surge season Nov-Feb.
+Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height due to the omission of the next leading order tidal constituents such as J$_1$.
+Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can be up to 40 cm during times of high tidal range, and in particular, during the storm surge season Nov-Feb.
 As such, storm surge model predictions were corrected by accounting for this error.
@@
 In addition, the contribution to the surge due to several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed.
-We have also examined the spatial variability of the surge as it propagates through the domain and have considered a case with high-resolution atmospheric forcing.
+We have also examined the spatial variability of the surge as it propagates through the domain and have considered a case with a strong wind event but no siginificant observed surge.
 
@@
 where $M_i' = M_i-\bar{O}$ and $O_i'=O_i-\bar{O}$ and $\bar{O}$ is the mean of the observations. This score is bounded by zero and one. Values close to one indicate good model skill.
-
-%First, a study of a large surge event on Feb 4, 2006 is presented.
-This event caused significant damage and flooding in Boundary Bay, a community south of Richmond, (citation) and has been chosen as a case study to demonstrate the model's skill in reproducing a storm surge and to quantitatively assess forcing conditions that are important in storm surge hind casting.
-Second, an extreme wind event on Dec 15, 2006 is examined.
-This event produced winds over (x) (citation) but did not result in significant flooding because the storm did not occur at high tide, however, it did result in significant property damage due to high winds and waves.
-This event has been used to understand the spatial and temporal variability of the surge propagation.
-Next, a large wind event on Nov 18, 2009 that did not result in a significant surge or flooding is presented.
-This case was chosen to identify the model's capability in handling large wind events that do not result in high water levels.
-Next, a case with high resolution atmospheric forcing for a recent surge on Dec 17, 2012 is discussed.
-Lastly, these cases presented in detail will be combined with $x$ additional runs for statistical evaluation of model performance.
 
@@
 \label{fig:dec2006}
-\end{figure}%table instead of figure?
+\end{figure}%table instead of figure? or maybe just the surge and not water level?
 
@@
 \section{Discussion}\label{sec:diss}
+
 %future directions: running real time, biology and chemistry, model improvements?
@@
 
+\appendix
 \section{Appendix}\label{sec:appendix}
+This section details the procedure used to tune the tidal forcing for a good match with observations in the Strait of Georgia.
+Originallly, tidal elevation and currents were extracted from a tide prediction tool called Webtide \citep{foreman2000webtide} at the western Juan de Fuca boundary.
+At the northern boundary,  M$_2$ and K$_1$ tidal harmonics for both elevation and currents were taken from \citet{thomson1980johnstone}, as well as elevation harmonics for O$_1$ and S$_2$.
+The remaining constituents were extrapolated from Webtide based on the ratio between M$_2$/S$_2$ amplitudes and phase differences for the semi-diurnal constituents and K$_1$/O$_1$ for the diurnal constituents.
+However, these forcing conditions did not result in a good match between observed and modelled tidal elevation harmonics in the Strait of Georgia, the main region of interest (although the matches in the Strait of Juan de Fuca were adequate).
+As such, tidal forcing was tuned, by first adjusting the western boundary M$_2$ anc K$_1$ phase and K$_1$ amplitude for a good match in the Strait of Georgia (specifically at Point Atkinson, Gibsons Landing, Winchelsea and Halfmoon Bay).
+This tuning was performed on simulations that used only M$_2$ and K$_1$ forcing.
+Next, the M$_2$ and K$_1$ fluxes at the northern boundary were adjusted because the cross-sectional area of transect used by \citet{thomson1980johnstone} to measure tidal currents in Johnstone Strait is smaller than the cross-sectional area of our northern boundary.
+Finally, the remaining constituents were added, starting with O$_1$, then S$_2$, and then in pairs P$_1$/N$_2$ and Q$_1$/K$_2$.
+Adjustments were made if the addition of new constituents hindered performance of the main M$_2$ and K$_1$ components.
+
+Below, Table \ref{tab:comparison} compares the M$_2$ and K$_1$ elevation harmonics with data from \citet{foreman1995tidal}, \citet{foreman2004m} and \citet{foreman2012circulation}.
+Overall, the average complex difference for M$_2$ is 8.05 cm and for K$_1$ is 2.58 cm, but both constituents perform very well in the Strait of Georgia region.
+
 \begin{table}[h]
@@
 \end{table}
-%May have to include station numbers and locations on a map? Maybe on the contour map, overlay with stations and green/red dots (hopefully mostly green)
-%Should we use the new method of calculating tides?
 

### 2014-10-17 RoughWork/manuscript.tex
started conclusions

@@
 The numerical model employs the Nucleus for European Modelling of the Ocean (NEMO) architecture in a regional configuration.
-The model is evaluated through comparisons of tidal harmonics, tidal current ellipses, and storm surge water levels with observations.
+The model is evaluated through comparisons of tidal elevation harmonics and storm surge water levels with observations.
 Important forcing factors contributing to storm surges have been assessed, finding that the surge entering the domain from the Pacific Ocean contributes most significantly.
@@
 At the ocean surface, meteorological conditions and fresh water input from rivers are included.
-The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2013new}.
+The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2014new}.
 Forecasts from the period of 2002-2011 are available.
@@
 \section{Discussion}\label{sec:diss}
+In this article, we have configured a three-dimensional numerical ocean model of the Salish Sea for accurate hindcasts of storm surges in the Strait of Georgia.
+The model is forced with eight tidal constituents at the western and northern open boundaries, as well as sea surface height anomaly from de-tided observations at Tofino and Port Hardy.
+Additionaly, wind stress and an inverse barometer effect were forced at the ocean surface using an atmopsheric reforecasting product called CGRF \citep{smith2014new}.
+Storm surge simulations began with zero initial conditions for the currents and sea surface height and were commenced several days prior to the event of interest to spin up the velocity fields.
+
+Hindcasts of significant events in February 2006, December 2006 and Novemeber 2009 were simulated.
+In general, the model performs well at reproducing the observed water level heights and surge amplitudes, with Willmott skill scores greater than 0.88 in all cases.
+We have also assessed the relative importance of different forcing factors contributing to the storm surge.
+The sea surface height anomaly entering the system from the Pacific Ocean contributes most significantly to the surge amplitude (>50 \%), in agreement with previous studies in this region \citep{murty1995storm}.
+We have also found that the wind and atmospheric conditions can set up local differences in the surge amplitude, resulting in local variations in the spatial distribution of the surge.
+Generally, surge amplitudes in the Strait of Georgia are higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions.
+Results from an additional simulation with no tidal forcing but only surge propagation and atmopsheric forcing found that nonlinear effects due to tide-surge interaction were most important after the San Juan and Gulf Islands.
+
+Several factors considered in other storm surge modelling studies have been neglected.
+Primarly, effects due to wetting and drying have not been included.
+This is an important factor to consider in shallow, low lying areas where water levels retreat significantly to reveal land.
+In this article, the four locations used to assess storm surge accurcay are not in shallow regions and so wetting and drying is not considered to be important.
+However, several communities in the Strait of Gerogia, such as Boundary Bay and Crescent Beach are in shallow regions and are typically most affected by storm surge during the winter season.
+As such, efforts to include wetting and drying will be considered in the future.
+Additionally, sea surface height anomaly from Neah Bay will be used in future real time water leve predictions during the storm surge season.
+
+The Strait of Georgia is a semi-enclosed body of water
 

### 2014-10-18 RoughWork/manuscript.tex
Mostly complete draft

@@
 \usepackage{amsmath}
+\usepackage{setspace} \doublespacing
 
@@
 Other regions require modelling over a very large domain since non-local effects from the open sea may influence surge propagation \citep{weisberg2006hurricane, lane2009verification}.
-Such large domains typically use nested or unstructered grids to resolve the details of complicated coastlines.
+Such large domains typically use nested or unstructured grids to resolve the details of complicated coastlines.
 Appropriate treatment of rivers \citep{flather1994storm}, accurate bathymetry and the inclusion surface waves \citep{xu2010storm} have also been identified as important issues.
@@
 This article presents the development of a three-dimensional numerical ocean model for the Strait of Georgia and its accuracy in simulating storm surge hindcasts.
-This baraoclinic model includes the effect of stratification and rivers, however, wetting and drying is not represented.
+This baroclinic model includes the effect of stratification and rivers, however, wetting and drying is not represented.
 Model evaluation has focused on regions where the near-coast slopes are steep.
@@
 The remaining constituents were extrapolated from Webtide.
-At the western boundary, the tidal elevation and currents forcing parameters were tuned for a good match wiht observations at locations within the Strait of Georgia, including Point Atkinson, Gibsons Landing, Winchelsea and Halfmoon Bay.
-At the nothern boundary, the tides were tuned for good matches at Yorke Island and Kelsey Bay.
+At the western boundary, the tidal elevation and currents forcing parameters were tuned for a good match Whit observations at locations within the Strait of Georgia, including Point Atkinson, Gibsons Landing, Winchelsea and Halfmoon Bay.
+At the northern boundary, the tides were tuned for good matches at Yorke Island and Kelsey Bay.
 Further details on the tuning are provided in the appendix \ref{sec:appendix}.
@@
 The modelled residual is calculated as the difference in model output between a simulation with all forcing conditions and a simulation forced only with tides and rivers.
-Water level observations are taken from Fisheries and Oceans Canada (citation) and are used to calculate the osberved residual, defined as the observed water level minus tidal predictions.
+Water level observations are taken from Fisheries and Oceans Canada (citation) and are used to calculate the observed residual, defined as the observed water level minus tidal predictions.
 Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this case, from the year prior surge event of interest.
@@
 In addition, the contribution to the surge due to several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed.
-We have also examined the spatial variability of the surge as it propagates through the domain and have considered a case with a strong wind event but no siginificant observed surge.
+We have also examined the spatial variability of the surge as it propagates through the domain and have considered a case with a strong wind event but no significant observed surge.
 
@@
 Smaller values of $\gamma^2$ indicate better model skill.
-This quantity was used by \citet{bernier2006predicting} and \citet{bernier2010tide} in storm surge predictions in Eastern Canada and is caclulated here as a point of comparison.
+This quantity was used by \citet{bernier2006predicting} and \citet{bernier2010tide} in storm surge predictions in Eastern Canada and is calculated here as a point of comparison.
 A third measure, called the Willmott skill score \citep{willmott1982some}, is also used to assess overall model skill in predicting the water level and residual.
@@
 There is a good match in the water level maximum at Victoria and Patricia Bay, with a relative error less than 2\%.
-There is only a 4.2 cm difference between observed and modelled water level elevation at Victoria and a 6.9 cm difference at Patricia Bay, the model slightly underpredicting the maximum water level in each case.
-The error at Point Atkinson and Campbell River is slightly more significant, reaching 2.5\% and 2.7\% respectively, close to 14 cm overpredicted in each case.
-It is suprising that Victoria and Patricia Bay agree most closely with observations because the M$_2$ tidal amplitudes in the model are known to be too low in this area.
-In conjunction with the overpredicted water levels at Point Atkinson and Campbell River, it is likely that one of the forcing factors is overestimated in this example.
+There is only a 4.2 cm difference between observed and modelled water level elevation at Victoria and a 6.9 cm difference at Patricia Bay, the model slightly under predicting the maximum water level in each case.
+The error at Point Atkinson and Campbell River is slightly more significant, reaching 2.5\% and 2.7\% respectively, close to 14 cm over predicted in each case.
+It is sup rising that Victoria and Patricia Bay agree most closely with observations because the M$_2$ tidal amplitudes in the model are known to be too low in this area.
+In conjunction with the over predicted water levels at Point Atkinson and Campbell River, it is likely that one of the forcing factors is overestimated in this example.
 
 A comparison between observed and modelled residuals is also displayed in the panels on the right.
-In general, there is a very good agreement between the observed and modelled residules.
+In general, there is a very good agreement between the observed and modelled residuals.
 The most noticeable difference is a 4-6 hour delay in the model's timing of the maximum residual except at Victoria where the maximum residual in sync with observations.
@@
 The skill scores degrade slightly for the residuals but are still reasonable and indicate very good agreement.
-Note that the mean absolute error ($\bar{e}$) is lower for the residuals, but this measure does not account for the larger variablity in water levels due to the tides.
+Note that the mean absolute error ($\bar{e}$) is lower for the residuals, but this measure does not account for the larger variability in water levels due to the tides.
 
@@
 \caption{Evaluation of model performance for the Feb 2006 simulation.
-Mean absolute error ($\bar{e}$), observed mean (subscript o), modelled mean (subscript m), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and resdiuals ($\delta$) during the period displayed in Figure \ref{fig:feb2006}.}
+Mean absolute error ($\bar{e}$), observed mean (subscript o), modelled mean (subscript m), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and residuals ($\delta$) during the period displayed in Figure \ref{fig:feb2006}.}
 \label{tab:feb2006stat} 
 \end{table} 
-%anything else to include? statistics calculated in Feb 2006 -paper prodcution.ipynb and saved in statisitcs_feb2006.csv 
+%anything else to include? statistics calculated in Feb 2006 -paper production.ipynb and saved in statistics_feb2006.CV 
 %I also have standard deviations, correlations, ...
@@
 The contribution to the surge due to atmospheric forcing appears to be geographically dependent, its removal leading to a drop of about 2.5 cm in the maximum surge at Point Atkinson and 2.3 cm at Campbell River.
-Interestingly, the surge amplitude increases by 4.8 cm at Victoria and 3.6 cm at Patricia Bay when atmospheric forces are negelcted.
+Interestingly, the surge amplitude increases by 4.8 cm at Victoria and 3.6 cm at Patricia Bay when atmospheric forces are neglected.
 The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
-Additionally, the surge amplitude decreases by a few centimtres if the inverse barometer effect due to atmospheric pressure is not included.
+Additionally, the surge amplitude decreases by a few centimetres if the inverse barometer effect due to atmospheric pressure is not included.
 
@@
 \includegraphics[scale=0.6]{Figures/dec2006.pdf}
-\caption{Comparison of observed and modelled water level and residuls during the Dec 15, 2006 storm surge simulation. }
+\caption{Comparison of observed and modelled water level and residuals during the Dec 15, 2006 storm surge simulation. }
 \label{fig:dec2006}
@@
 The evolution and spatial distribution of the surge is highlighted in Figure \ref{fig:spatial} where the modelled residual is plotted every hour beginning on Dec 15 at 05:30 UTC through to Dec 15 at 14:30 UTC.
-The surge enters the system from the Strait of Juna de Fuca, peaking between Dec 15 at 9:30 and Dec 15 at 11:30 and then begins to subside throughout the entire domain.
+The surge enters the system from the Strait of Juan de Fuca, peaking between Dec 15 at 9:30 and Dec 15 at 11:30 and then begins to subside throughout the entire domain.
 Notably, the Strait of Georgia experiences the largest surge values, upwards of 60 cm, whereas the Juan de Fuca region experiences a maximum surge around 45 cm.
@@
 Time advances from top left to bottom right.
-Wind vectors from the atmopsheric forcing are overlaid in black.}
+Wind vectors from the atmospheric forcing are overlaid in black.}
 \label{fig:spatial}
@@
 An evaluation of the spatial distribution of the surge in a simulation without atmospheric forcing is provided in Figure \ref{fig:spatial_sshonly}.
-In this case, the surge amplitude across the Strait of Georgia, is uniform with no pronouced difference between the mainland side and Vancouver Island side.
+In this case, the surge amplitude across the Strait of Georgia, is uniform with no pronounced difference between the mainland side and Vancouver Island side.
 While the surge entering the domain from the Pacific Ocean is the leading order contribution, the winds may induce localized changes to the surge amplitude.
 Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland.
-Even without atmospheric forcing, there is still a siginificant difference in surge amplitude between the Strait of Georgria and the Strait of Juan de Fuca, perhaps because the Strait of Georgia is also affected by the surge entering the domain from Johnstone Strait at the northern boundary.
+Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the Strait of Georgia is also affected by the surge entering the domain from Johnstone Strait at the northern boundary.
 
@@
 \includegraphics[scale=0.6]{Figures/dec2006_spatial_sshonly.pdf}
-\caption{As in Figure \ref{fig:spatial} but for a simulation that does not include atmopsheric forcing. }
+\caption{As in Figure \ref{fig:spatial} but for a simulation that does not include atmospheric forcing. }
 \label{fig:spatial_sshonly}
@@
 The largest surge did not coincide with the strong wind event which peaked at Nov 19 01:00 UTC, reflected in both the model and observations.
-At Point Atkinsons, the highest observed surge reached 62.5 cm on Nov 20 at 10:30 UTC, while the model recorded a maximum surge of 50.5 cm five hours later.
+At Point Atkinson, the highest observed surge reached 62.5 cm on Nov 20 at 10:30 UTC, while the model recorded a maximum surge of 50.5 cm five hours later.
 On Nov 19, the observed surge reached upwards of 50 cm, which is low when compared to the other cases considered.
-Neither of the surges occured during a high tide and so no high water level events were recorded.
+Neither of the surges occurred during a high tide and so no high water level events were recorded.
 Although not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and another surge entering the system at Tofino and Port Hardy.
@@
 The model is forced with eight tidal constituents at the western and northern open boundaries, as well as sea surface height anomaly from de-tided observations at Tofino and Port Hardy.
-Additionaly, wind stress and an inverse barometer effect were forced at the ocean surface using an atmopsheric reforecasting product called CGRF \citep{smith2014new}.
+Additionally, wind stress and an inverse barometer effect were forced at the ocean surface using an atmospheric reforecasting product called CGRF \citep{smith2014new}.
 Storm surge simulations began with zero initial conditions for the currents and sea surface height and were commenced several days prior to the event of interest to spin up the velocity fields.
 
-Hindcasts of significant events in February 2006, December 2006 and Novemeber 2009 were simulated.
+Hindcasts of significant events in February 2006, December 2006 and November 2009 were simulated.
 In general, the model performs well at reproducing the observed water level heights and surge amplitudes, with Willmott skill scores greater than 0.88 in all cases.
@@
 Generally, surge amplitudes in the Strait of Georgia are higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions.
-Results from an additional simulation with no tidal forcing but only surge propagation and atmopsheric forcing found that nonlinear effects due to tide-surge interaction were most important after the San Juan and Gulf Islands.
+Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to tide-surge interaction were most important after the San Juan and Gulf Islands.
 
 Several factors considered in other storm surge modelling studies have been neglected.
-Primarly, effects due to wetting and drying have not been included.
+Primarily, effects due to wetting and drying have not been included.
 This is an important factor to consider in shallow, low lying areas where water levels retreat significantly to reveal land.
-In this article, the four locations used to assess storm surge accurcay are not in shallow regions and so wetting and drying is not considered to be important.
-However, several communities in the Strait of Gerogia, such as Boundary Bay and Crescent Beach are in shallow regions and are typically most affected by storm surge during the winter season.
+In this article, the four locations used to assess storm surge accuracy are not in shallow regions and so wetting and drying is not considered to be important.
+However, several communities in the Strait of Georgia, such as Boundary Bay and Crescent Beach are in shallow regions and are typically most affected by storm surge during the winter season.
 As such, efforts to include wetting and drying will be considered in the future.
-Additionally, sea surface height anomaly from Neah Bay will be used in future real time water leve predictions during the storm surge season.
+Additionally, sea surface height anomaly from Neah Bay will be used in future real time water level predictions during the storm surge season.
 
-The Strait of Georgia is a semi-enclosed body of water
+Since the Strait of Georgia is a large, semi-enclosed body of water storm surges can behave and evolve differently than in other coastal regions. To be continued...
 
@@
 This section details the procedure used to tune the tidal forcing for a good match with observations in the Strait of Georgia.
-Originallly, tidal elevation and currents were extracted from a tide prediction tool called Webtide \citep{foreman2000webtide} at the western Juan de Fuca boundary.
+Originally, tidal elevation and currents were extracted from a tide prediction tool called Webtide \citep{foreman2000webtide} at the western Juan de Fuca boundary.
 At the northern boundary,  M$_2$ and K$_1$ tidal harmonics for both elevation and currents were taken from \citet{thomson1980johnstone}, as well as elevation harmonics for O$_1$ and S$_2$.
@@
 However, these forcing conditions did not result in a good match between observed and modelled tidal elevation harmonics in the Strait of Georgia, the main region of interest (although the matches in the Strait of Juan de Fuca were adequate).
-As such, tidal forcing was tuned, by first adjusting the western boundary M$_2$ anc K$_1$ phase and K$_1$ amplitude for a good match in the Strait of Georgia (specifically at Point Atkinson, Gibsons Landing, Winchelsea and Halfmoon Bay).
+As such, tidal forcing was tuned, by first adjusting the western boundary M$_2$ and K$_1$ phase and K$_1$ amplitude for a good match in the Strait of Georgia (specifically at Point Atkinson, Gibsons Landing, Winchelsea and Halfmoon Bay).
 This tuning was performed on simulations that used only M$_2$ and K$_1$ forcing.
@@
 This project is funded by the Marine Environmental Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada.
-%I'm likely missing a lot more (Keith, Youyu, Vasily, JP, Rich, Mark). Not sure about authorship but something to think about for later....
+Others like Keith, Youyu, Vasily, JP, Rich, Mark?
 

### 2014-10-20 RoughWork/manuscript.tex
Expanded on discussion section. Added other authors

@@
 \title{Storm surge modelling in the Strait of Georgia}
-\author{Nancy Soontiens, Susan Allen, Doug Latornell, Kate Le Sou\"{e}f, Idalia Machuca}
+\author{Nancy Soontiens, Susan Allen, Doug Latornell, Kate Le Sou\"{e}f, Idalia Machuca,
+\\ Youyu Lu, Keith Thompson, Jean-Philippe Paquin, Vasily Korabel}
 \date{\today}
@@
 Important forcing factors contributing to storm surges have been assessed, finding that the surge entering the domain from the Pacific Ocean contributes most significantly.
-In addition, local wind patterns cause a slight increase in surge amplitude on the mainland side of the Strait of Georgia when compared to the Vancouver Island coastal areas.
-Surge amplitudes are found to be greater within the Strait of Georgia when compared to amplitudes in the Strait of Juan de Fuca.
+In addition, local wind patterns cause a slight increase in surge amplitude on the mainland side of the Strait of Georgia when compared to the Vancouver Island coastal areas during a major wind storm on Dec 15, 2006.
+Generalll, surge amplitudes are found to be greater within the Strait of Georgia when compared to amplitudes in the Strait of Juan de Fuca.
 %needs a little more work.
@@
 As such, the split-explicit time stepping algorithm is employed, where the free surface and barotropic equations are solved with a smaller time step than that used for the other variables.
-The model time step is and barotropic time step are 10 s and 2 s respectively.
-Small vertical grid spacing and large vertical velocities in the Boundary Pass and Haro Strait region impose a fairly restrictive step through the vertical CFL condition.
+The model time step and barotropic time step are 10 s and 2 s respectively.
+Small vertical grid spacing and large vertical velocities in the Boundary Pass and Haro Strait region impose a fairly restrictive time step through the vertical CFL condition.
 
@@
 
+\subsection{An Additional Hindcast: November 16, 2006}
+
 \subsection{Spatial Extent of Surge: Dec 15, 2006}
@@
 
-Hindcasts of significant events in February 2006, December 2006 and November 2009 were simulated.
+Hindcasts of significant events in February 2006, November 2006, December 2006 and November 2009 were simulated.
 In general, the model performs well at reproducing the observed water level heights and surge amplitudes, with Willmott skill scores greater than 0.88 in all cases.
@@
 The sea surface height anomaly entering the system from the Pacific Ocean contributes most significantly to the surge amplitude (>50 \%), in agreement with previous studies in this region \citep{murty1995storm}.
-We have also found that the wind and atmospheric conditions can set up local differences in the surge amplitude, resulting in local variations in the spatial distribution of the surge.
-Generally, surge amplitudes in the Strait of Georgia are higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions.
-Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to tide-surge interaction were most important after the San Juan and Gulf Islands.
+We have also found that the wind and atmospheric conditions can set up spatial differences in the surge amplitude, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season.
+Additionally, surge amplitudes in the Strait of Georgia are usually higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions.
+Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important after the San Juan and Gulf Islands.
+The modelled surge amplitude increased by up to 10 \% at Point Atkinson when the tide-surge interaction was neglected.
 
-Several factors considered in other storm surge modelling studies have been neglected.
+Several factors considered in other storm surge modelling studies have been neglected in these simulations.
 Primarily, effects due to wetting and drying have not been included.
@@
 Additionally, sea surface height anomaly from Neah Bay will be used in future real time water level predictions during the storm surge season.
+Lastly, flooding effects due to wave run up is left for a more detailed inundation study with higher resolution in the near-shore areas.
 
-Since the Strait of Georgia is a large, semi-enclosed body of water storm surges can behave and evolve differently than in other coastal regions. To be continued...
+Since the Strait of Georgia is a large, semi-enclosed body of water storm surges can behave and evolve differently than in other coastal regions.
+For instance, the complicated coastlines and local wind patterns result in important spatial variations in the surge amplitude.
+As such, strong winds caused by a large storm is not the only factor to be considered in community preparedness for surge conditions.
+The geometry of the Strait and local wind patterns render communities on the mainland side of the southern Strait of Georgia more susceptible to surges.
+Additionally, commmunities such as Campbell River are at risk due to the long fetch and predominately southerly winds during the winter months.
+
+Finally, difficulties in reproducing accurate M$_2$ tidal harmonics in both the Strait of Georgia and the Strait of Juan de Fuca have been highlighted in this study among others \citep{stronach1993update, foreman2004m}.
+Future work will attempt to improve model accuracy by examining mixing parametrizations and energy dissipation due to bottom friction and their effect on tidal haromnics.
+Additionally, a biological-chemical modelling compenent will be added.
 
@@
 $D$ is the complex difference.
+The mean error and RMS error are calculated as the difference between the modelled quantity and observered quantity.
 }
@@
 \section{Acknowledgements}\label{sec:ack}
-We would like to thank Michael Foreman, Diane Masson, and Luc Fillion for providing data used in model set up and evaluation.
+We would like to thank Michael Foreman, Diane Masson, Luc Fillion Rich Pawlowicz, and Mark Halverson for providing data used in model set up and evaluation.
 This project is funded by the Marine Environmental Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada.
-Others like Keith, Youyu, Vasily, JP, Rich, Mark?
 

### 2014-10-22 RoughWork/manuscript.tex
Brief overview of the Nov 2006 case

@@
 \subsection{An Additional Hindcast: November 16, 2006}
+An additional storm surge hindcast for November 2006 is provided as another means of assessing model performance and is mentioned briefly here.
+On November 16, 2006 communities in the Strait of Georgia experienced a strong wind storm with speeds upwards of 20 m/s near Point Atkinson resulting in observed water levels of 5.1 m and a maximum surge amplitude of 83.7 cm.
+The surge amplitude was lower at the other stations, reaching no more than 68 cm.
+The model performs reasonably well in this hindcast, Willmott skill scores less greater than 0.94 as outlined in Table \ref{tab:nov2006stat}.
+The best match in maximum surge occured at Campbell River 6.2 cm underestimate by the model.
+The timing of the maximum surge at Campbell River was identical with observations and slightly earlier at the other stations.
+
+\begin{table}[h]
+\centering 
+\begin{tabular}{|l |c c c c c | c c c c c|} 
+\hline 
+& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
+\hline 
+Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
+\hline 
+Point Atkinson &  9.8           &  3.42                & 3.43                 &   0.034    & 0.991  &  7.7           &  0.290                 & 0.247                  &  0.128     & 0.950 \\
+Victoria       &  9.1           &  2.16                & 2.11                 &   0.070    & 0.978  &  6.2           &  0.222                 & 0.247                  &  0.171     & 0.946 \\
+Patricia Bay   & 10.0           &  2.57                & 2.53                 &   0.049    & 0.985  &  6.6           &  0.235                 & 0.203                  &  0.144     & 0.947 \\
+Campbell River & 10.7           &  3.20                & 3.26                 &   0.037    & 0.982  &  6.2           &  0.254                 & 0.236                  &  0.140     & 0.952 \\
+\hline 
+\end{tabular}
+\caption{Evaluation of model performance for the Nov 2006 hindcast.}
+\label{tab:nov2006stat} 
+\end{table}
 

### 2014-10-28 RoughWork/manuscript.tex
Fixed one sentence in discussion
Added Susan's corrections and  suggestions in comments.

@@
 The Strait of Georgia is a strongly stratified, semi-enclosed body of water located between Vancouver Island and the mainland of British Columbia.
-It is part of a larger system of waterways collectively known as the Salish Sea and is connected to the Pacific Ocean via the Strait of Juan de Fuca and Johnstone Strait.
-It is on average 222 km in length and 28 km wide and its maximum depth is 420 m \citep{thomson1981oceanography}.
+It is part of a larger system of waterways collectively known as the Salish Sea and is connected to the Pacific Ocean via the Strait of Juan de Fuca and Johnstone Strait (Figure \ref{fig:domain}).
+It is on average 220 km in length and 28 km wide and its maximum depth is 420 m \citep{thomson1981oceanography}.
 At its southern end, the Strait of Georgia is connected to the Strait of Juan de Fuca through several passages, the largest being Haro Strait.
@@
 \citet{foreman2004m} used data assimilation techniques to study M$_2$ tidal dissipation without local changes to the bottom friction in a two-dimensional finite element model.
+%were Crean/Stronach modelled phases/amps too high or low?
 
@@
 These studies have led to an improvement in understanding circulation properties in the Salish Sea and have highlighted important features that should be considered in future modelling work.
+%What was interesting about Yang et al?
+%Tides in Sutherland? JdF surface salinity in Giddings
+%Add more about reproduction of tides in other models 
+%Last sentance is lame
 
@@
 Additionally, there is a small contribution to increased sea level due to the thermal expansion of water during warmer years set by the El Nino Southern Oscillation (ENSO) \citep{abeys2011extreme}.
+%discuss this as a Kelvin wave
 
@@
 The highest water levels are observed in the winter around the perigean spring tide or ``King" tide, when the sun and moon are aligned with the Earth and the moon is at its perigee.
-As such, storm surges in the Strait of Georgia are typically observed between November and February.
-Several historical examples have been documented by \citet{forseth2006adaptation}.
+As such, storm surges in the Strait of Georgia are typically observed between November and February \citep{forseth2006adaptation}.
 
@@
 \section{Model Configuration}\label{sec:config}
-%To be included: NEMO overview, tides, rivers, bathymetry, atmospheric forcing, vertical/lateral mixing, boundary conditions, grid.
 
 %A section about how we configured the model for SoG 
-We have used the Nucleus for European Modelling of the Ocean (NEMO) framework in its regional configuration to develop an ocean model for the Strait of Georgia and Salish Sea.
+We have used the Nucleus for European Modelling of the Ocean (NEMO) framework in its regional configuration to develop an ocean model for the Strait of Georgia and Salish Sea \citep{madec2012nemo}.
 NEMO is a highly modularized tool used for studying ocean physics, ocean-ice interactions, and the bio-geochemical properties of the ocean.
@@
 %image may need more work
+%Add Johnstone Strait, Vancouver, Fraser River, Haro Strait, San Juan/Gulf Islands
 
@@
 The remaining constituents were extrapolated from Webtide.
-At the western boundary, the tidal elevation and currents forcing parameters were tuned for a good match Whit observations at locations within the Strait of Georgia, including Point Atkinson, Gibsons Landing, Winchelsea and Halfmoon Bay.
+At the western boundary, the tidal elevation and currents forcing parameters were tuned for a good match with observations at locations within the Strait of Georgia, including Point Atkinson, Gibsons Landing, Winchelsea and Halfmoon Bay.
 At the northern boundary, the tides were tuned for good matches at Yorke Island and Kelsey Bay.
@@
 The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}.
-The baroclinic velocities at the boundary were set to be equal to the values inside the boundary (zero-gradient boundary conditions).
+Ath the west, the baroclinic velocities at the boundary were set to be equal to the values inside the boundary (zero-gradient boundary conditions).
 This scheme is not part of core NEMO.
 Zero gradient conditions were chosen because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain.
+At the north baroclinic velocities were set to zero.
 
@@
 Partial slip allows one to include the frictional effects of lateral boundaries without the restrictive resolution required to represent the lateral boundary layer under no slip conditions.
-A lateral eddy viscosity of 20 m $^2$ s $^{-1}$ parametrizes horizontal friction and a lateral eddy diffusivity of 20.5 m $^2$ s $^{-1}$ is used.
+A lateral eddy viscosity of 20 m$^2$ s$^{-1}$ parametrizes horizontal friction and a lateral eddy diffusivity of 20.5 m$^2$ s$^{-1}$ is used.
 Bottom friction is represented by a quadratic law for the bottom momentum flux with drag coefficient $C_D = 5\times 10^{-3}$.
@@
 The modelled sea surface height adjusted to forcing in less than one day.
+%citation for Compute Canada resources?
 
@@
 \subsection{Tidal evaluation}
-The model was initially evaluated qualitatively by comparing patterns of tidal amplitude and phase to results from \citep{foreman1995tidal}.
+The model was initially evaluated qualitatively by comparing patterns of tidal amplitude and phase to results from \citet{foreman1995tidal}.
 For example, the amphidrome around Victoria was produced in the M$_2$ results, as well as the monotonic increase in $K_1$ amplitude moving northwards along the Strait of Georgia.
 Once the model was reproducing observed tidal patterns, model results were quantitatively evaluated by comparing modelled harmonic constituents to measured harmonic constituents at tidal measuring stations throughout the domain.
-Comparisons were made using the complex difference ($D$), defined by \citep{foreman1995tidal} as:
+Comparisons were made using the complex difference ($D$), defined as \citep{foreman1995tidal}:
 
@@
 \section{Storm Surge Hindcasts}\label{sec:storm}
-%outline: 
-%1. factors contributing to storm surge with focus on Feb 2006 event
-%2. Strong wind event but no surge (maybe Dec 2006 or Nov 2009)
-%3. High resolution wind (hopefully Dec 2012)
 A series of storm surge hindcasts is presented next.
@@
 Several significant surge or wind events have been considered in detail, as well as an array of hindcasts to asses model ability in a statistical measure.
-These cases have been used to asses the model's skill in reproducing water level height and surge amplitude and timing during storm events.
+These cases have been used to assess the model's skill in reproducing water level height and surge amplitude and timing during storm events.
 In addition, the contribution to the surge due to several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed.
@@
 
-\subsection{Factors Contributing to Storm Surges: Feb 4, 2006}
+\subsection{Example Storm Surge Hindcast: Feb 4, 2006}
 
-On Feb 4, 2006, a strong storm system with winds over 75 km/hr coinciding with an unusually high tide led to significant flooding in coastal communities of the Strait of Georgia \citep{romanowski2010storm}.
+On Feb 4, 2006, a strong storm system with winds over 75 km hr$^{-1}$ coinciding with an unusually high tide led to significant flooding in coastal communities of the Strait of Georgia \citep{romanowski2010storm}.
 The maximum water level observed at Point Atkinson on this day was 5.46 m, which is 0.82 m higher than the predicted tides, highlighting the significance of this event as it resulted in a large anomaly and also reached water levels close to the highest level ever recorded (5.61 m).
@@
 The error at Point Atkinson and Campbell River is slightly more significant, reaching 2.5\% and 2.7\% respectively, close to 14 cm over predicted in each case.
-It is sup rising that Victoria and Patricia Bay agree most closely with observations because the M$_2$ tidal amplitudes in the model are known to be too low in this area.
+It is surprising that Victoria and Patricia Bay agree most closely with observations because the M$_2$ tidal amplitudes in the model are known to be too low in this area.
 In conjunction with the over predicted water levels at Point Atkinson and Campbell River, it is likely that one of the forcing factors is overestimated in this example.
@@
 The most noticeable difference is a 4-6 hour delay in the model's timing of the maximum residual except at Victoria where the maximum residual in sync with observations.
-The errors in maximum residual range from 6.0 cm to high at Victoria to 15.9 cm too low at Campbell River.
+The errors in maximum residual range from 6.0 cm too high at Victoria to 15.9 cm too low at Campbell River.
 Point Atkinson and Patricia Bay both experience a very good match with observed residuals with errors less than 2 cm.
@@
 \end{table} 
-%anything else to include? statistics calculated in Feb 2006 -paper production.ipynb and saved in statistics_feb2006.CV 
+%anything else to include? statistics calculated in Feb 2006 -paper production.ipynb and saved in statisitics_feb2006.CV 
 %I also have standard deviations, correlations, ...
+
+\subsection{Factors Contributing to Storm Surges}
 
@@
 \end{figure}
+%Fix ylabels!!!!
 
@@
 \end{figure}
-
-\subsection{An Additional Hindcast: November 16, 2006}
-An additional storm surge hindcast for November 2006 is provided as another means of assessing model performance and is mentioned briefly here.
-On November 16, 2006 communities in the Strait of Georgia experienced a strong wind storm with speeds upwards of 20 m/s near Point Atkinson resulting in observed water levels of 5.1 m and a maximum surge amplitude of 83.7 cm.
-The surge amplitude was lower at the other stations, reaching no more than 68 cm.
-The model performs reasonably well in this hindcast, Willmott skill scores less greater than 0.94 as outlined in Table \ref{tab:nov2006stat}.
-The best match in maximum surge occured at Campbell River 6.2 cm underestimate by the model.
-The timing of the maximum surge at Campbell River was identical with observations and slightly earlier at the other stations.
-
-\begin{table}[h]
-\centering 
-\begin{tabular}{|l |c c c c c | c c c c c|} 
-\hline 
-& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
-\hline 
-Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
-\hline 
-Point Atkinson &  9.8           &  3.42                & 3.43                 &   0.034    & 0.991  &  7.7           &  0.290                 & 0.247                  &  0.128     & 0.950 \\
-Victoria       &  9.1           &  2.16                & 2.11                 &   0.070    & 0.978  &  6.2           &  0.222                 & 0.247                  &  0.171     & 0.946 \\
-Patricia Bay   & 10.0           &  2.57                & 2.53                 &   0.049    & 0.985  &  6.6           &  0.235                 & 0.203                  &  0.144     & 0.947 \\
-Campbell River & 10.7           &  3.20                & 3.26                 &   0.037    & 0.982  &  6.2           &  0.254                 & 0.236                  &  0.140     & 0.952 \\
-\hline 
-\end{tabular}
-\caption{Evaluation of model performance for the Nov 2006 hindcast.}
-\label{tab:nov2006stat} 
-\end{table}
+%Fix ylabels!!!!!
 
@@
 Notably, the Strait of Georgia experiences the largest surge values, upwards of 60 cm, whereas the Juan de Fuca region experiences a maximum surge around 45 cm.
-Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges (\~ 10 cm) than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast.
+Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges ($\approx$ 10 cm) than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast.
 At about 11:30 UTC, the winds switch to northwesterly and then westerly, further amplifying the surge along the mainland coast.
@@
 Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland.
-Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the Strait of Georgia is also affected by the surge entering the domain from Johnstone Strait at the northern boundary.
+Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the Strait of Georgia is also affected by the surge entering the domain from Johnstone Strait at the northern boundary or the incomoing surge wave is reflected from the mainland side.
+%fix up reflection bit
 
@@
 Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered.
-As recorded in the Environment Canada Historical Climate Data archives, the maximum hourly wind speed at Point Atkinson was 18.6 m/s and the largest gust observed was 25.8 m/s or 93 km/hr.
-In addition, large wind speeds over 15 m/s persisted for over 9 hours.
+As recorded in the Environment Canada Historical Climate Data archives, the maximum hourly wind speed at Point Atkinson was 18.6 m s$^{-1}$ and the largest gust observed was 25.8 m s$^{-1}$ or 93 km hr$^{-1}$.
+In addition, large wind speeds over 15 m s$^{-1}$ persisted for over 9 hours.
 Yet, the surge associated with this event is rather small when compared to the previous two cases as shown in Figure \ref{fig:nov2009} where we have plotted the wind speed at Point Atkinson, sea surface height anomaly forcing at the open boundaries, and observed and modelled residuals at Point Atkinson.
@@
 \end{table}
+%Fix label for CGRF: Atmospheric model
+
+\subsection{An Additional Hindcast: Nov 16, 2006}
+An additional storm surge hindcast for Nov 2006 is provided as another means of assessing model performance and is mentioned briefly here.
+On Nov 16, 2006 communities in the Strait of Georgia experienced a strong wind storm with speeds upwards of 20 m s$^{-1}$ near Point Atkinson resulting in observed water levels of 5.1 m and a maximum surge amplitude of 83.7 cm.
+The surge amplitude was lower at the other stations, reaching no more than 68 cm.
+The model performs reasonably well in this hindcast, Willmott skill scores less greater than 0.94 as outlined in Table \ref{tab:nov2006stat}.
+The best match in maximum surge occured at Campbell River with a 6.2 cm underestimate by the model.
+The timing of the maximum surge at Campbell River was identical with observations and slightly earlier at the other stations.
+
+\begin{table}[h]
+\centering 
+\begin{tabular}{|l |c c c c c | c c c c c|} 
+\hline 
+& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
+\hline 
+Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
+\hline 
+Point Atkinson &  9.8           &  3.42                & 3.43                 &   0.034    & 0.991  &  7.7           &  0.290                 & 0.247                  &  0.128     & 0.950 \\
+Victoria       &  9.1           &  2.16                & 2.11                 &   0.070    & 0.978  &  6.2           &  0.222                 & 0.247                  &  0.171     & 0.946 \\
+Patricia Bay   & 10.0           &  2.57                & 2.53                 &   0.049    & 0.985  &  6.6           &  0.235                 & 0.203                  &  0.144     & 0.947 \\
+Campbell River & 10.7           &  3.20                & 3.26                 &   0.037    & 0.982  &  6.2           &  0.254                 & 0.236                  &  0.140     & 0.952 \\
+\hline 
+\end{tabular}
+\caption{Evaluation of model performance for the Nov 2006 hindcast.}
+\label{tab:nov2006stat} 
+\end{table}
 
 %\subsection{High Resolution Atmospheric Forcing: Dec 17, 2012}
-%not to be included
-
-%I have one other simulation for November 2006. Skill scores are comparable to the Feb 2006 case, so I'm not sure what this case adds to the overall article.
 
@@
 
-Hindcasts of significant events in February 2006, November 2006, December 2006 and November 2009 were simulated.
+Hindcasts of four significant events in February 2006, November 2006, December 2006 and November 2009 were simulated.
 In general, the model performs well at reproducing the observed water level heights and surge amplitudes, with Willmott skill scores greater than 0.88 in all cases.
 We have also assessed the relative importance of different forcing factors contributing to the storm surge.
-The sea surface height anomaly entering the system from the Pacific Ocean contributes most significantly to the surge amplitude (>50 \%), in agreement with previous studies in this region \citep{murty1995storm}.
+The sea surface height anomaly entering the system from the Pacific Ocean contributes most significantly to the surge amplitude ($>50$ \%), in agreement with previous studies in this region \citep{murty1995storm}.
 We have also found that the wind and atmospheric conditions can set up spatial differences in the surge amplitude, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season.

### 2014-11-05 RoughWork/manuscript.tex
Fixed up models paragraph. Still need to add Giddings ref.

@@
 Several features related to the physical oceanography of this region render it a challenge to model numerically.
-The outflow from the Fraser River in the late spring is the primary source fresh water within the Strait of Georgia and leads to large vertical density gradients in the summer months.
+The outflow from the Fraser River in the late spring is the primary source fresh water within the Strait of Georgia and leads to large vertical density gradients in the summer months \citep{thomson1981oceanography}.
 Additionally, strong tidal mixing through the San Juan and Gulf Islands produces horizontal density gradients that separate the saline waters of the Strait of Juan de Fuca and the fresher water of the Strait of Georgia.
-It is important to accurately represent the vertical mixing in this region as it sets the rate of export of fresh water through estuarine circulation.
+It is important to accurately represent the vertical mixing in this region as it sets the rate of export of fresh water through estuarine circulation \citep{masson2004modelling}.
 The strong and fast tidal currents through the narrow channels of the north, such as Discovery Passage and Johnstone Strait, pose a particular challenge for numerical models.
@@
 This model employed a sigma or terrain-following vertical coordinate system leading to spurious diapycnal mixing since, numerically, the lateral mixing is computed on sigma levels which may intersect isopycnals in the presence of steep topography.
-\citet{yang2010multi} developed a model of southern Strait of Georgia and Puget Sound using finite volume techniques on an unstructured grid.
-\citet{sutherland2011model} employed the Regional Ocean Modelling System (ROMS) to study estuarine circulation in the Salish Sea with a model domain that extended outside the Strait of Juan de Fuca into the Pacific Ocean and up to the northern Strait of Georgia.
+A finite volume model on an unstrcutured grid by \citet{yang2010multi} used a truncated domain with an open boundary in the northern Strait of Georgia, where under predictions in the M$_2$ tidal phase over most of their domain were suggested to be caused by an inaccurate representation of the tides at their northern boundary.
+\citet{sutherland2011model} employed the Regional Ocean Modelling System (ROMS) to study estuarine circulation in the Salish Sea with a model domain that extended outside the Strait of Juan de Fuca into the Pacific Ocean and up to the northern Strait of Georgia and noted that their semidiurnal amplitudes were generally too low.
 The passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.
-These studies have led to an improvement in understanding circulation properties in the Salish Sea and have highlighted important features that should be considered in future modelling work.
-%What was interesting about Yang et al?
-%Tides in Sutherland? JdF surface salinity in Giddings
+Accurate reproduction of observed tidal amplitudes and phases is a recurring difficulty in modelling of this region.
+% JdF surface salinity in Giddings
 %Add more about reproduction of tides in other models
-%Last sentance is lame
 
@@
 %storm surge modelling
-%murty, danard, crean, non BC models (xu, flather)
 Coastal communities can prepare for and respond to storm surge hazards using predictions from storm surge models to determine if flooding conditions are probable.
@@
 \end{figure}
-%image may need more work
-%Add Johnstone Strait, Vancouver, Fraser River, Haro Strait, San Juan/Gulf Islands
 
@@
 
-\subsection{Example Storm Surge Hindcast: Feb 4, 2006}
+\subsection{Example Storm Surge Hindcasts}
 
@@
 
+An additional storm surge hindcast for Nov 2006 is provided as another means of assessing model performance and is mentioned briefly here.
+On Nov 16, 2006 communities in the Strait of Georgia experienced a strong wind storm with speeds upwards of 20 m s$^{-1}$ near Point Atkinson resulting in observed water levels of 5.1 m and a maximum surge amplitude of 83.7 cm.
+The surge amplitude was lower at the other stations, reaching no more than 68 cm.
+The model performs reasonably well in this hindcast, Willmott skill scores are greater than 0.94 as outlined in Table \ref{tab:nov2006stat}.
+The best match in maximum surge occured at Campbell River with a 6.2 cm underestimate by the model.
+The timing of the maximum surge at Campbell River was identical with observations and slightly earlier at the other stations.
+
+\begin{table}[h]
+\centering 
+\begin{tabular}{|l |c c c c c | c c c c c|} 
+\hline 
+& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
+\hline 
+Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
+\hline 
+Point Atkinson &  9.8           &  3.42                & 3.43                 &   0.034    & 0.991  &  7.7           &  0.290                 & 0.247                  &  0.128     & 0.950 \\
+Victoria       &  9.1           &  2.16                & 2.11                 &   0.070    & 0.978  &  6.2           &  0.222                 & 0.247                  &  0.171     & 0.946 \\
+Patricia Bay   & 10.0           &  2.57                & 2.53                 &   0.049    & 0.985  &  6.6           &  0.235                 & 0.203                  &  0.144     & 0.947 \\
+Campbell River & 10.7           &  3.20                & 3.26                 &   0.037    & 0.982  &  6.2           &  0.254                 & 0.236                  &  0.140     & 0.952 \\
+\hline 
+\end{tabular}
+\caption{Evaluation of model performance for the Nov 2006 hindcast.}
+\label{tab:nov2006stat} 
+\end{table}
+
 \subsection{Factors Contributing to Storm Surges}
@@
 
-\subsection{Spatial Extent of Surge: Dec 15, 2006}
+\subsection{Spatial Extent of Surge}
 On Dec 15, 2006, a major storm with wind gusts over 120 km/hr caused significant damage to many coastal communities on the Pacific Northwest.
@@
 
-\subsection{Strong Wind Event: Nov 19, 2009}
+\subsection{Strong Wind Event}
 Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered.
@@
 %Fix label for CGRF: Atmospheric model
-
-\subsection{An Additional Hindcast: Nov 16, 2006}
-An additional storm surge hindcast for Nov 2006 is provided as another means of assessing model performance and is mentioned briefly here.
-On Nov 16, 2006 communities in the Strait of Georgia experienced a strong wind storm with speeds upwards of 20 m s$^{-1}$ near Point Atkinson resulting in observed water levels of 5.1 m and a maximum surge amplitude of 83.7 cm.
-The surge amplitude was lower at the other stations, reaching no more than 68 cm.
-The model performs reasonably well in this hindcast, Willmott skill scores less greater than 0.94 as outlined in Table \ref{tab:nov2006stat}.
-The best match in maximum surge occured at Campbell River with a 6.2 cm underestimate by the model.
-The timing of the maximum surge at Campbell River was identical with observations and slightly earlier at the other stations.
-
-\begin{table}[h]
-\centering 
-\begin{tabular}{|l |c c c c c | c c c c c|} 
-\hline 
-& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
-\hline 
-Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
-\hline 
-Point Atkinson &  9.8           &  3.42                & 3.43                 &   0.034    & 0.991  &  7.7           &  0.290                 & 0.247                  &  0.128     & 0.950 \\
-Victoria       &  9.1           &  2.16                & 2.11                 &   0.070    & 0.978  &  6.2           &  0.222                 & 0.247                  &  0.171     & 0.946 \\
-Patricia Bay   & 10.0           &  2.57                & 2.53                 &   0.049    & 0.985  &  6.6           &  0.235                 & 0.203                  &  0.144     & 0.947 \\
-Campbell River & 10.7           &  3.20                & 3.26                 &   0.037    & 0.982  &  6.2           &  0.254                 & 0.236                  &  0.140     & 0.952 \\
-\hline 
-\end{tabular}
-\caption{Evaluation of model performance for the Nov 2006 hindcast.}
-\label{tab:nov2006stat} 
-\end{table}
 

### 2014-11-05 RoughWork/manuscript.tex
Added Giddings citation.

@@
 This model employed a sigma or terrain-following vertical coordinate system leading to spurious diapycnal mixing since, numerically, the lateral mixing is computed on sigma levels which may intersect isopycnals in the presence of steep topography.
+Another group of modellers studying a larger domain noted high temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}.
 A finite volume model on an unstrcutured grid by \citet{yang2010multi} used a truncated domain with an open boundary in the northern Strait of Georgia, where under predictions in the M$_2$ tidal phase over most of their domain were suggested to be caused by an inaccurate representation of the tides at their northern boundary.
@@
 The passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.
-Accurate reproduction of observed tidal amplitudes and phases is a recurring difficulty in modelling of this region.
-% JdF surface salinity in Giddings
+Accurate reproduction of observed tidal amplitudes and phases and appropriate treatment of mixing is a recurring difficulty in modelling of this region.
 %Add more about reproduction of tides in other models

### 2014-11-18 run_files/runs_dec2006.py
Added run files to storm-surge repo

@@ -1,0 +1,58 @@
+"""Example script to define and execute a series of Salish Sea NEMO model runs.
+"""
+from __future__ import absolute_import
+ 
+import os
+ 
+import salishsea_cmd.api
+ 
+ 
+def main():
+    run_desc = base_run_description()
+    runs = ('all_forcing', 'tidesonly', 'ssh_only')
+    tides= ('lateral', 'lateral.tidesonly','lateral')
+    surface=('surface','surface.nosurge','surface.nosurge')
+    for run_id,tide_id,surface_id in zip(runs,tides,surface):
+        do_run(run_id, run_desc, tide_id, surface_id)
+ 
+ 
+def do_run(run_id, run_desc, tide_id, surface_id):
+    run_desc['run_id'] = run_id
+    run_desc['namelists'][3] = (
+        'namelist.{}'.format(tide_id))
+    run_desc['namelists'][2] = (
+        'namelist.{}'.format(surface_id))
+    salishsea_cmd.api.run_in_subprocess(
+        run_id,
+        run_desc,
+        'iodef.xml',
+        os.path.join('/home/nksoonti/MEOPAR/SalishSea/results/storm_surges/final/dec2006/', run_id))
+ 
+ 
+def base_run_description():
+# Relative paths from SS-run-sets/SalishSea/storm_surges/new_config
+    run_desc = salishsea_cmd.api.run_description(
+        walltime='20:00:00',
+        NEMO_code='../../../../NEMO-code/',
+        forcing='../../../../NEMO-forcing/',
+        runs_dir='../../../../SalishSea/',
+        init_conditions=(
+            '/home/nksoonti/MEOPAR/SalishSea/results/spin-up/7dec16dec'),
+        )
+    run_desc['email'] = 'nsoontie@eos.ubc.ca'
+    # Relative paths to namelist section files
+    run_desc['namelists'] = [
+        'namelist.dec2006.time',
+        'namelist.dec2006.domain',
+        'namelist.surface',
+        'namelist.lateral',
+        'namelist.bottom',
+        'namelist.tracers',
+        'namelist.dynamics',
+        'namelist.compute.6x14',
+        ]
+    return run_desc
+ 
+ 
+if __name__ == '__main__':
+    main()

### 2014-11-18 run_files/runs_dec2006_freq.py

@@ -1,0 +1,52 @@
+"""Example script to define and execute a series of Salish Sea NEMO model runs.
+"""
+from __future__ import absolute_import
+ 
+import os
+ 
+import salishsea_cmd.api
+ 
+ 
+def main():
+    run_id = '30min'
+    iodefs= 'iodef_30min.xml'
+    run_desc = base_run_description(run_id)
+    do_run(run_id, run_desc, iodefs)
+       
+ 
+def do_run(run_id, run_desc, iodefid):
+    run_desc['run_id'] = run_id        
+    salishsea_cmd.api.run_in_subprocess(
+        run_id,
+        run_desc,
+        iodefid,
+        os.path.join('/home/nksoonti/MEOPAR/SalishSea/results/storm_surges/final/dec2006/restart',run_id))
+ 
+ 
+def base_run_description(run_id):
+# Relative paths from SS-run-sets/SalishSea/storm_surges/new_config
+    run_desc = salishsea_cmd.api.run_description(
+        walltime='7:00:00',
+        NEMO_code='../../../../NEMO-code/',
+        forcing='../../../../NEMO-forcing/',
+        runs_dir='../../../../SalishSea/',
+        init_conditions=(
+            '/home/nksoonti/MEOPAR/SalishSea/results/storm_surges/final/dec2006/all_forcing/SalishSea_00025920_restart.nc')
+        )
+    run_desc['email'] = 'nsoontie@eos.ubc.ca'
+    # Relative paths to namelist section files
+    run_desc['namelists'] = [
+        'namelist.dec2006.rtime',
+        'namelist.dec2006.domain',
+        'namelist.surface',
+        'namelist.lateral',
+        'namelist.bottom',
+        'namelist.tracers',
+        'namelist.dynamics',
+        'namelist.compute.6x14',
+        ]
+    return run_desc
+ 
+ 
+if __name__ == '__main__':
+    main()

### 2014-11-18 run_files/runs_dec2006_restart.py

@@ -1,0 +1,58 @@
+"""Example script to define and execute a series of Salish Sea NEMO model runs.
+"""
+from __future__ import absolute_import
+ 
+import os
+ 
+import salishsea_cmd.api
+ 
+ 
+def main():
+    runs = ('all_forcing', 'tidesonly','ssh_only')
+    tides= ('lateral', 'lateral.tidesonly','lateral')
+    surface=('surface','surface.nosurge','surface.nosurge')
+    for run_id,tide_id,surface_id in zip(runs,tides,surface):
+        run_desc = base_run_description(run_id)
+        do_run(run_id, run_desc, tide_id, surface_id)
+       
+ 
+def do_run(run_id, run_desc, tide_id, surface_id):
+    run_desc['run_id'] = run_id
+    run_desc['namelists'][3] = (
+        'namelist.{}'.format(tide_id))
+    run_desc['namelists'][2] = (
+        'namelist.{}'.format(surface_id))        
+    salishsea_cmd.api.run_in_subprocess(
+        run_id,
+        run_desc,
+        'iodef_1h.xml',
+        os.path.join('/home/nksoonti/MEOPAR/SalishSea/results/storm_surges/final/dec2006/restart',run_id))
+ 
+ 
+def base_run_description(run_id):
+# Relative paths from SS-run-sets/SalishSea/storm_surges/new_config
+    run_desc = salishsea_cmd.api.run_description(
+        walltime='7:00:00',
+        NEMO_code='../../../../NEMO-code/',
+        forcing='../../../../NEMO-forcing/',
+        runs_dir='../../../../SalishSea/',
+        init_conditions=(
+            os.path.join('/home/nksoonti/MEOPAR/SalishSea/results/storm_surges/final/dec2006/',run_id,'SalishSea_00025920_restart.nc'))
+        )
+    run_desc['email'] = 'nsoontie@eos.ubc.ca'
+    # Relative paths to namelist section files
+    run_desc['namelists'] = [
+        'namelist.dec2006.rtime',
+        'namelist.dec2006.domain',
+        'namelist.surface',
+        'namelist.lateral',
+        'namelist.bottom',
+        'namelist.tracers',
+        'namelist.dynamics',
+        'namelist.compute.6x14',
+        ]
+    return run_desc
+ 
+ 
+if __name__ == '__main__':
+    main()

### 2014-11-18 run_files/runs_dec2012.py

@@ -1,0 +1,58 @@
+"""Example script to define and execute a series of Salish Sea NEMO model runs.
+"""
+from __future__ import absolute_import
+ 
+import os
+ 
+import salishsea_cmd.api
+ 
+ 
+def main():
+    run_desc = base_run_description()
+    runs = ('all_forcing','weather_only','no_pressure',)
+    tides= ('lateral','lateral.tidesonly','lateral.tidesonly',)
+    surface=('surface.ops','surface.ops','surface.nopressure.ops',)
+    for run_id,tide_id,surface_id in zip(runs,tides,surface):
+        do_run(run_id, run_desc, tide_id, surface_id)
+ 
+ 
+def do_run(run_id, run_desc, tide_id, surface_id):
+    run_desc['run_id'] = run_id
+    run_desc['namelists'][3] = (
+        'namelist.{}'.format(tide_id))
+    run_desc['namelists'][2] = (
+        'namelist.{}'.format(surface_id))
+    salishsea_cmd.api.run_in_subprocess(
+        run_id,
+        run_desc,
+        'iodef.xml',
+        os.path.join('/home/nksoonti/MEOPAR/SalishSea/results/storm_surges/final/dec2012/ops/', run_id))
+ 
+ 
+def base_run_description():
+# Relative paths from SS-run-sets/SalishSea/storm_surges/new_config
+    run_desc = salishsea_cmd.api.run_description(
+        walltime='15:00:00',
+        NEMO_code='../../../../NEMO-code/',
+        forcing='../../../../NEMO-forcing/',
+        runs_dir='../../../../SalishSea/',
+        init_conditions=(
+            '/home/nksoonti/MEOPAR/SalishSea/results/spin-up/7dec16dec'),
+        )
+    run_desc['email'] = 'nsoontie@eos.ubc.ca'
+    # Relative paths to namelist section files
+    run_desc['namelists'] = [
+        'namelist.dec2012.time',
+        'namelist.dec2012.domain',
+        'namelist.surface',
+        'namelist.lateral',
+        'namelist.bottom',
+        'namelist.tracers',
+        'namelist.dynamics',
+        'namelist.compute.6x14',
+        ]
+    return run_desc
+ 
+ 
+if __name__ == '__main__':
+    main()

### 2014-11-18 run_files/runs_dec2012_CGRF.py

@@ -1,0 +1,58 @@
+"""Example script to define and execute a series of Salish Sea NEMO model runs.
+"""
+from __future__ import absolute_import
+ 
+import os
+ 
+import salishsea_cmd.api
+ 
+ 
+def main():
+    run_desc = base_run_description()
+    runs = ('all_forcing','tidesonly','weather_only','no_pressure',)
+    tides= ('lateral','lateral.tidesonly','lateral.tidesonly','lateral.tidesonly',)
+    surface=('surface','surface.nosurge','surface','surface.nopressure',)
+    for run_id,tide_id,surface_id in zip(runs,tides,surface):
+        do_run(run_id, run_desc, tide_id, surface_id)
+ 
+ 
+def do_run(run_id, run_desc, tide_id, surface_id):
+    run_desc['run_id'] = run_id
+    run_desc['namelists'][3] = (
+        'namelist.{}'.format(tide_id))
+    run_desc['namelists'][2] = (
+        'namelist.{}'.format(surface_id))
+    salishsea_cmd.api.run_in_subprocess(
+        run_id,
+        run_desc,
+        'iodef.xml',
+        os.path.join('/home/nksoonti/MEOPAR/SalishSea/results/storm_surges/final/dec2012/CGRF/', run_id))
+ 
+ 
+def base_run_description():
+# Relative paths from SS-run-sets/SalishSea/storm_surges/new_config
+    run_desc = salishsea_cmd.api.run_description(
+        walltime='15:00:00',
+        NEMO_code='../../../../NEMO-code/',
+        forcing='../../../../NEMO-forcing/',
+        runs_dir='../../../../SalishSea/',
+        init_conditions=(
+            '/home/nksoonti/MEOPAR/SalishSea/results/spin-up/7dec16dec'),
+        )
+    run_desc['email'] = 'nsoontie@eos.ubc.ca'
+    # Relative paths to namelist section files
+    run_desc['namelists'] = [
+        'namelist.dec2012.time',
+        'namelist.dec2012.domain',
+        'namelist.surface',
+        'namelist.lateral',
+        'namelist.bottom',
+        'namelist.tracers',
+        'namelist.dynamics',
+        'namelist.compute.6x14',
+        ]
+    return run_desc
+ 
+ 
+if __name__ == '__main__':
+    main()

### 2014-11-18 run_files/runs_feb2006.py

@@ -1,0 +1,58 @@
+"""Example script to define and execute a series of Salish Sea NEMO model runs.
+"""
+from __future__ import absolute_import
+ 
+import os
+ 
+import salishsea_cmd.api
+ 
+ 
+def main():
+    run_desc = base_run_description()
+    runs = ('all_forcing', 'tidesonly', 'weather_only', 'ssh_only')
+    tides= ('lateral', 'lateral.tidesonly', 'lateral.tidesonly','lateral')
+    surface=('surface','surface.nosurge','surface','surface.nosurge')
+    for run_id,tide_id,surface_id in zip(runs,tides,surface):
+        do_run(run_id, run_desc, tide_id, surface_id)
+ 
+ 
+def do_run(run_id, run_desc, tide_id, surface_id):
+    run_desc['run_id'] = run_id
+    run_desc['namelists'][3] = (
+        'namelist.{}'.format(tide_id))
+    run_desc['namelists'][2] = (
+        'namelist.{}'.format(surface_id))
+    salishsea_cmd.api.run_in_subprocess(
+        run_id,
+        run_desc,
+        'iodef.xml',
+        os.path.join('/home/nksoonti/MEOPAR/SalishSea/results/storm_surges/final/feb2006/', run_id))
+ 
+ 
+def base_run_description():
+# Relative paths from SS-run-sets/SalishSea/storm_surges/new_config
+    run_desc = salishsea_cmd.api.run_description(
+        walltime='20:00:00',
+        NEMO_code='../../../../NEMO-code/',
+        forcing='../../../../NEMO-forcing/',
+        runs_dir='../../../../SalishSea/',
+        init_conditions=(
+            '/home/nksoonti/MEOPAR/SalishSea/results/spin-up/31jan9feb'),
+        )
+    run_desc['email'] = 'nsoontie@eos.ubc.ca'
+    # Relative paths to namelist section files
+    run_desc['namelists'] = [
+        'namelist.feb2006.time',
+        'namelist.feb2006.domain',
+        'namelist.surface',
+        'namelist.lateral',
+        'namelist.bottom',
+        'namelist.tracers',
+        'namelist.dynamics',
+        'namelist.compute.6x14',
+        ]
+    return run_desc
+ 
+ 
+if __name__ == '__main__':
+    main()

### 2014-11-18 run_files/runs_feb2006_supp.py

@@ -1,0 +1,58 @@
+"""Example script to define and execute a series of Salish Sea NEMO model runs.
+"""
+from __future__ import absolute_import
+ 
+import os
+ 
+import salishsea_cmd.api
+ 
+ 
+def main():
+    run_desc = base_run_description()
+    runs = ('no_pressure', 'surge_only','harmonics8')
+    tides= ('lateral.tidesonly', 'lateral.surgeonly','lateral.harmonics8')
+    surface=('surface.nopressure','surface','surface')
+    for run_id,tide_id,surface_id in zip(runs,tides,surface):
+        do_run(run_id, run_desc, tide_id, surface_id)
+ 
+ 
+def do_run(run_id, run_desc, tide_id, surface_id):
+    run_desc['run_id'] = run_id
+    run_desc['namelists'][3] = (
+        'namelist.{}'.format(tide_id))
+    run_desc['namelists'][2] = (
+        'namelist.{}'.format(surface_id))
+    salishsea_cmd.api.run_in_subprocess(
+        run_id,
+        run_desc,
+        'iodef.xml',
+        os.path.join('/home/nksoonti/MEOPAR/SalishSea/results/storm_surges/final/feb2006/', run_id))
+ 
+ 
+def base_run_description():
+# Relative paths from SS-run-sets/SalishSea/storm_surges/new_config
+    run_desc = salishsea_cmd.api.run_description(
+        walltime='12:00:00',
+        NEMO_code='../../../../NEMO-code/',
+        forcing='../../../../NEMO-forcing/',
+        runs_dir='../../../../SalishSea/',
+        init_conditions=(
+            '/home/nksoonti/MEOPAR/SalishSea/results/spin-up/31jan9feb'),
+        )
+    run_desc['email'] = 'nsoontie@eos.ubc.ca'
+    # Relative paths to namelist section files
+    run_desc['namelists'] = [
+        'namelist.feb2006.time',
+        'namelist.feb2006.domain',
+        'namelist.surface',
+        'namelist.lateral',
+        'namelist.bottom',
+        'namelist.tracers',
+        'namelist.dynamics',
+        'namelist.compute.6x14',
+        ]
+    return run_desc
+ 
+ 
+if __name__ == '__main__':
+    main()

### 2014-11-18 run_files/runs_nov2006.py

@@ -1,0 +1,58 @@
+"""Example script to define and execute a series of Salish Sea NEMO model runs.
+"""
+from __future__ import absolute_import
+ 
+import os
+ 
+import salishsea_cmd.api
+ 
+ 
+def main():
+    run_desc = base_run_description()
+    runs = ('all_forcing', 'tidesonly')
+    tides= ('lateral', 'lateral.tidesonly')
+    surface=('surface','surface.nosurge')
+    for run_id,tide_id,surface_id in zip(runs,tides,surface):
+        do_run(run_id, run_desc, tide_id, surface_id)
+ 
+ 
+def do_run(run_id, run_desc, tide_id, surface_id):
+    run_desc['run_id'] = run_id
+    run_desc['namelists'][3] = (
+        'namelist.{}'.format(tide_id))
+    run_desc['namelists'][2] = (
+        'namelist.{}'.format(surface_id))
+    salishsea_cmd.api.run_in_subprocess(
+        run_id,
+        run_desc,
+        'iodef.xml',
+        os.path.join('/home/nksoonti/MEOPAR/SalishSea/results/storm_surges/final/nov2006/', run_id))
+ 
+ 
+def base_run_description():
+# Relative paths from SS-run-sets/SalishSea/storm_surges/new_config
+    run_desc = salishsea_cmd.api.run_description(
+        walltime='12:00:00',
+        NEMO_code='../../../../NEMO-code/',
+        forcing='../../../../NEMO-forcing/',
+        runs_dir='../../../../SalishSea/',
+        init_conditions=(
+            '/home/dlatorne/MEOPAR/SalishSea/spin-up/7nov16nov'),
+        )
+    run_desc['email'] = 'nsoontie@eos.ubc.ca'
+    # Relative paths to namelist section files
+    run_desc['namelists'] = [
+        'namelist.nov2006.time',
+        'namelist.nov2006.domain',
+        'namelist.surface',
+        'namelist.lateral',
+        'namelist.bottom',
+        'namelist.tracers',
+        'namelist.dynamics',
+        'namelist.compute.6x14',
+        ]
+    return run_desc
+ 
+ 
+if __name__ == '__main__':
+    main()

### 2014-11-18 run_files/runs_nov2009.py

@@ -1,0 +1,58 @@
+"""Example script to define and execute a series of Salish Sea NEMO model runs.
+"""
+from __future__ import absolute_import
+ 
+import os
+ 
+import salishsea_cmd.api
+ 
+ 
+def main():
+    run_desc = base_run_description()
+    runs = ('all_forcing', 'tidesonly')
+    tides= ('lateral', 'lateral.tidesonly')
+    surface=('surface','surface.nosurge')
+    for run_id,tide_id,surface_id in zip(runs,tides,surface):
+        do_run(run_id, run_desc, tide_id, surface_id)
+ 
+ 
+def do_run(run_id, run_desc, tide_id, surface_id):
+    run_desc['run_id'] = run_id
+    run_desc['namelists'][3] = (
+        'namelist.{}'.format(tide_id))
+    run_desc['namelists'][2] = (
+        'namelist.{}'.format(surface_id))
+    salishsea_cmd.api.run_in_subprocess(
+        run_id,
+        run_desc,
+        'iodef.xml',
+        os.path.join('/home/nksoonti/MEOPAR/SalishSea/results/storm_surges/final/nov2009/', run_id))
+ 
+ 
+def base_run_description():
+# Relative paths from SS-run-sets/SalishSea/storm_surges/new_config
+    run_desc = salishsea_cmd.api.run_description(
+        walltime='12:00:00',
+        NEMO_code='../../../../NEMO-code/',
+        forcing='../../../../NEMO-forcing/',
+        runs_dir='../../../../SalishSea/',
+        init_conditions=(
+            '/home/dlatorne/MEOPAR/SalishSea/spin-up/7nov16nov'),
+        )
+    run_desc['email'] = 'nsoontie@eos.ubc.ca'
+    # Relative paths to namelist section files
+    run_desc['namelists'] = [
+        'namelist.nov2009.time',
+        'namelist.nov2009.domain',
+        'namelist.surface',
+        'namelist.lateral',
+        'namelist.bottom',
+        'namelist.tracers',
+        'namelist.dynamics',
+        'namelist.compute.6x14',
+        ]
+    return run_desc
+ 
+ 
+if __name__ == '__main__':
+    main()

### 2014-11-20 RoughWork/manuscript.tex
Added dec2012 figures to text
Added observed residual to plot

@@
 
-%\subsection{High Resolution Atmospheric Forcing: Dec 17, 2012}
+\subsection{High Resolution Atmospheric Forcing}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/dec2012_surge.pdf}
+\caption{Comparison modelled residuals using a low resolution and high resolution atmospheric product during the Dec 2012 storm surge simulation. }
+\label{fig:dec2012_surge}
+\end{figure}
+%Add in observations?
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/dec2012_weather.pdf}
+\caption{Comparison observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson and Sandheads.}
+\label{fig:dec2013_weather}
+\end{figure}
 
@@
 For instance, the complicated coastlines and local wind patterns result in important spatial variations in the surge amplitude.
-As such, strong winds caused by a large storm is not the only factor to be considered in community preparedness for surge conditions.
+As such, the strong wind caused by a large storm are not the only factors to be considered in community preparedness for surge conditions.
 The geometry of the Strait and local wind patterns render communities on the mainland side of the southern Strait of Georgia more susceptible to surges.
@@
 %future directions: running real time, biology and chemistry, model improvements?
-%comments on wetting and drying (maybe most important in shallow regions)
 %summary of skill scores

### 2014-11-20 RoughWork/manuscript.tex
Adjusted new figure captions

@@
 \end{table}
-%Fix label for CGRF: Atmospheric model
 
@@
 \includegraphics[scale=0.6]{Figures/dec2012_surge.pdf}
-\caption{Comparison modelled residuals using a low resolution and high resolution atmospheric product during the Dec 2012 storm surge simulation. }
+\caption{Comparison observed (solid black) and modelled residuals for the Dec 2012 storm surge simulation.
+Modelled residuals from two simulations using a low resolution (dashed) and high resolution (gray) atmospheric product are shown.
+}
 \label{fig:dec2012_surge}
 \end{figure}
-%Add in observations?
 
@@
 \includegraphics[scale=0.6]{Figures/dec2012_weather.pdf}
-\caption{Comparison observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson and Sandheads.}
+\caption{Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson and Sandheads.}
 \label{fig:dec2013_weather}

### 2014-11-21 RoughWork/manuscript.tex
Started text for Dec 2012

@@
 \subsection{High Resolution Atmospheric Forcing}
+The above comparison between the modelled and observed winds at Point Atkinson suggests that the atmopsheric forcing product (CGRF) may not be adequate to produce accurate storm surge hindcasts since the wind speeds are substanitially lower than observed.
+CGRF is a global atmospheric reforecasting model with horizontal resolution approximately 33 km over the Salish Sea domain, which is problematic since the Strait of Georgia itself is only about 30 km wide.
+As such, there are only one or two grid cells that are entirely over water in our domain, resulting in overly damped wind speeds at grid points close to land such as Point Atkinson.
+To determine the effect of underresolved wind forcing on the storm surge hindcasts we compare simulations that have been forced with 1) the low resolution CGRF model and 2) a high resolution operational model from Environment Canada.
+
+%Where can I find a reference for the atmospheric model?
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/dec2012_weather.pdf}
+\caption{Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson and Sandheads.}
+\label{fig:dec2013_weather}
+\end{figure}
+%don't include direction?
 
@@
 \label{fig:dec2012_surge}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.6]{Figures/dec2012_weather.pdf}
-\caption{Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson and Sandheads.}
-\label{fig:dec2013_weather}
 \end{figure}

### 2014-11-22 RoughWork/manuscript.tex
Text for Dec 2012 case. Modifcations to abstract and discussion

@@
 Important forcing factors contributing to storm surges have been assessed, finding that the surge entering the domain from the Pacific Ocean contributes most significantly.
+Comparisons between simulations with high resolution and low resolution atmospheric forcing further emphasize this point.
 In addition, local wind patterns cause a slight increase in surge amplitude on the mainland side of the Strait of Georgia when compared to the Vancouver Island coastal areas during a major wind storm on Dec 15, 2006.
-Generalll, surge amplitudes are found to be greater within the Strait of Georgia when compared to amplitudes in the Strait of Juan de Fuca.
-%needs a little more work.
+Generally, surge amplitudes are found to be greater within the Strait of Georgia when compared to amplitudes in the Strait of Juan de Fuca.
 \end{abstract}
@@
 Accurate reproduction of observed tidal amplitudes and phases and appropriate treatment of mixing is a recurring difficulty in modelling of this region.
-%Add more about reproduction of tides in other models
 
@@
 During the high wind event on Nov 19 and Nov 20, the winds at Point Atkinson were directed to the west, explaining the lack of strong surge observed during this event since wind stresses would act to depress the water on the east side of the basin.
-In fact, the surge amplitude at Victoria on Nov 20 as 64.7 cm which is slightly higher than that observed at Point Atkinson (62.5 cm).
-Of the cases considered, this is the only example where the surge amplitude at Victoria was higher than that at Point Atkinson.
+In fact, the surge amplitude at Victoria on Nov 20 was 64.7 cm which is slightly higher than that observed at Point Atkinson (62.5 cm).
+Of the cases considered so far, this is the only example where the surge amplitude at Victoria was higher than that at Point Atkinson.
 In fact, this event observed a higher surge at Victoria than the Dec 2006 event, further highlighting the role that winds plays in setting up local variations in the surge.
@@
 As such, there are only one or two grid cells that are entirely over water in our domain, resulting in overly damped wind speeds at grid points close to land such as Point Atkinson.
-To determine the effect of underresolved wind forcing on the storm surge hindcasts we compare simulations that have been forced with 1) the low resolution CGRF model and 2) a high resolution operational model from Environment Canada.
+To determine the effect of underresolved wind forcing on the storm surge hindcasts we compare simulations that have been forced with 1) the low resolution CGRF model and 2) a high resolution (2.5 km) operational model from Environment Canada.
+These simulations are performed for a storm surge that occured on Dec 17, 2012 and cover the time period Dec 14-19, 2012.
+The high resolution product is not available for the hindcasts already discussed.
 
-%Where can I find a reference for the atmospheric model?
+First, a comparison of the wind speed and direction for the two models and observations is provided in Figure \ref{fig:dec2012_weather} at Point Atkinson and at Sandheads, a lighthouse station in the southern Strait of Georgia.
+For each of the models, the wind speed and direction at the closest grid point to the longitude/latitude of Point Atkinson and Sandheads is used for comparison.
+Although the low resolution CGRF model has a grid spacing of 33 km, two grid points align closely with these locations.
+At Point Atkinson, the high resolution model is in very good agreement with observations for both the wind speed and direction, however, the low resolution CGRF model significantly underestimates the wind speed during high wind periods on Dec 15, 2012 and Dec 17, 2012.
+At Sandheads, both the low resolution and high resolution models have good agreement with observations.
+The low resolution model, surprsingly, has a very good match likely because this grid cell happens to be mostly over water.
+Although not shown here, both models have a similar sea level atmospheric pressure at these two locations.
+
+%Where can I find a reference for the high res atmospheric model?
 
@@
 \includegraphics[scale=0.6]{Figures/dec2012_weather.pdf}
-\caption{Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson and Sandheads.}
-\label{fig:dec2013_weather}
+\caption{Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson (upper) and Sandheads (lower).
+Observed winds shown in solid black, low resolution CGRF model in dashed black and high resolution model in gray.
+}
+\label{fig:dec2012_weather}
 \end{figure}
 %don't include direction?
+
+Modelled and observed residuals for this simulation are displayed in Figure \ref{fig:dec2012_surge} where it is apparent that both of the models agree well with the observations.
+Willmott skill scores for both modelled residuals are above 0.92 at these four locations.
+Further, the modelled residuals deviate only slightly from each other, most noticeable at Point Atkinson during the large wind event on Dec 15, 2012 where simulation with the low resolution atmopsheric forcing has a residual that is 5 cm higher.
+Although the wind speeds on Dec 15, 2012 are quite high, this event did not result in any flooding or residuals greater than 50 cm.
+Significant flooding occured later, on Dec 17, 2012, where the observed surge amplitude at Point Atkinson was 68.3 cm.
+Even though the low resolution model significantly underpredicts the wind speed at some locations, it performs comparable to forcing with high resolution prdocuts.
+This result further demonstrates that the local wind is not the most important factor for setting up storm surges over this domain.
 
@@
 We have also assessed the relative importance of different forcing factors contributing to the storm surge.
-The sea surface height anomaly entering the system from the Pacific Ocean contributes most significantly to the surge amplitude ($>50$ \%), in agreement with previous studies in this region \citep{murty1995storm}.
+The sea surface height anomaly entering the system from the Pacific Ocean contributes most significantly to the surge amplitude ($>50$\%), in agreement with previous studies in this region \citep{murty1995storm}.
 We have also found that the wind and atmospheric conditions can set up spatial differences in the surge amplitude, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season.
-Additionally, surge amplitudes in the Strait of Georgia are usually higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions.
+Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions.
 Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important after the San Juan and Gulf Islands.
-The modelled surge amplitude increased by up to 10 \% at Point Atkinson when the tide-surge interaction was neglected.
+The modelled surge amplitude increased by about 10\% at Point Atkinson when the tide-surge interaction was neglected.
+
+Comparisons between high resolution and low resolution atmospheric forcing further suggest that the local winds are not the leading order factor in setting up large surge amplitudes.
+The low resolution model significantly underestimated wind speeds in certain areas of the domain, however, it performed just as accurately as the high resolution product.
+Although the high resolution model delivers fine scale detail and more accurate local wind patterns, it did not provide any improvement to the surge hindcast.
+Since both of these products provide the same level of accuracy in storm surge hindcasts, in the future we will use the high resolution atmospheric model to produce real-time storm surge predictions, even though most of out testing was performed with the low resolution model.
 
@@
 In this article, the four locations used to assess storm surge accuracy are not in shallow regions and so wetting and drying is not considered to be important.
-However, several communities in the Strait of Georgia, such as Boundary Bay and Crescent Beach are in shallow regions and are typically most affected by storm surge during the winter season.
+However, several communities in the Strait of Georgia, such as Boundary Bay and Crescent Beach, are in shallow regions and are typically most affected by storm surge during the winter season.
 As such, efforts to include wetting and drying will be considered in the future.
-Additionally, sea surface height anomaly from Neah Bay will be used in future real time water level predictions during the storm surge season.
+Additionally, the sea surface height anomaly from Neah Bay will be used in future real time water level predictions during the storm surge season.
 Lastly, flooding effects due to wave run up is left for a more detailed inundation study with higher resolution in the near-shore areas.
@@
 For instance, the complicated coastlines and local wind patterns result in important spatial variations in the surge amplitude.
-As such, the strong wind caused by a large storm are not the only factors to be considered in community preparedness for surge conditions.
+As such, the strong wind caused by a large storm are is not the only factor to be considered in community preparedness for surge conditions.
 The geometry of the Strait and local wind patterns render communities on the mainland side of the southern Strait of Georgia more susceptible to surges.
@@
 \section{Acknowledgements}\label{sec:ack}
-We would like to thank Michael Foreman, Diane Masson, Luc Fillion Rich Pawlowicz, and Mark Halverson for providing data used in model set up and evaluation.
+We would like to thank Michael Foreman, Diane Masson, Luc Fillion, Rich Pawlowicz, and Mark Halverson for providing data used in model set up and evaluation.
 This project is funded by the Marine Environmental Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada.

### 2014-11-25 RoughWork/manuscript.tex
Comments on local winds/eastern Canada

@@
 As such, storm surges in the Strait of Georgia are typically observed between November and February \citep{forseth2006adaptation}.
+Flooding caused by a storm surge is most likely to occur during a high spring tide \citep{abeys2011extreme}.
 
@@
 Storm surge modelling in the Strait of Georgia dates back to \citet{crean1988mathematical} in a barotropic finite difference model.
-Their work is highlighted in reviews of storm surges in Canadian waters by \citet{murty1995storm} and \citet{danard2003storm}.
-Compared to eastern Canada \citep{gray1984preliminary, bernier2006predicting,bobanovic2006forecasting}, relatively little has been published on storm surge modelling on the west coast of Canada.
+Their work is highlighted in reviews of storm surges in Canadian waters \citep{murty1995storm, danard2003storm}.
+Compared to eastern Canada \citep{gray1984preliminary, bernier2006predicting,bobanovic2006forecasting, bernier2007tide, bernier2010tide}, relatively little has been published on storm surge modelling on the west coast of Canada in recent years.
+A statisical overview of past surge events in British Columbia waters suggests an increase in extreme water level events during warm ENSO periods \citep{abeys2011extreme}.
 
@@
 
-Hindcasts of four significant events in February 2006, November 2006, December 2006 and November 2009 were simulated.
+Hindcasts of five significant events in February 2006, November 2006, December, November 2009, and December 2012 were simulated.
 In general, the model performs well at reproducing the observed water level heights and surge amplitudes, with Willmott skill scores greater than 0.88 in all cases.
 We have also assessed the relative importance of different forcing factors contributing to the storm surge.
-The sea surface height anomaly entering the system from the Pacific Ocean contributes most significantly to the surge amplitude ($>50$\%), in agreement with previous studies in this region \citep{murty1995storm}.
-We have also found that the wind and atmospheric conditions can set up spatial differences in the surge amplitude, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season.
+The sea surface height anomaly entering the system from the Pacific Ocean contributes most significantly to the surge amplitude, in agreement with previous studies in this region \citep{murty1995storm}.
+Although the large scale wind patterns and atmopsheric conditions important in setting up the surge on the Pacific Ocean boundaries, the local winds within the Strait of Georgia are not the leading order contribution to surge amplitude.
+However, the local wind patterns can set up small spatial differences in the surge amplitude, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season.
 Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions.
@@
 Comparisons between high resolution and low resolution atmospheric forcing further suggest that the local winds are not the leading order factor in setting up large surge amplitudes.
-The low resolution model significantly underestimated wind speeds in certain areas of the domain, however, it performed just as accurately as the high resolution product.
+The low resolution model significantly underestimated wind speeds in certain areas of the domain, however, it performed just as accurately in reproducing surge amplitudes as the high resolution product.
 Although the high resolution model delivers fine scale detail and more accurate local wind patterns, it did not provide any improvement to the surge hindcast.

### 2014-11-26 RoughWork/manuscript.tex
Small edits

@@
 Many rivers feed fresh water into the Salish Sea system, the most significant source being from the Fraser River.
-Mixed diurnal and semi-diurnal tides are found throughout most of the Salish Sea but a semi-diurnal amphidrome is present nearby Victoria, where the tides are mainly diurnal.
+Mixed diurnal and semi-diurnal tides are found throughout most of the Salish Sea but a semi-diurnal amphidrome is present nearby Victoria, where the tides are mainly diurnal \citep{thomson1981oceanography}.
 This system is home to several large coastal communities, such as Vancouver and Victoria, serves as a waterway for commercial traffic and recreational use, and supports a large and active fishing industry.
@@
 Several features related to the physical oceanography of this region render it a challenge to model numerically.
-The outflow from the Fraser River in the late spring is the primary source fresh water within the Strait of Georgia and leads to large vertical density gradients in the summer months \citep{thomson1981oceanography}.
+The outflow from the Fraser River in the late spring is the primary source of fresh water within the Strait of Georgia and leads to large vertical density gradients in the summer months \citep{thomson1981oceanography}.
 Additionally, strong tidal mixing through the San Juan and Gulf Islands produces horizontal density gradients that separate the saline waters of the Strait of Juan de Fuca and the fresher water of the Strait of Georgia.
 It is important to accurately represent the vertical mixing in this region as it sets the rate of export of fresh water through estuarine circulation \citep{masson2004modelling}.
-The strong and fast tidal currents through the narrow channels of the north, such as Discovery Passage and Johnstone Strait, pose a particular challenge for numerical models.
+The strong and fast tidal currents through the narrow channels of the north, such as Discovery Passage and Johnstone Strait, set a fairly restrictive time step and impose a particular challenge for numerical models.
 Yet, several modelling efforts for this region exist, ranging in choices of grid structure, mixing parametrizations, and domain size and extent.
@@
 %overview oF SoG/Salish Sea models
-Early work by \citet{crean1988mathematical,crean1988numerical} and \citet{stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia and Puget Sound using a finite difference discretization of the equations of motion.
+Early work by \citet{crean1988mathematical,crean1988numerical} and \citet{stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia, and Puget Sound using a finite difference discretization of the equations of motion.
 A barotropic version of their model was used to reproduce storm surge water level elevations with reasonable accuracy, however, this barotropic version did not include the effects of fresh water entering the system from rivers.
@@
 This early work pointed to an important connection between the model's bottom friction parametrization and the location and amplitude of the M$_2$ amphidrome.
-They found that a local increase in the bottom friction coefficient, especially in the area of Haro Strait, was necessary in order to adequately reproduce the observed tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}.
+They found that a local increase in the bottom friction coefficient, especially in the area of Haro Strait, was necessary in order to adequately reproduce the observed M$_2$ tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}.
+Without local adjustments to the bottom friction.
+thei M$_2$ amplitudes were too high in the Strait of Juan de Fuca and too low in the Strait of Georgia.
 Other studies over a similar domain, such as the barotropic finite element model of \citet{foreman1995tidal}, have avoided local increases in bottom friction.
-\citet{foreman2004m} used data assimilation techniques to study M$_2$ tidal dissipation without local changes to the bottom friction in a two-dimensional finite element model.
-%were Crean/Stronach modelled phases/amps too high or low?
+\citet{foreman2004m} used data assimilation techniques to study M$_2$ tidal dissipation without local changes to the bottom friction in a two-dimensional finite element model that did not include baroclinic effects.
 
@@
 This model employed a sigma or terrain-following vertical coordinate system leading to spurious diapycnal mixing since, numerically, the lateral mixing is computed on sigma levels which may intersect isopycnals in the presence of steep topography.
-Another group of modellers studying a larger domain noted high temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}.
 A finite volume model on an unstrcutured grid by \citet{yang2010multi} used a truncated domain with an open boundary in the northern Strait of Georgia, where under predictions in the M$_2$ tidal phase over most of their domain were suggested to be caused by an inaccurate representation of the tides at their northern boundary.
 \citet{sutherland2011model} employed the Regional Ocean Modelling System (ROMS) to study estuarine circulation in the Salish Sea with a model domain that extended outside the Strait of Juan de Fuca into the Pacific Ocean and up to the northern Strait of Georgia and noted that their semidiurnal amplitudes were generally too low.
-The passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.
+An updated version of this model produced high temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}.
+Smaller domains have also been considered, notably, the passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.
 Accurate reproduction of observed tidal amplitudes and phases and appropriate treatment of mixing is a recurring difficulty in modelling of this region.
@@
 This effect in combination with strong winds pushing water up against the coast can cause flooding, particularly if the storm occurs during an unusually high tide.
-Additionally, there is a small contribution to increased sea level due to the thermal expansion of water during warmer years set by the El Nino Southern Oscillation (ENSO) \citep{abeys2011extreme}.
+There can also be a small contribution to increased sea level due to the thermal expansion of water during warmer years set by the El Nino Southern Oscillation (ENSO) \citep{abeys2011extreme}.
 %discuss this as a Kelvin wave
@@
 This baroclinic model includes the effect of stratification and rivers, however, wetting and drying is not represented.
-Model evaluation has focused on regions where the near-coast slopes are steep.
-This paper is organized as follows: an overview of the model configuration and domain including a description of the ocean model we have employed is provided in section \ref{sec:config}.
+Model evaluation has focused on regions where the near-coast slopes are steep and so wetting and drying effects are believed to be unimportant.
+A detailed overview of the model configuration and domain including a description of the ocean model we have employed is provided in section \ref{sec:config}.
 This section also describes open and surface boundary conditions as well as choices in grid structure and mixing parametrizations.
@@
 As such, efforts to include wetting and drying will be considered in the future.
-Additionally, the sea surface height anomaly from Neah Bay will be used in future real time water level predictions during the storm surge season.
+Additionally, the sea surface height anomaly from Neah Bay will be used in future real-time water level predictions during the storm surge season.
 Lastly, flooding effects due to wave run up is left for a more detailed inundation study with higher resolution in the near-shore areas.

### 2014-11-27 RoughWork/manuscript.tex
minor edits

@@
 %A section about how we configured the model for SoG 
-We have used the Nucleus for European Modelling of the Ocean (NEMO) framework in its regional configuration to develop an ocean model for the Strait of Georgia and Salish Sea \citep{madec2012nemo}.
+We have used the Nucleus for European Modelling of the Ocean (NEMO) framework in its regional configuration \citep{madec2012nemo} to develop an ocean model for the Strait of Georgia and Salish Sea.
 NEMO is a highly modularized tool used for studying ocean physics, ocean-ice interactions, and the bio-geochemical properties of the ocean.
@@
 The modelled domain extends from the Strait of Juan de Fuca to Puget Sound to Johnstone Strait as shown in Figure \ref{fig:domain}.
-Bathymetry from the Cascadia physiography data set \citep{haugerud1999digital} was smoothed to limit the difference in depth across grid cells.
+Bathymetry from the Cascadia physiography data set \citep{haugerud1999digital} was smoothed to limit large changes in depth across grid cells, so that $\Delta h/\bar{h} \leq 0.8$, where $\Delta h$ is the difference in depth between two adjacent grid cells and $\bar{h}$ is their average depth.
 For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells.
@@
 The 40 vertical $z$-levels were stretched gradually in order to achieve higher resolution in the surface layer, with 1 m vertical grid spacing down to about 10 m in depth.
-Below 10m the grid was stretched to a maximum grid spacing of 27 m at the lowest layer.
+Below 10 m the grid was stretched to a maximum grid spacing of 27 m at the lowest layer.
 At the bottom boundary, partial $z$-levels were utilized in order to limit large changes in bathymetry across grid cells \citep{madec2012nemo}.
@@
 Temperature and salinity at the Juan de Fuca boundary were taken from a weekly climatology which was created from results from a model covering the Salish Sea and the west coast of Vancouver Island \citep{massonfine2012}.
-Their results, originally on s-levels were interpolated onto z-levels and then onto the NEMO horizontal grid.
+Their results, originally on $s$-levels were interpolated onto $z$-levels and then onto the NEMO horizontal grid.
 To prepare the climatology all years (1995-2008) were averaged and results, approximately every 15 days, were interpolated to a weekly climatology.
@@
 The storm surge simulations also used the sea surface height anomaly from Port Hardy, forced at the northern boundary in Johnstone Strait.
-The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}.
-Ath the west, the baroclinic velocities at the boundary were set to be equal to the values inside the boundary (zero-gradient boundary conditions).
-This scheme is not part of core NEMO.
+The tidal forcing and sea surface height was applied to the barotropic velocity forcing with the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}.
+At the west, the baroclinic velocities at the boundary were set equal to the values inside the boundary (zero-gradient boundary conditions).
+This scheme is not part of core NEMO version 3.4 used in this work.
 Zero gradient conditions were chosen because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain.
@@
 Bottom friction is represented by a quadratic law for the bottom momentum flux with drag coefficient $C_D = 5\times 10^{-3}$.
-Vertical turbulence and mixing is calculated through the $k-\epsilon$ configuration of the generic length scale (GLS) turbulence closure \citep{umlauf2003generic} with background vertical eddy viscosity and diffusivity set to $1\times10^{-4}$ and $1\times10^{-5}$ m$^2$ s$^{-1}$ respectively.
+Vertical turbulence and mixing is calculated through the $k-\epsilon$ configuration of the generic length scale (GLS) turbulence closure \citep{umlauf2003generic} with background vertical eddy viscosity and diffusivity set to $1\times10^{-4}$ m$^2$ s$^{-1}$ and $1\times10^{-5}$ m$^2$ s$^{-1}$ respectively.
 Details on the NEMO implementation of the partial slip lateral boundary condition, quadratic bottom friction law, and GLS turbulence closure scheme are provided by \citet{madec2012nemo}.
@@
 The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2014new}.
-Forecasts from the period of 2002-2011 are available.
+Forecasts from the period of 2002-2012 are available.
 Additionally, the inverse barometric effect of the atmospheric pressure is included, an important consideration in storm surge modelling.
+One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada.
+%citation?
 
-River input provides a significant volume of freshwater to the Salish Sea and can influence stratification, circulation and primary productivity.
+River input provides a significant volume of freshwater to the Salish Sea and can influence stratification, circulation, and primary productivity.
 However, most rivers in the domain are not gauged so parametrizations were required to represent river flow.
@@
 The watersheds included in our model were Fraser (which represents approximately 44\% of the freshwater input into our domain), Skagit (12\%), East Vancouver Island (North and South) (12\%), Howe (7\%), Bute (7\%), Puget (6\%), Juan de Fuca (5\%), Jervis (4\%) and Toba (3\%).
-
-The monthly flow from each river was input as a point source in the three grid points closest to the surface at the model point closest to the mouth of each river.
+The monthly flow from each river was input as a point source at a depth of 3 metres at the model point closest to the mouth of each river.
 Incoming water was assumed to be fresh and at surface temperature.
@@
 The model was spun up for a 15.5 months from the initial conditions above, starting Sept 16, 2002, using atmospheric forcing from 2002-2003, climatological temperature and salinity and sea surface height at the boundaries, with tides and climatological river output.
-All storm surge runs were started three days prior to the event of interest with zero initial velocities and sea surface height and a stratification profile from model spin up.
+All storm surge runs were started at least three days prior to the event of interest with zero initial velocities and sea surface height and a stratification profile from model spin up.
 The modelled sea surface height adjusted to forcing in less than one day.
-%citation for Compute Canada resources?
 
-\section{Model Evaluation}\label{sec:model}
+\section{Tidal Evaluation}\label{sec:model}
 
-\subsection{Tidal evaluation}
-The model was initially evaluated qualitatively by comparing patterns of tidal amplitude and phase to results from \citet{foreman1995tidal}.
-For example, the amphidrome around Victoria was produced in the M$_2$ results, as well as the monotonic increase in $K_1$ amplitude moving northwards along the Strait of Georgia.
+The model was initially evaluated qualitatively by comparing patterns of tidal amplitude and phase to results from \citet{foreman1995tidal}, focusing on the major M$_2$ and K$_1$ constituents.
+For example, the M$_2$ amphidrome around Victoria was reproduced as well as the monotonic increase in $K_1$ amplitude moving northwards along the Strait of Georgia.
 Once the model was reproducing observed tidal patterns, model results were quantitatively evaluated by comparing modelled harmonic constituents to measured harmonic constituents at tidal measuring stations throughout the domain.
@@
 \end{equation}\label{eq:compdiff}
-where $A_0$, $A_m$, $g_0$ and $g_m$ are observed and modelled amplitudes and phases.
+where $A_0$, $A_m$, $g_0$ and $g_m$ are the observed and modelled amplitudes and phases.
 Figure \ref{fig:tides} compares modelled and observed amplitude and phase for the $M_2$ and $K_1$ tidal constituents for stations along a transect from the Strait of Juan de Fuca through to the Strait of Georgia and Johnstone Strait.
-Complex differences for these two constituents are also provided.
+Complex differences for these two constituents are also plotted.
 Comparisons have been made with data from \citet{foreman1995tidal}, \citet{foreman2004m} and \citet{foreman2012circulation}.
-Table  \ref{tab:comparison} lists station names and numbers in the Appendix \ref{sec:appendix}.
+Station names and numbers are listed in Table  \ref{tab:comparison} in Appendix \ref{sec:appendix}.
 
@@
 As such, the M$_2$ amplitude and phase are also poorly reproduced here.
-In addition, M$_2$ amphidrome in the model, indicated by the drop in amplitude around station 5 (Victoria), is too strong resulting in lower than observed M$_2$ amplitudes in this region.
-Additionally, the M$_2$ phase errors in the Strait of Juan de Fuca region (stations 1-7) are quite large, resulting in large complex differences in this region.
-However, the match in the Strait of Georgia region (stations 10-24) is very good, with complex differences less than 6 cm, in part due to tidal tuning for a good match in the main area of interest.
-Details on the tidal tuning procedure are provided in the Appendix \ref{sec:appendix}.
-%still need to look up how other models of this region do with tides.
+The model's M$_2$ amphidrome, indicated by the drop in amplitude around station 5 (Victoria), is too strong resulting in lower than observed M$_2$ amplitudes in this region.
+Additionally, the M$_2$ phase errors in the Strait of Juan de Fuca (stations 1-7) are quite large, leading to large complex differences in this region.
+However, the match in the Strait of Georgia (stations 10-24) is very good, with complex differences less than 6 cm, in part due to tidal tuning for highly accurate tides in the main region of interest.
+Details on the tidal tuning procedure are provided in Appendix \ref{sec:appendix}.
 
@@
 On the left, a transect through the stations used for comparison.
-On the right, modelled M$_2$ (black) and K$_1$ (gray) amplitude (top) and phase (middle) compared with observations in the dashed curves of the same colour.
-The bottom plot displays the M$_2$ and K$_1$ complex differences.}
+On the right, modelled M$_2$ (black) and K$_1$ (grey) amplitude (top) and phase (middle) compared with observations in the dashed curves of the same colour.
+The bottom plot displays the M$_2$ and K$_1$ complex differences.
+Station names and numbers are listed Appendix \ref{sec:appendix} in Table \ref{tab:comparison}.}
 \label{fig:tides}
@@
 Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height due to the omission of the next leading order tidal constituents such as J$_1$.
-Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can be up to 40 cm during times of high tidal range, and in particular, during the storm surge season Nov-Feb.
+Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error approach 40 cm during times of high tidal range, and in particular, during the storm surge season Nov-Feb.
 As such, storm surge model predictions were corrected by accounting for this error.
@@
 The modelled residual is calculated as the difference in model output between a simulation with all forcing conditions and a simulation forced only with tides and rivers.
-Water level observations are taken from Fisheries and Oceans Canada (citation) and are used to calculate the observed residual, defined as the observed water level minus tidal predictions.
+Water level observations are taken from Fisheries and Oceans Canada (citation?) and are used to calculate the observed residual, defined as the observed water level minus tidal predictions.
 Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this case, from the year prior surge event of interest.
@@
 These cases have been used to assess the model's skill in reproducing water level height and surge amplitude and timing during storm events.
+The assessment includes comparisons at four permanent tidal measuring stations; Point Atkinson, Victoria, Patricia Bay, and Campbell River.
+These stations are indicated on the map in Figure \ref{fig:domain}.
 In addition, the contribution to the surge due to several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed.
@@
 The maximum water level observed at Point Atkinson on this day was 5.46 m, which is 0.82 m higher than the predicted tides, highlighting the significance of this event as it resulted in a large anomaly and also reached water levels close to the highest level ever recorded (5.61 m).
-A set of simulations from Feb 1 to 8, 2006 under a variety of forcing conditions has been completed in order to determine the contribution of each forcing factor to the surge.
+A set of hindcast simulations from Feb 1-8, 2006 have been completed.
 
-The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations and corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River.
-The timing of maximum water level matches well between the observations and the model at all four locations, the most significant contribution to high water level being from the tides which are generally well-represented by the model.
-There is a good match in the water level maximum at Victoria and Patricia Bay, with a relative error less than 2\%.
-There is only a 4.2 cm difference between observed and modelled water level elevation at Victoria and a 6.9 cm difference at Patricia Bay, the model slightly under predicting the maximum water level in each case.
-The error at Point Atkinson and Campbell River is slightly more significant, reaching 2.5\% and 2.7\% respectively, close to 14 cm over predicted in each case.
-It is surprising that Victoria and Patricia Bay agree most closely with observations because the M$_2$ tidal amplitudes in the model are known to be too low in this area.
-In conjunction with the over predicted water levels at Point Atkinson and Campbell River, it is likely that one of the forcing factors is overestimated in this example.
+The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations to corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River.
+The timing of maximum water level matches well between the observations and the model at all four locations, the most significant contribution to high water being from the tides which are generally well-represented by the model.
+There is good agreement in the water level maximum at both Victoria and Patricia Bay, with a relative error less than 2\%.
+In fact, there is only a 4.2 cm difference between observed and modelled maximum at Victoria and a 6.9 cm difference at Patricia Bay, the model slightly under predicting in each case.
+The error at Point Atkinson and Campbell River is slightly more significant, reaching 2.5\% and 2.7\% respectively, close to 14 cm over predicted at both locations.
+It is surprising that Victoria and Patricia Bay agree so closely with observations because the M$_2$ tidal amplitudes in the model are known to be too low in this area.
+When considering the over predicted water levels at Point Atkinson and Campbell River, it is likely that one of the forcing factors is overestimated in this example.
 
@@
 In general, there is a very good agreement between the observed and modelled residuals.
-The most noticeable difference is a 4-6 hour delay in the model's timing of the maximum residual except at Victoria where the maximum residual in sync with observations.
+The most noticeable difference is a 4-6 hour delay in the model's timing of the maximum residual except at Victoria where the maximum residual is in sync with observations.
 The errors in maximum residual range from 6.0 cm too high at Victoria to 15.9 cm too low at Campbell River.
-Point Atkinson and Patricia Bay both experience a very good match with observed residuals with errors less than 2 cm.
-Note that Point Atkinson and Campbell River observed the largest observed surges at 82 cm and 95 cm respectively.
+Both Point Atkinson and Patricia Bay experience a very good match with observed residuals with errors less than 2 cm.
+Note that Point Atkinson and Campbell River realized the largest observed surges at 82 cm and 95 cm respectively.
 The surges at Victoria and Patricia Bay were observed to be 66 cm and 74 cm.
@@
 
-The maximum water level and maximum residuals are important factors in storm surge hindcasts but a better indicator of overall model performance is provided by a group of statistical measures in Table \ref{tab:feb2006stat}.
-In general, $\gamma^2$ and $WS$ suggest good model agreement for the predicted water levels, most likely because the tidal signal dominates.
-The skill scores degrade slightly for the residuals but are still reasonable and indicate very good agreement.
-Note that the mean absolute error ($\bar{e}$) is lower for the residuals, but this measure does not account for the larger variability in water levels due to the tides.
+The maximum water level and maximum residual are important factors in storm surge hindcasts but a better indicator of overall model performance is provided by a group of statistical measures,including the mean absolute error, the observed mean, the modelled mean, $\gamma^2$, and the Willmott skill score for both the water level and residuals, in Table \ref{tab:feb2006stat}.
+The statistics are calculated with hourly output over the three-day period displayed in Figure \ref{fig:feb2006}.
+To begin with, the mean absolute error in the water level ranges from 7.6 cm at Victoria to 13.1 cm at Campbell River, whereas, the mean absolute error in the residuals is smaller at all stations, however, this measure does not account for the large variability in water levels due to the tides.
+A statistic that does take into account this variablity is $\gamma^2$ which suggests very good agreement in both the predicted water levels and residuals.
+Good agreement is also reflected in high Willmott skill scores at all locations.
+Note that the $\gamma^2$ measures and Willmott skill scores degrade slightly for the residuals, likely because the well-tuned tidal signal dominates the water level predictions.
 
@@
 \hline 
-Point Atkinson & 10.5           & 3.46                 & 3.54                 &   0.012    & 0.994  & 6.2            &  0.273                 & 0.309                  & 0.065     & 0.979     \\
-Victoria       &  8.6           & 2.18                 & 2.19                 &   0.032    & 0.992  & 6.1            &  0.235                 & 0.278                  & 0.085     & 0.970     \\
-Patricia Bay   &  9.1           & 2.60                 & 2.64                 &   0.024    & 0.992  & 6.9            &  0.223                 & 0.278                  & 0.070     & 0.969     \\
-Campbell River & 13.1           & 3.20                 & 3.21                 &   0.041    & 0.990  & 5.8            &  0.303                 & 0.300                  & 0.079     & 0.961     \\
+Point Atkinson & 10.5           & 3.46                 & 3.54                 &   0.012    & 0.996  & 6.2            &  0.273                 & 0.309                  & 0.065     & 0.979     \\
+Victoria       &  7.6           & 2.18                 & 2.19                 &   0.032    & 0.992  & 6.1            &  0.235                 & 0.278                  & 0.085     & 0.970     \\
+Patricia Bay   &  9.1           & 2.60                 & 2.64                 &   0.024    & 0.992  & 6.8            &  0.223                 & 0.278                  & 0.070     & 0.969     \\
+Campbell River & 13.1           & 3.20                 & 3.21                 &   0.041    & 0.990  & 5.8            &  0.303                 & 0.300                  & 0.079     & 0.978     \\
 \hline 
@@
 
+Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow reigons due to nonlinear effects and bottom friction, is presented for this hindcast.
+A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}.
+The most significant differences occur at Point Atkinson where the maximum of the surge-only case is 8.7 cm higher than maximum modelled residual, about 10\% of the total surge amplitude.
+In contrast, the tide-surge interaction at Victoria is hardly noticeable.
+Indeed, a spatial map of the difference between the modelled residual and surge-only difference (not shown here) indicates that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands.
+The passages between these islands are known for vigorous tidal mixing, which may have a nonlinear effect on the surge propagation.
+However, the effect is a relatively small overestimate of the surge elevation and no changes to the timing of the surge.
+So, in operational settings, performing a surge-only simulation and adding the predicted tides \textit{a posteriori} could be more computationally efficient with only a small decrease in the accuracy of the prediction.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/feb2006_tidesurge.pdf}
+\caption{Modelled residuals (solid) compared with a simulation of the surge propagation without tidal forcing (dashed) for the Feb 2006 hindcast. }
+\label{fig:tidesurge}
+\end{figure}
+
 An additional storm surge hindcast for Nov 2006 is provided as another means of assessing model performance and is mentioned briefly here.
@@
 The surge amplitude was lower at the other stations, reaching no more than 68 cm.
-The model performs reasonably well in this hindcast, Willmott skill scores are greater than 0.94 as outlined in Table \ref{tab:nov2006stat}.
+The model performs reasonably well in this hindcast; Willmott skill scores are greater than 0.94 as outlined in Table \ref{tab:nov2006stat}.
 The best match in maximum surge occured at Campbell River with a 6.2 cm underestimate by the model.
-The timing of the maximum surge at Campbell River was identical with observations and slightly earlier at the other stations.
+The timing of the maximum surge at Campbell River was identical with observations and to early at the other the stations.
 
@@
 Point Atkinson &  9.8           &  3.42                & 3.43                 &   0.034    & 0.991  &  7.7           &  0.290                 & 0.247                  &  0.128     & 0.950 \\
-Victoria       &  9.1           &  2.16                & 2.11                 &   0.070    & 0.978  &  6.2           &  0.222                 & 0.247                  &  0.171     & 0.946 \\
+Victoria       &  9.1           &  2.16                & 2.11                 &   0.070    & 0.978  &  6.2           &  0.222                 & 0.202                  &  0.171     & 0.946 \\
 Patricia Bay   & 10.0           &  2.57                & 2.53                 &   0.049    & 0.985  &  6.6           &  0.235                 & 0.203                  &  0.144     & 0.947 \\
-Campbell River & 10.7           &  3.20                & 3.26                 &   0.037    & 0.982  &  6.2           &  0.254                 & 0.236                  &  0.140     & 0.952 \\
+Campbell River & 10.7           &  3.20                & 3.26                 &   0.037    & 0.989  &  6.2           &  0.254                 & 0.236                  &  0.140     & 0.952 \\
 \hline 
@@
 
-Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented in Figure \ref{fig:factors}.
+Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented for the Feb 2006 storm surge in Figure \ref{fig:factors}.
 The modelled residual from four simulations have been plotted.
 First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly, is shown in the dashed black curve.
-Additionally, a simulation with all forcing except the atmospheric forcing is presented in solid grey and a simulation with all forcing except the sea surface height anomaly is shown solid black.
+Next, a simulation with all forcing except the atmospheric forcing is presented in solid grey and a simulation with all forcing except the sea surface height anomaly is shown solid black.
 Finally, a simulation with only wind forcing, that is no anomaly and no inverse barometer effect due to pressure, is shown in the dashed grey curve.
@@
 A similar drop in the maximum surge and delay in timing of the maximum surge is observed at the other locations.
-The atmospheric model indicates southeasterly winds near Point Atkinson during the onset of the surge but after the peak surge has passed, the winds shift to westerly.
+The atmospheric model indicates southeasterly winds near Point Atkinson during the onset of the surge but after the peak surge has passed the winds shift to westerly.
 A strong westerly wind would cause an elevated sea surface height at Point Atkinson as water piles against the coast, explaining the delay in peak surge when only atmospheric forcing is used since the winds shift to westerly after the peak surge has passed.
 A simulation with only wind forcing produces a surge of 7.8 cm at Point Atkinson, a small fraction of the total observed surge, and only centimeters at the other locations.
-Note that this wind-induced surge occurs well after the peak surge from observations and simulations with full forcing.
+Note that this wind-induced surge occurs well after the peak surge occured in both the observations and simulations with full forcing.
 
@@
 Interestingly, the surge amplitude increases by 4.8 cm at Victoria and 3.6 cm at Patricia Bay when atmospheric forces are neglected.
-The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
+The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in a slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
 Additionally, the surge amplitude decreases by a few centimetres if the inverse barometer effect due to atmospheric pressure is not included.
@@
 \includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
-\caption{Comparison of the modelled residual for four simulations: A simulation with all forcing (solid black), a simulation without atmospheric forcing (solid grey), a simulation without the sea surface height anomaly (dashed black) and a simulation with only wind forcing, i.e.\ no sea surface height anomaly or pressure (dashed grey).
+\caption{Comparison of the modelled residual for four simulations in Feb 2006: A simulation with all forcing (solid black), a simulation without atmospheric forcing (solid grey), a simulation without the sea surface height anomaly (dashed black), and a simulation with only wind forcing, i.e.\ no sea surface height anomaly or pressure (dashed grey).
 }
@@
 \end{figure}
-%Fix ylabels!!!!
-
-Finally, an assessment of the importance of including the tide-surge interaction is presented next.
-A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}.
-Recall, the modelled residual is defined as the difference between the sea surface height in a simulation with all forcing and a simulation with tides and rivers only.
-A comparison between the modelled residual and a surge-only run can determine the significance of the tide-surge interaction.
-The most significant differences occur at Point Atkinson where the maximum of the surge-only case is 8.7 cm higher than maximum modelled residual, about 10\% higher of the surge amplitude.
-In contrast, the tide-surge interaction at Victoria is hardly noticeable.
-Indeed, a spatial map of the difference between the modelled residual and surge-only difference (not shown here) indicates that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands.
-The passages between these islands are known for vigorous tidal mixing, which may have a nonlinear effect on the surge propagation.
-However, the effect is a relatively small overestimate of the surge elevation and no changes to the timing of the surge.
-So, in operational settings, performing a surge-only simulation and adding the predicted tides a posteriori could be more computationally efficient with only a small decrease in the accuracy of the prediction.
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.6]{Figures/feb2006_tidesurge.pdf}
-\caption{Modelled residuals (solid) compared with a simulation of the surge propagation without tidal forcing (dashed). }
-\label{fig:tidesurge}
-\end{figure}
-%Fix ylabels!!!!!
 
@@
 The best agreement occurs at Victoria where the observed surge was 62 cm and the model returned 44 cm.
-At all of the locations, the model predicted the maximum surge to occur within 3.5 hours of the observed surge.
+At all of the locations, the model predicted the maximum surge to occur within 3 hours of the observed surge.
 Typically, the modelled surge was too early, in contrast to the later predictions in the Feb 4, 2006 case.
 Error and skill scores for the modelled water level and the modelled residual are provided in Table \ref{tab:dec2006stat}.
-All skill scores rank above 0.89.
+All Willmott skill scores rank above 0.89 and the $\gamma^2$ measure also indicates good agreement between model and observations.
 The modelled residuals are on average too low and have a lower skill score than the Feb 2006 case.
@@
 Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the case.
-%Skill score between CGRF and observed winds?
 
@@
 Patricia Bay   &  9.2           &  2.63                & 2.58                 &   0.031    & 0.990  &  9.5           &  0.193                 & 0.149                  &  0.260     & 0.889 \\
-Campbell River & 10.6           &  3.23                & 3.28                 &   0.030    & 0.992  & 10.6           &  0.191                 & 0.163                  &  0.247     & 0.905 \\
+Campbell River & 10.6           &  3.23                & 3.28                 &   0.030    & 0.992  &  9.4           &  0.191                 & 0.163                  &  0.247     & 0.905 \\
 \hline 
@@
 The evolution and spatial distribution of the surge is highlighted in Figure \ref{fig:spatial} where the modelled residual is plotted every hour beginning on Dec 15 at 05:30 UTC through to Dec 15 at 14:30 UTC.
-The surge enters the system from the Strait of Juan de Fuca, peaking between Dec 15 at 9:30 and Dec 15 at 11:30 and then begins to subside throughout the entire domain.
+The surge enters the system from the Strait of Juan de Fuca, peaking between Dec 15 at 9:30 UTC and Dec 15 at 11:30 UTC and then begins to subside throughout the entire domain.
 Notably, the Strait of Georgia experiences the largest surge values, upwards of 60 cm, whereas the Juan de Fuca region experiences a maximum surge around 45 cm.
-Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges ($\approx$ 10 cm) than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast.
+Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges ($\sim$10 cm) than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast.
 At about 11:30 UTC, the winds switch to northwesterly and then westerly, further amplifying the surge along the mainland coast.
@@
 Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland.
-Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the Strait of Georgia is also affected by the surge entering the domain from Johnstone Strait at the northern boundary or the incomoing surge wave is reflected from the mainland side.
-%fix up reflection bit
+Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the Strait of Georgia is also affected by the surge entering the domain from Johnstone Strait at the northern boundary or the incomoing surge wave is partially reflected from the mainland coast.
 
@@
 \hline 
-Point Atkinson & 14.0           &  3.45                & 3.55                 &   0.012    & 0.995  &  6.1           &  0.313                 & 0.331                  &  0.237     & 0.913 \\
-Victoria       &  9.9           &  2.25                & 2.24                 &   0.023    & 0.989  &  4.8           &  0.323                 & 0.293                  &  0.245     & 0.899 \\
-Patricia Bay   & 11.2           &  2.50                & 2.65                 &   0.020    & 0.997  &  6.0           &  0.277                 & 0.290                  &  0.272     & 0.901 \\
-Campbell River & 13.5           &  3.22                & 3.27                 &   0.019    & 0.991  &  4.9           &  0.326                 & 0.314                  &  0.217     & 0.924 \\
+Point Atkinson & 14.0           &  3.44                & 3.55                 &   0.012    & 0.995  &  6.1           &  0.313                 & 0.331                  &  0.237     & 0.914 \\
+Victoria       &  9.9           &  2.26                & 2.24                 &   0.023    & 0.994  &  4.8           &  0.323                 & 0.293                  &  0.245     & 0.899 \\
+Patricia Bay   & 11.2           &  2.60                & 2.65                 &   0.020    & 0.993  &  6.0           &  0.277                 & 0.290                  &  0.272     & 0.901 \\
+Campbell River & 13.5           &  3.22                & 3.27                 &   0.019    & 0.995  &  4.9           &  0.326                 & 0.314                  &  0.217     & 0.924 \\
 \hline 
@@
 As such, there are only one or two grid cells that are entirely over water in our domain, resulting in overly damped wind speeds at grid points close to land such as Point Atkinson.
-To determine the effect of underresolved wind forcing on the storm surge hindcasts we compare simulations that have been forced with 1) the low resolution CGRF model and 2) a high resolution (2.5 km) operational model from Environment Canada.
+To determine the effect of underresolved wind forcing on the storm surge hindcasts we compare simulations that have been forced with 1) the low resolution CGRF model and 2) a high resolution (2.5 km) HRDPS model from Environment Canada.
 These simulations are performed for a storm surge that occured on Dec 17, 2012 and cover the time period Dec 14-19, 2012.
@@
 At Sandheads, both the low resolution and high resolution models have good agreement with observations.
-The low resolution model, surprsingly, has a very good match likely because this grid cell happens to be mostly over water.
+The low resolution model, surprsingly, has a very good match at Sandheads likely because this grid cell happens to be mostly over water.
 Although not shown here, both models have a similar sea level atmospheric pressure at these two locations.
@@
 \caption{Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson (upper) and Sandheads (lower).
-Observed winds shown in solid black, low resolution CGRF model in dashed black and high resolution model in gray.
+Observed winds shown in solid black, low resolution CGRF model in dashed black and high resolution model in grey.
 }
@@
 
-Modelled and observed residuals for this simulation are displayed in Figure \ref{fig:dec2012_surge} where it is apparent that both of the models agree well with the observations.
+Modelled and observed residuals for this simulation are displayed in Figure \ref{fig:dec2012_surge} where it is apparent that both of the simulations agree well with the observations.
 Willmott skill scores for both modelled residuals are above 0.92 at these four locations.
@@
 \includegraphics[scale=0.6]{Figures/dec2012_surge.pdf}
-\caption{Comparison observed (solid black) and modelled residuals for the Dec 2012 storm surge simulation.
-Modelled residuals from two simulations using a low resolution (dashed) and high resolution (gray) atmospheric product are shown.
+\caption{Comparison of observed (solid black) and modelled residuals for the Dec 2012 storm surge simulation.
+Modelled residuals from two simulations using a low resolution (dashed) and high resolution (grey) atmospheric products are shown.
 }
@@
 Hindcasts of five significant events in February 2006, November 2006, December, November 2009, and December 2012 were simulated.
-In general, the model performs well at reproducing the observed water level heights and surge amplitudes, with Willmott skill scores greater than 0.88 in all cases.
+In general, the model performs well at reproducing the observed water level heights and residulas, with Willmott skill scores greater than 0.88 in all cases.
 We have also assessed the relative importance of different forcing factors contributing to the storm surge.
 The sea surface height anomaly entering the system from the Pacific Ocean contributes most significantly to the surge amplitude, in agreement with previous studies in this region \citep{murty1995storm}.
-Although the large scale wind patterns and atmopsheric conditions important in setting up the surge on the Pacific Ocean boundaries, the local winds within the Strait of Georgia are not the leading order contribution to surge amplitude.
-However, the local wind patterns can set up small spatial differences in the surge amplitude, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season.
-Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions.
-Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important after the San Juan and Gulf Islands.
-The modelled surge amplitude increased by about 10\% at Point Atkinson when the tide-surge interaction was neglected.
+Although the large scale wind patterns and atmospheric conditions are important in setting up the surge on the Pacific Ocean boundaries, the local winds within the Strait of Georgia are not the leading order contribution to surge amplitude.
+However, the local wind patterns can set up small spatial differences in the surge, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season.
+Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions or a partially reflection from the mainland coast.
+Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important north the San Juan and Gulf Islands.
+The modelled surge amplitude increased by about 10\% at Point Atkinson when the tide-surge interaction was neglected in one example.
 
@@
 Although the high resolution model delivers fine scale detail and more accurate local wind patterns, it did not provide any improvement to the surge hindcast.
-Since both of these products provide the same level of accuracy in storm surge hindcasts, in the future we will use the high resolution atmospheric model to produce real-time storm surge predictions, even though most of out testing was performed with the low resolution model.
+Since both of these products provide the same level of accuracy for storm surge hindcasts, in the future we will use the high resolution atmospheric model to produce real-time storm surge predictions, even though most of out testing was performed with the low resolution model.
 
@@
 For instance, the complicated coastlines and local wind patterns result in important spatial variations in the surge amplitude.
-As such, the strong wind caused by a large storm are is not the only factor to be considered in community preparedness for surge conditions.
+As such, the strong wind caused by a large storm is not the only factor to be considered in community preparedness for surge conditions.
 The geometry of the Strait and local wind patterns render communities on the mainland side of the southern Strait of Georgia more susceptible to surges.
@@
 This section details the procedure used to tune the tidal forcing for a good match with observations in the Strait of Georgia.
-Originally, tidal elevation and currents were extracted from a tide prediction tool called Webtide \citep{foreman2000webtide} at the western Juan de Fuca boundary.
+Originally, eight tidal elevation and currents were extracted from a tide prediction tool called Webtide \citep{foreman2000webtide} at the western Juan de Fuca boundary.
 At the northern boundary,  M$_2$ and K$_1$ tidal harmonics for both elevation and currents were taken from \citet{thomson1980johnstone}, as well as elevation harmonics for O$_1$ and S$_2$.
@@
 This project is funded by the Marine Environmental Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada.
+This research was enabled in part by support provided by WestGrid (www.westgrid.ca) and Compute Canada Calcul Canada (www.computecanada.ca).
 

### 2014-11-27 RoughWork/manuscript.tex
Looked for typos and spelling

@@
 They found that a local increase in the bottom friction coefficient, especially in the area of Haro Strait, was necessary in order to adequately reproduce the observed M$_2$ tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}.
-Without local adjustments to the bottom friction.
-thei M$_2$ amplitudes were too high in the Strait of Juan de Fuca and too low in the Strait of Georgia.
+Without local adjustments to the bottom friction their M$_2$ amplitudes were too high in the Strait of Juan de Fuca and too low in the Strait of Georgia.
 Other studies over a similar domain, such as the barotropic finite element model of \citet{foreman1995tidal}, have avoided local increases in bottom friction.
@@
 This model employed a sigma or terrain-following vertical coordinate system leading to spurious diapycnal mixing since, numerically, the lateral mixing is computed on sigma levels which may intersect isopycnals in the presence of steep topography.
-A finite volume model on an unstrcutured grid by \citet{yang2010multi} used a truncated domain with an open boundary in the northern Strait of Georgia, where under predictions in the M$_2$ tidal phase over most of their domain were suggested to be caused by an inaccurate representation of the tides at their northern boundary.
-\citet{sutherland2011model} employed the Regional Ocean Modelling System (ROMS) to study estuarine circulation in the Salish Sea with a model domain that extended outside the Strait of Juan de Fuca into the Pacific Ocean and up to the northern Strait of Georgia and noted that their semidiurnal amplitudes were generally too low.
+A finite volume model on an unstructured grid by \citet{yang2010multi} involved a truncated domain with an open boundary in the northern Strait of Georgia, where under predictions in the M$_2$ tidal phase over most of their domain were suggested to be caused by an inaccurate representation of the tides at their northern boundary.
+\citet{sutherland2011model} employed the Regional Ocean Modelling System (ROMS) to study estuarine circulation in the Salish Sea with a model domain that extended outside the Strait of Juan de Fuca into the Pacific Ocean and up to the northern Strait of Georgia and noted that their semi-diurnal amplitudes were generally too low.
 An updated version of this model produced high temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}.
@@
 Most of these models neglect the effects of stratification, either by treating the ocean as a constant temperature and salinity or by using a two-dimensional depth-averaged set of equations, an assumption that is assumed to be valid.
+Further, an accurate representation of the meteorological conditions at the ocean surface is important.
 
@@
 Compared to eastern Canada \citep{gray1984preliminary, bernier2006predicting,bobanovic2006forecasting, bernier2007tide, bernier2010tide}, relatively little has been published on storm surge modelling on the west coast of Canada in recent years.
-A statisical overview of past surge events in British Columbia waters suggests an increase in extreme water level events during warm ENSO periods \citep{abeys2011extreme}.
+A statistical overview of past surge events in British Columbia waters suggests an increase in extreme water level events during warm ENSO periods \citep{abeys2011extreme}.
 
@@
 Model evaluation has focused on regions where the near-coast slopes are steep and so wetting and drying effects are believed to be unimportant.
-A detailed overview of the model configuration and domain including a description of the ocean model we have employed is provided in section \ref{sec:config}.
+A detailed overview of the model configuration and domain including a description of the ocean model configuration is provided in section \ref{sec:config}.
 This section also describes open and surface boundary conditions as well as choices in grid structure and mixing parametrizations.
@@
 At the northern boundary, the tides were tuned for good matches at Yorke Island and Kelsey Bay.
-Further details on the tuning are provided in the appendix \ref{sec:appendix}.
+Further details on the tuning are provided in Appendix \ref{sec:appendix}.
 
@@
 
-Sea surface height at the mouth of Juan de Fuca was set using values from the Tofino tide gauge.
+Sea surface height at the mouth of Juan de Fuca was set using values from the Tofino tide gauge which is outside of but close to the western boundary of the model domain.
 A monthly climatology was produced using daily averages from 2000-2010, binning them by month, averaging and setting the yearly mean to zero.
@@
 At the west, the baroclinic velocities at the boundary were set equal to the values inside the boundary (zero-gradient boundary conditions).
-This scheme is not part of core NEMO version 3.4 used in this work.
+This scheme is not part of core NEMO 3.4, the version used here.
 Zero gradient conditions were chosen because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain.
@@
 Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height due to the omission of the next leading order tidal constituents such as J$_1$.
-Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error approach 40 cm during times of high tidal range, and in particular, during the storm surge season Nov-Feb.
+Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can approach 40 cm during times of high tidal range, and in particular, during the storm surge season Nov-Feb.
 As such, storm surge model predictions were corrected by accounting for this error.
-Additionally, since the sea surface height is calculated about a zero reference level, the modelled water level is determined by adding a long term average of mean sea level at the specified location to the modelled sea surface height.
+Additionally, since the sea surface height is calculated about a zero reference level, the modelled water level is determined by adding a long term mean sea level at the specified location to the modelled sea surface height.
 
@@
 These cases have been used to assess the model's skill in reproducing water level height and surge amplitude and timing during storm events.
-The assessment includes comparisons at four permanent tidal measuring stations; Point Atkinson, Victoria, Patricia Bay, and Campbell River.
-These stations are indicated on the map in Figure \ref{fig:domain}.
+The assessment includes comparisons at four permanent tidal measuring stations; Point Atkinson, Victoria, Patricia Bay, and Campbell River (locations indicated in Figure \ref{fig:domain}).
 In addition, the contribution to the surge due to several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed.
 We have also examined the spatial variability of the surge as it propagates through the domain and have considered a case with a strong wind event but no significant observed surge.
+Finally, comparisons between hindcasts using high resolution and low resolution atmospheric forcing have been completed.
 
@@
 The errors in maximum residual range from 6.0 cm too high at Victoria to 15.9 cm too low at Campbell River.
-Both Point Atkinson and Patricia Bay experience a very good match with observed residuals with errors less than 2 cm.
+Both Point Atkinson and Patricia Bay experience a very good match with observed residuals with errors in the maximum less than 2 cm.
 Note that Point Atkinson and Campbell River realized the largest observed surges at 82 cm and 95 cm respectively.
@@
 
-The maximum water level and maximum residual are important factors in storm surge hindcasts but a better indicator of overall model performance is provided by a group of statistical measures,including the mean absolute error, the observed mean, the modelled mean, $\gamma^2$, and the Willmott skill score for both the water level and residuals, in Table \ref{tab:feb2006stat}.
+The maximum water level and maximum residual are important factors in storm surge hindcasts but a better indicator of overall model performance is provided by a group of statistical measures, including the mean absolute error, the observed mean, the modelled mean, $\gamma^2$, and the Willmott skill score for both the water level and residuals, in Table \ref{tab:feb2006stat}.
 The statistics are calculated with hourly output over the three-day period displayed in Figure \ref{fig:feb2006}.
 To begin with, the mean absolute error in the water level ranges from 7.6 cm at Victoria to 13.1 cm at Campbell River, whereas, the mean absolute error in the residuals is smaller at all stations, however, this measure does not account for the large variability in water levels due to the tides.
-A statistic that does take into account this variablity is $\gamma^2$ which suggests very good agreement in both the predicted water levels and residuals.
+A statistic that does take into account this variability is $\gamma^2$ which suggests very good agreement in both the predicted water levels and residuals.
 Good agreement is also reflected in high Willmott skill scores at all locations.
@@
 \caption{Evaluation of model performance for the Feb 2006 simulation.
-Mean absolute error ($\bar{e}$), observed mean (subscript o), modelled mean (subscript m), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and residuals ($\delta$) during the period displayed in Figure \ref{fig:feb2006}.}
+Mean absolute error ($\bar{e}$), observed mean (subscript o), modelled mean (subscript m), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and residuals ($\delta$) during the period displayed in Figure \ref{fig:feb2006} from Feb 3, 2006 to Feb 6, 2006.}
 \label{tab:feb2006stat} 
 \end{table} 
-%anything else to include? statistics calculated in Feb 2006 -paper production.ipynb and saved in statisitics_feb2006.CV 
+%anything else to include? statistics calculated in Feb 2006 -paper production.ipynb and saved in statistics_feb2006.CV 
 %I also have standard deviations, correlations, ...
 
-Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow reigons due to nonlinear effects and bottom friction, is presented for this hindcast.
+Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow regions due to nonlinear effects and bottom friction, is presented for this hindcast.
 A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}.
@@
 The model performs reasonably well in this hindcast; Willmott skill scores are greater than 0.94 as outlined in Table \ref{tab:nov2006stat}.
-The best match in maximum surge occured at Campbell River with a 6.2 cm underestimate by the model.
-The timing of the maximum surge at Campbell River was identical with observations and to early at the other the stations.
+The best match in maximum surge occurred at Campbell River with a 6.2 cm underestimate by the model.
+The timing of the maximum surge at Campbell River was identical with observations and too early at the other the stations.
 
@@
 \end{tabular}
-\caption{Evaluation of model performance for the Nov 2006 hindcast.}
+\caption{Evaluation of model performance for the Nov 2006 hindcast. Statistics were calculated over the period Nov 14, 2006 to Nov 17, 2006.}
 \label{tab:nov2006stat} 
@@
 A strong westerly wind would cause an elevated sea surface height at Point Atkinson as water piles against the coast, explaining the delay in peak surge when only atmospheric forcing is used since the winds shift to westerly after the peak surge has passed.
-A simulation with only wind forcing produces a surge of 7.8 cm at Point Atkinson, a small fraction of the total observed surge, and only centimeters at the other locations.
-Note that this wind-induced surge occurs well after the peak surge occured in both the observations and simulations with full forcing.
+A simulation with only wind forcing produces a surge of 7.8 cm at Point Atkinson, a small fraction of the total observed surge, and only centimetres at the other locations.
+Note that this wind-induced surge occurs well after the peak surge occurred in both the observations and simulations with full forcing.
 
@@
 \subsection{Spatial Extent of Surge}
-On Dec 15, 2006, a major storm with wind gusts over 120 km/hr caused significant damage to many coastal communities on the Pacific Northwest.
+On Dec 15, 2006, a major storm with wind gusts over 120 km hr$^{-1}$ caused significant damage to many coastal communities on the Pacific Northwest.
 Although the storm was very strong and the observed residual at Point Atkinson reached 84 cm, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area \citep{forseth2006adaptation}.
@@
 This could be due to a number of factors.
-First,the very large wind speeds observed during this storm are significantly underrepresented by the atmospheric model leading to lower wind stress and less piling of water against the coast.
+First, the very large wind speeds observed during this storm are significantly underrepresented by the atmospheric model leading to lower wind stress and less piling of water against the coast.
 Second, the anomaly forcing at the open boundaries could be too low.
@@
 \end{tabular}
-\caption{Evaluation of model performance for the Dec 2006 hindcast.}
+\caption{Evaluation of model performance for the Dec 2006 hindcast. Statistics were calculated over the period Dec 14, 2006 to Dec 17, 2006.}
 \label{tab:dec2006stat} 
@@
 Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges ($\sim$10 cm) than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast.
-At about 11:30 UTC, the winds switch to northwesterly and then westerly, further amplifying the surge along the mainland coast.
+At about 11:30 UTC, the winds at Point Atkinson switch to northwesterly and then westerly, further amplifying the surge along the mainland coast.
 
@@
 An evaluation of the spatial distribution of the surge in a simulation without atmospheric forcing is provided in Figure \ref{fig:spatial_sshonly}.
-In this case, the surge amplitude across the Strait of Georgia, is uniform with no pronounced difference between the mainland side and Vancouver Island side.
+In this case, the surge amplitude across the Strait of Georgia is uniform with no pronounced difference between the mainland side and Vancouver Island side.
 While the surge entering the domain from the Pacific Ocean is the leading order contribution, the winds may induce localized changes to the surge amplitude.
 Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland.
-Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the Strait of Georgia is also affected by the surge entering the domain from Johnstone Strait at the northern boundary or the incomoing surge wave is partially reflected from the mainland coast.
+Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the Strait of Georgia is also affected by the surge entering the domain from Johnstone Strait at the northern boundary or the incoming surge wave is partially reflected from the mainland coast.
 
@@
 Yet, the surge associated with this event is rather small when compared to the previous two cases as shown in Figure \ref{fig:nov2009} where we have plotted the wind speed at Point Atkinson, sea surface height anomaly forcing at the open boundaries, and observed and modelled residuals at Point Atkinson.
-Two significant surge events are noted, the first occurring on Nov 19 and a second, larger surge on Nov 20.
-The largest surge did not coincide with the strong wind event which peaked at Nov 19 01:00 UTC, reflected in both the model and observations.
-At Point Atkinson, the highest observed surge reached 62.5 cm on Nov 20 at 10:30 UTC, while the model recorded a maximum surge of 50.5 cm five hours later.
-On Nov 19, the observed surge reached upwards of 50 cm, which is low when compared to the other cases considered.
+Two significant surge events are noted, the first occurring on Nov 19 with an observed amplitude of about 50 cm at Point Atkinson, and a second, larger surge of 62.5 cm on Nov 20.
+Surprisingly, the strong wind event of Nov 19 was associated with the smaller of the two surges.
 Neither of the surges occurred during a high tide and so no high water level events were recorded.
@@
 These changes in water level can also contribute to storm surges.
-During the high wind event on Nov 19 and Nov 20, the winds at Point Atkinson were directed to the west, explaining the lack of strong surge observed during this event since wind stresses would act to depress the water on the east side of the basin.
+During the high wind events on Nov 19 and Nov 20, the winds at Point Atkinson were directed to the west, explaining the lack of strong surge observed during this event since wind stresses would act to depress the water on the east side of the basin.
 In fact, the surge amplitude at Victoria on Nov 20 was 64.7 cm which is slightly higher than that observed at Point Atkinson (62.5 cm).
 Of the cases considered so far, this is the only example where the surge amplitude at Victoria was higher than that at Point Atkinson.
-In fact, this event observed a higher surge at Victoria than the Dec 2006 event, further highlighting the role that winds plays in setting up local variations in the surge.
+This event observed a higher surge at Victoria than the Dec 2006 event, further highlighting the role that winds plays in setting up local variations in the surge.
 In this case, the model results do not reflect a higher surge at Victoria (46.1 cm versus 50.5 cm at Point Atkinson), perhaps due to an underestimate in the wind speed and some discrepancies in the wind direction of the atmospheric forcing.
@@
 \end{tabular}
-\caption{Evaluation of model performance for the Nov 2009 hindcast.}
+\caption{Evaluation of model performance for the Nov 2009 hindcast. Statistics were generated over the period Nov 17, 2009 to Nov 22, 2009.}
 \label{tab:nov2009stat} 
@@
 \subsection{High Resolution Atmospheric Forcing}
-The above comparison between the modelled and observed winds at Point Atkinson suggests that the atmopsheric forcing product (CGRF) may not be adequate to produce accurate storm surge hindcasts since the wind speeds are substanitially lower than observed.
+The above comparison between the modelled and observed winds at Point Atkinson suggests that the atmospheric forcing product (CGRF) may not be adequate to produce accurate storm surge hindcasts since the wind speeds are substantially lower than observed.
 CGRF is a global atmospheric reforecasting model with horizontal resolution approximately 33 km over the Salish Sea domain, which is problematic since the Strait of Georgia itself is only about 30 km wide.
 As such, there are only one or two grid cells that are entirely over water in our domain, resulting in overly damped wind speeds at grid points close to land such as Point Atkinson.
-To determine the effect of underresolved wind forcing on the storm surge hindcasts we compare simulations that have been forced with 1) the low resolution CGRF model and 2) a high resolution (2.5 km) HRDPS model from Environment Canada.
-These simulations are performed for a storm surge that occured on Dec 17, 2012 and cover the time period Dec 14-19, 2012.
+To determine the effect of under-resolved wind forcing on the storm surge hindcasts we compare simulations that have been forced with 1) the low resolution CGRF model and 2) a high resolution (2.5 km) HRDPS model from Environment Canada.
+These simulations are performed for a storm surge that occurred on Dec 17, 2012 and cover the time period Dec 14-19, 2012.
 The high resolution product is not available for the hindcasts already discussed.
@@
 At Sandheads, both the low resolution and high resolution models have good agreement with observations.
-The low resolution model, surprsingly, has a very good match at Sandheads likely because this grid cell happens to be mostly over water.
+The low resolution model has a very good match at Sandheads likely because this grid cell happens to be mostly over water.
 Although not shown here, both models have a similar sea level atmospheric pressure at these two locations.
@@
 Willmott skill scores for both modelled residuals are above 0.92 at these four locations.
-Further, the modelled residuals deviate only slightly from each other, most noticeable at Point Atkinson during the large wind event on Dec 15, 2012 where simulation with the low resolution atmopsheric forcing has a residual that is 5 cm higher.
+Further, the modelled residuals deviate only slightly from each other, most noticeable at Point Atkinson during the large wind event on Dec 15, 2012 where simulation with the low resolution atmospheric forcing has a residual that is 5 cm higher.
 Although the wind speeds on Dec 15, 2012 are quite high, this event did not result in any flooding or residuals greater than 50 cm.
-Significant flooding occured later, on Dec 17, 2012, where the observed surge amplitude at Point Atkinson was 68.3 cm.
-Even though the low resolution model significantly underpredicts the wind speed at some locations, it performs comparable to forcing with high resolution prdocuts.
+Significant flooding occurred later, on Dec 17, 2012, where the observed surge amplitude at Point Atkinson was 68.3 cm.
+Even though the low resolution model significantly underpredicts the wind speed at some locations, it performs comparable to forcing with high resolution products.
 This result further demonstrates that the local wind is not the most important factor for setting up storm surges over this domain.
@@
 Hindcasts of five significant events in February 2006, November 2006, December, November 2009, and December 2012 were simulated.
-In general, the model performs well at reproducing the observed water level heights and residulas, with Willmott skill scores greater than 0.88 in all cases.
+In general, the model performs well at reproducing the observed water level heights and residuals, with Willmott skill scores greater than 0.88 in all cases.
 We have also assessed the relative importance of different forcing factors contributing to the storm surge.
@@
 However, the local wind patterns can set up small spatial differences in the surge, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season.
-Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions or a partially reflection from the mainland coast.
+Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions or a partial reflection from the mainland coast.
 Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important north the San Juan and Gulf Islands.
@@
 As such, efforts to include wetting and drying will be considered in the future.
-Additionally, the sea surface height anomaly from Neah Bay will be used in future real-time water level predictions during the storm surge season.
+Additionally, the sea surface height anomaly from Neah Bay, forecasted by the National Oceanic and Atmospheric Administration (NOAA), will be used in future real-time water level predictions during the storm surge season.
 Lastly, flooding effects due to wave run up is left for a more detailed inundation study with higher resolution in the near-shore areas.
@@
 The geometry of the Strait and local wind patterns render communities on the mainland side of the southern Strait of Georgia more susceptible to surges.
-Additionally, commmunities such as Campbell River are at risk due to the long fetch and predominately southerly winds during the winter months.
+Additionally, communities such as Campbell River are at risk due to the long fetch and predominately southerly winds during the winter months.
 
 Finally, difficulties in reproducing accurate M$_2$ tidal harmonics in both the Strait of Georgia and the Strait of Juan de Fuca have been highlighted in this study among others \citep{stronach1993update, foreman2004m}.
-Future work will attempt to improve model accuracy by examining mixing parametrizations and energy dissipation due to bottom friction and their effect on tidal haromnics.
-Additionally, a biological-chemical modelling compenent will be added.
+Future work will attempt to improve model accuracy by examining mixing parametrizations and energy dissipation due to bottom friction and their effect on tidal harmonics.
+Additionally, a biological-chemical modelling component will be added.
 
@@
 This section details the procedure used to tune the tidal forcing for a good match with observations in the Strait of Georgia.
-Originally, eight tidal elevation and currents were extracted from a tide prediction tool called Webtide \citep{foreman2000webtide} at the western Juan de Fuca boundary.
+Originally, the eight most important tidal elevation and currents harmonics were extracted from a tide prediction tool called Webtide \citep{foreman2000webtide} at the western Juan de Fuca boundary.
 At the northern boundary,  M$_2$ and K$_1$ tidal harmonics for both elevation and currents were taken from \citet{thomson1980johnstone}, as well as elevation harmonics for O$_1$ and S$_2$.
@@
 $D$ is the complex difference.
-The mean error and RMS error are calculated as the difference between the modelled quantity and observered quantity.
+The mean error and root mean square (RMS) error are calculated as the difference between the modelled quantity and observed quantity.
 }

### 2014-12-01 RoughWork/manuscript.tex
Added more acknolwedgements

@@
 \section{Acknowledgements}\label{sec:ack}
-We would like to thank Michael Foreman, Diane Masson, Luc Fillion, Rich Pawlowicz, and Mark Halverson for providing data used in model set up and evaluation.
+We would like to thank Michael Foreman, Diane Masson, Luc Fillion, Rich Pawlowicz, Mark Halverson, Kao-Shen Chung, and Pramod Thupaki for providing data used in model set up and evaluation.
+For providing guidance with stakeholder engagement, we also thank Stephanie Chang, Jackie Yip, and Shona van Zijil de jong.
 This project is funded by the Marine Environmental Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada.

### 2014-12-03 RoughWork/manuscript.tex
Added environment canada references

@@
 Additionally, the inverse barometric effect of the atmospheric pressure is included, an important consideration in storm surge modelling.
-One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada.
-%citation?
+One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.
 
@@
 Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered.
-As recorded in the Environment Canada Historical Climate Data archives, the maximum hourly wind speed at Point Atkinson was 18.6 m s$^{-1}$ and the largest gust observed was 25.8 m s$^{-1}$ or 93 km hr$^{-1}$.
+As recorded in the Environment Canada Climate Data archives \citep{ECClimateArchive}, the maximum hourly wind speed at Point Atkinson was 18.6 m s$^{-1}$ and the largest gust observed was 25.8 m s$^{-1}$ or 93 km hr$^{-1}$.
 In addition, large wind speeds over 15 m s$^{-1}$ persisted for over 9 hours.
@@
 Although not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and another surge entering the system at Tofino and Port Hardy.
-%citation for Environment Canada data?
 
@@
 \caption{Wind and storm surge conditions at Point Atkinson for a strong wind event on Nov 19, 2009.
-Top: Wind speed from Environment Canada observations at Point Atkinson (solid) and from the atmospheric model (CGRF) at a grid point near Point Atkinson (dashed).
+Top: Wind speed from Environment Canada observations \citep{ECClimateArchive} at Point Atkinson (solid) and from the atmospheric model (CGRF) at a grid point near Point Atkinson (dashed).
 Middle: Sea surface height anomaly at Tofino (solid) and Port Hardy (dashed) used as forcing conditions at the western and northern open boundaries.
@@
 Observed winds shown in solid black, low resolution CGRF model in dashed black and high resolution model in grey.
-}
+Observations are from the Environment Canada Climate Data archives \citep{ECClimateArchive}.}
 \label{fig:dec2012_weather}

### 2014-12-03 RoughWork/manuscript.tex
Wordy corrections from Susan

@@
 An updated version of this model produced high temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}.
-Smaller domains have also been considered, notably, the passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.
-Accurate reproduction of observed tidal amplitudes and phases and appropriate treatment of mixing is a recurring difficulty in modelling of this region.
+Smaller domains have also been considered; notably, the passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.
+Accurate reproduction of observed tidal amplitudes and phases and appropriate treatment of mixing is a recurring difficulty in modelling this region.
 
@@
 Although not used in the present work, NEMO's options for grid nesting and bio-geochemical coupling make it a useful tool for studying the complex physics and bio-geochemical interactions within the Strait of Georgia.
-This work focuses on validating the physical set up of the Salish Sea model, in particular, determining appropriate forcing and boundary conditions for accurate reproduction of tidal amplitudes and phases as well as storm surge elevations.
+This work focuses on validating the physical configuration of the Salish Sea model, in particular, determining appropriate forcing and boundary conditions for accurate reproduction of tidal amplitudes and phases as well as storm surge elevations.
 Future work will include bio-geochemical coupling and data assimilation.
@@
 As such, the split-explicit time stepping algorithm is employed, where the free surface and barotropic equations are solved with a smaller time step than that used for the other variables.
-The model time step and barotropic time step are 10 s and 2 s respectively.
-Small vertical grid spacing and large vertical velocities in the Boundary Pass and Haro Strait region impose a fairly restrictive time step through the vertical CFL condition.
+The model baroclinic time step and barotropic time step are 10 s and 2 s respectively.
+Small vertical grid spacing and large vertical velocities in the Boundary Pass and Haro Strait region impose a fairly restrictive baroclinic time step through the vertical Courant-Friedrichs-Lewy (CFL) condition.
 
@@
 
-Sea surface height at the mouth of Juan de Fuca was set using values from the Tofino tide gauge which is outside of but close to the western boundary of the model domain.
+Sea surface height at the mouth of Juan de Fuca was set using values from the Tofino tide gauge which is outside of, but close to, the western boundary of the model domain.
 A monthly climatology was produced using daily averages from 2000-2010, binning them by month, averaging and setting the yearly mean to zero.
@@
 The watersheds included in our model were Fraser (which represents approximately 44\% of the freshwater input into our domain), Skagit (12\%), East Vancouver Island (North and South) (12\%), Howe (7\%), Bute (7\%), Puget (6\%), Juan de Fuca (5\%), Jervis (4\%) and Toba (3\%).
-The monthly flow from each river was input as a point source at a depth of 3 metres at the model point closest to the mouth of each river.
-Incoming water was assumed to be fresh and at surface temperature.
+The monthly flow from each river was input as a point source from 0-3 metres at the model point closest to the mouth of each river.
+Incoming water was assumed to be fresh and at surface air temperature.
 A total of 150 rivers were parametrized by this method and their locations are indicated by the blue dots in Figure \ref{fig:domain}.
@@
 For example, the M$_2$ amphidrome around Victoria was reproduced as well as the monotonic increase in $K_1$ amplitude moving northwards along the Strait of Georgia.
-Once the model was reproducing observed tidal patterns, model results were quantitatively evaluated by comparing modelled harmonic constituents to measured harmonic constituents at tidal measuring stations throughout the domain.
+As the model was reproducing observed tidal patterns, model results were quantitatively evaluated by comparing modelled harmonic constituents to measured harmonic constituents at tidal measuring stations throughout the domain.
 Comparisons were made using the complex difference ($D$), defined as \citep{foreman1995tidal}:
@@
 \section{Storm Surge Hindcasts}\label{sec:storm}
-A series of storm surge hindcasts is presented next.
+A series of five storm surge hindcasts is presented next.
 Storm surge simulations were started from rest at least three days prior to the event of interest to allow for adjustment of model currents.
@@
 Water level observations are taken from Fisheries and Oceans Canada (citation?) and are used to calculate the observed residual, defined as the observed water level minus tidal predictions.
-Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this case, from the year prior surge event of interest.
+Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this case, from the year prior to the surge event of interest.
 Additionally, several simulations with varying forcing conditions such as removing atmospheric forcing or removing the sea surface height anomaly at the open boundaries have been completed.
 
-Several significant surge or wind events have been considered in detail, as well as an array of hindcasts to asses model ability in a statistical measure.
+Several significant surge or wind events have been considered in detail, as well as an array of hindcasts to assess model ability in a statistical measure.
 These cases have been used to assess the model's skill in reproducing water level height and surge amplitude and timing during storm events.
@@
 In fact, there is only a 4.2 cm difference between observed and modelled maximum at Victoria and a 6.9 cm difference at Patricia Bay, the model slightly under predicting in each case.
-The error at Point Atkinson and Campbell River is slightly more significant, reaching 2.5\% and 2.7\% respectively, close to 14 cm over predicted at both locations.
+The error at Point Atkinson and Campbell River is slightly more significant, reaching 2.5\% and 2.7\% respectively, close to 14~cm over predicted at both locations.
 It is surprising that Victoria and Patricia Bay agree so closely with observations because the M$_2$ tidal amplitudes in the model are known to be too low in this area.
@@
 
-A comparison between observed and modelled residuals is also displayed in the panels on the right.
-In general, there is a very good agreement between the observed and modelled residuals.
+In general, there is a very good agreement between the observed and modelled residuals \ref{fig:feb2006}.
 The most noticeable difference is a 4-6 hour delay in the model's timing of the maximum residual except at Victoria where the maximum residual is in sync with observations.
@@
 Both Point Atkinson and Patricia Bay experience a very good match with observed residuals with errors in the maximum less than 2 cm.
-Note that Point Atkinson and Campbell River realized the largest observed surges at 82 cm and 95 cm respectively.
-The surges at Victoria and Patricia Bay were observed to be 66 cm and 74 cm.
 
@@
 
-The maximum water level and maximum residual are important factors in storm surge hindcasts but a better indicator of overall model performance is provided by a group of statistical measures, including the mean absolute error, the observed mean, the modelled mean, $\gamma^2$, and the Willmott skill score for both the water level and residuals, in Table \ref{tab:feb2006stat}.
+The maximum water level and maximum residual are important factors in storm surge hindcasts but a better indicator of overall model performance is provided by a group of statistical measures, including the mean absolute error, the observed mean, the modelled mean, $\gamma^2$, and the Willmott skill score for both the water level and residuals (Table \ref{tab:feb2006stat}).
 The statistics are calculated with hourly output over the three-day period displayed in Figure \ref{fig:feb2006}.
@@
 
-Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow regions due to nonlinear effects and bottom friction, is presented for this hindcast.
-A simulation of the surge propagation without including any tidal forcing is compared with the modelled residual in Figure \ref{fig:tidesurge}.
+Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow regions due to nonlinear effects and bottom friction, is presented for this hindcast (Figure \ref{fig:tidesurge}).
 The most significant differences occur at Point Atkinson where the maximum of the surge-only case is 8.7 cm higher than maximum modelled residual, about 10\% of the total surge amplitude.
@@
 The surge amplitude was lower at the other stations, reaching no more than 68 cm.
-The model performs reasonably well in this hindcast; Willmott skill scores are greater than 0.94 as outlined in Table \ref{tab:nov2006stat}.
+The model performs reasonably well in this hindcast; Willmott skill scores are greater than 0.94 (Table \ref{tab:nov2006stat}).
 The best match in maximum surge occurred at Campbell River with a 6.2 cm underestimate by the model.
@@
 
-Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented for the Feb 2006 storm surge in Figure \ref{fig:factors}.
-The modelled residual from four simulations have been plotted.
-First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly, is shown in the dashed black curve.
-Next, a simulation with all forcing except the atmospheric forcing is presented in solid grey and a simulation with all forcing except the sea surface height anomaly is shown solid black.
-Finally, a simulation with only wind forcing, that is no anomaly and no inverse barometer effect due to pressure, is shown in the dashed grey curve.
-It is clear that the sea surface height anomaly contributes most significantly to the surge at each of these locations.
+Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented for the Feb 2006 storm surge (Figure \ref{fig:factors}).Four simulations were done: First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly.
+Second, a simulation with all forcing except the atmospheric forcing.
+Third, a simulation with all forcing except the sea surface height anomaly.
+Fourth, a simulation with only wind forcing, that is no anomaly and no inverse barometer effect due to pressure.
+It is clear that the sea surface height anomaly contributes most significantly to the surge at each of these locations (Figure \ref{fig:factors}).
 When the sea surface height anomaly is not included as a forcing condition, the surge drops from a maximum 82 cm to 11 cm at Point Atkinson and its maximum occurs eight hours later.
 A similar drop in the maximum surge and delay in timing of the maximum surge is observed at the other locations.
-The atmospheric model indicates southeasterly winds near Point Atkinson during the onset of the surge but after the peak surge has passed the winds shift to westerly.
+The atmospheric model indicates southeasterly winds near Point Atkinson during the onset of the surge but after the peak surge has passed the winds shifted to westerly.
 A strong westerly wind would cause an elevated sea surface height at Point Atkinson as water piles against the coast, explaining the delay in peak surge when only atmospheric forcing is used since the winds shift to westerly after the peak surge has passed.
@@
 Previous simulations by \citet{murty1995storm} suggest that the inverse barometer effect and the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.
-A detailed assessment of the effect of the local atmospheric conditions and winds should be performed with high resolution atmospheric forcing data.
 
@@
 Although the storm was very strong and the observed residual at Point Atkinson reached 84 cm, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area \citep{forseth2006adaptation}.
-This event has been used to examine the propagation of the surge throughout the domain and spatial variations in the surge amplitude.
+Because of the high winds, this event has been used to examine the propagation of the surge throughout the domain and spatial variations in the surge amplitude.
 First, an assessment of the accuracy of the hindcast is provided.
@@
 Typically, the modelled surge was too early, in contrast to the later predictions in the Feb 4, 2006 case.
-Error and skill scores for the modelled water level and the modelled residual are provided in Table \ref{tab:dec2006stat}.
-All Willmott skill scores rank above 0.89 and the $\gamma^2$ measure also indicates good agreement between model and observations.
+All Willmott skill scores rank above 0.89 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:dec2006stat}).
 The modelled residuals are on average too low and have a lower skill score than the Feb 2006 case.
@@
 The observed residual at Tofino, BC (a point near by but outside of the domain) is up to 10 cm lower than the observed residual at Neah Bay, WA (a point near by but inside of the domain).
-Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the case.
+Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the cause.
 
@@
 
-An evaluation of the spatial distribution of the surge in a simulation without atmospheric forcing is provided in Figure \ref{fig:spatial_sshonly}.
-In this case, the surge amplitude across the Strait of Georgia is uniform with no pronounced difference between the mainland side and Vancouver Island side.
+In a simulation without atmospheric forcing, the surge amplitude across the Strait of Georgia is uniform with no pronounced difference between the mainland side and Vancouver Island side (Figure \ref{fig:spatial_sshonly}).
 While the surge entering the domain from the Pacific Ocean is the leading order contribution, the winds may induce localized changes to the surge amplitude.
@@
 In addition, large wind speeds over 15 m s$^{-1}$ persisted for over 9 hours.
-Yet, the surge associated with this event is rather small when compared to the previous two cases as shown in Figure \ref{fig:nov2009} where we have plotted the wind speed at Point Atkinson, sea surface height anomaly forcing at the open boundaries, and observed and modelled residuals at Point Atkinson.
+Yet, the surge associated with this event is rather small when compared to the previous two cases (Figure \ref{fig:nov2009}).
 Two significant surge events are noted, the first occurring on Nov 19 with an observed amplitude of about 50 cm at Point Atkinson, and a second, larger surge of 62.5 cm on Nov 20.
@@
 This event observed a higher surge at Victoria than the Dec 2006 event, further highlighting the role that winds plays in setting up local variations in the surge.
-In this case, the model results do not reflect a higher surge at Victoria (46.1 cm versus 50.5 cm at Point Atkinson), perhaps due to an underestimate in the wind speed and some discrepancies in the wind direction of the atmospheric forcing.
+In this case, the model results do not reflect a higher surge at Victoria (46.1 cm versus 50.5~cm at Point Atkinson), perhaps due to an underestimate in the wind speed and some discrepancies in the wind direction of the atmospheric forcing.
 
@@
 
-Once again, an evaluation of the model skill for this hindcast is provided in Table \ref{tab:nov2009stat}.
-The modelled water levels and residuals are too high, except at Victoria.
+The modelled water levels and residuals are too high for this hindcast, except at Victoria (Table \ref{tab:nov2009stat}).
 Although Victoria has the smallest mean error in both the water level and residuals, its Willmott skill score and $\gamma^2$ values rank low when compared to the other stations.
@@
 
-Modelled and observed residuals for this simulation are displayed in Figure \ref{fig:dec2012_surge} where it is apparent that both of the simulations agree well with the observations.
+Both of the simulations agree well with the observations (\ref{fig:dec2012_surge}).
 Willmott skill scores for both modelled residuals are above 0.92 at these four locations.
@@
 Since both of these products provide the same level of accuracy for storm surge hindcasts, in the future we will use the high resolution atmospheric model to produce real-time storm surge predictions, even though most of out testing was performed with the low resolution model.
-
+Additionally, the sea surface height anomaly from Neah Bay, forecasted by the National Oceanic and Atmospheric Administration (NOAA), will be used in future real-time water level predictions during the storm surge season.
 Several factors considered in other storm surge modelling studies have been neglected in these simulations.
@@
 As such, efforts to include wetting and drying will be considered in the future.
-Additionally, the sea surface height anomaly from Neah Bay, forecasted by the National Oceanic and Atmospheric Administration (NOAA), will be used in future real-time water level predictions during the storm surge season.
 Lastly, flooding effects due to wave run up is left for a more detailed inundation study with higher resolution in the near-shore areas.
@@
 This section details the procedure used to tune the tidal forcing for a good match with observations in the Strait of Georgia.
-Originally, the eight most important tidal elevation and currents harmonics were extracted from a tide prediction tool called Webtide \citep{foreman2000webtide} at the western Juan de Fuca boundary.
+Originally, the eight most important tidal constituents were extracted from a tide prediction tool called Webtide \citep{foreman2000webtide} at the western Juan de Fuca boundary.
 At the northern boundary,  M$_2$ and K$_1$ tidal harmonics for both elevation and currents were taken from \citet{thomson1980johnstone}, as well as elevation harmonics for O$_1$ and S$_2$.
-The remaining constituents were extrapolated from Webtide based on the ratio between M$_2$/S$_2$ amplitudes and phase differences for the semi-diurnal constituents and K$_1$/O$_1$ for the diurnal constituents.
+The remaining constituents were extrapolated from Webtide based on the ratio between S$_2$/M$_2$ amplitudes and phase differences for the semi-diurnal constituents and O$_1$/K$_1$ for the diurnal constituents.
 However, these forcing conditions did not result in a good match between observed and modelled tidal elevation harmonics in the Strait of Georgia, the main region of interest (although the matches in the Strait of Juan de Fuca were adequate).
@@
 Finally, the remaining constituents were added, starting with O$_1$, then S$_2$, and then in pairs P$_1$/N$_2$ and Q$_1$/K$_2$.
-Adjustments were made if the addition of new constituents hindered performance of the main M$_2$ and K$_1$ components.
+Final adjustments were made by slightly varying amplitude and phase, evaluating the changes to M$_2$, K$_1$ and the ratios S$_2$/M$_2$, N$_2$/M$_2$, K$_2$/M$_2$, O$_1$/K$_1$, P$_1$/K$_1$, Q$_1$/K$_1$ and then choosing new values to move closer to observations by linear interpolation.
 

### 2014-12-03 RoughWork/manuscript.tex
Fixed up citations

@@
 This early work pointed to an important connection between the model's bottom friction parametrization and the location and amplitude of the M$_2$ amphidrome.
-They found that a local increase in the bottom friction coefficient, especially in the area of Haro Strait, was necessary in order to adequately reproduce the observed M$_2$ tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}.
+They found that a local increase in the bottom friction coefficient, typically from 0.002 to 0.004, especially in the area of Haro Strait, was necessary in order to adequately reproduce the observed M$_2$ tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}.
 Without local adjustments to the bottom friction their M$_2$ amplitudes were too high in the Strait of Juan de Fuca and too low in the Strait of Georgia.
@@
 This effect in combination with strong winds pushing water up against the coast can cause flooding, particularly if the storm occurs during an unusually high tide.
-There can also be a small contribution to increased sea level due to the thermal expansion of water during warmer years set by the El Nino Southern Oscillation (ENSO) \citep{abeys2011extreme}.
-%discuss this as a Kelvin wave
+There is also a small contribution to increased sea level due to positive sea surface height anomalies propagating northward from the equator along the eastern Pacific coast during El Nino years \citep{strub2002altimeter}.
 
@@
 Compared to eastern Canada \citep{gray1984preliminary, bernier2006predicting,bobanovic2006forecasting, bernier2007tide, bernier2010tide}, relatively little has been published on storm surge modelling on the west coast of Canada in recent years.
-A statistical overview of past surge events in British Columbia waters suggests an increase in extreme water level events during warm ENSO periods \citep{abeys2011extreme}.
+However, a storm surge forecasting program is currently operating over a domain that covers the entire British Columbia coastline \citep{StormBC}.
+Further, a statistical overview of past surge events in British Columbia waters suggests an increase in extreme water level events during warm El Nino periods \citep{abeys2011extreme}.
 
-%outline
 This article presents the development of a three-dimensional numerical ocean model for the Strait of Georgia and its accuracy in simulating storm surge hindcasts.
@@
 
-%A section about how we configured the model for SoG 
 We have used the Nucleus for European Modelling of the Ocean (NEMO) framework in its regional configuration \citep{madec2012nemo} to develop an ocean model for the Strait of Georgia and Salish Sea.
@@
 
-Sea surface height at the mouth of Juan de Fuca was set using values from the Tofino tide gauge which is outside of, but close to, the western boundary of the model domain.
+Sea surface height at the mouth of Juan de Fuca was set using values from the Tofino tide gauge \citep{DFOObservations} which is outside of, but close to, the western boundary of the model domain.
 A monthly climatology was produced using daily averages from 2000-2010, binning them by month, averaging and setting the yearly mean to zero.
@@
 The modelled residual is calculated as the difference in model output between a simulation with all forcing conditions and a simulation forced only with tides and rivers.
-Water level observations are taken from Fisheries and Oceans Canada (citation?) and are used to calculate the observed residual, defined as the observed water level minus tidal predictions.
+Water level observations are taken from the Canadian Tides and Water Level Data Archive \citep{DFOObservations} and are used to calculate the observed residual, defined as the observed water level minus tidal predictions.
 Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this case, from the year prior to the surge event of interest.
@@
 
-Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow regions due to nonlinear effects and bottom friction, is presented for this hindcast (Figure \ref{fig:tidesurge}).
+Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow regions due to nonlinear effects and bottom friction \citep{bernier2007tide}, is presented for this hindcast (Figure \ref{fig:tidesurge}).
 The most significant differences occur at Point Atkinson where the maximum of the surge-only case is 8.7 cm higher than maximum modelled residual, about 10\% of the total surge amplitude.

### 2014-12-03 RoughWork/manuscript.tex
Merged tables into one

@@
 
-The maximum water level and maximum residual are important factors in storm surge hindcasts but a better indicator of overall model performance is provided by a group of statistical measures, including the mean absolute error, the observed mean, the modelled mean, $\gamma^2$, and the Willmott skill score for both the water level and residuals (Table \ref{tab:feb2006stat}).
+The maximum water level and maximum residual are important factors in storm surge hindcasts but a better indicator of overall model performance is provided by a group of statistical measures, including the mean absolute error, the observed mean, the modelled mean, $\gamma^2$, and the Willmott skill score for both the water level and residuals (Table \ref{tab:statistics}).
 The statistics are calculated with hourly output over the three-day period displayed in Figure \ref{fig:feb2006}.
@@
 \hline 
+Feb 2006       &                &                      &                      &            &        &                &                        &                        &           &           \\
+\hline 
 Point Atkinson & 10.5           & 3.46                 & 3.54                 &   0.012    & 0.996  & 6.2            &  0.273                 & 0.309                  & 0.065     & 0.979     \\
@@
 \hline 
+Nov 2006       &                &                      &                      &            &        &                &                        &                        &           &           \\
+\hline 
+Point Atkinson &  9.8           &  3.42                & 3.43                 &   0.034    & 0.991  &  7.7           &  0.290                 & 0.247                  &  0.128     & 0.950 \\
+Victoria       &  9.1           &  2.16                & 2.11                 &   0.070    & 0.978  &  6.2           &  0.222                 & 0.202                  &  0.171     & 0.946 \\
+Patricia Bay   & 10.0           &  2.57                & 2.53                 &   0.049    & 0.985  &  6.6           &  0.235                 & 0.203                  &  0.144     & 0.947 \\
+Campbell River & 10.7           &  3.20                & 3.26                 &   0.037    & 0.989  &  6.2           &  0.254                 & 0.236                  &  0.140     & 0.952 \\
+\hline 
+Dec 2006       &                &                      &                      &            &        &                &                        &                        &           &           \\
+\hline 
+Point Atkinson & 10.5           &  3.45                & 3.47                 &   0.025    & 0.994  & 11.1           &  0.212                 & 0.186                  &  0.275     & 0.891 \\
+Victoria       &  9.6           &  2.23                & 2.16                 &   0.042    & 0.984  &  7.9           &  0.190                 & 0.156                  &  0.223     & 0.911 \\
+Patricia Bay   &  9.2           &  2.63                & 2.58                 &   0.031    & 0.990  &  9.5           &  0.193                 & 0.149                  &  0.260     & 0.889 \\
+Campbell River & 10.6           &  3.23                & 3.28                 &   0.030    & 0.992  &  9.4           &  0.191                 & 0.163                  &  0.247     & 0.905 \\
+\hline  
+Nov 2009       &                &                      &                      &            &        &                &                        &                        &           &           \\
+\hline 
+Point Atkinson & 14.0           &  3.44                & 3.55                 &   0.012    & 0.995  &  6.1           &  0.313                 & 0.331                  &  0.237     & 0.914 \\
+Victoria       &  9.9           &  2.26                & 2.24                 &   0.023    & 0.994  &  4.8           &  0.323                 & 0.293                  &  0.245     & 0.899 \\
+Patricia Bay   & 11.2           &  2.60                & 2.65                 &   0.020    & 0.993  &  6.0           &  0.277                 & 0.290                  &  0.272     & 0.901 \\
+Campbell River & 13.5           &  3.22                & 3.27                 &   0.019    & 0.995  &  4.9           &  0.326                 & 0.314                  &  0.217     & 0.924 \\
+\hline 
+Dec 2012-CGRF &                &                      &                      &            &        &                &                        &                        &           &           \\
+\hline 
+Point Atkinson &  9.1           &  3.36                & 3.40                 &   0.006    & 0.998  &  5.0           &  0.300                 & 0.290                  &  0.161     & 0.949 \\
+Victoria       & 14.6           &  2.15                & 2.08                 &   0.039    & 0.988  &  4.8           &  0.309                 & 0.274                  &  0.150     & 0.937 \\
+Patricia Bay   & 14.2           &  2.55                & 2.51                 &   0.026    & 0.992  &  4.7           &  0.300                 & 0.274                  &  0.140     & 0.945 \\
+Campbell River & 11.6           &  3.06                & 3.09                 &   0.013    & 0.996  &  5.7           &  0.241                 & 0.278                  &  0.182     & 0.921 \\
+\hline 
+Dec 2012-HRDPS &                &                      &                      &            &        &                &                        &                        &           &           \\
+\hline 
+Point Atkinson &  8.8           &  3.35                & 3.39                 &   0.006    & 0.998  &  5.4           &  0.300                 & 0.278                  &  0.163     & 0.943 \\
+Victoria       & 14.3           &  2.15                & 2.09                 &   0.038    & 0.988  &  4.6           &  0.309                 & 0.277                  &  0.145     & 0.941 \\
+Patricia Bay   & 14.0           &  2.55                & 2.52                 &   0.025    & 0.993  &  4.5           &  0.300                 & 0.278                  &  0.130     & 0.953 \\
+Campbell River & 11.6           &  3.06                & 3.09                 &   0.014    & 0.993  &  5.6           &  0.241                 & 0.275                  &  0.183     & 0.924 \\
+\hline 
 \end{tabular}
-\caption{Evaluation of model performance for the Feb 2006 simulation.
-Mean absolute error ($\bar{e}$), observed mean (subscript o), modelled mean (subscript m), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and residuals ($\delta$) during the period displayed in Figure \ref{fig:feb2006} from Feb 3, 2006 to Feb 6, 2006.}
-\label{tab:feb2006stat} 
+\caption{Evaluation of model performance for all hindcasts presented.
+Mean absolute error ($\bar{e}$), observed mean (subscript o), modelled mean (subscript m), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and residuals ($\delta$).
+Statisitics are calculated over the time period displayed in esch simulation's corresponding figure.}
+\label{tab:statistics} 
 \end{table} 
@@
 The surge amplitude was lower at the other stations, reaching no more than 68 cm.
-The model performs reasonably well in this hindcast; Willmott skill scores are greater than 0.94 (Table \ref{tab:nov2006stat}).
+The model performs reasonably well in this hindcast; Willmott skill scores are greater than 0.94 (Table \ref{tab:statistics}).
 The best match in maximum surge occurred at Campbell River with a 6.2 cm underestimate by the model.
@@
 
-\begin{table}[h]
-\centering 
-\begin{tabular}{|l |c c c c c | c c c c c|} 
-\hline 
-& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
-\hline 
-Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
-\hline 
-Point Atkinson &  9.8           &  3.42                & 3.43                 &   0.034    & 0.991  &  7.7           &  0.290                 & 0.247                  &  0.128     & 0.950 \\
-Victoria       &  9.1           &  2.16                & 2.11                 &   0.070    & 0.978  &  6.2           &  0.222                 & 0.202                  &  0.171     & 0.946 \\
-Patricia Bay   & 10.0           &  2.57                & 2.53                 &   0.049    & 0.985  &  6.6           &  0.235                 & 0.203                  &  0.144     & 0.947 \\
-Campbell River & 10.7           &  3.20                & 3.26                 &   0.037    & 0.989  &  6.2           &  0.254                 & 0.236                  &  0.140     & 0.952 \\
-\hline 
-\end{tabular}
-\caption{Evaluation of model performance for the Nov 2006 hindcast. Statistics were calculated over the period Nov 14, 2006 to Nov 17, 2006.}
-\label{tab:nov2006stat} 
-\end{table}
+%\begin{table}[h]
+%\centering 
+%\begin{tabular}{|l |c c c c c | c c c c c|} 
+%\hline 
+%& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
+%\hline 
+%Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
+%\hline 
+%Point Atkinson &  9.8           &  3.42                & 3.43                 &   0.034    & 0.991  &  7.7           &  0.290                 & 0.247                  &  0.128     & 0.950 \\
+%Victoria       &  9.1           &  2.16                & 2.11                 &   0.070    & 0.978  &  6.2           &  0.222                 & 0.202                  &  0.171     & 0.946 \\
+%Patricia Bay   & 10.0           &  2.57                & 2.53                 &   0.049    & 0.985  &  6.6           &  0.235                 & 0.203                  &  0.144     & 0.947 \\
+%Campbell River & 10.7           &  3.20                & 3.26                 &   0.037    & 0.989  &  6.2           &  0.254                 & 0.236                  &  0.140     & 0.952 \\
+%\hline 
+%\end{tabular}
+%\caption{Evaluation of model performance for the Nov 2006 hindcast. Statistics were calculated over the period Nov 14, 2006 to Nov 17, 2006.}
+%\label{tab:nov2006stat} 
+%\end{table}
 
@@
 Typically, the modelled surge was too early, in contrast to the later predictions in the Feb 4, 2006 case.
-All Willmott skill scores rank above 0.89 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:dec2006stat}).
+All Willmott skill scores rank above 0.89 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}).
 The modelled residuals are on average too low and have a lower skill score than the Feb 2006 case.
@@
 
-\begin{table}[h]
-\centering 
-\begin{tabular}{|l |c c c c c | c c c c c|} 
-\hline 
-& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
-\hline 
-Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
-\hline 
-Point Atkinson & 10.5           &  3.45                & 3.47                 &   0.025    & 0.994  & 11.1           &  0.212                 & 0.186                  &  0.275     & 0.891 \\
-Victoria       &  9.6           &  2.23                & 2.16                 &   0.042    & 0.984  &  7.9           &  0.190                 & 0.156                  &  0.223     & 0.911 \\
-Patricia Bay   &  9.2           &  2.63                & 2.58                 &   0.031    & 0.990  &  9.5           &  0.193                 & 0.149                  &  0.260     & 0.889 \\
-Campbell River & 10.6           &  3.23                & 3.28                 &   0.030    & 0.992  &  9.4           &  0.191                 & 0.163                  &  0.247     & 0.905 \\
-\hline 
-\end{tabular}
-\caption{Evaluation of model performance for the Dec 2006 hindcast. Statistics were calculated over the period Dec 14, 2006 to Dec 17, 2006.}
-\label{tab:dec2006stat} 
-\end{table}
+%\begin{table}[h]
+%\centering 
+%\begin{tabular}{|l |c c c c c | c c c c c|} 
+%\hline 
+%& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
+%\hline 
+%Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
+%\hline 
+%Point Atkinson & 10.5           &  3.45                & 3.47                 &   0.025    & 0.994  & 11.1           &  0.212                 & 0.186                  &  0.275     & 0.891 \\
+%Victoria       &  9.6           &  2.23                & 2.16                 &   0.042    & 0.984  &  7.9           &  0.190                 & 0.156                  &  0.223     & 0.911 \\
+%Patricia Bay   &  9.2           &  2.63                & 2.58                 &   0.031    & 0.990  &  9.5           &  0.193                 & 0.149                  &  0.260     & 0.889 \\
+%Campbell River & 10.6           &  3.23                & 3.28                 &   0.030    & 0.992  &  9.4           &  0.191                 & 0.163                  &  0.247     & 0.905 \\
+%\hline 
+%\end{tabular}
+%\caption{Evaluation of model performance for the Dec 2006 hindcast. Statistics were calculated over the period Dec 14, 2006 to Dec 17, 2006.}
+%\label{tab:dec2006stat} 
+%\end{table}
 
@@
 
-The modelled water levels and residuals are too high for this hindcast, except at Victoria (Table \ref{tab:nov2009stat}).
+The modelled water levels and residuals are too high for this hindcast, except at Victoria (Table \ref{tab:statistics}).
 Although Victoria has the smallest mean error in both the water level and residuals, its Willmott skill score and $\gamma^2$ values rank low when compared to the other stations.
@@
 
-\begin{table}[h]
-\centering 
-\begin{tabular}{|l |c c c c c | c c c c c|} 
-\hline 
-& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
-\hline 
-Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
-\hline 
-Point Atkinson & 14.0           &  3.44                & 3.55                 &   0.012    & 0.995  &  6.1           &  0.313                 & 0.331                  &  0.237     & 0.914 \\
-Victoria       &  9.9           &  2.26                & 2.24                 &   0.023    & 0.994  &  4.8           &  0.323                 & 0.293                  &  0.245     & 0.899 \\
-Patricia Bay   & 11.2           &  2.60                & 2.65                 &   0.020    & 0.993  &  6.0           &  0.277                 & 0.290                  &  0.272     & 0.901 \\
-Campbell River & 13.5           &  3.22                & 3.27                 &   0.019    & 0.995  &  4.9           &  0.326                 & 0.314                  &  0.217     & 0.924 \\
-\hline 
-\end{tabular}
-\caption{Evaluation of model performance for the Nov 2009 hindcast. Statistics were generated over the period Nov 17, 2009 to Nov 22, 2009.}
-\label{tab:nov2009stat} 
-\end{table}
+%\begin{table}[h]
+%\centering 
+%\begin{tabular}{|l |c c c c c | c c c c c|} 
+%\hline 
+%& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
+%\hline 
+%Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
+%\hline 
+%Point Atkinson & 14.0           &  3.44                & 3.55                 &   0.012    & 0.995  &  6.1           &  0.313                 & 0.331                  &  0.237     & 0.914 \\
+%Victoria       &  9.9           &  2.26                & 2.24                 &   0.023    & 0.994  &  4.8           &  0.323                 & 0.293                  &  0.245     & 0.899 \\
+%Patricia Bay   & 11.2           &  2.60                & 2.65                 &   0.020    & 0.993  &  6.0           &  0.277                 & 0.290                  &  0.272     & 0.901 \\
+%Campbell River & 13.5           &  3.22                & 3.27                 &   0.019    & 0.995  &  4.9           &  0.326                 & 0.314                  &  0.217     & 0.924 \\
+%\hline 
+%\end{tabular}
+%\caption{Evaluation of model performance for the Nov 2009 hindcast. Statistics were generated over the period Nov 17, 2009 to Nov 22, 2009.}
+%\label{tab:nov2009stat} 
+%\end{table}
 

### 2014-12-04 RoughWork/manuscript.tex
Reworked overview of storm surge section

@@
 
-Model results have been compared with observations by calculating both the total modelled water level and the modelled residual.
+Model results have been compared with observations by calculating both the total modelled water level and the modelled residual at four permanent tidal measuring stations; Point Atkinson, Victoria, Patricia Bay, and Campbell River (locations indicated in Figure \ref{fig:domain}).
 The modelled residual is calculated as the difference in model output between a simulation with all forcing conditions and a simulation forced only with tides and rivers.
@@
 Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this case, from the year prior to the surge event of interest.
-Additionally, several simulations with varying forcing conditions such as removing atmospheric forcing or removing the sea surface height anomaly at the open boundaries have been completed.
-
-Several significant surge or wind events have been considered in detail, as well as an array of hindcasts to assess model ability in a statistical measure.
-These cases have been used to assess the model's skill in reproducing water level height and surge amplitude and timing during storm events.
-The assessment includes comparisons at four permanent tidal measuring stations; Point Atkinson, Victoria, Patricia Bay, and Campbell River (locations indicated in Figure \ref{fig:domain}).
-In addition, the contribution to the surge due to several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed.
-We have also examined the spatial variability of the surge as it propagates through the domain and have considered a case with a strong wind event but no significant observed surge.
-Finally, comparisons between hindcasts using high resolution and low resolution atmospheric forcing have been completed.
 
@@
 \end{equation} 
-where $M_i' = M_i-\bar{O}$ and $O_i'=O_i-\bar{O}$ and $\bar{O}$ is the mean of the observations. This score is bounded by zero and one. Values close to one indicate good model skill.
+where $M_i' = M_i-\bar{O}$ and $O_i'=O_i-\bar{O}$ and $\bar{O}$ is the mean of the observations.
+This score is bounded by zero and one.
+Values close to one indicate good model skill.
+These three statisitics, along with the mean observed values and mean modelled valuesm are reported in Table \ref{tab:statistics} for each hindcast.
+Further, comparisons between the maximum observed and modelled water level and residual are typically reported in the text.
 
-\subsection{Example Storm Surge Hindcasts}
+The five hindcast simulations have been used to assess the model's skill and to learn more about how the storm surge propagates through the Salish Sea domain.
+First, a general assessment of the model's skill is presented in two hindcasts during Feb 2006 and Nov 2006 in section \ref{sec:hind}.
+The Feb 2006 case is also used to determine the importance of modelling tidal forcing with the surge propogation.
+Second, the contribution to the surge due to several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed for the Feb 2006 case in section \ref{sec:factors}.
+Third, the spatial variability of the surge as it propagates through the domain is examined for a Dec 2006 storm surge in section \ref{sec:spatial}.
+Fourth, a high wind event that did not result in a large surge or flooding for a Nov 2009 case is discussed in section \ref{sec:wind}.
+Finally, comparisons between hindcasts using high resolution and low resolution atmospheric forcing is provided in section \ref{sec:res}.
+
+\subsection{Example Storm Surge Hindcasts}\label{sec:hind}
 
@@
 Feb 2006       &                &                      &                      &            &        &                &                        &                        &           &           \\
-\hline 
 Point Atkinson & 10.5           & 3.46                 & 3.54                 &   0.012    & 0.996  & 6.2            &  0.273                 & 0.309                  & 0.065     & 0.979     \\
@@
 Nov 2006       &                &                      &                      &            &        &                &                        &                        &           &           \\
-\hline 
 Point Atkinson &  9.8           &  3.42                & 3.43                 &   0.034    & 0.991  &  7.7           &  0.290                 & 0.247                  &  0.128     & 0.950 \\
@@
 Dec 2006       &                &                      &                      &            &        &                &                        &                        &           &           \\
-\hline 
 Point Atkinson & 10.5           &  3.45                & 3.47                 &   0.025    & 0.994  & 11.1           &  0.212                 & 0.186                  &  0.275     & 0.891 \\
@@
 Nov 2009       &                &                      &                      &            &        &                &                        &                        &           &           \\
-\hline 
 Point Atkinson & 14.0           &  3.44                & 3.55                 &   0.012    & 0.995  &  6.1           &  0.313                 & 0.331                  &  0.237     & 0.914 \\
@@
 Dec 2012-CGRF &                &                      &                      &            &        &                &                        &                        &           &           \\ 
-\hline 
 Point Atkinson &  9.1           &  3.36                & 3.40                 &   0.006    & 0.998  &  5.0           &  0.300                 & 0.290                  &  0.161     & 0.949 \\
@@
 Dec 2012-HRDPS &                &                      &                      &            &        &                &                        &                        &           &           \\
-\hline 
 Point Atkinson &  8.8           &  3.35                & 3.39                 &   0.006    & 0.998  &  5.4           &  0.300                 & 0.278                  &  0.163     & 0.943 \\
@@
 
-\subsection{Factors Contributing to Storm Surges}
+\subsection{Factors Contributing to Storm Surges}\label{sec:factors}
 
@@
 
-\subsection{Spatial Extent of Surge}
+\subsection{Spatial Extent of Surge}\label{sec:spatial}
 On Dec 15, 2006, a major storm with wind gusts over 120 km hr$^{-1}$ caused significant damage to many coastal communities on the Pacific Northwest.
@@
 
-\subsection{Strong Wind Event}
+\subsection{Strong Wind Event}\label{sec:wind}
 Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered.
@@
 
-\subsection{High Resolution Atmospheric Forcing}
+\subsection{High Resolution Atmospheric Forcing}\label{sec:res}
 The above comparison between the modelled and observed winds at Point Atkinson suggests that the atmospheric forcing product (CGRF) may not be adequate to produce accurate storm surge hindcasts since the wind speeds are substantially lower than observed.

### 2014-12-04 RoughWork/manuscript.tex
Consistent labels
new table with averages

@@
 Additionally, since the sea surface height is calculated about a zero reference level, the modelled water level is determined by adding a long term mean sea level at the specified location to the modelled sea surface height.
+All times are reported in UTC.
 
@@
 Values close to one indicate good model skill.
-These three statisitics, along with the mean observed values and mean modelled valuesm are reported in Table \ref{tab:statistics} for each hindcast.
+These three statistics, along with the mean observed values and mean modelled values are reported in Table \ref{tab:statistics} for each hindcast.
 Further, comparisons between the maximum observed and modelled water level and residual are typically reported in the text.
@@
 First, a general assessment of the model's skill is presented in two hindcasts during Feb 2006 and Nov 2006 in section \ref{sec:hind}.
-The Feb 2006 case is also used to determine the importance of modelling tidal forcing with the surge propogation.
+The Feb 2006 case is also used to determine the importance of modelling tidal forcing with the surge propagation.
 Second, the contribution to the surge due to several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed for the Feb 2006 case in section \ref{sec:factors}.
@@
 Fourth, a high wind event that did not result in a large surge or flooding for a Nov 2009 case is discussed in section \ref{sec:wind}.
-Finally, comparisons between hindcasts using high resolution and low resolution atmospheric forcing is provided in section \ref{sec:res}.
+Finally, comparisons between hindcasts using high resolution (HRDPS) and low resolution (CGRF) atmospheric forcing is provided in section \ref{sec:res}.
 
@@
 Nov 2009       &                &                      &                      &            &        &                &                        &                        &           &           \\
-Point Atkinson & 14.0           &  3.44                & 3.55                 &   0.012    & 0.995  &  6.1           &  0.313                 & 0.331                  &  0.237     & 0.914 \\
-Victoria       &  9.9           &  2.26                & 2.24                 &   0.023    & 0.994  &  4.8           &  0.323                 & 0.293                  &  0.245     & 0.899 \\
-Patricia Bay   & 11.2           &  2.60                & 2.65                 &   0.020    & 0.993  &  6.0           &  0.277                 & 0.290                  &  0.272     & 0.901 \\
-Campbell River & 13.5           &  3.22                & 3.27                 &   0.019    & 0.995  &  4.9           &  0.326                 & 0.314                  &  0.217     & 0.924 \\
+Point Atkinson & 14.0           &  3.43                & 3.54                 &   0.012    & 0.995  &  5.8           &  0.295                 & 0.316                  &  0.200     & 0.926 \\
+Victoria       &  9.9           &  2.25                & 2.23                 &   0.024    & 0.988  &  4.5           &  0.312                 & 0.280                  &  0.211     & 0.911 \\
+Patricia Bay   & 10.3           &  2.59                & 2.64                 &   0.019    & 0.997  &  5.6           &  0.265                 & 0.275                  &  0.230     & 0.918 \\
+Campbell River & 13.6           &  3.22                & 3.26                 &   0.019    & 0.991  &  4.5           &  0.313                 & 0.297                  &  0.182     & 0.934 \\
 \hline 
@@
 \hline 
+\hline
+Averages       &                &                      &                      &            &        &                &                        &                        &           &         \\
+Point Atkinson &	10.8	& 3.42	& 3.48	& 0.018	& 0.995	& 7.2	& 0.274	& 0.270	& 0.166	 & 0.939 \\
+Victoria       &	10.2	& 2.19	& 2.15	& 0.041	& 0.986	& 5.9	& 0.254	& 0.238	& 0.168	 & 0.935 \\
+Patricia Bay   &	10.6	& 2.59	& 2.58	& 0.030	& 0.991	& 6.6	& 0.243	& 0.236	& 0.169	 & 0.934 \\
+Campbell River &	11.9	& 3.18	& 3.22	& 0.028	& 0.992	& 6.3	& 0.260	& 0.255	& 0.166  & 0.938 \\
+\hline
 \end{tabular}
@@
 Mean absolute error ($\bar{e}$), observed mean (subscript o), modelled mean (subscript m), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and residuals ($\delta$).
-Statisitics are calculated over the time period displayed in esch simulation's corresponding figure.}
+Statistics are calculated over the time period displayed in each simulation's corresponding figure (three or four days).
+For the Nov 2006 case, the statistics were calculated over a three day period.
+The average values were calculated over all of the hindcasts, excluding the Dec 2012 HRDPS simulation.
+}
 \label{tab:statistics} 
@@
 The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in a slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
-Additionally, the surge amplitude decreases by a few centimetres if the inverse barometer effect due to atmospheric pressure is not included.
+Additionally, the surge amplitude decreases by a few centimetres if the inverse barometer effect due to atmospheric pressure is not included, in contrast to other findings that reported the inverse barometer effect contributes up to $^1/_3$ of the total surge amplitude \citep{murty1995storm}.
 
@@
 \includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
-\caption{Comparison of the modelled residual for four simulations in Feb 2006: A simulation with all forcing (solid black), a simulation without atmospheric forcing (solid grey), a simulation without the sea surface height anomaly (dashed black), and a simulation with only wind forcing, i.e.\ no sea surface height anomaly or pressure (dashed grey).
+\caption{Comparison of the modelled residual for four simulations in Feb 2006: A simulation with all forcing (solid black), a simulation without atmospheric forcing (dashed green), a simulation without the sea surface height anomaly (solid red), and a simulation with only wind forcing, i.e.\ no sea surface height anomaly or pressure (dashed blue).
 }
@@
 Significant flooding occurred later, on Dec 17, 2012, where the observed surge amplitude at Point Atkinson was 68.3 cm.
-Even though the low resolution model significantly underpredicts the wind speed at some locations, it performs comparable to forcing with high resolution products.
+Even though the low resolution model significantly under predicts the wind speed at some locations, it performs comparable to forcing with high resolution products.
 This result further demonstrates that the local wind is not the most important factor for setting up storm surges over this domain.
@@
 
-Hindcasts of five significant events in February 2006, November 2006, December, November 2009, and December 2012 were simulated.
-In general, the model performs well at reproducing the observed water level heights and residuals, with Willmott skill scores greater than 0.88 in all cases.
+Hindcasts of five significant events in February 2006, November 2006, December 2006, November 2009, and December 2012 were simulated.
+In general, the model performs well at reproducing the observed water level heights and residuals, with Willmott skill scores greater than 0.88 in all cases and generally lower skill scores for the residuals.
+Further, in a single simulation, the skill scores are fairly consistent between the four stations examined.
 We have also assessed the relative importance of different forcing factors contributing to the storm surge.
@@
 Although the high resolution model delivers fine scale detail and more accurate local wind patterns, it did not provide any improvement to the surge hindcast.
-Since both of these products provide the same level of accuracy for storm surge hindcasts, in the future we will use the high resolution atmospheric model to produce real-time storm surge predictions, even though most of out testing was performed with the low resolution model.
+Since both of these products provide the same level of accuracy for storm surge hindcasts, in the future we will use the high resolution atmospheric model to produce real-time storm surge predictions, even though most of our testing was performed with the low resolution model.
 Additionally, the sea surface height anomaly from Neah Bay, forecasted by the National Oceanic and Atmospheric Administration (NOAA), will be used in future real-time water level predictions during the storm surge season.
+
 Several factors considered in other storm surge modelling studies have been neglected in these simulations.

### 2014-12-05 RoughWork/manuscript.tex
final search for typos

@@
 They found that a local increase in the bottom friction coefficient, typically from 0.002 to 0.004, especially in the area of Haro Strait, was necessary in order to adequately reproduce the observed M$_2$ tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}.
-Without local adjustments to the bottom friction their M$_2$ amplitudes were too high in the Strait of Juan de Fuca and too low in the Strait of Georgia.
+Without local adjustments to the bottom friction their M$_2$ amplitudes were too low and the M$_2$ tidal phases were too large in the Strait of Georgia.
 Other studies over a similar domain, such as the barotropic finite element model of \citet{foreman1995tidal}, have avoided local increases in bottom friction.
@@
 The watersheds included in our model were Fraser (which represents approximately 44\% of the freshwater input into our domain), Skagit (12\%), East Vancouver Island (North and South) (12\%), Howe (7\%), Bute (7\%), Puget (6\%), Juan de Fuca (5\%), Jervis (4\%) and Toba (3\%).
-The monthly flow from each river was input as a point source from 0-3 metres at the model point closest to the mouth of each river.
+The monthly flow from each river was input as a point source from 0 to 3 metres at the model point closest to the mouth of each river.
 Incoming water was assumed to be fresh and at surface air temperature.
@@
 
-In general, there is a very good agreement between the observed and modelled residuals \ref{fig:feb2006}.
+In general, there is a very good agreement between the observed and modelled residuals (Figure \ref{fig:feb2006}).
 The most noticeable difference is a 4-6 hour delay in the model's timing of the maximum residual except at Victoria where the maximum residual is in sync with observations.
@@
 
-Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented for the Feb 2006 storm surge (Figure \ref{fig:factors}).Four simulations were done: First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly.
+Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented for the Feb 2006 storm surge (Figure \ref{fig:factors}).
+Four simulations were done: First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly.
 Second, a simulation with all forcing except the atmospheric forcing.
@@
 The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in a slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
-Additionally, the surge amplitude decreases by a few centimetres if the inverse barometer effect due to atmospheric pressure is not included, in contrast to other findings that reported the inverse barometer effect contributes up to $^1/_3$ of the total surge amplitude \citep{murty1995storm}.
+Additionally, the surge amplitude decreases by a few centimetres if the inverse barometer effect due to atmospheric pressure is not include.
+Other findings have reported the inverse barometer effect contributes up to $^1/_3$ of the total surge amplitude \citep{murty1995storm}, but only a minor effect is noticeable here.
 
@@
 In a model domain that does not include the Pacific Ocean, this effect must be included as a boundary condition.
-Previous simulations by \citet{murty1995storm} suggest that the inverse barometer effect and the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.
+Previous simulations by \citet{murty1995storm} suggest that the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.
 
@@
 Point Atkinson observed a maximum surge amplitude of 84 cm, which is approximately 28 cm higher than the modelled residual, however the timing of the surge agrees almost exactly as displayed in Figure \ref{fig:dec2006}.
-The best agreement occurs at Victoria where the observed surge was 62 cm and the model returned 44 cm.
+The best agreement occurs at Victoria where the observed surge was 62 cm and the model returned 44~cm.
 At all of the locations, the model predicted the maximum surge to occur within 3 hours of the observed surge.
 Typically, the modelled surge was too early, in contrast to the later predictions in the Feb 4, 2006 case.
-All Willmott skill scores rank above 0.89 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}).
+All Willmott skill scores rank above 0.88 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}).
 The modelled residuals are on average too low and have a lower skill score than the Feb 2006 case.
@@
 Although Victoria has the smallest mean error in both the water level and residuals, its Willmott skill score and $\gamma^2$ values rank low when compared to the other stations.
-Generally, the skills scores are very good, scoring above 0.89 at all stations.
+Generally, the skills scores are very good, scoring above 0.91 at all stations.
 
@@
 
-Both of the simulations agree well with the observations (\ref{fig:dec2012_surge}).
+Both of the simulations agree well with the observations (Figure \ref{fig:dec2012_surge}).
 Willmott skill scores for both modelled residuals are above 0.92 at these four locations.

### 2014-12-05 RoughWork/manuscript.tex
Edits from Susan and Kate.

@@
 %intro to storm surges
-Many coastal communities in the Strait of Georgia are at risk to flooding and property damaged caused by storm surges, a natural hazard arising from the combination of a strong wind storm and high tide.
+Many coastal communities in the Strait of Georgia are at risk to flooding and property damage caused by storm surges, a natural hazard arising from the combination of a strong wind storm and high tide.
 The low atmospheric pressure associated with a storm acts as an inverse barometer elevating the sea level.
@@
 Fourth, a high wind event that did not result in a large surge or flooding for a Nov 2009 case is discussed in section \ref{sec:wind}.
-Finally, comparisons between hindcasts using high resolution (HRDPS) and low resolution (CGRF) atmospheric forcing is provided in section \ref{sec:res}.
+Finally, comparisons between hindcasts using high resolution (HRDPS) and low resolution (CGRF) atmospheric forcing is provided for the Dec 2012 case in section \ref{sec:res}.
 
@@
 The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in a slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
-Additionally, the surge amplitude decreases by a few centimetres if the inverse barometer effect due to atmospheric pressure is not include.
-Other findings have reported the inverse barometer effect contributes up to $^1/_3$ of the total surge amplitude \citep{murty1995storm}, but only a minor effect is noticeable here.
+Additionally, the surge amplitude decreases by a few centimetres if the inverse barometer effect due to atmospheric pressure is not included.
+Other findings have reported the inverse barometer effect contributes up to $^1/_3$ of the total surge amplitude \citep{murty1995storm}, but only a minor effect is seen here.
 

### 2014-12-11 RoughWork/manuscript.tex
Fixed typo in legend.
Revisions in hindcast section. Mostly shortening

@@
 The monthly flow from each river was input as a point source from 0 to 3 metres at the model point closest to the mouth of each river.
-Incoming water was assumed to be fresh and at surface air temperature.
+Incoming water was assumed to be fresh and at surface temperature.
 A total of 150 rivers were parametrized by this method and their locations are indicated by the blue dots in Figure \ref{fig:domain}.
@@
 These three statistics, along with the mean observed values and mean modelled values are reported in Table \ref{tab:statistics} for each hindcast.
-Further, comparisons between the maximum observed and modelled water level and residual are typically reported in the text.
 
@@
 
-On Feb 4, 2006, a strong storm system with winds over 75 km hr$^{-1}$ coinciding with an unusually high tide led to significant flooding in coastal communities of the Strait of Georgia \citep{romanowski2010storm}.
-The maximum water level observed at Point Atkinson on this day was 5.46 m, which is 0.82 m higher than the predicted tides, highlighting the significance of this event as it resulted in a large anomaly and also reached water levels close to the highest level ever recorded (5.61 m).
-A set of hindcast simulations from Feb 1-8, 2006 have been completed.
-
+A hindcast for a storm surge that occurred on Feb 4, 2006 is presented next.
+The simulation ran from Feb 1-8, 2006.
 The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations to corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River.
-The timing of maximum water level matches well between the observations and the model at all four locations, the most significant contribution to high water being from the tides which are generally well-represented by the model.
-There is good agreement in the water level maximum at both Victoria and Patricia Bay, with a relative error less than 2\%.
-In fact, there is only a 4.2 cm difference between observed and modelled maximum at Victoria and a 6.9 cm difference at Patricia Bay, the model slightly under predicting in each case.
-The error at Point Atkinson and Campbell River is slightly more significant, reaching 2.5\% and 2.7\% respectively, close to 14~cm over predicted at both locations.
+First, the timing of maximum water level matches well between the observations and the model at all four locations, the most significant contribution to high water being from the tides which are generally well-represented by the model.
+There is good agreement in the water level maximum at both Victoria and Patricia Bay, although it is slightly under predicted at these two locations.
+The maximum water level at Point Atkinson and Campbell River is slightly over predicted and the error is more significant, but the tidal range is also larger.
 It is surprising that Victoria and Patricia Bay agree so closely with observations because the M$_2$ tidal amplitudes in the model are known to be too low in this area.
@@
 
-In general, there is a very good agreement between the observed and modelled residuals (Figure \ref{fig:feb2006}).
+In general, there is also very good agreement between the observed and modelled residuals (Figure \ref{fig:feb2006}).
 The most noticeable difference is a 4-6 hour delay in the model's timing of the maximum residual except at Victoria where the maximum residual is in sync with observations.
-The errors in maximum residual range from 6.0 cm too high at Victoria to 15.9 cm too low at Campbell River.
-Both Point Atkinson and Patricia Bay experience a very good match with observed residuals with errors in the maximum less than 2 cm.
+The surge manifests as a broad peak in both the modelled and observed residual leading to a high probability that large residuals coincide with the high tide.
+As such, it is of minor consequence that the timing of the observed and modelled maximum residual do not match perfectly.
 
@@
 
-The maximum water level and maximum residual are important factors in storm surge hindcasts but a better indicator of overall model performance is provided by a group of statistical measures, including the mean absolute error, the observed mean, the modelled mean, $\gamma^2$, and the Willmott skill score for both the water level and residuals (Table \ref{tab:statistics}).
+A more detailed evaluation of the model performance is provided by a group of statistical measures, including the mean absolute error, the observed mean, the modelled mean, $\gamma^2$, and the Willmott skill score for both the water level and residuals (Table \ref{tab:statistics}).
 The statistics are calculated with hourly output over the three-day period displayed in Figure \ref{fig:feb2006}.
-To begin with, the mean absolute error in the water level ranges from 7.6 cm at Victoria to 13.1 cm at Campbell River, whereas, the mean absolute error in the residuals is smaller at all stations, however, this measure does not account for the large variability in water levels due to the tides.
+The mean absolute error in the water level ranges from 7.6 cm at Victoria to 13.1 cm at Campbell River, whereas, the mean absolute error in the residuals is smaller at all stations, however, this measure does not account for the large variability in water levels due to the tides.
 A statistic that does take into account this variability is $\gamma^2$ which suggests very good agreement in both the predicted water levels and residuals.
@@
 
-An additional storm surge hindcast for Nov 2006 is provided as another means of assessing model performance and is mentioned briefly here.
-On Nov 16, 2006 communities in the Strait of Georgia experienced a strong wind storm with speeds upwards of 20 m s$^{-1}$ near Point Atkinson resulting in observed water levels of 5.1 m and a maximum surge amplitude of 83.7 cm.
-The surge amplitude was lower at the other stations, reaching no more than 68 cm.
+An additional storm surge hindcast for a Nov 16, 2006 storm surge is provided as another means of assessing model performance and is mentioned briefly here.
 The model performs reasonably well in this hindcast; Willmott skill scores are greater than 0.94 (Table \ref{tab:statistics}).
@@
 The contribution to the surge due to atmospheric forcing appears to be geographically dependent, its removal leading to a drop of about 2.5 cm in the maximum surge at Point Atkinson and 2.3 cm at Campbell River.
-Interestingly, the surge amplitude increases by 4.8 cm at Victoria and 3.6 cm at Patricia Bay when atmospheric forces are neglected.
+The surge amplitude increases by 4.8 cm at Victoria and 3.6 cm at Patricia Bay when atmospheric forces are neglected.
 The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in a slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
@@
 \subsection{Spatial Extent of Surge}\label{sec:spatial}
-On Dec 15, 2006, a major storm with wind gusts over 120 km hr$^{-1}$ caused significant damage to many coastal communities on the Pacific Northwest.
-Although the storm was very strong and the observed residual at Point Atkinson reached 84 cm, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area \citep{forseth2006adaptation}.
+A major wind storm on Dec 15, 2006 is hindcasted next.
+Although the storm was very strong (gusts over 120 km hr$^{-1}$) and the observed residual at Point Atkinson reached 84 cm, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area \citep{forseth2006adaptation}.
 Because of the high winds, this event has been used to examine the propagation of the surge throughout the domain and spatial variations in the surge amplitude.
@@
 
-Point Atkinson observed a maximum surge amplitude of 84 cm, which is approximately 28 cm higher than the modelled residual, however the timing of the surge agrees almost exactly as displayed in Figure \ref{fig:dec2006}.
-The best agreement occurs at Victoria where the observed surge was 62 cm and the model returned 44~cm.
-At all of the locations, the model predicted the maximum surge to occur within 3 hours of the observed surge.
-Typically, the modelled surge was too early, in contrast to the later predictions in the Feb 4, 2006 case.
-All Willmott skill scores rank above 0.88 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}).
-The modelled residuals are on average too low and have a lower skill score than the Feb 2006 case.
+Modelled residuals are generally too low in this hindcast (Figure \ref{fig:dec2006}) however the timing of the surge agrees well.
+All Willmott skill scores rank above 0.88 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}) but the performance has degraded when compared to the Feb 2006 and Nov 2006 cases.
 This could be due to a number of factors.
@@
 Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered.
-As recorded in the Environment Canada Climate Data archives \citep{ECClimateArchive}, the maximum hourly wind speed at Point Atkinson was 18.6 m s$^{-1}$ and the largest gust observed was 25.8 m s$^{-1}$ or 93 km hr$^{-1}$.
-In addition, large wind speeds over 15 m s$^{-1}$ persisted for over 9 hours.
-Yet, the surge associated with this event is rather small when compared to the previous two cases (Figure \ref{fig:nov2009}).
+Although large wind speeds were observed, the surge associated with this event is rather small when compared to the previous cases (Figure \ref{fig:nov2009}).
 Two significant surge events are noted, the first occurring on Nov 19 with an observed amplitude of about 50 cm at Point Atkinson, and a second, larger surge of 62.5 cm on Nov 20.
@@
 Of the cases considered so far, this is the only example where the surge amplitude at Victoria was higher than that at Point Atkinson.
-This event observed a higher surge at Victoria than the Dec 2006 event, further highlighting the role that winds plays in setting up local variations in the surge.
-In this case, the model results do not reflect a higher surge at Victoria (46.1 cm versus 50.5~cm at Point Atkinson), perhaps due to an underestimate in the wind speed and some discrepancies in the wind direction of the atmospheric forcing.
+The model results do not reflect a higher surge at Victoria (46.1 cm versus 50.5~cm at Point Atkinson), perhaps due to an underestimate in the wind speed and some discrepancies in the wind direction of the atmospheric forcing.
 
@@
 
-Both of the simulations agree well with the observations (Figure \ref{fig:dec2012_surge}).
+Hindcasts with both atmospheric forcing products agree well with the observations (Figure \ref{fig:dec2012_surge}).
 Willmott skill scores for both modelled residuals are above 0.92 at these four locations.
@@
 Lastly, flooding effects due to wave run up is left for a more detailed inundation study with higher resolution in the near-shore areas.
+Local winds have a significant impact on wave action which is a primary concern for flooding management.
 

### 2014-12-11 RoughWork/manuscript.tex
a few typoographical edits before sending to east coast

@@
 Model evaluation has focused on regions where the near-coast slopes are steep and so wetting and drying effects are believed to be unimportant.
-A detailed overview of the model configuration and domain including a description of the ocean model configuration is provided in section \ref{sec:config}.
+A detailed overview of the model configuration and domain is provided in section \ref{sec:config}.
 This section also describes open and surface boundary conditions as well as choices in grid structure and mixing parametrizations.
@@
 To prepare the climatology all years (1995-2008) were averaged and results, approximately every 15 days, were interpolated to a weekly climatology.
-The model is relaxed to the forced temperature and salinity over the 10 grid points (about 5km) closest to the open boundaries, using a flow relaxation scheme \citep{engedahl1995use}.
+The model is relaxed to the forced temperature and salinity over the 10 grid points (about 5 km) closest to the open boundaries, using a flow relaxation scheme \citep{engedahl1995use}.
 %other references?
@@
 For the storm surge simulations, hourly variations in sea surface height were used.
-These values are the Tofino tide gauge values, de-tided and with the zero reset as for the climatology.
+These values are the Tofino tide gauge values with the tides removed.
+Tidal predictions were generated with a tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year prior to the surge event to avoid over predicted constituents during high surge years.
 The storm surge simulations also used the sea surface height anomaly from Port Hardy, forced at the northern boundary in Johnstone Strait.
@@
 
-River input provides a significant volume of freshwater to the Salish Sea and can influence stratification, circulation, and primary productivity.
+River input provides a significant volume of fresh water to the Salish Sea and can influence stratification, circulation, and primary productivity.
 However, most rivers in the domain are not gauged so parametrizations were required to represent river flow.
@@
 
-Modelled residuals are generally too low in this hindcast (Figure \ref{fig:dec2006}) however the timing of the surge agrees well.
+Modelled residuals are generally too low in this hindcast (Figure \ref{fig:dec2006}), however, the timing of the surge agrees well.
 All Willmott skill scores rank above 0.88 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}) but the performance has degraded when compared to the Feb 2006 and Nov 2006 cases.
@@
 Significant flooding occurred later, on Dec 17, 2012, where the observed surge amplitude at Point Atkinson was 68.3 cm.
-Even though the low resolution model significantly under predicts the wind speed at some locations, it performs comparable to forcing with high resolution products.
+Even though the low resolution model significantly under predicts the wind speed at some locations, it performs comparably to forcing with high resolution products.
 This result further demonstrates that the local wind is not the most important factor for setting up storm surges over this domain.

### 2014-12-13 AO_guides/manuscriptAO.tex
Started converting to AO style. Need to figure out citations!

--- 2014-12-11 RoughWork/manuscript.tex
@@
-\documentclass[pdftex,10pt]{article}
-\usepackage{fullpage}
-\usepackage{graphicx}
-\usepackage{natbib}
-\usepackage{amsmath}
-\usepackage{setspace} \doublespacing
+% tATOguide.tex
+% v1.2 released April 2014
+
+\documentclass{tATO2e}
+
+\usepackage{subfigure}% Support for small, `sub' figures and tables
+
+\usepackage[longnamesfirst,sort]{natbib}% Citation support using natbib.sty
+\bibpunct[, ]{(}{)}{;}{a}{,}{,}% Citation support using natbib.sty
+
+%\usepackage[natbibapa]{apacite}% Citation support using apacite.sty. Commands using natbib.sty MUST be deactivated first!
+
+\begin{document}
+
+%\jvol{00} \jnum{00} \jyear{2014} \jmonth{April}
+
+%\articletype{GUIDE}% NOT REQUIRED IN A RESEARCH ARTICLE
 
 \title{Storm surge modelling in the Strait of Georgia}
-\author{Nancy Soontiens, Susan Allen, Doug Latornell, Kate Le Sou\"{e}f, Idalia Machuca,
-\\ Youyu Lu, Keith Thompson, Jean-Philippe Paquin, Vasily Korabel}
-\date{\today}
 
-\begin{document}
+\author{Nancy Soontiens$^{1}$$^{\ast}$\thanks{$^\ast$Corresponding author's email: nsoontie@eos.ubc.ca
+\vspace{6pt}}, Susan Allen$^{1}$,  Doug Latornell$^{1}$,  Kate Le Sou\"{e}f$^{1}$, Idalia Machuca$^{1}$ \\ 
+Youyu Lu$^{2}$, Keith Thompson$^{3}$, Jean-Philippe Paquin$^{3}$, Vasily Korabel$^{3}$
+\\\vspace{6pt}  $^{1}${\em{University of British Columbia}};
+$^{2}${\em{Bedford Insitute of Oceanography}}; 
+$^{3}${\em{Dalhousie University}}\\\received{v1.2 released April 2014} }
+
 \maketitle
@@
 \end{abstract}
+
+\begin{keywords}storm surge; Strait of Georgia; Salish Sea; numerical modelling; spatial variation
+\end{keywords}
 
@@
 \centering 
-\begin{tabular}{|l |c c c c c | c c c c c|} 
+\tbl{Evaluation of model performance for all hindcasts presented.
+Mean absolute error ($\bar{e}$), observed mean (subscript o), modelled mean (subscript m), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and residuals ($\delta$).
+Statistics are calculated over the time period displayed in each simulation's corresponding figure (three or four days).
+For the Nov 2006 case, the statistics were calculated over a three day period.
+The average values were calculated over all of the hindcasts, excluding the Dec 2012 HRDPS simulation.
+}
+{\begin{tabular}{|l |c c c c c | c c c c c|} 
 \hline 
@@
 \hline
-\end{tabular}
-\caption{Evaluation of model performance for all hindcasts presented.
-Mean absolute error ($\bar{e}$), observed mean (subscript o), modelled mean (subscript m), $\gamma^2$, and the Willmott skill score ($WS$) are provided for water level predictions ($\eta$) and residuals ($\delta$).
-Statistics are calculated over the time period displayed in each simulation's corresponding figure (three or four days).
-For the Nov 2006 case, the statistics were calculated over a three day period.
-The average values were calculated over all of the hindcasts, excluding the Dec 2012 HRDPS simulation.
-}
+\end{tabular}}
 \label{tab:statistics}
@@
 The timing of the maximum surge at Campbell River was identical with observations and too early at the other the stations.
-
-%\begin{table}[h]
-%\centering 
-%\begin{tabular}{|l |c c c c c | c c c c c|} 
-%\hline 
-%& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
-%\hline 
-%Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
-%\hline 
-%Point Atkinson &  9.8           &  3.42                & 3.43                 &   0.034    & 0.991  &  7.7           &  0.290                 & 0.247                  &  0.128     & 0.950 \\
-%Victoria       &  9.1           &  2.16                & 2.11                 &   0.070    & 0.978  &  6.2           &  0.222                 & 0.202                  &  0.171     & 0.946 \\
-%Patricia Bay   & 10.0           &  2.57                & 2.53                 &   0.049    & 0.985  &  6.6           &  0.235                 & 0.203                  &  0.144     & 0.947 \\
-%Campbell River & 10.7           &  3.20                & 3.26                 &   0.037    & 0.989  &  6.2           &  0.254                 & 0.236                  &  0.140     & 0.952 \\
-%\hline 
-%\end{tabular}
-%\caption{Evaluation of model performance for the Nov 2006 hindcast. Statistics were calculated over the period Nov 14, 2006 to Nov 17, 2006.}
-%\label{tab:nov2006stat} 
-%\end{table}
 
@@
 \end{figure}%table instead of figure? or maybe just the surge and not water level?
-
-%\begin{table}[h]
-%\centering 
-%\begin{tabular}{|l |c c c c c | c c c c c|} 
-%\hline 
-%& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
-%\hline 
-%Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
-%\hline 
-%Point Atkinson & 10.5           &  3.45                & 3.47                 &   0.025    & 0.994  & 11.1           &  0.212                 & 0.186                  &  0.275     & 0.891 \\
-%Victoria       &  9.6           &  2.23                & 2.16                 &   0.042    & 0.984  &  7.9           &  0.190                 & 0.156                  &  0.223     & 0.911 \\
-%Patricia Bay   &  9.2           &  2.63                & 2.58                 &   0.031    & 0.990  &  9.5           &  0.193                 & 0.149                  &  0.260     & 0.889 \\
-%Campbell River & 10.6           &  3.23                & 3.28                 &   0.030    & 0.992  &  9.4           &  0.191                 & 0.163                  &  0.247     & 0.905 \\
-%\hline 
-%\end{tabular}
-%\caption{Evaluation of model performance for the Dec 2006 hindcast. Statistics were calculated over the period Dec 14, 2006 to Dec 17, 2006.}
-%\label{tab:dec2006stat} 
-%\end{table}
 
@@
 
-%\begin{table}[h]
-%\centering 
-%\begin{tabular}{|l |c c c c c | c c c c c|} 
-%\hline 
-%& \multicolumn{5}{|c|}{Water Level ($\eta$)}        & \multicolumn{5}{|c|}{Residual ($\delta$)} \\ 
-%\hline 
-%Location       & $\bar{e}$ (cm) & $\bar{\eta_{o}}$ (m) & $\bar{\eta_{m}}$ (m) & $\gamma^2$ & $WS$   & $\bar{e}$ (cm) & $\bar{\delta_{o}}$ (m) & $\bar{\delta_{m}}$ (m) & $\gamma^2$ & $WS$ \\
-%\hline 
-%Point Atkinson & 14.0           &  3.44                & 3.55                 &   0.012    & 0.995  &  6.1           &  0.313                 & 0.331                  &  0.237     & 0.914 \\
-%Victoria       &  9.9           &  2.26                & 2.24                 &   0.023    & 0.994  &  4.8           &  0.323                 & 0.293                  &  0.245     & 0.899 \\
-%Patricia Bay   & 11.2           &  2.60                & 2.65                 &   0.020    & 0.993  &  6.0           &  0.277                 & 0.290                  &  0.272     & 0.901 \\
-%Campbell River & 13.5           &  3.22                & 3.27                 &   0.019    & 0.995  &  4.9           &  0.326                 & 0.314                  &  0.217     & 0.924 \\
-%\hline 
-%\end{tabular}
-%\caption{Evaluation of model performance for the Nov 2009 hindcast. Statistics were generated over the period Nov 17, 2009 to Nov 22, 2009.}
-%\label{tab:nov2009stat} 
-%\end{table}
-
 \subsection{High Resolution Atmospheric Forcing}\label{sec:res}
@@
 Although not shown here, both models have a similar sea level atmospheric pressure at these two locations.
-
-%Where can I find a reference for the high res atmospheric model?
 
@@
 
-\appendix
+\appendices
 \section{Appendix}\label{sec:appendix}
@@
 \centering 
-\begin{tabular}{l l l l c c c c c c} 
+\tbl{Comparisons of modelled and observed amplitudes and phases.
+$R$ is the ratio of modelled to observed amplitude.
+$\Delta \phi$ is the difference between modelled and observed phase.
+$D$ is the complex difference.
+The mean error and root mean square (RMS) error are calculated as the difference between the modelled quantity and observed quantity.
+}
+{\begin{tabular}{l l l l c c c c c c} 
 \hline 
@@
 \hline  
-\end{tabular}
-\caption{Comparisons of modelled and observed amplitudes and phases.
-$R$ is the ratio of modelled to observed amplitude.
-$\Delta \phi$ is the difference between modelled and observed phase.
-$D$ is the complex difference.
-The mean error and root mean square (RMS) error are calculated as the difference between the modelled quantity and observed quantity.
-}
+\end{tabular}}
 \label{tab:comparison} 
@@
 
-\section{Acknowledgements}\label{sec:ack}
+\section*{Acknowledgements}\label{sec:ack}
 We would like to thank Michael Foreman, Diane Masson, Luc Fillion, Rich Pawlowicz, Mark Halverson, Kao-Shen Chung, and Pramod Thupaki for providing data used in model set up and evaluation.

### 2014-12-13 AO_guides/tATOguide.tex
Started converting to AO style. Need to figure out citations!

% tATOguide.tex
% v1.2 released April 2014

\documentclass{tATO2e}

\usepackage{subfigure}% Support for small, `sub' figures and tables

\usepackage[longnamesfirst,sort]{natbib}% Citation support using natbib.sty
\bibpunct[, ]{(}{)}{;}{a}{,}{,}% Citation support using natbib.sty

%\usepackage[natbibapa]{apacite}% Citation support using apacite.sty. Commands using natbib.sty MUST be deactivated first!

\begin{document}

%\jvol{00} \jnum{00} \jyear{2014} \jmonth{April}

\articletype{GUIDE}% NOT REQUIRED IN A RESEARCH ARTICLE

\title{{\itshape Atmosphere-Ocean} \LaTeX\ Style Guide for Authors \break (Own Layout Style + American Psychological Association Reference Style)}

\author{A. N. Author$^{1}$$^{\ast}$\thanks{$^\ast$Corresponding author's email: latex.helpdesk@tandf.co.uk
\vspace{6pt}} and I. T. Consultant$^{2}$\\\vspace{6pt}  $^{1}${\em{Taylor \& Francis, 4 Park Square, Milton Park, Abingdon, UK}};
$^{2}${\em{Institut f\"{u}r Informatik, Albert-Ludwigs-Universit\"{a}t, Freiburg,
Germany}}\\\received{v1.2 released April 2014} }

\maketitle

\begin{abstract}
This guide is for authors who are preparing papers for the Taylor \& Francis journal {\em Atmosphere-Ocean (tATO)}
using the \LaTeX\ document preparation system and the class file {\tt tATO2e.cls}, which
is available via the journal's homepage on the Taylor \& Francis website. Authors planning
to submit papers prepared in \LaTeX\ are advised to use {\tt tATO2e.cls} as early as possible in the creation of their files.
\end{abstract}

\begin{keywords}submission instructions; source file coding;
references citation; fonts; numbering {\bf{(Please provide four to eight keywords taken
from terms used in your manuscript}})
\end{keywords}

\section{Introduction}

In order to assist authors in the process of preparing a manuscript for {\itshape Atmosphere-Ocean} ({\it tATO}), the journal's layout style has been implemented as a \LaTeXe\ class file based on the {\tt article} document class. Commands that differ from or are provided in addition to the standard \LaTeXe\ interface are explained in this guide. The guide alone is not intended as a substitute for an appropriate \LaTeXe\ manual.

The \verb"tATOguide".tex file can also be used as a template for composing an article for submission by cutting, pasting, inserting and
deleting text as appropriate, using the \LaTeX\ environments provided (e.g. \verb"\begin{equation}", \verb"\begin{enumerate}").

\subsection{The {\em tATO} Document Class}\label{S1.1}

The \texttt{tATO2e} class file preserves the standard \LaTeXe\ interface such that any document that can
be produced using {\tt article.cls} can also be produced using the {\it tATO} document class.
However, the measure (the width of the text on a page) differs from the default for {\tt article.cls}, therefore line breaks
will change and some long equations may need to be reformatted accordingly.

If your article is accepted for publication in the journal, it will be typeset in Monotype Times. As most authors do not own this font, the page make-up would inevitably alter with the change of font. Moreover, the class file produces single-column output, which will be converted to two-column format by the typesetter. This reduces formatting problems during preparation of manuscripts by authors due to long lines and equations spanning more than one column. Line endings would change anyway during preparation of proofs from two-column manuscripts because typesetters' character sets differ slightly in size from those available on most PCs and laptops. Please therefore ignore details such as slightly long lines of text, page stretching, or figures falling out of synchronization with their citations in the text: these details will be dealt with by the typesetter. Similarly, it is unnecessary to spend time addressing warnings in the log file---if your .tex file compiles to produce a PDF file that correctly shows how you wish your paper to appear, such warnings will not prevent your source files being imported into the typesetter's program.

\subsection{Submission of \LaTeX\ Articles to the Journal}\label{S1.2}

Manuscripts for possible publication in the journal should be submitted to the Editors for review as directed in the journal's Instructions for Authors, which may be found at {\tt{http://www.tandf.co.uk/journals/authors/tatoauth.asp}}.

Manuscripts created using \LaTeX\ should be converted to PDF format prior to submission. The \LaTeX\ source files and any graphics files will be required in addition to the final PDF version when final, revised versions of accepted manuscripts are submitted.

`Open-source' \LaTeXe\ should be used in preference to proprietary systems such as TCILaTeX or Scientific WorkPlace; similarly, class files such as REVTeX~4.1 that produce a document in the style of a different publisher and journal should not be used for preference.

Authors who wish to incorporate Encapsulated PostScript artwork directly in their articles can do so by using
Tomas~Rokicki's {\tt EPSF} macros (which are supplied with the DVIPS PostScript driver). See Section~\ref{eps},
which also demonstrates how to treat landscape pages. Please remember to supply any additional figure macros you
use with your article in the preamble before \verb"\begin{document}". Authors should not attempt to use
implementation-specific \verb"\special"s directly.

Ensure that any author-defined macros are gathered together in the source file, just before the
\verb"\begin{document}" command.

Please note that if serious problems are encountered with the coding of a paper (missing author-defined macros,
for example), it may prove necessary to divert the paper to conventional typesetting, i.e. it will be re-keyed.

\section{Using the {\bi tATO} class file}

If the file {\tt tATO2e.cls} is not already in the appropriate system directory for \LaTeXe\ files, either
arrange for it to be put there, or copy it to your working folder. In order to use the {\it tATO} document class, replace the command
\verb"\documentclass{article}" at the beginning of your document with the command
\verb"\documentclass{tATO2e}".

The following document-class options should {\em not\/} be used with the {\it tATO} class file:
%
\begin{itemize}
   \item {\tt 10pt}, {\tt 11pt}, {\tt 12pt}---unavailable;
   \item {\tt oneside}, {\tt twoside}---not necessary: oneside is the default;
   \item {\tt leqno} and {\tt titlepage}---should not be used;
   \item {\tt onecolumn}---not necessary as it is the default style;
   \item {\tt twocolumn}---should not be used for reasons explained in Section~\ref{S1.1}.
\end{itemize}
%
The \texttt{geometry} package and commands associated with it should also not be used to adjust the page dimensions.

\section{Additional features}

\subsection{Footnotes to Article Titles and Authors' Names}

On the title page, the \verb"\thanks" control sequence may be used to produce a footnote to either the title or authors' names. Footnote symbols for this purpose should be used in the order:
$\dagger$ (coded as \verb"\dagger"), $\ddagger$ (\verb"\ddagger"), $\S$ (\verb"\S"), $\P$ (\verb"\P"), $\|$ (\verb"\|"), $\dagger\dagger$ (\verb"\dagger\dagger"), $\ddagger\ddagger$ (\verb"\ddagger\ddagger"),
$\S\S$ (\verb"\S\S"), $\P\P$ (\verb"\P\P"), $\|\|$ (\verb"\|\|").

Any \verb"footnote"s to the main text will automatically be assigned the superscript
 symbols 1, 2, 3, etc. by the class file.\footnote{If preferred, the \texttt{endnotes} package
 may be used to set the notes at the end of your text, before the bibliography.
 The symbols will be changed to match the style of the journal if necessary by the typesetter.}

The title, author(s) and affiliation(s) should be followed by the \verb"\maketitle" command. If preparing an anonymized version for peer review, \verb"\maketitle" may follow
directly after the title in order to shield the authors' identities from the reviewers.

\subsection{Abstracts}

At the beginning of your article, the title should be generated in the usual way using the {\verb"\maketitle"}
command. Immediately following the title you should include an abstract. The abstract should be enclosed within
an {\tt abstract} environment. For example, the titles for this guide were produced by the following source code:
%
\begin{verbatim}
\title{{\itshape Atmosphere-Ocean} \LaTeX\ Style Guide for Authors
(Own Layout Style + American Psychological Association Reference Style)}

\author{A.N. Author$^{1}$$^{\ast}$\thanks{$^\ast$Corresponding
author's email: latex.helpdesk@tandf.co.uk \vspace{6pt}} and I.T.
Consultant$^{2}$\\\vspace{6pt}  $^{1}${\em{Taylor \& Francis,
4 Park Square, Milton Park, Abingdon, UK}}; $^{2}${\em{Institut
f\"{u}r Informatik, Albert-Ludwigs-Universit\"{a}t, Freiburg,
Germany}}\\\received{v1.2 released April 2014} }

\maketitle

\begin{abstract}
This guide is for authors who are preparing papers for the Taylor
\& Francis journal {\em Atmosphere-Ocean (tATO)} using the \LaTeX\
document preparation system and the class file {\tt tATO2e.cls},
which is available via the journal's homepage on the Taylor \&
Francis website. Authors planning to submit papers prepared in
\LaTeX\ are advised to use {\tt tATO2e.cls} as early as possible
in the creation of their files.
\end{abstract}
\end{verbatim}

\subsection{Lists}

The {\it tATO} class file provides numbered and unnumbered lists using the {\tt enumerate} environment and bulleted
lists  using the {\tt itemize} environment.

The enumerated list will number each list item with arabic numerals by default, for example:
%
\begin{enumerate}
  \item first item
  \item second item
  \item third item
\end{enumerate}
%
was produced by
%
\begin{verbatim}
\begin{enumerate}
  \item first item
  \item second item
  \item third item
\end{enumerate}
\end{verbatim}
%
Alternative numbering styles can be achieved by inserting an optional argument in square brackets to each \verb"item", e.g. \verb"\item[(i)] first item" to create a list numbered with roman numerals.

Unnumbered lists are also provided using the {\tt enumerate} environment. For example,
%
\begin{enumerate}
  \item[] First unnumbered indented item without label.
  \item[] Second unnumbered item.
  \item[] Third unnumbered item.
\end{enumerate}
%
was produced by:
%
\begin{verbatim}
\begin{enumerate}
  \item[] First unnumbered indented item without label.
  \item[] Second unnumbered item.
  \item[] Third unnumbered item.
\end{enumerate}
\end{verbatim}

Bulleted lists are provided using the {\tt itemize} environment. For example,
%
\begin{itemize}
  \item First bulleted item
  \item Second bulleted item
  \item Third bulleted item
\end{itemize}
%
was produced by:
%
\begin{verbatim}
\begin{itemize}
  \item First bulleted item
  \item Second bulleted item
  \item Third bulleted item
\end{itemize}
\end{verbatim}

\subsection{Landscape Pages}\label{eps}

If a table or illustration is too wide to fit the standard measure, it must be turned, with its caption,
through 90$^{\circ}$ anticlockwise. Landscape illustrations and/or tables can be produced using
the \verb"rotating" package, which is called by the \textit{tATO} class file. The following commands can
be used to produce such pages.
%
\begin{verbatim}
\setcounter{figure}{0}
\begin{sidewaysfigure}
\centerline{\epsfbox{fig1.eps}}
\caption{An example of a landscape figure caption.}
\label{landfig}
\end{sidewaysfigure}
\end{verbatim}

\begin{verbatim}
\setcounter{table}{0}
\begin{sidewaystable}
 \tbl{The Largest Optical Telescopes.}
  {\begin{tabular}{@{}llllcll}
    .
    .
    .
  \end{tabular}}\label{tab1}
\end{sidewaystable}
\end{verbatim}
%
Before any float environment, use the \verb"\setcounter" command
as above to fix the numbering of the caption. Subsequent captions
will then be automatically renumbered accordingly.

\section{Some guidelines for using standard features}

The following notes are intended to help you achieve the best effects with the tATO2e class file.

\subsection{Sections}

\LaTeXe\ provides five levels of section heading, all of which are defined in the tATO2e class file:
\begin{enumerate}
   \item[(A)] \verb"\section"
   \item[(B)] \verb"\subsection"
   \item[(C)] \verb"\subsubsection"
   \item[(D)] \verb"\paragraph"
   \item[(E)] \verb"\subparagraph"
\end{enumerate}
Numbering is automatically generated for section, subsection, subsubsection and paragraph headings. If you need
additional text styles in the headings, see the examples in Section~\ref{headings}.

\subsection{Illustrations (Figures)}

The \textit{tATO} class file will cope with most positioning of your illustrations and you should not normally need to use the optional placement specifiers of the \texttt{figure} environment. See `Instructions for Authors' in the journal's homepage on the Taylor \& Francis website for how to submit artwork (note that requests to supply figures and tables separately from text are for the benefit of authors using Microsoft Word; authors using \LaTeX\ may include these at the appropriate locations in their PDF file). The original source files of any illustrations will be required when the final, revised version is submitted. Authors should ensure that their figures are suitable (in terms of lettering size, etc.) for the reductions they intend.

Figure captions should appear below the figure itself, therefore the \verb"\caption" command should appear after the
figure. For example, Fig.~\ref{sample-figure} with caption and sub-captions is produced using the following
commands:
%
\begin{verbatim}
\begin{figure}
\begin{center}
\subfigure[An example of an individual figure sub-caption.]{
\resizebox*{6cm}{!}{\includegraphics{senu_gr1.eps}}}\hspace{5pt}
\subfigure[A slightly shorter sub-caption.]{
\resizebox*{6cm}{!}{\includegraphics{senu_gr2.eps}}}
\caption{\label{fig1} Example of a two-part figure with individual
 sub-captions showing that captions are flush left and justified if
 greater than one line of text, otherwise centred under the figure.}
\label{sample-figure}
\end{center}
\end{figure}
\end{verbatim}

\begin{figure}
\begin{center}
\subfigure[An example of an individual figure sub-caption.]
{\resizebox*{6cm}{!}{\includegraphics{senu_gr1.eps}}}\hspace{5pt}
\subfigure[A slightly shorter sub-caption.]
{\resizebox*{6cm}{!}{\includegraphics{senu_gr2.eps}}}
\caption{\label{fig1} Example of a two-part figure with individual
 sub-captions showing that captions are flush left and justified if
 greater than one line of text, otherwise centred under the figure.}
\label{sample-figure}
\end{center}
\end{figure}

The control sequences \verb"\subfigure{}" and \verb"\includegraphics{}" require subfigure.sty and graphicx.sty.
The former is called in the preamble of the \texttt{tATOguide.tex} file (in order to allow your choice of alternative if preferred)
and the latter by the \texttt{tATO2e} class file; both are included with the \LaTeX\ style guide package for this journal for convenience.

To ensure that your figures are correctly numbered automatically, the \verb"\label{}" command should be inserted just
after \verb"\caption{}".

\subsection{Tables}

The \textit{tATO} class file will cope with most positioning of your tables and you should not normally need to use the optional
placement specifiers of the \texttt{table} environment.

The \texttt{tabular} environment can be used as illustrated here to produce tables with single thin horizontal rules at the
head and foot and elsewhere, as per the style of the journal. Commands to redefine quantities such as \verb"\arraystretch"
should be omitted.

The table caption appears above the body of the table in \textit{tATO} style, therefore the \verb"\tbl" command should appear before
the body of the table. For example, Table~\ref{symbols} is produced using the following commands. Note that \verb"\rm" will
produce a roman character in math mode. There are also \verb"\bf" and \verb"\it", which produce bold face and text italic in math mode.

\begin{table}
\tbl{Example of a table showing that its caption is as wide as the table
 itself and justified.}
{\begin{tabular}[l]{@{}lcccccc}\toprule
  Class$^{\rm a}$ & $\gamma _1$ & $\gamma _2$$^{\rm b}$
         & $\langle \gamma \rangle$ & $G$ & $|{\bm f}|$ & $\theta _{c}$ \\
\colrule
  BL Lacs &5 & 36 & 7 & $-4.0$ & $1.0\times 10^{-2}$ & 10$^\circ$ \\
  FSRQs & 5 & 40 & 11 & $-2.3$ & $0.5\times 10^{-2}$ & 14$^\circ$ \\
\botrule
\end{tabular}}
\tabnote{$^{\rm a}$This footnote shows what footnote symbols to use.}
\tabnote{$^{\rm b}$This footnote shows the text turning over when a long footnote is added.}
\label{symbols}
\end{table}

\begin{verbatim}
\begin{table}
\tbl{Example of a table showing that its caption is as wide as the table
 itself and justified.}
{\begin{tabular}{@{}lcccccc}\toprule
  Class$^{\rm a}$ & $\gamma _1$ & $\gamma _2$$^{\rm b}$
         & $\langle \gamma \rangle$ & $G$ & $|{\bm f}|$ & $\theta _{c}$ \\
\colrule
  BL Lacs &5 & 36 & 7 & $-4.0$ & $1.0\times 10^{-2}$ & 10$^\circ$ \\
  FSRQs & 5 & 40 & 11 & $-2.3$ & $0.5\times 10^{-2}$ & 14$^\circ$ \\
\botrule
\end{tabular}}
\tabnote{$^{\rm a}$This footnote shows what footnote symbols to use.}
\tabnote{$^{\rm b}$This footnote shows the text turning over when a
 long footnote is added.}
\label{symbols}
\end{table}
\end{verbatim}
%
To ensure that your tables are correctly numbered automatically, the \verb"\label{}" command should be inserted just before
\verb"\end{table}".

Tables produced using the \texttt{booktabs} package of macros for typesetting tables are also compatible with the \textit{tATO} class file.

\subsection{Typesetting Mathematics}\label{TMth}

\subsubsection{Displayed mathematics}

The {\it tATO} class file will set displayed mathematics centred on the measure without equation numbers, provided
that you use the \LaTeXe\ standard control sequences open (\verb"\[") and close (\verb"\]") square brackets as
delimiters. The equation
\[
  \sum_{i=1}^p \lambda_i = {\rm trace}({\textrm{\bf S}})\qquad
  i\in {\mathbb R}
\]
\normalfont was typeset using the commands
%
\begin{verbatim}
\[
  \sum_{i=1}^p \lambda_i = {\rm trace}({\textrm{\bf S}})\qquad
  i\in {\mathbb R}
\]
\end{verbatim}

For those of your equations that you wish to be automatically
numbered sequentially throughout the text, use the {\tt{equation}}
environment, e.g.

\begin{equation}
  \sum_{i=1}^p \lambda_i = {\rm trace}({\textrm{\bf S}})\qquad
  i\in {\mathbb R}
\end{equation}
%
was typeset using the commands

\begin{verbatim}
\begin{equation}
  \sum_{i=1}^p \lambda_i = {\rm trace}({\textrm{\bf S}})quad
  i\in {\mathbb R}
\end{equation}
\end{verbatim}

Part numbers for sets of equations may be generated using the
{\tt{subequations}} environment, e.g.
\begin{subequations} \label{subeqnexample}
\begin{equation}
        \varepsilon \rho w_{tt}(s,t)
        =
        N[w_{s}(s,t),w_{st}(s,t)]_{s},
        \label{subeqnpart}
\end{equation}
\begin{equation}
        w_{tt}(1,t)+N[w_{s}(1,t),w_{st}(1,t)] = 0,
\end{equation}
\end{subequations}
which was generated using the control sequences

\begin{verbatim}
\begin{subequations} \label{subeqnexample}
\begin{equation}
        \varepsilon \rho w_{tt}(s,t)
        =
        N[w_{s}(s,t),w_{st}(s,t)]_{s},
        \label{subeqnpart}
\end{equation}
\begin{equation}
        w_{tt}(1,t)+N[w_{s}(1,t),w_{st}(1,t)] = 0,
\end{equation}
\end{subequations}
\end{verbatim}
This is made possible by the {\tt{subeqn}} package, which is called
by the class file. If you put the \verb"\label{}" just after the
\verb"\begin{subequations}" line, references will be to the
collection of equations, `(\ref{subeqnexample})' in the example
above. Or, like the example code above, you can reference each
equation individually---e.g. `(\ref{subeqnpart})'.

\subsubsection{Bold math italic symbols}

To get bold math italic you can use \verb"\bm", which works for
all sizes, e.g.
%
\begin{verbatim}
\sffamily
\begin{equation}
   {\rm d}({\bm s_{t_{\bm u}}) = \langle{\bm\alpha({\sf{\textbf L}})}
   [RM({\bm X}_y + {\bm s}_t) - RM({\bm x}_y)]^2 \rangle
\end{equation}
\normalfont
\end{verbatim}
%
produces\sffamily
\begin{equation}
   {\rm d}({\bm s_{t_{\bm u}}}) = \langle {\bm\alpha({\sf{\textbf L}})}[RM({\bm X}_y
   + {\bm s}_t) - RM({\bm x}_y)]^2 \rangle
\end{equation}\normalfont
Note that subscript, superscript, subscript to subscript, etc.
sizes will take care of themselves and are italic, not bold,
unless coded individually. \verb"\bm" produces the same effect as
\verb"\boldmath". \verb"\sffamily"...\verb"\normalfont" allows
upright sans serif fonts to be created in math mode by using the
control sequence `\verb"\sf"'.

\subsubsection{Bold Greek}\label{boldgreek}

Bold lowercase as well as uppercase Greek characters can be
obtained by \verb"{\bm \gamma}", which gives ${\bm \gamma}$, and
\verb"{\bm \Gamma}", which gives ${\bm \Gamma}$.

\subsubsection{Upright lowercase Greek characters and the upright partial derivative sign}\label{upgreek}

Upright lowercase Greek characters can be obtained with the \textit{tATO} class file by inserting the letter `u' in the control
code for the character, e.g. \verb"\umu" and \verb"\upi" produce $\umu$ (used, for example, in the symbol for the
unit microns---$\umu{\rm m}$) and $\upi$ (the ratio of the circumference to the diameter of a circle). Similarly,
the control code for the upright partial derivative $\upartial$ is \verb"\upartial".

\subsection{Acknowledgements}

An unnumbered section, e.g. \verb"\section*{Acknowledgement(s)}", should be used for thanks, etc.
and placed before any Notes or References sections.

\subsection{Funding}

An unnumbered section, e.g. \verb"\section*{Funding}", should be used for grant details, etc.
and placed before any Notes or References sections.

\subsection{Notes}

An unnumbered section, e.g. \verb"\section*{Note(s)}", may be placed before the References section.

\subsection{Supplemental Material}

Supplemental material should be referenced within your article where appropriate. An unnumbered section, e.g. \verb"\section*{Supplemental material}", detailing the supplemental material available should be placed immediately before the list of references, and should include a brief description of each supplemental file.

\subsection{Appendices}

Any appendices should be set before the list of references, beginning with the
command \verb"\appendices" followed by the command \verb"\section"
for each appendix title, e.g.

\begin{verbatim}
\appendices
\section{This is the title of the first appendix}
\section{This is the title of the second appendix}
\end{verbatim}

\noindent produces:\medskip

\noindent \textbf{Appendix A. This is the title of the first appendix}\vspace{10pt}

\noindent \textbf{Appendix B. This is the title of the second appendix}

\medskip
Subsections, equations, figures, tables, etc. within
appendices will then be automatically numbered as appropriate.

\subsection{References}\label{refs}

\subsubsection{References cited in the text}

References should be cited in the text in accordance with American Psychological Association style, i.e. in alphabetical order separated by semi-colons, e.g. (Green, 2002; Harlow, 1983; Wegener \& Petty, 1994) or `... see Smith (1985, p.~75)'. If there are two authors with the same surname, use the first initials with the surnames, e.g. (E.~Johnson, 2001; L.~Johnson, 1998). If there are three to five authors, list all the authors in the first citation, e.g. (Kernis, Cornell, Sun, Berry, \& Harlow, 1993). In subsequent citations, use only the first author's surname followed by et al., e.g. (Kernis et al., 1993). For six or more authors, cite the first author's name followed by et al. For two or more sources by the same author(s) in the same year, use lower-case letters (a,~b,~c,...) with the year to order the entries in the References list and use these lower-case letters with the year in the in-text citations, e.g. (Green, 1981a,b). Further details on this reference style can be found in the journal's Instructions for Authors.

Each bibliographical entry has a key, which is assigned by the author and used to refer to that entry in the text. In this document, the key \verb"ev94" in the citation form \verb"\citep{ev94}" produces `\citep{ev94}', and the keys \verb"Eri1984", \verb"Bar91" and \verb"Mil93" in the citation form \verb"\citep{Eri1984,Bar91,Mil93}" produce `\citep{Eri1984,Mil93,Bar91}'. The citation \verb"\citep{GloRib51}" produces `\citep{GloRib51}' where the citation first appears in the text, and `\citep{GloRib51}' in any subsequent citations. The appropriate citation style for different situations can be obtained, for example, by \verb"\citet{hk97}" for `\citet{hk97}', \verb"\citet{Maz91,PeaEtAl76,aiex00}" for `\citet{Maz91,PeaEtAl76,aiex00}', and \verb"\citealt{Kor95}" for `\citealt{Kor95}'. Optional notes may be included at the beginning and end of a citation by the use of square brackets, e.g. \verb"\citep[see][chap. 3]{Agu95}" produces `\citep[see][chap. 3]{Agu95}'. Citation of the year alone may be produced by \verb"\citeyear{cwm73}", i.e. `\citeyear{cwm73}', or \verb"\citeyearpar{cwm73}", i.e. `\citeyearpar{cwm73}'.

\subsubsection{The list of references}

References should be listed at the end of the main text in alphabetical order, then chronologically (earliest first), with full page ranges (where appropriate) and issue numbers (for journals paginated by issue). If a reference has more than seven named authors, list the first six, followed by an ellipsis (...), then the last author. The following list shows some references prepared in the style of the journal.
%
\begin{thebibliography}{12}

\bibitem[Agutter(1995)]{Agu95}
Agutter, A. J. (1995). {\em{The linguistic significance of current
 British slang}}. Unpublished doctoral dissertation, Edinburgh
 University, UK.

\bibitem[Aiex et al.(2000)]{aiex00}
Aiex, R. M., Pierce, I. F., Donizetti, G., {{v}on~Weber}, C. M., Bizet, G.,
 Bach, C. P. E., \ldots Dukas, P. (2000). {\em{Computing tools for
 modelling orchestral performance}} (Tech. Rep. No. DAMTP 2000/NA10).
 Cambridge, UK: University of Cambridge.

\bibitem[Barlow(1991)]{Bar91}
Barlow, D. H. (Ed.). (1991). Diagnoses, dimensions, and DSM-IV [special
 issue]. {\em{Journal of Abnormal Psychology}}, {\em{100}\,}(3).

\bibitem[Ericsson \& Simon(1984)]{Eri1984}
Ericsson, K. A., \& Simon, H. A. (1984). {\em{Protocol analysis:
 Verbal reports as data}}. Cambridge, MA: MIT Press.

\bibitem[Evans(1994)]{ev94}
Evans, W. A. (1994). Approaches to intelligent information retrieval.
 {\em{Information Processing and Management}}, {\em{7\,}}(2), 147--168.

\bibitem[Glover et al.(1951)Glover, Bizet, Bach, Strauss, \& Ribeiro]{GloRib51}
Glover, F., Bizet, G., Bach, C. P. E., Strauss, R., \& Ribeiro, C. C. (Eds.).
 (1951). {\em{Lessons of the British war economy}} (2nd ed., Vol. 5, pp. 1--24).
 Westport, CT: Greenwood Press.

\bibitem[Kern(1997)]{hk97}
Kern, H. (1997). The resurgent Japanese economy and a Japan--United
 States free trade agreement. In C. Lambert \& G. Holst (Eds.), {\em{4th
 international conference on the restructuring of the economic and
 political system in Japan and Europe}}, Milan, Italy, 21--25 May 1996 (pp.
 147--156). Singapore: World Scientific.

\bibitem[Korb(1995)]{Kor95}
Korb, K. B. (1995). Persons and things: Book review of Bringsjord on
 robot-consciousness. {\em Psycholoquy}, {\em{6\,}}(15). Retrieved
 May 20, 2004, from http://psycprints.ecs.soton.ac.uk/archive/00000462/

\bibitem[Mazzeo(1991)]{Maz91}
Mazzeo, J. (1991) {\em{Comparability of computer and paper-and-pencil
 scores}} (College Board Rep. No. 91). Princeton, NJ: Educational
 Testing Service.

\bibitem[Miller(1993)]{Mil93}
Miller, M. E. (1993). The Interactive Tester (Version 4.0) [Computer
 software]. Westminster, CA: Psytek Services.

\bibitem[Misner(1973)]{cwm73}
Misner, C. W. (1973). Efficient algorithms for layer assignment
 problems. In I. Gottlob (Ed.), {\em Gravitation in a collapsing
 Universe} (2nd ed., Vol. 5, pp. 63--83). San Francisco, CA: Freeman.

\bibitem[Pierce et al.(1976)]{PeaEtAl76}
Pierce, I. F., Donizetti, G., von Weber, C. M., Glover, F., Ogilvy,
 A. T., Clements, A. R. S., \& Ribeiro, C. C. (1976). A model of output,
 employment, wages and prices in the UK. In M. Laguna \& J. L.
 Gonz\'{a}les-Velarde (Eds.), {\em{Computing tools for modeling,
 optimization and simulation: Interfaces in computer science and
 operations research}} (2nd  ed., pp. 1--24). Boston, MA: Cambridge
 University Press.

\end{thebibliography}

\bigskip
\noindent This was produced by typing:

\begin{verbatim}
\begin{thebibliography}{12}

\bibitem[Agutter(1995)]{Agu95}
Agutter, A. J. (1995). {\em{The linguistic significance of current British
 slang}}. Unpublished doctoral dissertation, Edinburgh University, UK.

\bibitem[Aiex et al.(2000)]{aiex00}
Aiex, R. M., Pierce, I. F., Donizetti, G., {{v}on~Weber}, C. M., Bizet, G.,
 Bach, C. P. E., \ldots Dukas, P. (2000). {\em{Computing tools for
 modelling orchestral performance}} (Tech. Rep. No. DAMTP 2000/NA10).
 Cambridge, UK: University of Cambridge.

\bibitem[Barlow(1991)]{Bar91}
Barlow, D. H. (Ed.). (1991). Diagnoses, dimensions, and DSM-IV [special
issue]. {\em{Journal of Abnormal Psychology}}, {\em{100}\,}(3).

\bibitem[Ericsson \& Simon(1984)]{Eri1984}
Ericsson, K. A., \& Simon, H. A. (1984). {\em{Protocol analysis: Verbal
 reports as data}}. Cambridge, MA: MIT Press.

\bibitem[Evans(1994)]{ev94}
Evans, W. A. (1994). Approaches to intelligent information retrieval.
 {\em{Information Processing and Management}}, {\em{7\,}}(2), 147--168.

\bibitem[Glover et al.(1951)Glover, Bizet, Bach, Strauss, \& Ribeiro]{GloRib51}
Glover, F., Bizet, G., Bach, C. P. E., Strauss, R., \& Ribeiro, C. C.
 (Eds.). (1951). {\em{Lessons of the British war economy}} (2nd ed.,
 Vol. 5, pp. 1--24). Westport, CT: Greenwood Press.

\bibitem[Kern(1997)]{hk97}
Kern, H. (1997). The resurgent Japanese economy and a Japan--United
 States free trade agreement. In C. Lambert \& G. Holst (Eds.),
 {\em{4th international conference on the restructuring of the
 economic and political system in Japan and Europe}}, Milan, Italy,
 21--25 May 1996 (pp. 147--156). Singapore: World Scientific.

\bibitem[Korb(1995)]{Kor95}
Korb, K. B. (1995). Persons and things: Book review of Bringsjord on
 robot-consciousness. {\em Psycholoquy}, {\em{6\,}}(15). Retrieved
 May 20, 2004, from http://psycprints.ecs.soton.ac.uk/archive/00000462/

\bibitem[Mazzeo(1991)]{Maz91}
Mazzeo, J. (1991) {\em{Comparability of computer and paper-and-pencil
 scores}} (College Board Rep. No. 91). Princeton, NJ: Educational
 Testing Service.

\bibitem[Miller(1993)]{Mil93}
Miller, M. E. (1993). The Interactive Tester (Version 4.0) [Computer
 software]. Westminster, CA: Psytek Services.

\bibitem[Misner(1973)]{cwm73}
Misner, C. W. (1973). Efficient algorithms for layer assignment
 problems. In I. Gottlob (Ed.), {\em Gravitation in a collapsing
 Universe} (2nd ed., Vol. 5, pp. 63--83). San Francisco, CA: Freeman.

\bibitem[Pierce et al.(1976)]{PeaEtAl76}
Pierce, I. F., Donizetti, G., von Weber, C. M., Glover, F., Ogilvy,
 A. T., Clements, A. R. S., \& Ribeiro, C. C. (1976). A model of
 output, employment, wages and prices in the UK. In M. Laguna \&
 J. L. Gonz\'{a}les-Velarde (Eds.), {\em{Computing tools for
 modeling, optimization and simulation: Interfaces in computer
 science and operations research}} (2nd  ed., pp. 1--24). Boston, MA:
 Cambridge University Press.

\end{thebibliography}
\end{verbatim}
\medskip
\noindent Each entry takes the form:

\begin{verbatim}
\bibitem[short list of authors' surnames(date of publication)long list of
 authors' surnames]{key} Bibliography entry
\end{verbatim}

\noindent where `\texttt{long list of authors' surnames}' is the \emph{optional} `long' list of three, four or five surnames which enables them all to appear where the \verb"bibitem" is first cited in the text (if the long list is missing, the short list will be used instead), and `{\tt key}' is the tag that is to be used as an argument for the \verb"\cite{}" commands in the text of the article. The {\tt Bibliography entry} should be the material that is to appear in the list of references, suitably formatted. The commands \verb"\usepackage[longnamesfirst,sort]{natbib}" and \verb"\bibpunct[, ]{(}{)}{;}{a}{,}{,}" need to appear in the preamble of your .tex file in order to generate the in-text citations.

Instead of typing the bibliography by hand, you may prefer to create the list of references using a \textsc{Bib}\TeX\ database. For this we suggest using Erik Meijer's \texttt{apacite} package, which is available via CTAN. The \verb"apacite.sty", \verb"apacite.bst" and \verb"english.apc" files need to be placed in your working folder or in an appropriate directory, and the lines
%
\begin{verbatim}
\bibliographystyle{apacite}
\bibliography{tATOguide}
\end{verbatim}
%
\noindent included where the list of references should appear, where tATOguide.bib is the database of bibliographic details for the references section included with the {\itshape tATO} \LaTeX\ style guide package (to be replaced with the name of your own .bib file). The \verb"\usepackage[]{natbib}" and \verb"\bibpunct" commands are \emph{not} then required in your preamble and must be replaced with \verb"\usepackage[]{apacite}". The \verb"[natbibapa]" option is added in order to enable citation commands of the type \verb"\citep" and \verb"\citet".

The \LaTeX\ source file of your paper will extract from your .bib file only those references that are cited in that paper and list them in the References section of it. Please include a copy of your .bib file and the final generated .bbl file among your source files if your .tex file does not contain a reference list in a \texttt{thebibliography} environment.

\subsection{{\em tATO} Macros}

Table~\ref{macros} gives a list of macros for use with {\it tATO}. The list displays each macro's code and a
description/demonstration of its function.

\begin{table}
\tbl{{\it tATO} macros.}
{\begin{tabular}{@{}ll}
\toprule
$\backslash$thanks\{title-page footnote to article title & e.g. `Corresponding author's email:\\
or author\} & A.N.Author@uiowa.edu'\\\cr
$\backslash$begin\{abstract\}...$\backslash$end\{abstract\} & for
abstract on titlepage\\\\ $\backslash$bm\{math and symbols\} &
bold italic $\bm{math\;and\;symbols}$\\\cr $\backslash$bi\{text\}
& bold italic \bi{text}\\\cr $\backslash$sf\{text or upright
symbols in math mode\} & sans serif \sf{text} or
$\sf{upright\;symbols\;in\;math\;mode}$
\\\botrule
\end{tabular}}
\label{macros}
\end{table}

\section{Example of a section heading including
   {\fontencoding{T1}\scshape{small caps}}, {\bi italic},
   and bold Greek such as ${\bm\kappa}$}\label{headings}

The following code shows how to achieve this section heading:

\begin{verbatim}
\section{Example of section heading including
   {\fontencoding{T1}\scshape{small caps}}, {\bi italic},
   and bold Greek such as ${\bm\kappa}$}\label{headings}
\end{verbatim}

\section{{\textit{tATO}} journal style}

The notes given here relate to common style errors found in manuscripts, but are {\itshape not\/}
intended to be exhaustive.

\subsection{Hyphens, n-rules, m-rules and Minus Signs}\label{dashes}

\begin{enumerate}
\item[(i)] Hyphens (one dash in \TeX/\LaTeX). {\it tATO} uses hyphens for compound adjectives (e.g.\ low-density gas, least-squares fit,
two-component  model) but not for complex  units  or ranges, which could become cumbersome (e.g.\ 15~km~s$^{-1}$
feature, 100--200~$\umu$m observations).

\item[(ii)] n-rules (two dashes in \TeX/\LaTeX). These are used
(a) to denote a range (e.g.\ 1.6--2.2~$\umu$m);
(b) to denote the joining of two words of equal standing (e.g.\ Kolmogorov--Smirnov  test, Herbig--Haro object).

\item[(iii)] The  m-rule (three dashes in \TeX/\LaTeX) is used as an alternative to parentheses (e.g.\ `the results---assuming no temperature gradient---are indicative of \ldots').

\item[(iv)] The minus sign (one dash in \TeX/\LaTeX) is produced
automatically in math mode by the use of a single dash, e.g.
\begin{equation}
y_{i} \in \{-1, 1 \} \quad \forall i \in V,
\end{equation}
\noindent where $|-V|=A^2+B^2.$\medskip

\noindent is produced by

\begin{verbatim}
\begin{equation}
y_{i} \in \{-1, 1 \} \quad \forall i \in V,
\end{equation}
\noindent where $|-V|=A^2+B^2.$
\end{verbatim}

\end{enumerate}

\subsection{References}

It is important to use the correct reference style, details  of which can be found in Section~\ref{refs}.

\subsection{Maths Fonts}

Scalar  variables should be mediumface italic (e.g. $s$ for
speed); vectors should be bold italic (e.g. $\bm v$ for velocity);
matrices should be bold roman (upright) (e.g. $\bf A$), and
tensors should be bold upright sans serif (e.g. {\sffamily{\textbf
L}}). Differential d, partial differential $\upartial$, complex i,
exponential e, superscript T for `transpose', sin, cos, tan, log,
etc., should all be roman. Openface, or `blackboard', fonts can be
used, for example, for the integers $\mathbb Z$ and the reals
$\mathbb R$. Sub/superscripts that are physical variables should
be italic, while those  that are labels should be roman (e.g.\
$C_p$, $T_{\rm eff}$). Displayed equations should have end-of-line
punctuation appropriate to the running text sentence of which they
form a part.

\section{Troubleshooting}

Authors may from time to time encounter problems with the  preparation
of their papers in \LaTeX\/. The appropriate  action  to
take will depend on the nature of the problem---the following is
intended to act as a guide.
%
\begin{enumerate}
\item[(i)] If the problem is with \LaTeX\ itself, rather than with the
actual macros, please refer to an appropriate handbook for
initial advice. If the solution cannot be found, or if you
suspect that the problem lies with the macros, then please contact
Taylor \& Francis for assistance ({\tt latex.helpdesk@tandf.co.uk}).

\item[(ii)] Problems with page make-up (e.g.\ large spaces between paragraphs,
above headings, or below figures; figures/tables appearing out of order):
please do not attempt to remedy these using `hard' page make-up
commands---the typesetter will deal with such problems. (You may, if you
wish, draw attention to particular problems when submitting the final version
of your paper.)

\item[(iii)] If a required font is not available at your site, allow \TeX\
to substitute the font and specify which font you require in the
covering letter accompanying your file(s).
\end{enumerate}

\section{Fixes for coding problems}

This guide has been designed to minimize the need for user-defined macros to create special symbols. Authors
are urged, wherever possible, to use the following coding rather than create their own. This will minimize
the danger of author-defined macros being accidentally  `overridden' when the paper is typeset (see
Section~\ref{TMth}, `Typesetting mathematics' above). In cases where it is essential to create your own macros,
these should be displayed in the preamble of the source file before \verb"\begin{document}".

\begin{enumerate}
\item[(i)] Fonts in section headings and paper titles. The following are examples
of styles that sometimes prove difficult to code.

\subsection*{Paper titles:}

\hsize380pt\bf{\noindent Generalized Flory Theory at ${\bm\delta >
{\bf 50}^\circ}$}

    \noindent\normalfont is produced by
\begin{verbatim}
\title{Generalized Flory Theory at
        ${\bm\delta > {\bfseries 50}^\circ}$}
\end{verbatim}
\medskip

{\bf{\noindent Ion--ion Correlations in H\,{\sc ii} Regions}}

\noindent\normalfont is produced by
%
\begin{verbatim}
\title{Ion--ion Correlations in H\,{\sc ii} Regions}
\end{verbatim}

\stepcounter{enumi}

\item[(ii)] n-rules, m-rules, hyphens and minus signs (see Section~\ref{dashes} for
correct usage). To create the correct symbols in the sentence

\begin{quote}
The high-resolution observations were made along a line at an
angle of $-15^\circ$ (East from North) from the axis of the
jet---which runs North--South
\end{quote}

\noindent you would use the following code:

\begin{verbatim}
The high-resolution observations were made along a line at an
angle of $-15^\circ$ (East from North) from the axis of the
jet---which runs North--South
\end{verbatim}

\item[(iii)] Fonts in superscripts and subscripts. Subscripts and superscripts will automatically come  out in the correct font
and size in a math environment (e.g. enclosed by `\verb"$"'
delimiters in running text or within \verb"\[...\]" or the
`equation' environment for displayed equations). You can create
the output ${\bm k_x}$ by typing \verb"${\bm k_x}$". If the
subscripts or superscripts need to be other than italic, they
should be coded individually---see (vi) below.

\item[(iv)] Calligraphic letters (uppercase only).
Normal calligraphic can be produced with \verb"\cal" as usual (in
math mode).

\item[(v)] Automatic scaling of brackets. The codes \verb"\left" and
\verb"\right" should  be used to scale brackets automatically to
fit the equation being set. For example, to get
\[
   v = x \left( \frac{N+2}{N} \right)
\]
use the code

\begin{verbatim}
\[
   v = x \left( \frac{N+2}{N} \right)
\]
\end{verbatim}

\item[(vi)] Roman font in equations. It is often necessary to make some
symbols roman in an equation (e.g.\ units, non-variable
subscripts). For example, to get
\[
   \sigma \simeq (r/13~h^{-1}~{\rm Mpc})^{-0.9},
   \qquad \omega = \frac{N-N_{\rm s}}{N_{\rm R}}
\]
\noindent use the code

\begin{verbatim}
\[
   \sigma \simeq (r/13~h^{-1}
   ~{\rm Mpc})^{-0.9}, \qquad \omega
   =\frac{N-N_{{\rm s}}}{N_{{\rm R}}}
\]
\end{verbatim}
The \texttt{siunits} package of macros for typesetting units is also compatible with the \textit{tATO} class file.
\end{enumerate}

\section{Obtaining the tATO2e class file}\label{FTP}

\subsection{Via the Taylor \& Francis Website}

This Guide for Authors and the \texttt{tATO2e} class file may be obtained via the Instructions for Authors
on the Taylor \& Francis homepage for the journal.

Please note that the class file calls up the following open-source \LaTeX\ packages, which will, for convenience,
unpack with the downloaded Guide for Authors and class file: amsbsy.sty; amsfonts.sty; amsmath.sty; amssymb.sty; epsfig.sty; graphicx.sty; rotating.sty.
The Guide for Authors optionally calls for natbib.sty and subfigure.sty, which are also supplied for convenience.

\subsection{Via e-mail}

This Guide for Authors, the class file and the associated open-source \LaTeX\ packages are also available by
e-mail. Requests should be addressed to {\tt latex.helpdesk@tandf.co.uk} clearly stating for which journal you
require the Guide for Authors and/or class file.

\end{document}

### 2014-12-15 AO_guides/manuscriptAO.tex
Added unding section

@@
 
-\usepackage[longnamesfirst,sort]{natbib}% Citation support using natbib.sty
-\bibpunct[, ]{(}{)}{;}{a}{,}{,}% Citation support using natbib.sty
+%\usepackage[longnamesfirst,sort]{natbib}% Citation support using natbib.sty
+%\bibpunct[, ]{(}{)}{;}{a}{,}{,}% Citation support using natbib.sty
 
-%\usepackage[natbibapa]{apacite}% Citation support using apacite.sty. Commands using natbib.sty MUST be deactivated first!
+\usepackage[natbibapa]{apacite}% Citation support using apacite.sty. Commands using natbib.sty MUST be deactivated first!
 
@@
 For providing guidance with stakeholder engagement, we also thank Stephanie Chang, Jackie Yip, and Shona van Zijil de jong.
-This project is funded by the Marine Environmental Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada.
 This research was enabled in part by support provided by WestGrid (www.westgrid.ca) and Compute Canada Calcul Canada (www.computecanada.ca).
 
-\bibliographystyle{chicago}
+\section*{Funding}
+This project is funded by the Marine Environmental Observation Prediction and Response Network (MEOPAR), a Network of Centres of Excellence of Canada.
+
+\bibliographystyle{apacite}
 \bibliography{ref}

### 2015-01-02 AO_guides/manuscriptAO.tex
modified title, author order and added webtide website

@@
 
-\title{Storm surge modelling in the Strait of Georgia}
+\title{Storm surges in the Strait of Georgia simulated with a regional model}
 
 \author{Nancy Soontiens$^{1}$$^{\ast}$\thanks{$^\ast$Corresponding author's email: nsoontie@eos.ubc.ca
-\vspace{6pt}}, Susan Allen$^{1}$,  Doug Latornell$^{1}$,  Kate Le Sou\"{e}f$^{1}$, Idalia Machuca$^{1}$ \\ 
-Youyu Lu$^{2}$, Keith Thompson$^{3}$, Jean-Philippe Paquin$^{3}$, Vasily Korabel$^{3}$
+\vspace{6pt}}, Susan Allen$^{1}$,  Doug Latornell$^{1}$,  Kate Le Sou\"{e}f$^{1}$, Idalia Machuca$^{1}$,\\ Jean-Philippe Paquin$^{2}$,
+Youyu Lu$^{3}$, Keith Thompson$^{2}$, Vasily Korabel$^{2}$
 \\\vspace{6pt}  $^{1}${\em{University of British Columbia}};
-$^{2}${\em{Bedford Insitute of Oceanography}}; 
-$^{3}${\em{Dalhousie University}}\\\received{v1.2 released April 2014} }
+$^{2}${\em{Dalhousie University}};
+$^{3}${\em{Bedford Insitute of Oceanography}}\\\received{v1.2 released April 2014} }
 
@@
 
-\section{Model Configuration}\label{sec:config}
+\section{Model Description}\label{sec:config}
 
@@
 Future work will include bio-geochemical coupling and data assimilation.
+NEMO version 3.4 was used in this study.
 
@@
 The model includes two open boundaries that connect to the Pacific Ocean, the western boundary of the Strait of Juan de Fuca as well as Johnstone Strait at the north, both of which are forced with eight tidal constituents (K$_1$ ,O$_1$, P$_1$, Q$_1$, M$_2$, K$_2$, N$_2$, S$_2$), temperature and salinity climatologies, and the sea surface height anomaly.
-Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on work by \citet{foreman2000webtide}.
+Tidal heights and currents at grid points along the Juan de Fuca boundary were extracted from Webtide \citep{webtide}, an online web prediction model that provides tidal predictions for the northeast Pacific Ocean based on work by \citet{foreman2000webtide}.
 The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major M$_2$ and K$_1$ constituents.
@@
 Temperature and salinity at the Juan de Fuca boundary were taken from a weekly climatology which was created from results from a model covering the Salish Sea and the west coast of Vancouver Island \citep{massonfine2012}.
-Their results, originally on $s$-levels were interpolated onto $z$-levels and then onto the NEMO horizontal grid.
+Their results, originally on $s$-levels were interpolated onto $z$-levels and then onto the NEMO horizontal grid to give a two-dimensional forcing that varies with time.
 To prepare the climatology all years (1995-2008) were averaged and results, approximately every 15 days, were interpolated to a weekly climatology.
@@
 
-Sea surface height at the mouth of Juan de Fuca was set using values from the Tofino tide gauge \citep{DFOObservations} which is outside of, but close to, the western boundary of the model domain.
+Sea surface height was set uniformly across the mouth of Juan de Fuca using values from the Tofino tide gauge \citep{DFOObservations} which is outside of, but close to, the western boundary of the model domain.
 A monthly climatology was produced using daily averages from 2000-2010, binning them by month, averaging and setting the yearly mean to zero.
@@
 The tidal forcing and sea surface height was applied to the barotropic velocity forcing with the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}.
+The barotropic velocity induced by the surge amplitude is set to zero in the Flather scheme since this data is unavailable, which is a resonable choice since the surge amplitudes are generally much smaller than the tidal range.
 At the west, the baroclinic velocities at the boundary were set equal to the values inside the boundary (zero-gradient boundary conditions).
@@
 
-\section{Discussion}\label{sec:diss}
+\section{Conclusion and discussion}\label{sec:diss}
 In this article, we have configured a three-dimensional numerical ocean model of the Salish Sea for accurate hindcasts of storm surges in the Strait of Georgia.
@@
 This section details the procedure used to tune the tidal forcing for a good match with observations in the Strait of Georgia.
-Originally, the eight most important tidal constituents were extracted from a tide prediction tool called Webtide \citep{foreman2000webtide} at the western Juan de Fuca boundary.
+Originally, the eight most important tidal constituents were extracted from a tide prediction tool called Webtide \citep{webtide} at the western Juan de Fuca boundary based on work by \citet{foreman2000webtide}.
 At the northern boundary,  M$_2$ and K$_1$ tidal harmonics for both elevation and currents were taken from \citet{thomson1980johnstone}, as well as elevation harmonics for O$_1$ and S$_2$.

### 2015-01-02 AO_guides/manuscriptAO.tex
Clarifified inversge barometer forcing

@@
 Forecasts from the period of 2002-2012 are available.
-Additionally, the inverse barometric effect of the atmospheric pressure is included, an important consideration in storm surge modelling.
+Forcing due to the local atmospheric pressure is also included by calculating an inverse barometer sea surface height,
+\[ \eta_{IB} = -\frac{1}{g\rho_{0}}\left(P_{s}-P_0\right),\]
+where $g$ is the accelaration due to gravity, $\rho_{0}$ is a reference density set to 1035 kg m$^{-3}$, $P_{s}$ is the sea level atmospheric pressure from the atmospheric model, and $P_0$ is a reference sea level pressure set to 101,000 Pa.
+The gradient of $\eta_{IB}$  is scaled by $g$ and added to the forcing terms in the ocean momentum equations \citep{madec2012nemo}.
+As such, the contribution due to local atmospheric pressure is usually small since the spatial gradient in the sea level pressure is small compared to the model domain.
+Since the atmospheric model employs a terrain following vertical coordinate the model surface pressure has been corrected to sea level using \citep{holton1992introduction}
+\[ P_s = P_1\left(\gamma\frac{z_1}{T_1} +1 \right)^\frac{g}{\gamma R},\]
+where $P_1$ and $T_1$ are the model pressure and temperature at height $z_1$, $R$ is the ideal gas constant, and $\gamma$ is an temperature lapse rate set to 0.0098 K m$^{-1}$.
 One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.

### 2015-01-02 AO_guides/manuscriptAO.tex
edited IB comments

@@
 
-At the ocean surface, meteorological conditions and fresh water input from rivers are included.
 The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2014new}.
 Forecasts from the period of 2002-2012 are available.
-Forcing due to the local atmospheric pressure is also included by calculating an inverse barometer sea surface height,
+One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.
+In addition to momentum and heat fluxes, forcing due to the local atmospheric pressure is also included by calculating an inverse barometer sea surface height,
 \[ \eta_{IB} = -\frac{1}{g\rho_{0}}\left(P_{s}-P_0\right),\]
@@
 The gradient of $\eta_{IB}$  is scaled by $g$ and added to the forcing terms in the ocean momentum equations \citep{madec2012nemo}.
-As such, the contribution due to local atmospheric pressure is usually small since the spatial gradient in the sea level pressure is small compared to the model domain.
 Since the atmospheric model employs a terrain following vertical coordinate the model surface pressure has been corrected to sea level using \citep{holton1992introduction}
@@
 where $P_1$ and $T_1$ are the model pressure and temperature at height $z_1$, $R$ is the ideal gas constant, and $\gamma$ is an temperature lapse rate set to 0.0098 K m$^{-1}$.
-One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.
 
@@
 Monthly runoff volumes for each watershed for each year from 1970 to 2012 were acquired from \citet{morrison2011rivers}, as well as monthly averages.
-
 Freshwater runoff from each watershed was divided between the rivers in that watershed.

### 2015-01-03 AO_guides/manuscriptAO.tex
Modified presentation of Factors contributing to storm syurge section

@@
 In addition to momentum and heat fluxes, forcing due to the local atmospheric pressure is also included by calculating an inverse barometer sea surface height,
-\[ \eta_{IB} = -\frac{1}{g\rho_{0}}\left(P_{s}-P_0\right),\]
+\begin{equation}
+ \eta_{IB} = -\frac{1}{g\rho_{0}}\left(P_{s}-P_0\right), \label{eq:inverse}
+\end{equation}
 where $g$ is the accelaration due to gravity, $\rho_{0}$ is a reference density set to 1035 kg m$^{-3}$, $P_{s}$ is the sea level atmospheric pressure from the atmospheric model, and $P_0$ is a reference sea level pressure set to 101,000 Pa.
@@
 First, a general assessment of the model's skill is presented in two hindcasts during Feb 2006 and Nov 2006 in section \ref{sec:hind}.
-The Feb 2006 case is also used to determine the importance of modelling tidal forcing with the surge propagation.
 Second, the contribution to the surge due to several forcing conditions, such as atmospheric forcing and open boundary forcing, has been assessed for the Feb 2006 case in section \ref{sec:factors}.
+This case is also used to determine the importance of modelling tidal forcing with the surge propagation.
 Third, the spatial variability of the surge as it propagates through the domain is examined for a Dec 2006 storm surge in section \ref{sec:spatial}.
@@
 
+An additional storm surge hindcast for a Nov 16, 2006 storm surge is provided as another means of assessing model performance and is mentioned briefly here.
+The model performs reasonably well in this hindcast; Willmott skill scores are greater than 0.94 (Table \ref{tab:statistics}).
+The best match in maximum surge occurred at Campbell River with a 6.2 cm underestimate by the model.
+The timing of the maximum surge at Campbell River was identical with observations and too early at the other the stations.
+
+\subsection{Factors Contributing to Storm Surges}\label{sec:factors}
+
+Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented for the Feb 2006 storm surge (Figure \ref{fig:factors}).
+Several simulations were done: First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly.
+Second, a simulation with all forcing except the atmospheric forcing.
+Third, a simulation with all forcing except the sea surface height anomaly.
+Fourth, a simulation with only wind forcing, that is no anomaly and no inverse barometer effect due to pressure.
+Comparisons between these simulations indicate the relative importance the the open boundary forcing, the local wind forcing, and the local atmospheric pressure forcing.
+Finally, a simulation without tidal forcing was performed to study the importance of the tide-surge interaction.
+Contributions due to stratification and bathymetric resolution will be left to a future study.
+
+First, it is clear that the sea surface height anomaly at the open boundary contributes most significantly to the surge at each of these locations (Figure \ref{fig:factors}).
+When the sea surface height anomaly is not included as a forcing condition, the surge drops from a maximum 82 cm to 11 cm at Point Atkinson and its maximum occurs eight hours later.
+A similar drop in the maximum surge and delay in timing of the maximum surge is observed at the other locations.
+When atmospheric forcing is neglected, almost all of the surge amplitude is accounted for in the sea surface height anomaly forcing at the open boundary.
+
+%Local winds
+The contribution to the surge due to atmospheric forcing is small and appears to be geographically dependent, most likely due to the effects of the local winds.
+Removal of the atmosperhic forcing leads to a drop of about 2.5 cm in the maximum surge at Point Atkinson and 2.3 cm at Campbell River.
+By contrast, the surge amplitude increases by 4.8 cm at Victoria and 3.6 cm at Patricia Bay when atmospheric forces are neglected.
+The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in a slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
+A simulation with only wind forcing produces a surge of 7.8 cm at Point Atkinson, a small fraction of the total observed surge, and only centimetres at the other locations.
+Note that this wind-induced surge occurs well after the peak surge occurred in both the observations and simulations with full forcing.
+
+%local pressure
+Next, the surge amplitude decreases by a few centimetres at all four locations if local atmospheric pressure is not included in the surface forcing.
+Other findings have reported the inverse barometer effect due to atmospheric pressure contributes up to $^1/_3$ of the total surge amplitude \citep{murty1995storm}.
+Note that in these simulations the removal of the atmospheric pressure in the surface forcing does not entirely remove the inverese barometer effect from the model results since the sea surface height anomaly at the open boundary includes an inverse barometer component.
+As such, these simulations indicate the importance of the local atmospheric pressure forcing but not the inverse barometer entirely.
+The contribution due to the inverse barometer can be calculated with equation \ref{eq:inverse} or by the approximation that 1 hPa drop in pressure corresponds to 1 cm rise in sea level.
+In this event, the pressure at the Vancouver International Airport dropped to about 99 000 Pa \citep{ECClimateArchive}, corresponding to a 20 cm rise in sea level.
+This inverse barometer component is more inline with the findings of \citep{murty1995storm} but does not alter the conclusion that the open boundary forcing contributes most significantly.
+Further, the model calculates the local atmopsheric pressure forcing using a spatial gradient, and since the spatial scale of the low pressure system is large compared to the domain the local atmospheric pressure should not contribute greatly.
+
+These comparisons point to the importance of including, as a forcing condition at the open boundary, the surge entering the domain from the Pacific Ocean.
+Storms travelling over the Pacific Ocean towards the west coast of the northern United States and Canada can induce elevated sea levels along the coastline which then enter the Salish Sea system through the Strait of Juan de Fuca.
+In a model domain that does not include the Pacific Ocean, this effect must be included as a boundary condition.
+Previous simulations by \citet{murty1995storm} suggest that the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
+\caption{Comparison of the modelled residual for four simulations in Feb 2006: A simulation with all forcing (solid black), a simulation without atmospheric forcing (dashed green), a simulation without the sea surface height anomaly (solid red), and a simulation with only wind forcing, i.e.\ no sea surface height anomaly or pressure (dashed blue).
+}
+\label{fig:factors}
+\end{figure}
+
 Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow regions due to nonlinear effects and bottom friction \citep{bernier2007tide}, is presented for this hindcast (Figure \ref{fig:tidesurge}).
@@
 \end{figure}
-
-An additional storm surge hindcast for a Nov 16, 2006 storm surge is provided as another means of assessing model performance and is mentioned briefly here.
-The model performs reasonably well in this hindcast; Willmott skill scores are greater than 0.94 (Table \ref{tab:statistics}).
-The best match in maximum surge occurred at Campbell River with a 6.2 cm underestimate by the model.
-The timing of the maximum surge at Campbell River was identical with observations and too early at the other the stations.
-
-\subsection{Factors Contributing to Storm Surges}\label{sec:factors}
-
-Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented for the Feb 2006 storm surge (Figure \ref{fig:factors}).
-Four simulations were done: First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly.
-Second, a simulation with all forcing except the atmospheric forcing.
-Third, a simulation with all forcing except the sea surface height anomaly.
-Fourth, a simulation with only wind forcing, that is no anomaly and no inverse barometer effect due to pressure.
-It is clear that the sea surface height anomaly contributes most significantly to the surge at each of these locations (Figure \ref{fig:factors}).
-When the sea surface height anomaly is not included as a forcing condition, the surge drops from a maximum 82 cm to 11 cm at Point Atkinson and its maximum occurs eight hours later.
-A similar drop in the maximum surge and delay in timing of the maximum surge is observed at the other locations.
-The atmospheric model indicates southeasterly winds near Point Atkinson during the onset of the surge but after the peak surge has passed the winds shifted to westerly.
-A strong westerly wind would cause an elevated sea surface height at Point Atkinson as water piles against the coast, explaining the delay in peak surge when only atmospheric forcing is used since the winds shift to westerly after the peak surge has passed.
-A simulation with only wind forcing produces a surge of 7.8 cm at Point Atkinson, a small fraction of the total observed surge, and only centimetres at the other locations.
-Note that this wind-induced surge occurs well after the peak surge occurred in both the observations and simulations with full forcing.
-
-The contribution to the surge due to atmospheric forcing appears to be geographically dependent, its removal leading to a drop of about 2.5 cm in the maximum surge at Point Atkinson and 2.3 cm at Campbell River.
-The surge amplitude increases by 4.8 cm at Victoria and 3.6 cm at Patricia Bay when atmospheric forces are neglected.
-The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in a slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
-Additionally, the surge amplitude decreases by a few centimetres if the inverse barometer effect due to atmospheric pressure is not included.
-Other findings have reported the inverse barometer effect contributes up to $^1/_3$ of the total surge amplitude \citep{murty1995storm}, but only a minor effect is seen here.
-
-These comparisons point to the importance of including, as a forcing condition at the open boundary, the surge entering the domain from the Pacific Ocean.
-Storms travelling over the Pacific Ocean towards the west coast of the northern United States and Canada can induce elevated sea levels along the coastline which then enter the Salish Sea system through the Strait of Juan de Fuca.
-In a model domain that does not include the Pacific Ocean, this effect must be included as a boundary condition.
-Previous simulations by \citet{murty1995storm} suggest that the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
-\caption{Comparison of the modelled residual for four simulations in Feb 2006: A simulation with all forcing (solid black), a simulation without atmospheric forcing (dashed green), a simulation without the sea surface height anomaly (solid red), and a simulation with only wind forcing, i.e.\ no sea surface height anomaly or pressure (dashed blue).
-}
-\label{fig:factors}
-\end{figure}
+%modify figure!
 

### 2015-01-03 AO_guides/manuscriptAO.tex
reordered intro inline with Youyu's comments

@@
 It is part of a larger system of waterways collectively known as the Salish Sea and is connected to the Pacific Ocean via the Strait of Juan de Fuca and Johnstone Strait (Figure \ref{fig:domain}).
-It is on average 220 km in length and 28 km wide and its maximum depth is 420 m \citep{thomson1981oceanography}.
 At its southern end, the Strait of Georgia is connected to the Strait of Juan de Fuca through several passages, the largest being Haro Strait.
 The San Juan and Gulf Islands are scattered throughout these passages where complex bathymetry and steep topographic slopes induce vigorous tidal mixing.
-Many rivers feed fresh water into the Salish Sea system, the most significant source being from the Fraser River.
 Mixed diurnal and semi-diurnal tides are found throughout most of the Salish Sea but a semi-diurnal amphidrome is present nearby Victoria, where the tides are mainly diurnal \citep{thomson1981oceanography}.
@@
 %my attempt at physical oceanography in SoG- Susan feel free to edit.
-Several features related to the physical oceanography of this region render it a challenge to model numerically.
+%Several features related to the physical oceanography of this region render it a challenge to model numerically.
 The outflow from the Fraser River in the late spring is the primary source of fresh water within the Strait of Georgia and leads to large vertical density gradients in the summer months \citep{thomson1981oceanography}.
@@
 
+%intro to storm surges
+Many coastal communities in the Strait of Georgia are at risk to flooding and property damage caused by storm surges, a natural hazard arising from the combination of a strong wind storm and high tide.
+The low atmospheric pressure associated with a storm acts as an inverse barometer elevating the sea level.
+This effect in combination with strong winds pushing water up against the coast can cause flooding, particularly if the storm occurs during an unusually high tide.
+There is also a small contribution to increased sea level due to positive sea surface height anomalies propagating northward from the equator along the eastern Pacific coast during El Nino years \citep{strub2002altimeter}.
+Communities in the Strait of Georgia are particularly susceptible to storm surges in the winter months when there is an increased propensity for storms and a higher mean sea level caused by southeasterly winds and geostrophic adjustment \citep{danard2003storm}.
+In the summer months, the wind direction changes to northwesterly and mean sea level is lower.
+In British Columbia waters, the largest tidal ranges occur in June and December at the summer and winter solstices \citep{thomson1981oceanography}.
+The highest water levels are observed in the winter around the perigean spring tide or ``King" tide, when the sun and moon are aligned with the Earth and the moon is at its perigee.
+As such, storm surges in the Strait of Georgia are typically observed between November and February \citep{forseth2006adaptation}.
+Flooding caused by a storm surge is most likely to occur during a high spring tide \citep{abeys2011extreme}.
+
+Coastal communities can prepare for and respond to storm surge hazards using predictions from storm surge models to determine if flooding conditions are probable.
+Like coastal ocean models, some of the complexity of a storm surge model depends on the properties of the geographic location of interest.
+In shallow regions, it is important to include the effects of wetting and drying which allows grid cells to alternate between land and water depending on the modelled sea surface height and depth \citep{hubbert1999storm, weisberg2006hurricane}.
+Other regions require modelling over a very large domain since non-local effects from the open sea may influence surge propagation \citep{weisberg2006hurricane, lane2009verification}.
+Such large domains typically use nested or unstructured grids to resolve the details of complicated coastlines.
+Appropriate treatment of rivers \citep{flather1994storm}, accurate bathymetry and the inclusion surface waves \citep{xu2010storm} have also been identified as important issues.
+Most of these models neglect the effects of stratification, either by treating the ocean as a constant temperature and salinity or by using a two-dimensional depth-averaged set of equations, an assumption that is assumed to be valid.
+Further, an accurate representation of the meteorological conditions at the ocean surface is important.
+
+%storm surge modelling
+Storm surge modelling in the Strait of Georgia dates back to \citet{crean1988mathematical} in a barotropic finite difference model.
+Their work is highlighted in reviews of storm surges in Canadian waters \citep{murty1995storm, danard2003storm}.
+Compared to eastern Canada \citep{gray1984preliminary, bernier2006predicting,bobanovic2006forecasting, bernier2007tide, bernier2010tide}, relatively little has been published on storm surge modelling on the west coast of Canada in recent years.
+However, a storm surge forecasting program is currently operating over a domain that covers the entire British Columbia coastline \citep{StormBC}.
+Further, a statistical overview of past surge events in British Columbia waters suggests an increase in extreme water level events during warm El Nino periods \citep{abeys2011extreme}.
+
 %overview oF SoG/Salish Sea models
-Early work by \citet{crean1988mathematical,crean1988numerical} and \citet{stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia, and Puget Sound using a finite difference discretization of the equations of motion.
-A barotropic version of their model was used to reproduce storm surge water level elevations with reasonable accuracy, however, this barotropic version did not include the effects of fresh water entering the system from rivers.
+The early modelling work of \citet{crean1988mathematical,crean1988numerical} and \citet{stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia, and Puget Sound using a finite difference discretization of the equations of motion.
+%A barotropic version of their model was used to reproduce storm surge water level elevations with reasonable accuracy, however, this barotropic version did not include the effects of fresh water entering the system from rivers.
 Later, a baroclinic version of their model examined residual circulation and fresh water flushing rates \citep{stronach1993update, marinone1996model}.
-This early work pointed to an important connection between the model's bottom friction parametrization and the location and amplitude of the M$_2$ amphidrome.
+These models pointed to an important connection between the model's bottom friction parametrization and the location and amplitude of the M$_2$ amphidrome.
 They found that a local increase in the bottom friction coefficient, typically from 0.002 to 0.004, especially in the area of Haro Strait, was necessary in order to adequately reproduce the observed M$_2$ tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}.
@@
 
-%intro to storm surges
-Many coastal communities in the Strait of Georgia are at risk to flooding and property damage caused by storm surges, a natural hazard arising from the combination of a strong wind storm and high tide.
-The low atmospheric pressure associated with a storm acts as an inverse barometer elevating the sea level.
-This effect in combination with strong winds pushing water up against the coast can cause flooding, particularly if the storm occurs during an unusually high tide.
-There is also a small contribution to increased sea level due to positive sea surface height anomalies propagating northward from the equator along the eastern Pacific coast during El Nino years \citep{strub2002altimeter}.
-
-Communities in the Strait of Georgia are particularly susceptible to storm surges in the winter months when there is an increased propensity for storms and a higher mean sea level caused by southeasterly winds and geostrophic adjustment \citep{danard2003storm}.
-In the summer months, the wind direction changes to northwesterly and mean sea level is lower.
-In British Columbia waters, the largest tidal ranges occur in June and December at the summer and winter solstices \citep{thomson1981oceanography}.
-The highest water levels are observed in the winter around the perigean spring tide or ``King" tide, when the sun and moon are aligned with the Earth and the moon is at its perigee.
-As such, storm surges in the Strait of Georgia are typically observed between November and February \citep{forseth2006adaptation}.
-Flooding caused by a storm surge is most likely to occur during a high spring tide \citep{abeys2011extreme}.
-
-%storm surge modelling
-Coastal communities can prepare for and respond to storm surge hazards using predictions from storm surge models to determine if flooding conditions are probable.
-Like coastal ocean models, some of the complexity of a storm surge model depends on the properties of the geographic location of interest.
-In shallow regions, it is important to include the effects of wetting and drying which allows grid cells to alternate between land and water depending on the modelled sea surface height and depth \citep{hubbert1999storm, weisberg2006hurricane}.
-Other regions require modelling over a very large domain since non-local effects from the open sea may influence surge propagation \citep{weisberg2006hurricane, lane2009verification}.
-Such large domains typically use nested or unstructured grids to resolve the details of complicated coastlines.
-Appropriate treatment of rivers \citep{flather1994storm}, accurate bathymetry and the inclusion surface waves \citep{xu2010storm} have also been identified as important issues.
-Most of these models neglect the effects of stratification, either by treating the ocean as a constant temperature and salinity or by using a two-dimensional depth-averaged set of equations, an assumption that is assumed to be valid.
-Further, an accurate representation of the meteorological conditions at the ocean surface is important.
-
-Storm surge modelling in the Strait of Georgia dates back to \citet{crean1988mathematical} in a barotropic finite difference model.
-Their work is highlighted in reviews of storm surges in Canadian waters \citep{murty1995storm, danard2003storm}.
-Compared to eastern Canada \citep{gray1984preliminary, bernier2006predicting,bobanovic2006forecasting, bernier2007tide, bernier2010tide}, relatively little has been published on storm surge modelling on the west coast of Canada in recent years.
-However, a storm surge forecasting program is currently operating over a domain that covers the entire British Columbia coastline \citep{StormBC}.
-Further, a statistical overview of past surge events in British Columbia waters suggests an increase in extreme water level events during warm El Nino periods \citep{abeys2011extreme}.
-
 This article presents the development of a three-dimensional numerical ocean model for the Strait of Georgia and its accuracy in simulating storm surge hindcasts.
-This baroclinic model includes the effect of stratification and rivers, however, wetting and drying is not represented.
+To compliment previous barotropic storm surge modelling in this region, this model employs a higher resolution and includes the effect of stratification and rivers.
+However, wetting and drying is not represented.
 Model evaluation has focused on regions where the near-coast slopes are steep and so wetting and drying effects are believed to be unimportant.
@@
 The model's accuracy in producing storm surge hindcasts is also assessed in section \ref{sec:storm}.
+Several simulations comparing the effects of different forcing conditions are also discussed.
 Finally, a summary and discussion of future research goals is provided in section \ref{sec:diss}.

### 2015-01-03 AO_guides/manuscriptAO.tex
Keith't typographical comments

@@
 It is important to accurately represent the vertical mixing in this region as it sets the rate of export of fresh water through estuarine circulation \citep{masson2004modelling}.
-The strong and fast tidal currents through the narrow channels of the north, such as Discovery Passage and Johnstone Strait, set a fairly restrictive time step and impose a particular challenge for numerical models.
-Yet, several modelling efforts for this region exist, ranging in choices of grid structure, mixing parametrizations, and domain size and extent.
+The strong tidal currents through the narrow channels of the north, such as Discovery Passage and Johnstone Strait, set a fairly restrictive time step and impose a particular challenge for numerical models.
+Several modelling efforts for this region exist, varying in choices of grid structure, mixing parametrizations, and domain.
 
@@
 More recently, modelling studies have focused on baroclinic features.
-\citet{masson2004modelling} used the Princeton Ocean Model (POM) in a regional mode to examine the seasonal variability of deep water properties in the Strait of Georgia noting a strong sensitivity to vertical mixing over the sills.
+\citet{masson2004modelling} used the Princeton Ocean Model (POM) in a regional mode to examine the seasonal variability of deep water properties in the Strait of Georgia and noted a strong sensitivity to vertical mixing over the sills.
 This model employed a sigma or terrain-following vertical coordinate system leading to spurious diapycnal mixing since, numerically, the lateral mixing is computed on sigma levels which may intersect isopycnals in the presence of steep topography.
-A finite volume model on an unstructured grid by \citet{yang2010multi} involved a truncated domain with an open boundary in the northern Strait of Georgia, where under predictions in the M$_2$ tidal phase over most of their domain were suggested to be caused by an inaccurate representation of the tides at their northern boundary.
-\citet{sutherland2011model} employed the Regional Ocean Modelling System (ROMS) to study estuarine circulation in the Salish Sea with a model domain that extended outside the Strait of Juan de Fuca into the Pacific Ocean and up to the northern Strait of Georgia and noted that their semi-diurnal amplitudes were generally too low.
+%A finite volume model on an unstructured grid by \citet{yang2010multi} involved a truncated domain with an open boundary in the northern Strait of Georgia, where under predictions in the M$_2$ tidal phase over most of their domain were suggested to be caused by an inaccurate representation of the tides at their northern boundary.
+\citet{sutherland2011model} employed the Regional Ocean Modelling System (ROMS) to study estuarine circulation in the Salish Sea with a model domain that extended outside the Strait of Juan de Fuca into the Pacific Ocean and up to the northern Strait of Georgia.
+They noted that their semi-diurnal amplitudes were generally too low.
 An updated version of this model produced high temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}.
@@
 Monthly runoff volumes for each watershed for each year from 1970 to 2012 were acquired from \citet{morrison2011rivers}, as well as monthly averages.
-Freshwater runoff from each watershed was divided between the rivers in that watershed.
+Freshwater runoff from each watershed was divided among the rivers in that watershed.
 The area drained by each river was estimated from Toporama maps by the Atlas of Canada and watershed maps available on the Washington State government website.
@@
 
-Model results have been compared with observations by calculating both the total modelled water level and the modelled residual at four permanent tidal measuring stations; Point Atkinson, Victoria, Patricia Bay, and Campbell River (locations indicated in Figure \ref{fig:domain}).
+Model results have been compared with observations by calculating both the total modelled water level and the modelled residual at four permanent tidal measuring stations: Point Atkinson, Victoria, Patricia Bay, and Campbell River (locations indicated in Figure \ref{fig:domain}).
 The modelled residual is calculated as the difference in model output between a simulation with all forcing conditions and a simulation forced only with tides and rivers.
@@
 In general, there is also very good agreement between the observed and modelled residuals (Figure \ref{fig:feb2006}).
-The most noticeable difference is a 4-6 hour delay in the model's timing of the maximum residual except at Victoria where the maximum residual is in sync with observations.
+The most noticeable difference is a 4-6 hour delay in the model's timing of the maximum residual except at Victoria where the maximum residual is in phase with observations.
 The surge manifests as a broad peak in both the modelled and observed residual leading to a high probability that large residuals coincide with the high tide.
@@
 
-Next, an assessment of the factors that are most important in scenarios leading up to storm surges is presented for the Feb 2006 storm surge (Figure \ref{fig:factors}).
+Next, an assessment of the factors that are most important in the storm surge forcing conditions is presented for the Feb 2006 storm surge (Figure \ref{fig:factors}).
 Several simulations were done: First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly.
@@
 In a model domain that does not include the Pacific Ocean, this effect must be included as a boundary condition.
-Previous simulations by \citet{murty1995storm} suggest that the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.
+Previous simulations by \citet{murty1995storm} suggest that the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain, consistent with the results of the present study.
 
@@
 In contrast, the tide-surge interaction at Victoria is hardly noticeable.
-Indeed, a spatial map of the difference between the modelled residual and surge-only difference (not shown here) indicates that the tide-surge interaction is more pronounced after the San Juan and Gulf Islands.
+Indeed, a spatial map of the difference between the modelled residual and surge-only difference (not shown here) indicates that the tide-surge interaction is more pronounced north of the San Juan and Gulf Islands.
 The passages between these islands are known for vigorous tidal mixing, which may have a nonlinear effect on the surge propagation.
@@
 $\Delta \phi$ is the difference between modelled and observed phase.
-$D$ is the complex difference.
+$D$ is amplitude of the complex difference.
 The mean error and root mean square (RMS) error are calculated as the difference between the modelled quantity and observed quantity.

### 2015-01-03 AO_guides/manuscriptAO.tex
Youyu's comment about nonIB oscillations. Added references he suggested

@@
 \[ P_s = P_1\left(\gamma\frac{z_1}{T_1} +1 \right)^\frac{g}{\gamma R},\]
-where $P_1$ and $T_1$ are the model pressure and temperature at height $z_1$, $R$ is the ideal gas constant, and $\gamma$ is an temperature lapse rate set to 0.0098 K m$^{-1}$.
+where $P_1$ and $T_1$ are the model pressure and temperature at height $z_1$, $R$ is the ideal gas constant, and $\gamma$ is a temperature lapse rate set to 0.0098 K m$^{-1}$.
+Since the model domain does not include the open Pacific Ocean, pressure gradients between the open ocean and the semi-enclosed Salish Sea and their effect in the sea surface height cannot be accounted for.
+Previous studies in the Hudson Bay and Laborador Shelf system have aruged that oscillations in the sea surface height can be expained, in part, by the response to the pressure gradient between these regions \citep{wright1987influence,young1995synoptic}.
 

### 2015-01-04 AO_guides/manuscriptAO.tex
Changed rivers to green dots for better visibility

@@
 \includegraphics[scale=0.5]{Figures/bathy.pdf}
-\caption{Model domain (highlighted area) including bathymetry, rivers (blue circles), and storm surge locations of interest  (red stars).}\label{fig:domain}
+\caption{Model domain (highlighted area) including bathymetry, rivers (green circles), and storm surge locations of interest  (red stars).}\label{fig:domain}
 \end{figure}
@@
 Incoming water was assumed to be fresh and at surface temperature.
-A total of 150 rivers were parametrized by this method and their locations are indicated by the blue dots in Figure \ref{fig:domain}.
+A total of 150 rivers were parametrized by this method and their locations are indicated by the green dots in Figure \ref{fig:domain}.
 

### 2015-01-05 AO_guides/manuscriptAO.tex
Clarified corrected model as per Vasily's comment

@@
 Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can approach 40 cm during times of high tidal range, and in particular, during the storm surge season Nov-Feb.
-As such, storm surge model predictions were corrected by accounting for this error.
-Additionally, since the sea surface height is calculated about a zero reference level, the modelled water level is determined by adding a long term mean sea level at the specified location to the modelled sea surface height.
+As such, storm surge model predictions were corrected by accounting for this error; for each storm surge location of intetest, the difference between a full tidal prediction and a tidal prediction with eight constiuents is added to the modelled sea surface height.
+This correction is only applied to the modelled water level and not the modelled residuals.
+Additionally, since the modelled sea surface height is calculated about a zero reference level, the modelled water level is determined by adding a long term mean sea level at the specified location to the modelled sea surface height.
 All times are reported in UTC.
@@
 In contrast, the tide-surge interaction at Victoria is hardly noticeable.
-Indeed, a spatial map of the difference between the modelled residual and surge-only difference (not shown here) indicates that the tide-surge interaction is more pronounced north of the San Juan and Gulf Islands.
+Indeed, a spatial map of the difference between the modelled residual and surge-only difference (Figure \ref{fig:tidesurge_spatial}) indicates that the tide-surge interaction is more pronounced north of the San Juan and Gulf Islands.
 The passages between these islands are known for vigorous tidal mixing, which may have a nonlinear effect on the surge propagation.
@@
 \end{figure}
-%modify figure!
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.6]{Figures/feb2006_tidesurge_spatial.pdf}
+\caption{Spatial dependency of tide-surge interaction Difference in sea surface height between simulations with tidal forcing and without tidal forcing. Each plot is an average over 4 hours. }
+\label{fig:tidesurge_spatial}
+\end{figure}
 

### 2015-01-05 AO_guides/manuscriptAO.tex
Keith's comment about gamma^2 in December 2006 case

@@
 %overview oF SoG/Salish Sea models
-The early modelling work of \citet{crean1988mathematical,crean1988numerical} and \citet{stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia, and Puget Sound using a finite difference discretization of the equations of motion.
+The early modelling work of \citet{crean1988mathematical}, \citet{crean1988numerical} and \citet{stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia, and Puget Sound using a finite difference discretization of the equations of motion.
 %A barotropic version of their model was used to reproduce storm surge water level elevations with reasonable accuracy, however, this barotropic version did not include the effects of fresh water entering the system from rivers.
@@
 Modelled residuals are generally too low in this hindcast (Figure \ref{fig:dec2006}), however, the timing of the surge agrees well.
-All Willmott skill scores rank above 0.88 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}) but the performance has degraded when compared to the Feb 2006 and Nov 2006 cases.
+All Willmott skill scores rank above 0.88 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}).
+The $\gamm^2$ measure suggests that the standard deviation of the error in the modelled resdiuals is roughly half of the standard deviation of the observed residuals.
+Note that the performance has degraded when compared to the Feb 2006 and Nov 2006 cases.
 This could be due to a number of factors.

### 2015-01-05 AO_guides/manuscriptAO.tex
Addressed Keith's comment about gamma^2 for Feb 2006. Also tried to clarify use of Willmott skill score

@@
 where $M_i' = M_i-\bar{O}$ and $O_i'=O_i-\bar{O}$ and $\bar{O}$ is the mean of the observations.
-This score is bounded by zero and one.
-Values close to one indicate good model skill.
-These three statistics, along with the mean observed values and mean modelled values are reported in Table \ref{tab:statistics} for each hindcast.
+This score is bounded by zero and on and values close to one indicate good model skill.
+Since the Willmott skill score is nondimensional, it is an approriate statistic to report when comparing skills of different variables, in this case, the total water level and the residuals.
+Since the total water level has a larger variability than the residuals the $\gamma^2$ measure should be much lower for the total water levels.
+These three statistics, along with the mean observed values and mean modelled values (which indicate the model's bias) are reported in Table \ref{tab:statistics} for each hindcast.
 
@@
 The model performs reasonably well in this hindcast; Willmott skill scores are greater than 0.94 (Table \ref{tab:statistics}).
+Further, the $\gamma^2$ statistic suggests that the standard deviation of the error in the modelled residuals is about 40\% ($\sqrt{0.16}$) of the standard deviation of the observed residuals.
 The best match in maximum surge occurred at Campbell River with a 6.2 cm underestimate by the model.
@@
 In this event, the pressure at the Vancouver International Airport dropped to about 99 000 Pa \citep{ECClimateArchive}, corresponding to a 20 cm rise in sea level.
-This inverse barometer component is more inline with the findings of \citep{murty1995storm} but does not alter the conclusion that the open boundary forcing contributes most significantly.
+This inverse barometer contribution is more inline with the findings of \citep{murty1995storm} but does not alter the conclusion that the open boundary forcing contributes most significantly.
 Further, the model calculates the local atmopsheric pressure forcing using a spatial gradient, and since the spatial scale of the low pressure system is large compared to the domain the local atmospheric pressure should not contribute greatly.
@@
 All Willmott skill scores rank above 0.88 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}).
-The $\gamm^2$ measure suggests that the standard deviation of the error in the modelled resdiuals is roughly half of the standard deviation of the observed residuals.
+The $\gamma^2$ measure suggests that the standard deviation of the error in the modelled resdiuals is roughly half of the standard deviation of the observed residuals.
 Note that the performance has degraded when compared to the Feb 2006 and Nov 2006 cases.

### 2015-01-05 AO_guides/manuscriptAO.tex
Added youyu's comment about how surge enters domain from open water. Couldn't find the Greatbatch reference though

@@
 Next, an assessment of the factors that are most important in the storm surge forcing conditions is presented for the Feb 2006 storm surge (Figure \ref{fig:factors}).
-Several simulations were done: First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly.
+Several simulations were done: First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly at the open boundary (or remote forcing).
 Second, a simulation with all forcing except the atmospheric forcing.
-Third, a simulation with all forcing except the sea surface height anomaly.
-Fourth, a simulation with only wind forcing, that is no anomaly and no inverse barometer effect due to pressure.
-Comparisons between these simulations indicate the relative importance the the open boundary forcing, the local wind forcing, and the local atmospheric pressure forcing.
+Third, a simulation with all forcing except the remote forcing.
+Fourth, a simulation with only wind forcing, that is no remote forcing and no local atmospheric pressure.
+Comparisons between these simulations indicate the relative importance of the remote forcing, the local wind forcing, and the local atmospheric pressure forcing.
 Finally, a simulation without tidal forcing was performed to study the importance of the tide-surge interaction.
@@
 A similar drop in the maximum surge and delay in timing of the maximum surge is observed at the other locations.
-When atmospheric forcing is neglected, almost all of the surge amplitude is accounted for in the sea surface height anomaly forcing at the open boundary.
+When atmospheric forcing is neglected, almost all of the surge amplitude is accounted for in remote forcing.
 
@@
 
+A simple steady state argument balancing the wind stress with the pressure gradient can be used to calculate the slope of the sea surface in a body of water with a constant depth $D$ \citep{pugh2004changing}:
+\begin{equation}
+\text{slope} = \frac{C\rho_{air}U^2}{g\rho D},
+\end{equation}
+where $C$ is the coefficient of friction, $\rho_{air}$ is the density of air, $g$ is the acceleration due to gravity, $U$ is the wind speed, and $\rho$ is the density of water.
+Given  $C=10^{-3}$, $\rho_{air}=1.23$ kg m$^{-3}$, $\rho= 1035$  kg m$^{-3}$, wind speed $U=20$ m s$^{-1}$, and depth $D=100$ m, the slope of the sea surface is approximately $5\times10^{-7}$.
+At Point Atkinson, a westerly wind can act over a small distance of approximately 50 km, giving a sea surface elevation on the order of centimetres in agreement with the model findings.
+Since the Strait of Georgia is not a constant depth and is up to 400 m deep, this calculation is only an approximation.
+Choosing a larger depth would result in an even smaller estimate for the sea surface elevation caused by wind stresses.
+Further, northwesterly winds would induce the largest elevations at Point Atkinson due to the large distance over which the winds can act.
+
 %local pressure
@@
 In this event, the pressure at the Vancouver International Airport dropped to about 99 000 Pa \citep{ECClimateArchive}, corresponding to a 20 cm rise in sea level.
-This inverse barometer contribution is more inline with the findings of \citep{murty1995storm} but does not alter the conclusion that the open boundary forcing contributes most significantly.
+This inverse barometer contribution is more inline with the findings of \citep{murty1995storm} but does not alter the conclusion that the remote forcing contributes most significantly.
 Further, the model calculates the local atmopsheric pressure forcing using a spatial gradient, and since the spatial scale of the low pressure system is large compared to the domain the local atmospheric pressure should not contribute greatly.
@@
 Although not shown here, both models have a similar sea level atmospheric pressure at these two locations.
+Note that the HRDPS model does not employ data assimilation.
 
@@
 \end{figure}
-%don't include direction?
 
@@
 Even though the low resolution model significantly under predicts the wind speed at some locations, it performs comparably to forcing with high resolution products.
+Much of the fine scale detail offered by the high resolution atmopsheric model is averaged out since, to first order, the storm surge is a spatial integral of the wind.
 This result further demonstrates that the local wind is not the most important factor for setting up storm surges over this domain.
@@
 We have also assessed the relative importance of different forcing factors contributing to the storm surge.
-The sea surface height anomaly entering the system from the Pacific Ocean contributes most significantly to the surge amplitude, in agreement with previous studies in this region \citep{murty1995storm}.
+The sea surface height anomaly entering the system from the Pacific Ocean, or remote forcing,  contributes most significantly to the surge amplitude, in agreement with previous studies in this region \citep{murty1995storm}.
 Although the large scale wind patterns and atmospheric conditions are important in setting up the surge on the Pacific Ocean boundaries, the local winds within the Strait of Georgia are not the leading order contribution to surge amplitude.
@@
 Local winds have a significant impact on wave action which is a primary concern for flooding management.
+Effects due to stratification will also be considered in future work.
+Further, the remote forcing is applied uniformly across the open boundary and so any spatial differences in the incoming surge are not accounted for in the model.
 
@@
 Additionally, a biological-chemical modelling component will be added.
+Stratification effects on storm surge will be considered in a future study to determine if a barotropic model is reasonably accurate.
 

### 2015-01-05 AO_guides/manuscriptAO.tex
Fixed up Flather description

@@
 
-Sea surface height was set uniformly across the mouth of Juan de Fuca using values from the Tofino tide gauge \citep{DFOObservations} which is outside of, but close to, the western boundary of the model domain.
+Sea surface height anomaly (or remote forcing) was set uniformly across the mouth of Juan de Fuca using values from the Tofino tide gauge \citep{DFOObservations} which is outside of, but close to, the western boundary of the model domain.
 A monthly climatology was produced using daily averages from 2000-2010, binning them by month, averaging and setting the yearly mean to zero.
@@
 The storm surge simulations also used the sea surface height anomaly from Port Hardy, forced at the northern boundary in Johnstone Strait.
-The tidal forcing and sea surface height was applied to the barotropic velocity forcing with the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}.
-The barotropic velocity induced by the surge amplitude is set to zero in the Flather scheme since this data is unavailable, which is a resonable choice since the surge amplitudes are generally much smaller than the tidal range.
+The tidal forcing and remote forcing were applied using the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}, a radiation scheme that allows disturbances to propagate out of the domain without reflections off of the open boundary.
+The barotropic velocity component of the remote forcing was set to zero.
 At the west, the baroclinic velocities at the boundary were set equal to the values inside the boundary (zero-gradient boundary conditions).

### 2015-01-06 AO_guides/manuscriptAO.tex
Minor typographical edits

@@
 $^{2}${\em{Dalhousie University}};
-$^{3}${\em{Bedford Insitute of Oceanography}}\\\received{v1.2 released April 2014} }
+$^{3}${\em{Bedford Institute of Oceanography}}\\\received{v1.2 released April 2014} }
 
@@
 They noted that their semi-diurnal amplitudes were generally too low.
-An updated version of this model produced high temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}.
+An updated version of this model produced high surface temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}.
 Smaller domains have also been considered; notably, the passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.
@@
 
-We have used the Nucleus for European Modelling of the Ocean (NEMO) framework in its regional configuration \citep{madec2012nemo} to develop an ocean model for the Strait of Georgia and Salish Sea.
+We have used the Nucleus for European Modelling of the Ocean (NEMO) framework in its regional configuration \citep{madec2012nemo} to develop an ocean model for the Salish Sea.
 NEMO is a highly modularized tool used for studying ocean physics, ocean-ice interactions, and the bio-geochemical properties of the ocean.
@@
 \end{equation}
-where $g$ is the accelaration due to gravity, $\rho_{0}$ is a reference density set to 1035 kg m$^{-3}$, $P_{s}$ is the sea level atmospheric pressure from the atmospheric model, and $P_0$ is a reference sea level pressure set to 101,000 Pa.
+where $g$ is the acceleration due to gravity, $\rho_{0}$ is a reference density set to 1035 kg m$^{-3}$, $P_{s}$ is the sea level atmospheric pressure from the atmospheric model, and $P_0$ is a reference sea level pressure set to 101,000 Pa.
 The gradient of $\eta_{IB}$  is scaled by $g$ and added to the forcing terms in the ocean momentum equations \citep{madec2012nemo}.
@@
 where $P_1$ and $T_1$ are the model pressure and temperature at height $z_1$, $R$ is the ideal gas constant, and $\gamma$ is a temperature lapse rate set to 0.0098 K m$^{-1}$.
-Since the model domain does not include the open Pacific Ocean, pressure gradients between the open ocean and the semi-enclosed Salish Sea and their effect in the sea surface height cannot be accounted for.
-Previous studies in the Hudson Bay and Laborador Shelf system have aruged that oscillations in the sea surface height can be expained, in part, by the response to the pressure gradient between these regions \citep{wright1987influence,young1995synoptic}.
+Since the model domain does not include the open Pacific Ocean, pressure gradients between the open ocean and the semi-enclosed Salish Sea and their effect on the sea surface height cannot be accounted for.
+Previous studies in the Hudson Bay and Labrador Shelf system have argued that oscillations in the sea surface height can be explained, in part, by the response to the pressure gradient between these regions \citep{wright1987influence,young1995synoptic}.
 
@@
 
-In general, the K$_1$ tidal harmonics are well represented by the model with complex differences below 5 cm at all stations except Seymour Narrows (station 27).
+In general, the K$_1$ tidal harmonics are well-represented by the model with complex differences below 5 cm at all stations except Seymour Narrows (station 27).
 Seymour Narrows is a very constricted region in Johnstone Strait and is not well resolved by the model.
@@
 Storm surge simulations were started from rest at least three days prior to the event of interest to allow for adjustment of model currents.
-Stratification conditions from model spin-up at a time close to the event (within five days) were chosen to initialize the temperature and salinity fields.
+Stratification conditions from model spin up at a time close to the event (within five days) were chosen to initialize the temperature and salinity fields.
 Boundary conditions include atmospheric forcing and fresh water input from rivers at the surface, sea surface height anomaly from observations at the open boundary, and tidal forcing with eight tidal constituents.
@@
 Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can approach 40 cm during times of high tidal range, and in particular, during the storm surge season Nov-Feb.
-As such, storm surge model predictions were corrected by accounting for this error; for each storm surge location of intetest, the difference between a full tidal prediction and a tidal prediction with eight constiuents is added to the modelled sea surface height.
+As such, storm surge model predictions were corrected by accounting for this error; for each storm surge location of interest, the difference between a full tidal prediction and a tidal prediction with eight constituents is added to the modelled sea surface height.
 This correction is only applied to the modelled water level and not the modelled residuals.
@@
 where $M_i' = M_i-\bar{O}$ and $O_i'=O_i-\bar{O}$ and $\bar{O}$ is the mean of the observations.
-This score is bounded by zero and on and values close to one indicate good model skill.
-Since the Willmott skill score is nondimensional, it is an approriate statistic to report when comparing skills of different variables, in this case, the total water level and the residuals.
-Since the total water level has a larger variability than the residuals the $\gamma^2$ measure should be much lower for the total water levels.
+This score is bounded by zero and one and values close to one indicate good model skill.
+Since the Willmott skill score is nondimensional, it is an appropriate statistic to report when comparing skills of different variables, in this case, the total water level and the residuals.
+Since the total water level has a larger variability than the residuals the $\gamma^2$ measure should be much lower for the total water levels, hence, the Willmott skill score provides an additional point of comparison.
 These three statistics, along with the mean observed values and mean modelled values (which indicate the model's bias) are reported in Table \ref{tab:statistics} for each hindcast.
@@
 The contribution to the surge due to atmospheric forcing is small and appears to be geographically dependent, most likely due to the effects of the local winds.
-Removal of the atmosperhic forcing leads to a drop of about 2.5 cm in the maximum surge at Point Atkinson and 2.3 cm at Campbell River.
+Removal of the atmospheric forcing leads to a drop of about 2.5 cm in the maximum surge at Point Atkinson and 2.3 cm at Campbell River.
 By contrast, the surge amplitude increases by 4.8 cm at Victoria and 3.6 cm at Patricia Bay when atmospheric forces are neglected.
 The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in a slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
-A simulation with only wind forcing produces a surge of 7.8 cm at Point Atkinson, a small fraction of the total observed surge, and only centimetres at the other locations.
+A simulation with only wind forcing produces a surge of 7.8~cm at Point Atkinson, a small fraction of the total observed surge, and only centimetres at the other locations.
 Note that this wind-induced surge occurs well after the peak surge occurred in both the observations and simulations with full forcing.
@@
 Other findings have reported the inverse barometer effect due to atmospheric pressure contributes up to $^1/_3$ of the total surge amplitude \citep{murty1995storm}.
-Note that in these simulations the removal of the atmospheric pressure in the surface forcing does not entirely remove the inverese barometer effect from the model results since the sea surface height anomaly at the open boundary includes an inverse barometer component.
-As such, these simulations indicate the importance of the local atmospheric pressure forcing but not the inverse barometer entirely.
-The contribution due to the inverse barometer can be calculated with equation \ref{eq:inverse} or by the approximation that 1 hPa drop in pressure corresponds to 1 cm rise in sea level.
+Note that in these simulations the removal of the atmospheric pressure in the surface forcing does not entirely remove the inverse barometer effect from the model results since the sea surface height anomaly at the open boundary includes an inverse barometer component.
+As such, these simulations indicate the effect of the local atmospheric pressure forcing but not the inverse barometer entirely.
+The contribution due to the inverse barometer can be calculated with equation (\ref{eq:inverse}) or by the approximation that 1 hPa drop in pressure corresponds to 1 cm rise in sea level.
 In this event, the pressure at the Vancouver International Airport dropped to about 99 000 Pa \citep{ECClimateArchive}, corresponding to a 20 cm rise in sea level.
-This inverse barometer contribution is more inline with the findings of \citep{murty1995storm} but does not alter the conclusion that the remote forcing contributes most significantly.
-Further, the model calculates the local atmopsheric pressure forcing using a spatial gradient, and since the spatial scale of the low pressure system is large compared to the domain the local atmospheric pressure should not contribute greatly.
+This inverse barometer contribution is more in line with the findings of \citep{murty1995storm} but does not alter the conclusion that the remote forcing contributes most significantly.
+Further, the model calculates the local atmospheric pressure forcing using a spatial gradient, and since the spatial scale of the low pressure system is large compared to the domain the local atmospheric pressure should not contribute greatly.
 
@@
 In contrast, the tide-surge interaction at Victoria is hardly noticeable.
-Indeed, a spatial map of the difference between the modelled residual and surge-only difference (Figure \ref{fig:tidesurge_spatial}) indicates that the tide-surge interaction is more pronounced north of the San Juan and Gulf Islands.
+Indeed, a spatial map of the difference between the modelled residual and surge-only simulation (Figure \ref{fig:tidesurge_spatial}) indicates that the tide-surge interaction is more pronounced north of the San Juan and Gulf Islands.
 The passages between these islands are known for vigorous tidal mixing, which may have a nonlinear effect on the surge propagation.
@@
 \includegraphics[scale=0.6]{Figures/feb2006_tidesurge_spatial.pdf}
-\caption{Spatial dependency of tide-surge interaction Difference in sea surface height between simulations with tidal forcing and without tidal forcing. Each plot is an average over 4 hours. }
+\caption{Spatial dependency of tide-surge interaction in Feb 2006 hindcast.
+Difference in sea surface height between simulations with tidal forcing and without tidal forcing.
+Each plot is an average over 4 hours.
+}
 \label{fig:tidesurge_spatial}
@@
 All Willmott skill scores rank above 0.88 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}).
-The $\gamma^2$ measure suggests that the standard deviation of the error in the modelled resdiuals is roughly half of the standard deviation of the observed residuals.
+The $\gamma^2$ measure suggests that the standard deviation of the error in the modelled residuals is roughly half of the standard deviation of the observed residuals.
 Note that the performance has degraded when compared to the Feb 2006 and Nov 2006 cases.
@@
 Even though the low resolution model significantly under predicts the wind speed at some locations, it performs comparably to forcing with high resolution products.
-Much of the fine scale detail offered by the high resolution atmopsheric model is averaged out since, to first order, the storm surge is a spatial integral of the wind.
+Much of the fine scale detail offered by the high resolution atmospheric model is averaged out since, to first order, the storm surge is a spatial integral of the wind.
 This result further demonstrates that the local wind is not the most important factor for setting up storm surges over this domain.
@@
 The model is forced with eight tidal constituents at the western and northern open boundaries, as well as sea surface height anomaly from de-tided observations at Tofino and Port Hardy.
-Additionally, wind stress and an inverse barometer effect were forced at the ocean surface using an atmospheric reforecasting product called CGRF \citep{smith2014new}.
+Additionally, wind stress and local atmospheric pressure forcing were used at the ocean surface using an atmospheric reforecasting product called CGRF \citep{smith2014new}.
 Storm surge simulations began with zero initial conditions for the currents and sea surface height and were commenced several days prior to the event of interest to spin up the velocity fields.
@@
 As such, efforts to include wetting and drying will be considered in the future.
-Lastly, flooding effects due to wave run up is left for a more detailed inundation study with higher resolution in the near-shore areas.
+Additionally, flooding effects due to wave run up is left for a more detailed inundation study with higher resolution in the near-shore areas.
 Local winds have a significant impact on wave action which is a primary concern for flooding management.
-Effects due to stratification will also be considered in future work.
-Further, the remote forcing is applied uniformly across the open boundary and so any spatial differences in the incoming surge are not accounted for in the model.
+Further, the remote forcing is applied uniformly across the open boundary and so any spatial variability in the incoming surge are not accounted for in the model.
+Finally, stratification effects on storm surge will be considered in a future study to determine if a barotropic model is reasonably accurate.
 
@@
 Additionally, a biological-chemical modelling component will be added.
-Stratification effects on storm surge will be considered in a future study to determine if a barotropic model is reasonably accurate.
 
@@
 \appendices
-\section{Appendix}\label{sec:appendix}
+\section{}\label{sec:appendix}
 This section details the procedure used to tune the tidal forcing for a good match with observations in the Strait of Georgia.

### 2015-01-06 AO_guides/manuscriptAO.tex
moved Floats to end
Removed endfloat because labelling was messed up with appendix

@@
 \usepackage{subfigure}% Support for small, `sub' figures and tables
-
+%\usepackage[nolists,nomarkers]{endfloat}
 %\usepackage[longnamesfirst,sort]{natbib}% Citation support using natbib.sty
@@
 \label{fig:dec2006}
-\end{figure}%table instead of figure? or maybe just the surge and not water level?
+\end{figure}
 
@@
 \appendices
-\section{}\label{sec:appendix}
+\section{Details on tidal tuning}\label{sec:appendix}
 This section details the procedure used to tune the tidal forcing for a good match with observations in the Strait of Georgia.

### 2015-01-07 AO_guides/manuscriptAO.tex
Last comment about forecasts and need of results from a larger scale system
second draft comments from Youyu

@@
 One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.
-In addition to momentum and heat fluxes, forcing due to the local atmospheric pressure is also included by calculating an inverse barometer sea surface height,
+In addition to momentum and heat fluxes, forcing due to the surface atmospheric pressure is also included in the ocean momentum equations \citep{madec2012nemo}.
+In the open ocean, the surface atmospheric pressure primarily induces an inverse barometer (IB) sea surface height,
 \begin{equation}
@@
 \end{equation}
-where $g$ is the acceleration due to gravity, $\rho_{0}$ is a reference density set to 1035 kg m$^{-3}$, $P_{s}$ is the sea level atmospheric pressure from the atmospheric model, and $P_0$ is a reference sea level pressure set to 101,000 Pa.
-The gradient of $\eta_{IB}$  is scaled by $g$ and added to the forcing terms in the ocean momentum equations \citep{madec2012nemo}.
+where $g$ is the acceleration due to gravity, $\rho_{0}$ is a reference density set to 1035 kg m$^{-3}$, $P_{s}$ is the sea level atmospheric pressure from the atmospheric model, and $P_0$ is a reference sea level pressure (set to 101,000 Pa).
 Since the atmospheric model employs a terrain following vertical coordinate the model surface pressure has been corrected to sea level using \citep{holton1992introduction}
@@
 where $P_1$ and $T_1$ are the model pressure and temperature at height $z_1$, $R$ is the ideal gas constant, and $\gamma$ is a temperature lapse rate set to 0.0098 K m$^{-1}$.
-Since the model domain does not include the open Pacific Ocean, pressure gradients between the open ocean and the semi-enclosed Salish Sea and their effect on the sea surface height cannot be accounted for.
-Previous studies in the Hudson Bay and Labrador Shelf system have argued that oscillations in the sea surface height can be explained, in part, by the response to the pressure gradient between these regions \citep{wright1987influence,young1995synoptic}.
+In coastal and shelf seas, the atmospheric pressure forcing can also induce significant non-IB sea surface height, as demonstrated in previous studies of the Hudson Bay and Labrador shelf system \citep{wright1987influence,young1995synoptic}.
 
@@
 Third, a simulation with all forcing except the remote forcing.
-Fourth, a simulation with only wind forcing, that is no remote forcing and no local atmospheric pressure.
+Fourth, a simulation with only wind forcing, that is, excluding remote forcing and local atmospheric pressure.
 Comparisons between these simulations indicate the relative importance of the remote forcing, the local wind forcing, and the local atmospheric pressure forcing.
@@
 Next, the surge amplitude decreases by a few centimetres at all four locations if local atmospheric pressure is not included in the surface forcing.
-Other findings have reported the inverse barometer effect due to atmospheric pressure contributes up to $^1/_3$ of the total surge amplitude \citep{murty1995storm}.
+This drop is smaller than the IB sea level change calculated by equation (\ref{eq:inverse}), approximately 1 cm rise in sea level associated with 1 hPa drop in atmospheric pressure.
+In this event, the pressure at the Vancouver International Airport dropped to about 99 000 Pa \citep{ECClimateArchive}, corresponding to a 20 cm rise in sea level.
+This magnitude is more in line with the findings of \citep{murty1995storm} that the IB contributes up to $^1/_3$ of the total surge amplitude.
+In our model, because the influence of remote atmospheric pressure forcing is included in the sea level at the open boundary, the above tests simply suggests that the local atmopsheric pressure generates a sea level changes that is smaller than the IB response.
+This is related to the small spatial scale of the model domain.
+
+Other findings have reported the inverse barometer effect due to atmospheric pressure contributes \citep{murty1995storm}.
 Note that in these simulations the removal of the atmospheric pressure in the surface forcing does not entirely remove the inverse barometer effect from the model results since the sea surface height anomaly at the open boundary includes an inverse barometer component.
@@
 The contribution due to the inverse barometer can be calculated with equation (\ref{eq:inverse}) or by the approximation that 1 hPa drop in pressure corresponds to 1 cm rise in sea level.
-In this event, the pressure at the Vancouver International Airport dropped to about 99 000 Pa \citep{ECClimateArchive}, corresponding to a 20 cm rise in sea level.
-This inverse barometer contribution is more in line with the findings of \citep{murty1995storm} but does not alter the conclusion that the remote forcing contributes most significantly.
 Further, the model calculates the local atmospheric pressure forcing using a spatial gradient, and since the spatial scale of the low pressure system is large compared to the domain the local atmospheric pressure should not contribute greatly.
@@
 Storms travelling over the Pacific Ocean towards the west coast of the northern United States and Canada can induce elevated sea levels along the coastline which then enter the Salish Sea system through the Strait of Juan de Fuca.
-In a model domain that does not include the Pacific Ocean, this effect must be included as a boundary condition.
 Previous simulations by \citet{murty1995storm} suggest that the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain, consistent with the results of the present study.
@@
 Further, the remote forcing is applied uniformly across the open boundary and so any spatial variability in the incoming surge are not accounted for in the model.
+Since the remote forcing is an important contributing factor, any forecasting with this regional model must make use of results from a larger scale prediction system.
 Finally, stratification effects on storm surge will be considered in a future study to determine if a barotropic model is reasonably accurate.

### 2015-01-08 AO_guides/manuscriptAO.tex
Minor typo
Final comments from Youyu and the Greatbach reference.

@@
 In shallow regions, it is important to include the effects of wetting and drying which allows grid cells to alternate between land and water depending on the modelled sea surface height and depth \citep{hubbert1999storm, weisberg2006hurricane}.
-Other regions require modelling over a very large domain since non-local effects from the open sea may influence surge propagation \citep{weisberg2006hurricane, lane2009verification}.
+Other regions require modelling over a very large domain since non-local effects from the open sea may influence surge propagation \citep{mercer2002barotropic, weisberg2006hurricane, lane2009verification}.
 Such large domains typically use nested or unstructured grids to resolve the details of complicated coastlines.
@@
 Next, an assessment of the factors that are most important in the storm surge forcing conditions is presented for the Feb 2006 storm surge (Figure \ref{fig:factors}).
-Several simulations were done: First, a simulation with all the forcing conditions, including tides, atmospheric forcing, rivers, and sea surface height anomaly at the open boundary (or remote forcing).
-Second, a simulation with all forcing except the atmospheric forcing.
-Third, a simulation with all forcing except the remote forcing.
-Fourth, a simulation with only wind forcing, that is, excluding remote forcing and local atmospheric pressure.
+Several simulations were done: First, a simulation with all the forcing conditions, including tides, local atmospheric forcing (both local winds and local atmospheric pressure), and sea surface height anomaly at the open boundary (also called remote forcing) in the black curve.
+Second, a simulation with all forcing except the local atmospheric forcing in the dashed green curve.
+Third, a simulation with all forcing except the remote forcing in the red curve.
+Fourth, a simulation with only tides and local wind forcing, that is, excluding remote forcing and local atmospheric pressure forcing in the dashed blue curve.
 Comparisons between these simulations indicate the relative importance of the remote forcing, the local wind forcing, and the local atmospheric pressure forcing.
-Finally, a simulation without tidal forcing was performed to study the importance of the tide-surge interaction.
+Finally, a simulation without tidal forcing was performed to study the importance of the tide-surge interaction (Figure \ref{fig:tidesurge}).
 Contributions due to stratification and bathymetric resolution will be left to a future study.
 
-First, it is clear that the sea surface height anomaly at the open boundary contributes most significantly to the surge at each of these locations (Figure \ref{fig:factors}).
-When the sea surface height anomaly is not included as a forcing condition, the surge drops from a maximum 82 cm to 11 cm at Point Atkinson and its maximum occurs eight hours later.
+First, it is clear that remote forcing at the open boundary contributes most significantly to the surge at each of these locations (Figure \ref{fig:factors}).
+When the remote forcing is not included as a forcing condition, the surge drops from a maximum 82 cm to 11 cm at Point Atkinson and its maximum occurs eight hours later.
 A similar drop in the maximum surge and delay in timing of the maximum surge is observed at the other locations.
-When atmospheric forcing is neglected, almost all of the surge amplitude is accounted for in remote forcing.
+When local atmospheric forcing is neglected, almost all of the surge amplitude is accounted for in the remote forcing.
 
 %Local winds
-The contribution to the surge due to atmospheric forcing is small and appears to be geographically dependent, most likely due to the effects of the local winds.
-Removal of the atmospheric forcing leads to a drop of about 2.5 cm in the maximum surge at Point Atkinson and 2.3 cm at Campbell River.
-By contrast, the surge amplitude increases by 4.8 cm at Victoria and 3.6 cm at Patricia Bay when atmospheric forces are neglected.
+The contribution to the surge due to local atmospheric forcing is small and appears to be geographically dependent, most likely due to the effects of the local winds.
+Removal of the local atmospheric forcing leads to a drop of about 2.5 cm in the maximum surge at Point Atkinson and 2.3 cm at Campbell River.
+By contrast, the surge amplitude increases by 4.8 cm at Victoria and 3.6 cm at Patricia Bay.
 The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in a slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
-A simulation with only wind forcing produces a surge of 7.8~cm at Point Atkinson, a small fraction of the total observed surge, and only centimetres at the other locations.
+A simulation with only tides and local wind forcing produces a surge of 7.8~cm at Point Atkinson, a small fraction of the total observed surge, and only centimetres at the other locations.
 Note that this wind-induced surge occurs well after the peak surge occurred in both the observations and simulations with full forcing.
@@
 In this event, the pressure at the Vancouver International Airport dropped to about 99 000 Pa \citep{ECClimateArchive}, corresponding to a 20 cm rise in sea level.
-This magnitude is more in line with the findings of \citep{murty1995storm} that the IB contributes up to $^1/_3$ of the total surge amplitude.
-In our model, because the influence of remote atmospheric pressure forcing is included in the sea level at the open boundary, the above tests simply suggests that the local atmopsheric pressure generates a sea level changes that is smaller than the IB response.
+This magnitude is in line with the findings of \citep{murty1995storm} that the IB contributes up to $^1/_3$ of the total surge amplitude.
+In our model, including the influence of remote atmospheric pressure forcing (in the sea level at the open boundary) produces realistic surge amplitudes compared with the runs only including local atmospheric forcing.
+Our results suggest that the local atmospheric pressure generates a sea level change that is smaller than the IB response.
 This is related to the small spatial scale of the model domain.
-
-Other findings have reported the inverse barometer effect due to atmospheric pressure contributes \citep{murty1995storm}.
-Note that in these simulations the removal of the atmospheric pressure in the surface forcing does not entirely remove the inverse barometer effect from the model results since the sea surface height anomaly at the open boundary includes an inverse barometer component.
-As such, these simulations indicate the effect of the local atmospheric pressure forcing but not the inverse barometer entirely.
-The contribution due to the inverse barometer can be calculated with equation (\ref{eq:inverse}) or by the approximation that 1 hPa drop in pressure corresponds to 1 cm rise in sea level.
-Further, the model calculates the local atmospheric pressure forcing using a spatial gradient, and since the spatial scale of the low pressure system is large compared to the domain the local atmospheric pressure should not contribute greatly.
 
@@
 \includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
-\caption{Comparison of the modelled residual for four simulations in Feb 2006: A simulation with all forcing (solid black), a simulation without atmospheric forcing (dashed green), a simulation without the sea surface height anomaly (solid red), and a simulation with only wind forcing, i.e.\ no sea surface height anomaly or pressure (dashed blue).
+\caption{Comparison of the modelled residual for four simulations in Feb 2006: A simulation with all forcing (solid black), a simulation without local atmospheric forcing (dashed green), a simulation without remote forcing (solid red), and a simulation with only tides and local wind forcing, i.e.\ no remote forcing or local atmospheric pressure (dashed blue).
 }

### 2015-01-08 AO_guides/manuscriptAO.tex
Modified paper size and "To appear" header
Susan's comments

@@
 
-\documentclass{tATO2e}
+\documentclass[letterpaper]{tATO2e}
 
@@
 \author{Nancy Soontiens$^{1}$$^{\ast}$\thanks{$^\ast$Corresponding author's email: nsoontie@eos.ubc.ca
-\vspace{6pt}}, Susan Allen$^{1}$,  Doug Latornell$^{1}$,  Kate Le Sou\"{e}f$^{1}$, Idalia Machuca$^{1}$,\\ Jean-Philippe Paquin$^{2}$,
+\vspace{6pt}}, Susan E. Allen$^{1}$,  Doug Latornell$^{1}$,  Kate Le Sou\"{e}f$^{1}$, Idalia Machuca$^{1}$,\\ Jean-Philippe Paquin$^{2}$,
 Youyu Lu$^{3}$, Keith Thompson$^{2}$, Vasily Korabel$^{2}$
@@
 %storm surge modelling
-Storm surge modelling in the Strait of Georgia dates back to \citet{crean1988mathematical} in a barotropic finite difference model.
+Storm surge modelling in the Strait of Georgia dates back to \citet{crean1988mathematical} using a barotropic finite difference model.
 Their work is highlighted in reviews of storm surges in Canadian waters \citep{murty1995storm, danard2003storm}.
@@
 Next, section \ref{sec:model} presents an evaluation of model performance through comparisons of tidal amplitudes and phases with observations.
-The model's accuracy in producing storm surge hindcasts is also assessed in section \ref{sec:storm}.
+The model's accuracy in producing storm surge hindcasts is assessed in section \ref{sec:storm}.
 Several simulations comparing the effects of different forcing conditions are also discussed.
@@
 Although not used in the present work, NEMO's options for grid nesting and bio-geochemical coupling make it a useful tool for studying the complex physics and bio-geochemical interactions within the Strait of Georgia.
-This work focuses on validating the physical configuration of the Salish Sea model, in particular, determining appropriate forcing and boundary conditions for accurate reproduction of tidal amplitudes and phases as well as storm surge elevations.
+This work focuses on the physical configuration of the Salish Sea model, in particular, determining appropriate forcing and boundary conditions for accurate reproduction of tidal amplitudes and phases as well as storm surge elevations.
 Future work will include bio-geochemical coupling and data assimilation.
@@
 
-The modelled domain extends from the Strait of Juan de Fuca to Puget Sound to Johnstone Strait as shown in Figure \ref{fig:domain}.
+The modelled domain extends from the Strait of Juan de Fuca to Puget Sound to Johnstone Strait (Figure \ref{fig:domain}).
 Bathymetry from the Cascadia physiography data set \citep{haugerud1999digital} was smoothed to limit large changes in depth across grid cells, so that $\Delta h/\bar{h} \leq 0.8$, where $\Delta h$ is the difference in depth between two adjacent grid cells and $\bar{h}$ is their average depth.
@@
 The 40 vertical $z$-levels were stretched gradually in order to achieve higher resolution in the surface layer, with 1 m vertical grid spacing down to about 10 m in depth.
-Below 10 m the grid was stretched to a maximum grid spacing of 27 m at the lowest layer.
+Below 10 m the grid was stretched to a maximum grid spacing of 27 m in the lowest layer.
 At the bottom boundary, partial $z$-levels were utilized in order to limit large changes in bathymetry across grid cells \citep{madec2012nemo}.
@@
 where $A_0$, $A_m$, $g_0$ and $g_m$ are the observed and modelled amplitudes and phases.
-Figure \ref{fig:tides} compares modelled and observed amplitude and phase for the $M_2$ and $K_1$ tidal constituents for stations along a transect from the Strait of Juan de Fuca through to the Strait of Georgia and Johnstone Strait.
-Complex differences for these two constituents are also plotted.
+Figure \ref{fig:tides} compares modelled and observed amplitude, phase, and complex differences for the $M_2$ and $K_1$ tidal constituents for stations along a transect from the Strait of Juan de Fuca through to the Strait of Georgia and Johnstone Strait.
 Comparisons have been made with data from \citet{foreman1995tidal}, \citet{foreman2004m} and \citet{foreman2012circulation}.
@@
 Although not shown here, both models have a similar sea level atmospheric pressure at these two locations.
-Note that the HRDPS model does not employ data assimilation.
+Note that data assimilation is used in the parent model of the nested HRDPS product.
 
@@
 \section{Conclusion and discussion}\label{sec:diss}
-In this article, we have configured a three-dimensional numerical ocean model of the Salish Sea for accurate hindcasts of storm surges in the Strait of Georgia.
+In this article, we describe a three-dimensional numerical ocean model of the Salish Sea for accurate hindcasts of storm surges in the Strait of Georgia.
 The model is forced with eight tidal constituents at the western and northern open boundaries, as well as sea surface height anomaly from de-tided observations at Tofino and Port Hardy.
@@
 Hindcasts of five significant events in February 2006, November 2006, December 2006, November 2009, and December 2012 were simulated.
-In general, the model performs well at reproducing the observed water level heights and residuals, with Willmott skill scores greater than 0.88 in all cases and generally lower skill scores for the residuals.
+In general, the model performs well at reproducing the observed water level heights and residuals, with Willmott skill scores greater than 0.88 in all cases and generally higher skill scores for the water level heights.
 Further, in a single simulation, the skill scores are fairly consistent between the four stations examined.
@@
 Since the remote forcing is an important contributing factor, any forecasting with this regional model must make use of results from a larger scale prediction system.
-Finally, stratification effects on storm surge will be considered in a future study to determine if a barotropic model is reasonably accurate.
 

### 2015-01-12 AO_guides/manuscriptAO.tex
Idalia;s comments

@@
 
-Temperature and salinity at the Juan de Fuca boundary were taken from a weekly climatology which was created from results from a model covering the Salish Sea and the west coast of Vancouver Island \citep{massonfine2012}.
+Temperature and salinity at the Juan de Fuca boundary were taken from a weekly climatology which was created from results of a model covering the Salish Sea and the west coast of Vancouver Island \citep{massonfine2012}.
 Their results, originally on $s$-levels were interpolated onto $z$-levels and then onto the NEMO horizontal grid to give a two-dimensional forcing that varies with time.
@@
 Since the Willmott skill score is nondimensional, it is an appropriate statistic to report when comparing skills of different variables, in this case, the total water level and the residuals.
-Since the total water level has a larger variability than the residuals the $\gamma^2$ measure should be much lower for the total water levels, hence, the Willmott skill score provides an additional point of comparison.
+Since the total water level has a larger variability than the residuals, the $\gamma^2$ measure should be much lower for the total water levels; hence, the Willmott skill score provides an additional point of comparison.
 These three statistics, along with the mean observed values and mean modelled values (which indicate the model's bias) are reported in Table \ref{tab:statistics} for each hindcast.
@@
 The statistics are calculated with hourly output over the three-day period displayed in Figure \ref{fig:feb2006}.
-The mean absolute error in the water level ranges from 7.6 cm at Victoria to 13.1 cm at Campbell River, whereas, the mean absolute error in the residuals is smaller at all stations, however, this measure does not account for the large variability in water levels due to the tides.
+The mean absolute error in the water level ranges from 7.6 cm at Victoria to 13.1 cm at Campbell River, whereas, the mean absolute error in the residuals is smaller at all stations; however, this measure does not account for the large variability in water levels due to the tides.
 A statistic that does take into account this variability is $\gamma^2$ which suggests very good agreement in both the predicted water levels and residuals.
@@
 In this event, the pressure at the Vancouver International Airport dropped to about 99 000 Pa \citep{ECClimateArchive}, corresponding to a 20 cm rise in sea level.
-This magnitude is in line with the findings of \citep{murty1995storm} that the IB contributes up to $^1/_3$ of the total surge amplitude.
+This magnitude is in line with the findings of \citet{murty1995storm} that the IB contributes up to $^1/_3$ of the total surge amplitude.
 In our model, including the influence of remote atmospheric pressure forcing (in the sea level at the open boundary) produces realistic surge amplitudes compared with the runs only including local atmospheric forcing.
@@
 Second, the anomaly forcing at the open boundaries could be too low.
-The observed residual at Tofino, BC (a point near by but outside of the domain) is up to 10 cm lower than the observed residual at Neah Bay, WA (a point near by but inside of the domain).
+The observed residual at Tofino, BC (a point nearby but outside of the domain) is up to 10 cm lower than the observed residual at Neah Bay, WA (a point nearby but inside of the domain).
 Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the cause.
@@
 Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions or a partial reflection from the mainland coast.
-Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important north the San Juan and Gulf Islands.
+Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important north of the San Juan and Gulf Islands.
 The modelled surge amplitude increased by about 10\% at Point Atkinson when the tide-surge interaction was neglected in one example.

### 2015-01-15 AO_guides/manuscriptAO.tex
Keith's comments round 2

@@
 The model is evaluated through comparisons of tidal elevation harmonics and storm surge water levels with observations.
-Important forcing factors contributing to storm surges have been assessed, finding that the surge entering the domain from the Pacific Ocean contributes most significantly.
+Important forcing factors contributing to storm surges have been assessed.
+It is shown that the surge entering the domain from the Pacific Ocean contributes most significantly to the variability.
 Comparisons between simulations with high resolution and low resolution atmospheric forcing further emphasize this point.
@@
 %intro to storm surges
-Many coastal communities in the Strait of Georgia are at risk to flooding and property damage caused by storm surges, a natural hazard arising from the combination of a strong wind storm and high tide.
+Many coastal communities in the Strait of Georgia are at risk of flooding and property damage caused by storm surges, a natural hazard arising from the combination of a strong wind storm and high tide.
 The low atmospheric pressure associated with a storm acts as an inverse barometer elevating the sea level.
@@
 The highest water levels are observed in the winter around the perigean spring tide or ``King" tide, when the sun and moon are aligned with the Earth and the moon is at its perigee.
-As such, storm surges in the Strait of Georgia are typically observed between November and February \citep{forseth2006adaptation}.
+Storm surges in the Strait of Georgia are typically observed between November and February \citep{forseth2006adaptation}.
 Flooding caused by a storm surge is most likely to occur during a high spring tide \citep{abeys2011extreme}.
@@
 Such large domains typically use nested or unstructured grids to resolve the details of complicated coastlines.
-Appropriate treatment of rivers \citep{flather1994storm}, accurate bathymetry and the inclusion surface waves \citep{xu2010storm} have also been identified as important issues.
-Most of these models neglect the effects of stratification, either by treating the ocean as a constant temperature and salinity or by using a two-dimensional depth-averaged set of equations, an assumption that is assumed to be valid.
-Further, an accurate representation of the meteorological conditions at the ocean surface is important.
+Appropriate treatment of rivers \citep{flather1994storm}, accurate bathymetry and the inclusion surface waves \citep{xu2010storm} have also been identified as important.
+Most of these models neglect the effects of stratification by treating the ocean as a constant temperature and salinity.
+An accurate forecast of the meteorological conditions at the ocean surface is of course essential in the generation of useful storm surge forecasts.
 
 %storm surge modelling
-Storm surge modelling in the Strait of Georgia dates back to \citet{crean1988mathematical} using a barotropic finite difference model.
+Storm surge modelling in the Strait of Georgia dates back to \citet{crean1988mathematical} who used a barotropic finite difference model.
 Their work is highlighted in reviews of storm surges in Canadian waters \citep{murty1995storm, danard2003storm}.
@@
 These models pointed to an important connection between the model's bottom friction parametrization and the location and amplitude of the M$_2$ amphidrome.
-They found that a local increase in the bottom friction coefficient, typically from 0.002 to 0.004, especially in the area of Haro Strait, was necessary in order to adequately reproduce the observed M$_2$ tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}.
-Without local adjustments to the bottom friction their M$_2$ amplitudes were too low and the M$_2$ tidal phases were too large in the Strait of Georgia.
-Other studies over a similar domain, such as the barotropic finite element model of \citet{foreman1995tidal}, have avoided local increases in bottom friction.
-\citet{foreman2004m} used data assimilation techniques to study M$_2$ tidal dissipation without local changes to the bottom friction in a two-dimensional finite element model that did not include baroclinic effects.
+They found that a local increase in the bottom friction coefficient, typically from 0.002 to 0.004 in the baroclinic model, especially in the vicinity of Haro Strait, was necessary in order to adequately reproduce the observed M$_2$ tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}.
+Without local adjustments to the bottom friction their M$_2$ amplitudes were too low and the M$_2$ tidal phases were too large in the Strait of Georgia \citet{crean1988mathematical} and local adjustments were applied in both their barotropic and baroclinic models.
+Other studies over a similar domain, such as the barotropic finite element model of \citet{foreman1995tidal}, have avoided local increases in bottom friction although their domain did not include Puget Sound and the northern Strait of Georgia.
+\citet{foreman2004m} assimilated tide gauge harmonics to study M$_2$ tidal dissipation without local changes to the bottom friction in a two-dimensional finite element model that did not include baroclinic effects.
 
@@
 
-This article presents the development of a three-dimensional numerical ocean model for the Strait of Georgia and its accuracy in simulating storm surge hindcasts.
-To compliment previous barotropic storm surge modelling in this region, this model employs a higher resolution and includes the effect of stratification and rivers.
+This article presents the development of a three-dimensional numerical ocean model for the Strait of Georgia and its skill in simulating storm surge hindcasts.
+To complement previous barotropic storm surge modelling in this region, this model employs a higher resolution and includes the effect of stratification and rivers.
 However, wetting and drying is not represented.
@@
 A detailed overview of the model configuration and domain is provided in section \ref{sec:config}.
-This section also describes open and surface boundary conditions as well as choices in grid structure and mixing parametrizations.
+That section also describes open and surface boundary conditions as well as choices in grid structure and mixing parametrizations.
 Next, section \ref{sec:model} presents an evaluation of model performance through comparisons of tidal amplitudes and phases with observations.
-The model's accuracy in producing storm surge hindcasts is assessed in section \ref{sec:storm}.
+The model's skill in producing storm surge hindcasts is assessed in section \ref{sec:storm}.
 Several simulations comparing the effects of different forcing conditions are also discussed.
@@
 We have used the Nucleus for European Modelling of the Ocean (NEMO) framework in its regional configuration \citep{madec2012nemo} to develop an ocean model for the Salish Sea.
-NEMO is a highly modularized tool used for studying ocean physics, ocean-ice interactions, and the bio-geochemical properties of the ocean.
+NEMO is a highly modular tool used for studying ocean physics, ocean-ice interactions, and the bio-geochemical properties of the ocean.
 NEMO's ocean core solves the three-dimensional hydrostatic equations of motion for an incompressible fluid under the Boussinesq approximation on a structured computational grid.
@@
 Below 10 m the grid was stretched to a maximum grid spacing of 27 m in the lowest layer.
-At the bottom boundary, partial $z$-levels were utilized in order to limit large changes in bathymetry across grid cells \citep{madec2012nemo}.
+At the bottom boundary, partial $z$-levels were used in order to limit large changes in bathymetry across grid cells \citep{madec2012nemo}.
 
@@
 Tidal predictions were generated with a tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year prior to the surge event to avoid over predicted constituents during high surge years.
-The storm surge simulations also used the sea surface height anomaly from Port Hardy, forced at the northern boundary in Johnstone Strait.
+The storm surge simulations also used the sea surface height anomaly from Port Hardy (50.72 $^\circ$ N, 127.49 $^\circ$ W), forced at the northern boundary in Johnstone Strait.
 The tidal forcing and remote forcing were applied using the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}, a radiation scheme that allows disturbances to propagate out of the domain without reflections off of the open boundary.
@@
 At the west, the baroclinic velocities at the boundary were set equal to the values inside the boundary (zero-gradient boundary conditions).
-This scheme is not part of core NEMO 3.4, the version used here.
+This scheme is not part of core NEMO 3.4 (the version used here).
 Zero gradient conditions were chosen because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain.
@@
 Forecasts from the period of 2002-2012 are available.
-One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.
+One additional simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.
 In addition to momentum and heat fluxes, forcing due to the surface atmospheric pressure is also included in the ocean momentum equations \citep{madec2012nemo}.
@@
 
-Initial conditions for temperature and salinity were taken from a CTD cast in the middle Strait of Georgia taken in Sept 2002 \citep{pawlowiczetal2007}.
+Initial conditions for temperature and salinity came from a CTD cast in the middle Strait of Georgia taken in Sept 2002 \citep{pawlowiczetal2007}.
 Conditions were initially uniform horizontally and velocity was initialized at zero.
-The model was spun up for a 15.5 months from the initial conditions above, starting Sept 16, 2002, using atmospheric forcing from 2002-2003, climatological temperature and salinity and sea surface height at the boundaries, with tides and climatological river output.
+The model was spun up for 15.5 months from the initial conditions above, starting Sept 16, 2002, using atmospheric forcing from 2002-2003, climatological temperature and salinity and sea surface height at the boundaries, with tides and climatological river output.
 All storm surge runs were started at least three days prior to the event of interest with zero initial velocities and sea surface height and a stratification profile from model spin up.
@@
 Much of the fine scale detail offered by the high resolution atmospheric model is averaged out since, to first order, the storm surge is a spatial integral of the wind.
-This result further demonstrates that the local wind is not the most important factor for setting up storm surges over this domain.
+This result further demonstrates that the local wind is not the most important factor for setting up storm surges over this domain; however, local wind patterns contribute to wave conditions which can impact flooding and are not accounted for in this model.
 
@@
 
-Comparisons between high resolution and low resolution atmospheric forcing further suggest that the local winds are not the leading order factor in setting up large surge amplitudes.
+Comparisons between high resolution and low resolution atmospheric forcing further suggest that local winds are not the leading order factor in setting up large surge amplitudes.
 The low resolution model significantly underestimated wind speeds in certain areas of the domain, however, it performed just as accurately in reproducing surge amplitudes as the high resolution product.

### 2015-01-19 AO_guides/manuscriptAO.tex
one last read through

@@
 They found that a local increase in the bottom friction coefficient, typically from 0.002 to 0.004 in the baroclinic model, especially in the vicinity of Haro Strait, was necessary in order to adequately reproduce the observed M$_2$ tidal amplitudes and phase lags in the Strait of Georgia \citep{stronach1993update}.
-Without local adjustments to the bottom friction their M$_2$ amplitudes were too low and the M$_2$ tidal phases were too large in the Strait of Georgia \citet{crean1988mathematical} and local adjustments were applied in both their barotropic and baroclinic models.
+Without local adjustments to the bottom friction their M$_2$ amplitudes were too low and the M$_2$ tidal phases were too large in the Strait of Georgia \citep{crean1988mathematical}.
+Local adjustments to bottom friction were applied in both their barotropic and baroclinic models.
 Other studies over a similar domain, such as the barotropic finite element model of \citet{foreman1995tidal}, have avoided local increases in bottom friction although their domain did not include Puget Sound and the northern Strait of Georgia.

### 2015-01-20 cover.tex
cover letter

\documentclass{letter}
\usepackage{fullpage}
\signature{Nancy Soontiens}
\address{Department of Earth, Ocean, and Atmospheric Sciences\\ University of British Columbia \\ Vancouver, BC
\\ nsoontie@eos.ubc.ca}
\begin{document}

\begin{letter}{}

\opening{Dear Editor,}

On my behalf of my co-authors I am submitting an article titled ``Storm surges in the Strait of Georgia simulated with a regional model" for consideration for publication in Atmosphere-Ocean. This article presents research undertaken by scientists collaborating across Canada to produce a new regional ocean model of the Salish Sea.  We believe that this topic is interesting and relevant to the readership of Atmosphere-Ocean, particularly those readers engaged in regional ocean modelling in Canada.

I look forward to hearing from you at your earliest convenience.

\closing{Sincerely,}
\end{letter}
\end{document}

### 2015-04-28 AO_guides/manuscriptAO.tex
Modifed weather figure caption

@@
 Observed winds shown in solid black, low resolution CGRF model in dashed black and high resolution model in grey.
-Observations are from the Environment Canada Climate Data archives \citep{ECClimateArchive}.}
+Observations {\color{red} from Sandheads and Point Atkinson} \are from the Environment Canada Climate Data archives \citep{ECClimateArchive}.
+{\color{red} Observations at the NOAA Buoy are found at this url: http://www.ndbc.noaa.gov/station_page.php?station=46087.
+Wind direction represents the direction that the wind is headed.} }
 \label{fig:dec2012_weather}

### 2015-05-04 AO_guides/manuscriptAO.tex
formatting data files

--- 2015-01-19 AO_guides/manuscriptAO.tex
@@
 \usepackage[natbibapa]{apacite}% Citation support using apacite.sty. Commands using natbib.sty MUST be deactivated first!
+\usepackage{color}
+\usepackage{hyperref}
 
@@
 \hline 
-Feb 2006       &                &                      &                      &            &        &                &                        &                        &           &           \\
-Point Atkinson & 10.5           & 3.46                 & 3.54                 &   0.012    & 0.996  & 6.2            &  0.273                 & 0.309                  & 0.065     & 0.979     \\
-Victoria       &  7.6           & 2.18                 & 2.19                 &   0.032    & 0.992  & 6.1            &  0.235                 & 0.278                  & 0.085     & 0.970     \\
-Patricia Bay   &  9.1           & 2.60                 & 2.64                 &   0.024    & 0.992  & 6.8            &  0.223                 & 0.278                  & 0.070     & 0.969     \\
-Campbell River & 13.1           & 3.20                 & 3.21                 &   0.041    & 0.990  & 5.8            &  0.303                 & 0.300                  & 0.079     & 0.978     \\
+Feb 2006& & & & & & & & & &\\ 
+Point Atkinson & 11.6& 3.458& 3.558& 0.012& 0.994& 5.5& 0.318& 0.33& 0.058& 0.985\\
+Victoria       & 8.2& 2.182& 2.205& 0.035& 0.99& 5.3& 0.276& 0.298& 0.075& 0.979\\
+Patricia Bay   & 10.4& 2.601& 2.678& 0.025& 0.99& 5.5& 0.275& 0.298& 0.066& 0.981\\
+Campbell River & 12.2& 3.202& 3.235& 0.034& 0.991& 7.0& 0.27& 0.317& 0.066& 0.973\\
 \hline
-Nov 2006       &                &                      &                      &            &        &                &                        &                        &           &           \\
-Point Atkinson &  9.8           &  3.42                & 3.43                 &   0.034    & 0.991  &  7.7           &  0.290                 & 0.247                  &  0.128     & 0.950 \\
-Victoria       &  9.1           &  2.16                & 2.11                 &   0.070    & 0.978  &  6.2           &  0.222                 & 0.202                  &  0.171     & 0.946 \\
-Patricia Bay   & 10.0           &  2.57                & 2.53                 &   0.049    & 0.985  &  6.6           &  0.235                 & 0.203                  &  0.144     & 0.947 \\
-Campbell River & 10.7           &  3.20                & 3.26                 &   0.037    & 0.989  &  6.2           &  0.254                 & 0.236                  &  0.140     & 0.952 \\
+Nov 2006& & & & & & & & & & \\
+Point Atkinson & 10.2& 3.422& 3.499& 0.026& 0.99& 6.7& 0.32& 0.305& 0.128& 0.96\\
+Victoria       & 7.8& 2.158& 2.182& 0.065& 0.982& 4.7& 0.246& 0.26& 0.134& 0.961\\
+Patricia Bay   & 9.4& 2.565& 2.634& 0.043& 0.983& 5.9& 0.258& 0.261& 0.152& 0.953\\
+Campbell River & 9.9& 3.201& 3.243& 0.032& 0.991& 6.3& 0.29& 0.293& 0.153& 0.951\\
 \hline
-Dec 2006       &                &                      &                      &            &        &                &                        &                        &           &           \\
-Point Atkinson & 10.5           &  3.45                & 3.47                 &   0.025    & 0.994  & 11.1           &  0.212                 & 0.186                  &  0.275     & 0.891 \\
-Victoria       &  9.6           &  2.23                & 2.16                 &   0.042    & 0.984  &  7.9           &  0.190                 & 0.156                  &  0.223     & 0.911 \\
-Patricia Bay   &  9.2           &  2.63                & 2.58                 &   0.031    & 0.990  &  9.5           &  0.193                 & 0.149                  &  0.260     & 0.889 \\
-Campbell River & 10.6           &  3.23                & 3.28                 &   0.030    & 0.992  &  9.4           &  0.191                 & 0.163                  &  0.247     & 0.905 \\
+Dec 2006& & & & & & & & & & \\
+Point Atkinson & 12.9& 3.452& 3.531& 0.028& 0.99& 10.0& 0.336& 0.33& 0.272& 0.895\\
+Victoria       & 8.1& 2.229& 2.226& 0.046& 0.988& 7.2& 0.306& 0.3& 0.227& 0.916\\
+Patricia Bay   & 10.2& 2.628& 2.674& 0.035& 0.989& 8.8& 0.308& 0.293& 0.255& 0.896\\
+Campbell River & 11.3& 3.234& 3.265& 0.034& 0.991& 8.2& 0.311& 0.304& 0.235& 0.912\\
 \hline
-Nov 2009       &                &                      &                      &            &        &                &                        &                        &           &           \\
-Point Atkinson & 14.0           &  3.43                & 3.54                 &   0.012    & 0.995  &  5.8           &  0.295                 & 0.316                  &  0.200     & 0.926 \\
-Victoria       &  9.9           &  2.25                & 2.23                 &   0.024    & 0.988  &  4.5           &  0.312                 & 0.280                  &  0.211     & 0.911 \\
-Patricia Bay   & 10.3           &  2.59                & 2.64                 &   0.019    & 0.997  &  5.6           &  0.265                 & 0.275                  &  0.230     & 0.918 \\
-Campbell River & 13.6           &  3.22                & 3.26                 &   0.019    & 0.991  &  4.5           &  0.313                 & 0.297                  &  0.182     & 0.934 \\
+Nov 2009& & & & & & & & & & \\
+Point Atkinson & 14.5& 3.426& 3.55& 0.01& 0.995& 8.3& 0.344& 0.397& 0.204& 0.897\\
+Victoria       & 10.9& 2.249& 2.24& 0.029& 0.992& 4.4& 0.365& 0.361& 0.201& 0.933\\
+Patricia Bay   & 12.0& 2.589& 2.685& 0.021& 0.991& 7.4& 0.307& 0.356& 0.216& 0.894\\
+Campbell River & 11.6& 3.219& 3.247& 0.016& 0.996& 6.9& 0.341& 0.38& 0.188& 0.915\\
 \hline
-Dec 2012-CGRF &                &                      &                      &            &        &                &                        &                        &           &           \\ 
-Point Atkinson &  9.1           &  3.36                & 3.40                 &   0.006    & 0.998  &  5.0           &  0.300                 & 0.290                  &  0.161     & 0.949 \\
-Victoria       & 14.6           &  2.15                & 2.08                 &   0.039    & 0.988  &  4.8           &  0.309                 & 0.274                  &  0.150     & 0.937 \\
-Patricia Bay   & 14.2           &  2.55                & 2.51                 &   0.026    & 0.992  &  4.7           &  0.300                 & 0.274                  &  0.140     & 0.945 \\
-Campbell River & 11.6           &  3.06                & 3.09                 &   0.013    & 0.996  &  5.7           &  0.241                 & 0.278                  &  0.182     & 0.921 \\
+Dec 2012-CGRF& & & & & & & & & & \\
+Point Atkinson & 11.0& 3.362& 3.432& 0.007& 0.997& 5.7& 0.274& 0.273& 0.219& 0.929\\
+Victoria       & 12.9& 2.146& 2.111& 0.036& 0.99& 4.7& 0.282& 0.256& 0.175& 0.934\\
+Patricia Bay   & 12.8& 2.551& 2.571& 0.024& 0.993& 5.2& 0.275& 0.255& 0.185& 0.934\\
+Campbell River & 12.1& 3.064& 3.106& 0.013& 0.996& 8.4& 0.188& 0.263& 0.216& 0.849\\
 \hline
-Dec 2012-HRDPS &                &                      &                      &            &        &                &                        &                        &           &           \\
-Point Atkinson &  8.8           &  3.35                & 3.39                 &   0.006    & 0.998  &  5.4           &  0.300                 & 0.278                  &  0.163     & 0.943 \\
-Victoria       & 14.3           &  2.15                & 2.09                 &   0.038    & 0.988  &  4.6           &  0.309                 & 0.277                  &  0.145     & 0.941 \\
-Patricia Bay   & 14.0           &  2.55                & 2.52                 &   0.025    & 0.993  &  4.5           &  0.300                 & 0.278                  &  0.130     & 0.953 \\
-Campbell River & 11.6           &  3.06                & 3.09                 &   0.014    & 0.993  &  5.6           &  0.241                 & 0.275                  &  0.183     & 0.924 \\
+Dec 2012-HRDPS& & &  &  &  &  &  &  &  &  \\
+Point Atkinson &  10.5&  3.362&  3.42&  0.007&  0.998&  6.2&  0.274&  0.26&  0.231&  0.921\\
+Victoria       &  12.6&  2.146&  2.113&  0.035&  0.99&  4.6&  0.282&  0.258&  0.171&  0.937\\
+Patricia Bay   &  12.6& 2.551& 2.575& 0.024& 0.993& 5.0& 0.275& 0.26& 0.175& 0.94\\
+Campbell River & 12.1& 3.064& 3.104& 0.014& 0.996& 8.1& 0.188& 0.26& 0.218& 0.854\\
 \hline
 \hline
-Averages       &                &                      &                      &            &        &                &                        &                        &           &         \\
-Point Atkinson &	10.8	& 3.42	& 3.48	& 0.018	& 0.995	& 7.2	& 0.274	& 0.270	& 0.166	 & 0.939 \\
-Victoria       &	10.2	& 2.19	& 2.15	& 0.041	& 0.986	& 5.9	& 0.254	& 0.238	& 0.168	 & 0.935 \\
-Patricia Bay   &	10.6	& 2.59	& 2.58	& 0.030	& 0.991	& 6.6	& 0.243	& 0.236	& 0.169	 & 0.934 \\
-Campbell River &	11.9	& 3.18	& 3.22	& 0.028	& 0.992	& 6.3	& 0.260	& 0.255	& 0.166  & 0.938 \\
+Averages& & & & & & & & & & \\
+Point Atkinson & 11.6&	3.401&	3.483&	0.013&	0.995&	6.7&	0.302&	0.307	&0.192	&0.932\\
+Victoria       &10.8&	2.175	&2.162	&0.04	&0.989	&5.1	&0.29	&0.28&	0.167&	0.94\\
+Patricia Bay   &11.6&	2.573&	2.619&	0.028&	0.991&	6.0&	0.281&	0.279&	0.179&	0.932\\
+Campbell River & 11.6&	3.139&	3.177&	0.021&	0.994&	7.8&	0.246&	0.293&	0.188&	0.893\\
 \hline
@@
 \includegraphics[scale=0.6]{Figures/dec2012_weather.pdf}
-\caption{Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson (upper) and Sandheads (lower).
+\caption{{\color{red}Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson (top), Sandheads (middle) and NOAA Buoy 46087 (bottom).
 Observed winds shown in solid black, low resolution CGRF model in dashed black and high resolution model in grey.
-Observations are from the Environment Canada Climate Data archives \citep{ECClimateArchive}.}
+Observations from Sandheads and Point Atkinson are from the Environment Canada Climate Data archives \citep{ECClimateArchive}.
+Observations at the NOAA Buoy are found at this url: \url{http://www.ndbc.noaa.gov/station_page.php?station=46087}.
+Wind direction represents the direction that the wind is headed.} }
 \label{fig:dec2012_weather}

### 2015-05-04 AO_guides/manuscriptAO.tex
added peak table

@@
 \usepackage{color}
-\usepackage{hyperref}
+\usepackage{nohyperref}
+\usepackage{url}
 
@@
 %I also have standard deviations, correlations, ...
+\begin{table}[h]
+\centering 
+\tbl{{\color{red}Summary of maximum water level, maximum surge amplitude and model delay for all simulations.
+The model time delay is the number of hours between the maximum in the observations and the maximum in the model.
+A negative number means the model's maximum was later than the observations.
+Values calucated over the same periods discusses in Table \ref{tab:statistics}.
+} }
+{\begin{tabular}{|l |c c c | c c c|} 
+\hline 
+& \multicolumn{3}{|c|}{Water Level ($\eta$)}        & \multicolumn{3}{|c|}{Residual ($\delta$)} \\ 
+\hline 
+Location       & Max obs (m) & Max mod (m) & Delay (hours) &Max obs (m) & Max mod (m) & Delay (hours) \\
+\hline
+Feb 2006&  &  &  &  &  & \\
+Point Atkinson  & 5.460 & 5.588 & 0 & 0.844 & 0.841 & -3\\
+Victoria        & 3.440 & 3.415 & -1 & 0.700 & 0.745 & 1\\
+Patricia Bay    & 4.250 & 4.179 & -1 & 0.781 & 0.773 & -2\\
+Campbell River  & 5.100 & 5.170 & -1 & 0.876 & 0.815 & -5\\
+\hline
+Nov 2006 &  &  &  &  &  & \\
+Point Atkinson  & 5.010 & 5.040 & 0 & 0.880 & 0.725 & 1\\
+Victoria        & 2.995 & 2.968 & 1 & 0.664 & 0.563 & 5\\
+Patricia Bay    & 3.840 & 3.743 & 0 & 0.709 & 0.607 & 2\\
+Campbell River  & 4.750 & 4.739 & 0 & 0.709 & 0.666 & -1\\
+\hline
+Dec 2006 &  &  &  &  &  & \\
+Point Atkinson  & 4.565 & 4.654 & 0 & 0.944 & 0.699 & 0\\
+Victoria        & 2.965 & 3.005 & -1 & 0.737 & 0.580 & 1\\
+Patricia Bay    & 3.505 & 3.480 & 0 & 0.859 & 0.613 & 1\\
+Campbell River  & 4.295 & 4.389 & -1 & 0.823 & 0.626 & 0\\
+\hline
+Nov 2009 &  &  &  &  &  & \\
+Point Atkinson  & 5.150 & 5.272 & 0 & 0.699 & 0.587 & -7\\
+Victoria        & 3.195 & 3.144 & -4 & 0.718 & 0.540 & -8\\
+Patricia Bay    & 3.985 & 3.902 & 0 & 0.637 & 0.541 & -8\\
+Campbell River  & 4.680 & 4.782 & 0 & 0.664 & 0.588 & -1\\
+\hline
+Dec 2012-CRGF &  &  &  &  &  & \\
+Point Atkinson  & 5.345 & 5.416 & 0 & 0.663 & 0.555 & -4\\
+Victoria        & 3.145 & 3.117 & 0 & 0.663 & 0.518 & -4\\
+Patricia Bay    & 4.115 & 3.975 & 0 & 0.632 & 0.533 & -4\\
+Campbell River  & 4.805 & 4.900 & 0 & 0.578 & 0.507 & -6\\
+\hline
+Dec 2012-HRDPS &  &  &  &  &  & \\
+Point Atkinson  & 5.345 & 5.414 & 0 & 0.663 & 0.555 & -4\\
+Victoria        & 3.145 & 3.123 & 0 & 0.663 & 0.518 & -4\\
+Patricia Bay    & 4.115 & 3.984 & 0 & 0.632 & 0.533 & -4\\
+Campbell River  & 4.805 & 4.905 & 0 & 0.578 & 0.511 & -6\\
+\hline
+\hline
+Averages  &  &  &  &  &  & \\
+Point Atkinson  & 5.196 & 5.277 & 0  & 0.752 & 0.634 & -3.125\\
+Victoria        & 3.147 & 3.125 & -0.625 & 0.684 & 0.563 & -2.125\\
+Patricia Bay    & 4.005 & 3.901 & -0.125 & 0.689 & 0.584 & -2.875\\
+Campbell River  & 4.756 & 4.835 & -0.250 & 0.673 & 0.590 & -3.875\\ 
+\hline 
+\end{tabular}}
+\label{tab:peak}
+\end{table}
 

### 2015-05-05 AO_guides/manuscriptAO.tex
removed commetn about northern forcing

@@
 Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland.
-Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the Strait of Georgia is also affected by the surge entering the domain from Johnstone Strait at the northern boundary or the incoming surge wave is partially reflected from the mainland coast.
+{\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps the incoming surge wave is partially reflected from the mainland coast in the Strait of Georgia}.
 
@@
 However, the local wind patterns can set up small spatial differences in the surge, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season.
-Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions or a partial reflection from the mainland coast.
-Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important north of the San Juan and Gulf Islands.
+{\color{red}Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due a partial reflection of the surge from the mainland coast.} Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important north of the San Juan and Gulf Islands.
 The modelled surge amplitude increased by about 10\% at Point Atkinson when the tide-surge interaction was neglected in one example.

### 2015-05-05 Revisions/tides/calculate_harmonics_filter_NOAA.m
Modified for readinf files from NOAA

@@ -1,0 +1,68 @@
+function  [tidestruc, lat, mean_wl] = calculate_harmonics_filter_NOAA(csvfilename,location)
+%caclulate_harmonics(csvfilename,location, start) Calculates tidal
+%harmonics at specifie location
+%   Uses t_tide to write a file with tidal harmonics from a time series in
+%   csvfilename.
+%   location is a string of the location. 
+%   note that for accurate harmonics the time series needs to be one year
+%   long. Also avoid using surge-heavy years as that could over estimate
+%   the harmonics.
+%   returns the tidestruc from the ttide analysis, and the lat and mean sea
+%   level from the data file.
+%   Saves the harmoics in a file. 
+%   This function is used by get_ttide_8.m
+
+%NKS June 2014
+
+%Read in the measured water level data the location
+fid = fopen(csvfilename);
+meas = textscan(fid,'%f-%f-%f %f:%f %f %f %f %f',...
+    'HeaderLines',3, 'EmptyValue',NaN,'Delimiter',',');
+lat = csvread(csvfilename,1,1,[1,1,1,1]);
+lat
+fclose(fid);
+
+%Calculate dates from columns of data
+time = datenum(meas{1},meas{2},meas{3},meas{4},meas{5},0);
+
+%Start date of measured water level record
+start_date = time(1); start_date(1,1)
+end_date = time(end);
+
+%Truncated time series based on a tidal filter
+wlev_modified = filter_tides_NOAA(csvfilename,0.3);
+figure
+plot(wlev_modified)
+
+clear time newmeas meas
+
+%Use t_tide to determine harmonic constituents. Needs to be at least one
+%year time series (366 days)
+analysis_file = [location  '_analysis_filter_' datestr(start_date) '_' datestr(end_date) ];
+[tidestruc,~] = t_tide(wlev_modified,'start time',start_date(1,1),'latitude',lat,'output',analysis_file);
+%should I output the tidal prediction too and pass it along in get_ttide_8?
+
+%mean water level
+mean_wl = nanmean(wlev_modified);
+%Save the harmonics
+harmonics_file = [location  '_harmonics_' datestr(start_date) '_' datestr(end_date) '_filter.csv'];
+fid = fopen(harmonics_file, 'w');
+%add some headers
[…]

### 2015-05-05 Revisions/tides/get_ttide_8_filter.m

@@ -1,4 +1,4 @@
-function [pred_all,pred_8,tim] = get_ttide_8_filter(csvfilename,location, starts, ends)
+function [pred_all,pred_8,tim] = get_ttide_8_filter(csvfilename,location, starts, ends, type)
 %returns the tidal predictions if only 8 constituents were used.
 %the 8 constituents are: M2,K1,O1,P1,Q1,N2,S2,K2
 %csvfilename contains DFO produced water level observations.
@@ -7,13 +7,18 @@ function [pred_all,pred_8,tim] = get_ttide_8_filter(csvfilename,location, starts
 %saves a spreadsheet with tim, pred_all, pred_8
 %starts is that start date for predictions (eg. 01-Jan-2006)
 %ends is the end date for the predictions (eg. 31-Dec-2006)
+%type representes the type of water level measurements (NOAA or DFO)
 
 % NKS May 2014
 
 %% Initial harmonics and prediction caclulations
 % Calculate harmonics
-[tidestruc,lat,msl]=calculate_harmonics_filter(csvfilename,location);
-
+if strcmp(type, 'DFO')
+    [tidestruc,lat,msl]=calculate_harmonics_filter(csvfilename,location);
+elseif strcmp(type,'NOAA')
+    [tidestruc,lat,msl]=calculate_harmonics_filter_NOAA(csvfilename,location);
+else print('Unrecognised type')
+end
 %Remove long period constituents. Do not use these in the predictions
 names_long = ['SA  '; 'SSA '; 'MSM '; 'MM  ';'MSF ';'MF  '];
 for i=1:length(names_long)

### 2015-05-05 Revisions/tides/calculate_harmonics_filter_NOAA.m
Redid Neah Bay tidal predictions to be like my other ones.

@@ -31,8 +31,6 @@ end_date = time(end);
 
 %Truncated time series based on a tidal filter
 wlev_modified = filter_tides_NOAA(csvfilename,0.3);
-figure
-plot(wlev_modified)
 
 clear time newmeas meas
 

### 2015-05-07 AO_guides/manuscriptAO.tex
Added new bathy images, resized zoom image and started discussion of large scale surge.

@@
 
-%some general intro paragraph here. Need something about SoG size, geography, islands, sills, rivers, tides....
 The Strait of Georgia is a strongly stratified, semi-enclosed body of water located between Vancouver Island and the mainland of British Columbia.
@@
 As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
+{\color{red} A magnified image of the complex passages through the San Jaun and Gulf Islands is provided in Figure \ref{fig:zoom}.}
 
@@
 \includegraphics[scale=0.5]{Figures/bathy.pdf}
-\caption{Model domain (highlighted area) including bathymetry, rivers (green circles), and storm surge locations of interest  (red stars).}\label{fig:domain}
+\caption{{\color{red} Model domain (highlighted area) including bathymetry, rivers (green circles), and storm surge locations of interest  (red stars).
+The black rectangle represents the region displayed in Figure \ref{fig:zoom}.}}\label{fig:domain}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.4]{Figures/bathy_zoom.pdf}
+\caption{{\color{red} Model bathymetry and coastlines in the passages between the San Juan and Gulf Islands.}}\label{fig:zoom}
 \end{figure}
@@
 
-Sea surface height anomaly (or remote forcing) was set uniformly across the mouth of Juan de Fuca using values from the Tofino tide gauge \citep{DFOObservations} which is outside of, but close to, the western boundary of the model domain.
+Sea surface height anomaly (or remote forcing) was set uniformly across the mouth of Juan de Fuca using values from the Tofino tide gauge \citep{DFOObservations} which is outside of, but close to, the western boundary of the model domain (Figure \ref{fig:domain}).
 A monthly climatology was produced using daily averages from 2000-2010, binning them by month, averaging and setting the yearly mean to zero.
@@
 These values are the Tofino tide gauge values with the tides removed.
-Tidal predictions were generated with a tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year prior to the surge event to avoid over predicted constituents during high surge years.
-The storm surge simulations also used the sea surface height anomaly from Port Hardy (50.72 $^\circ$ N, 127.49 $^\circ$ W), forced at the northern boundary in Johnstone Strait.
+{\color{red}Tidal predictions were generated with a tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year of the storm surge.
+In order to prevent over-predicted constituents due to high surge years, the time series is first filtered by applying a Doodson filter (NOAA citation) and removing periods with large non-tidal amplitudes.
+Long period constituents such as SA, SSA, etc are not used in the tidal predictions since these constituents are often contaminated by non-tidal energy from seasonal meteorological trends.
+Further, constituents with a low signal to noise ratio were neglected.
+} The storm surge simulations also used the sea surface height anomaly from Port Hardy, forced at the northern boundary in Johnstone Strait.
+
+{\color{red} Although Tofino is not the closest tide guage to the western boundary, we have found that the surge conditions on the western coast of Vancouver Island and near the mouth of the Strait of Juan de Fuca are fairly uniform.
+}
+
+\begin{table}[h]
+\centering 
+\tbl{Average surge amplitudes and delay from Tofino at western tide gauges. Averages obtained over the surge events discussed on this article.  } 
+{\begin{tabular}{|l |c c} 
+\hline
+
+\end{tabular}}
+\end{table}
 The tidal forcing and remote forcing were applied using the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}, a radiation scheme that allows disturbances to propagate out of the domain without reflections off of the open boundary.
@@
 All storm surge runs were started at least three days prior to the event of interest with zero initial velocities and sea surface height and a stratification profile from model spin up.
-The modelled sea surface height adjusted to forcing in less than one day.
+The modelled sea surface height adjusted to forcing in less than one day {\color{red} as determined by comparing a simulation with restart conditions to one with velocities and sea surface height initialized to zero}.
 

### 2015-05-07 AO_guides/manuscriptAO.tex
Redid 2009 Toke Point tides and finished table for Pacific Coast Surge

@@
 \centering 
-\tbl{Average surge amplitudes and delay from Tofino at western tide gauges. Averages obtained over the surge events discussed on this article.  } 
+\tbl{Average surge amplitudes and delay in peak timing from Tofino at several tide gauges on the Pacific coast.
+Averages obtained over the surge events discussed on this article.
+A negative delay indicates that the peak surge occurred later than at Tofino.}
 {\begin{tabular}{|l |c c} 
 \hline
-
+Location & Mean surge amplitude (m) & Mean delay from Tofino (hrs) \\
+Tofino  & 0.707 & -- \\
+Toke Point & 1.17 & -2.83 \\
+Neah Bay & 0.658 & -1.83 \\
+Bamfield & 0.658 & -0.5 \\
+Winter Harbour & 0.572 & 0.333 \\
+Port Hardy & 0.491 & -2.67
+\hline
 \end{tabular}}

### 2015-05-11 AO_guides/manuscriptAO.tex
Added Porlire pass

@@
 Bathymetry from the Cascadia physiography data set \citep{haugerud1999digital} was smoothed to limit large changes in depth across grid cells, so that $\Delta h/\bar{h} \leq 0.8$, where $\Delta h$ is the difference in depth between two adjacent grid cells and $\bar{h}$ is their average depth.
-For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells.
+For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells {\color{red} inward from the open boundary}.
 As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
@@
 At the bottom boundary, partial $z$-levels were used in order to limit large changes in bathymetry across grid cells \citep{madec2012nemo}.
+{\color{red} Model constraints require that all passages be at least two grid points wide.
+This restriction required choices in either widening or closing narrow straits.
+In particular, Active Pass was closed and Porlier Pass was widened.
+Shallow regions were represented by constant depth of 4 m.
+All of these compromises are expected to have significant effects locally but minimal effects on the large scale.}
 

### 2015-05-12 stormtools_revisions.py
moved storm_tools revisions here

@@ -1,0 +1,767 @@
+# Copyright 2013-2015 The Salish Sea MEOPAR contributors
+# and The University of British Columbia
+
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+
+#    http://www.apache.org/licenses/LICENSE-2.0
+
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+
+"""A collection of tools for storm surge results from the Salish Sea Model.
+"""
+from __future__ import division
+
+import datetime
+import netCDF4 as NC
+import numpy as np
+import arrow
+import cStringIO
+import requests
+from dateutil import tz
+from xml.etree import cElementTree as ElementTree
+import pandas as pd
+from salishsea_tools import tidetools
+import csv
+
+#msl over 1983-2001 in metres
+MSLS = {'PointAtkinson': 3.095, 'Victoria': 1.902, 'PatriciaBay': 2.298, 'CampbellRiver': 2.896}
+
+def convert_date_seconds(times, start):
+    """
+    This function converts model output time in seconds to datetime objects.
+    Note: Doug has a better version of this in nc_tools.timestamp
+
+    :arg times: array of seconds since the start date of a simulation. From time_counter in model output.
+    :type times: int
+
+    :arg start: string containing the start date of the simulation in format '01-Nov-2006'
+    :type start: str
+
+    :arg diff: string indicating the time step in the times data E.g. months, seconds, days
+    :type diff: str
+
+    :returns: array of datetime objects representing the time of model outputs.
+    """
+    arr_times=[]
+    for ii in range(0,len(times)):
+        arr_start =arrow.Arrow.strptime(start,'%d-%b-%Y')
+        arr_new=arr_start.replace(seconds=times[ii])
+        arr_times.append(arr_new.datetime)
+
+    return arr_times
+
+def convert_date_hours(times, start):
+    """
[…]

### 2015-05-12 Revisions/tides/get_ttide_8_filter.m
added some comments

@@ -4,6 +4,7 @@ function [pred_all,pred_8,tim] = get_ttide_8_filter(csvfilename,location, starts
 %csvfilename contains DFO produced water level observations.
 %pred_all is the t_tide prediction with all constituents
 %pred_8 is the t_tide prediction with only 8 constituents
+%pred_noshallow is pred_all without the shallow water constituents
 %saves a spreadsheet with tim, pred_all, pred_8
 %starts is that start date for predictions (eg. 01-Jan-2006)
 %ends is the end date for the predictions (eg. 31-Dec-2006)

### 2015-05-13 AO_guides/manuscriptAO.tex
Worked on revisions

@@
 The numerical model employs the Nucleus for European Modelling of the Ocean (NEMO) architecture in a regional configuration.
-The model is evaluated through comparisons of tidal elevation harmonics and storm surge water levels with observations.
+The model is evaluated through {\color{red} the} comparisons of tidal elevation harmonics and storm surge with observations.
 Important forcing factors contributing to storm surges have been assessed.
-It is shown that the surge entering the domain from the Pacific Ocean contributes most significantly to the variability.
-Comparisons between simulations with high resolution and low resolution atmospheric forcing further emphasize this point.
-In addition, local wind patterns cause a slight increase in surge amplitude on the mainland side of the Strait of Georgia when compared to the Vancouver Island coastal areas during a major wind storm on Dec 15, 2006.
+It is shown that the surge entering the domain from the Pacific Ocean contributes most significantly to the {\color{red} surge amplitude within the Strait of Georgia.
+Comparisons between simulations with high resolution and low resolution atmospheric forcing further emphasize tthat remote forcing is the dominant factor in surge amplitudes of this region.} In addition, local wind patterns cause a slight increase in surge amplitude on the mainland side of the Strait of Georgia when compared to the Vancouver Island coastal areas during a major wind storm on Dec 15, 2006.
 Generally, surge amplitudes are found to be greater within the Strait of Georgia when compared to amplitudes in the Strait of Juan de Fuca.
@@
 
-The Strait of Georgia is a strongly stratified, semi-enclosed body of water located between Vancouver Island and the mainland of British Columbia.
+The Strait of Georgia is a strongly stratified, semi-enclosed body of water located between Vancouver Island and the mainland of British Columbia {\color{red}\citep{thomson1981oceanography}}.
 It is part of a larger system of waterways collectively known as the Salish Sea and is connected to the Pacific Ocean via the Strait of Juan de Fuca and Johnstone Strait (Figure \ref{fig:domain}).
 At its southern end, the Strait of Georgia is connected to the Strait of Juan de Fuca through several passages, the largest being Haro Strait.
-The San Juan and Gulf Islands are scattered throughout these passages where complex bathymetry and steep topographic slopes induce vigorous tidal mixing.
+The San Juan and Gulf Islands are scattered throughout these passages where complex bathymetry and steep topographic slopes induce vigorous tidal mixing {\color{red}\citep{thomson1981oceanography}}.
 Mixed diurnal and semi-diurnal tides are found throughout most of the Salish Sea but a semi-diurnal amphidrome is present nearby Victoria, where the tides are mainly diurnal \citep{thomson1981oceanography}.
-This system is home to several large coastal communities, such as Vancouver and Victoria, serves as a waterway for commercial traffic and recreational use, and supports a large and active fishing industry.
+This system is {\color{red} a} home to several large coastal communities, such as Vancouver and Victoria, serves as a waterway for commercial traffic and recreational use, and supports a large and active fishing industry.
 
@@
 There is also a small contribution to increased sea level due to positive sea surface height anomalies propagating northward from the equator along the eastern Pacific coast during El Nino years \citep{strub2002altimeter}.
-Communities in the Strait of Georgia are particularly susceptible to storm surges in the winter months when there is an increased propensity for storms and a higher mean sea level caused by southeasterly winds and geostrophic adjustment \citep{danard2003storm}.
+Communities in the Strait of Georgia are particularly susceptible to storm surges in the winter months when there is an increased propensity for storms and a higher sea level caused by southeasterly winds and geostrophic adjustment \citep{danard2003storm}.
 In the summer months, the wind direction changes to northwesterly and mean sea level is lower.
@@
 Such large domains typically use nested or unstructured grids to resolve the details of complicated coastlines.
-Appropriate treatment of rivers \citep{flather1994storm}, accurate bathymetry and the inclusion surface waves \citep{xu2010storm} have also been identified as important.
+Appropriate treatment of rivers \citep{flather1994storm}, accurate bathymetry and the inclusion {\color{red} of} surface waves \citep{xu2010storm} have also been identified as important.
 Most of these models neglect the effects of stratification by treating the ocean as a constant temperature and salinity.
@@
 Their work is highlighted in reviews of storm surges in Canadian waters \citep{murty1995storm, danard2003storm}.
-Compared to eastern Canada \citep{gray1984preliminary, bernier2006predicting,bobanovic2006forecasting, bernier2007tide, bernier2010tide}, relatively little has been published on storm surge modelling on the west coast of Canada in recent years.
+Compared to eastern Canada \citep{gray1984preliminary, bernier2006predicting,bobanovic2006forecasting, bernier2007tide, bernier2010tide, ma2015oceanic}, relatively little has been published on storm surge modelling on the west coast of Canada in recent years.
 However, a storm surge forecasting program is currently operating over a domain that covers the entire British Columbia coastline \citep{StormBC}.
@@
 %overview oF SoG/Salish Sea models
-The early modelling work of \citet{crean1988mathematical}, \citet{crean1988numerical} and \citet{stronach1993update} resulted in a series of models covering the Strait of Juan de Fuca, the Strait of Georgia, and Puget Sound using a finite difference discretization of the equations of motion.
+The early modelling work of \citet{crean1988mathematical}, \citet{crean1988numerical} and \citet{stronach1993update} {\color{red} developed} a series of models covering the Strait of Juan de Fuca, the Strait of Georgia, and Puget Sound using a finite difference discretization of the equations of motion.
 %A barotropic version of their model was used to reproduce storm surge water level elevations with reasonable accuracy, however, this barotropic version did not include the effects of fresh water entering the system from rivers.
@@
 They noted that their semi-diurnal amplitudes were generally too low.
-An updated version of this model produced high surface temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}.
-Smaller domains have also been considered; notably, the passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.
+{\color{red} An updated version of this model with added biogeochemical components and an extended domain produced higher than observed surface temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}.} Smaller domains have also been considered; notably, the passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.
 Accurate reproduction of observed tidal amplitudes and phases and appropriate treatment of mixing is a recurring difficulty in modelling this region.
@@
 This article presents the development of a three-dimensional numerical ocean model for the Strait of Georgia and its skill in simulating storm surge hindcasts.
-To complement previous barotropic storm surge modelling in this region, this model employs a higher resolution and includes the effect of stratification and rivers.
+{\color{red} The main purposes are to identify the forcing factors that contribute most significantly to storm surges in the Strait of Georgia, to examine the spatial distribution of the storm surge, and to evaluate the performance of a new model of this region.}  To complement previous barotropic storm surge modelling in this region, this model employs a higher resolution and includes the effect of stratification and rivers.
 However, wetting and drying is not represented.
@@
 NEMO's ocean core solves the three-dimensional hydrostatic equations of motion for an incompressible fluid under the Boussinesq approximation on a structured computational grid.
-Although not used in the present work, NEMO's options for grid nesting and bio-geochemical coupling make it a useful tool for studying the complex physics and bio-geochemical interactions within the Strait of Georgia.
-This work focuses on the physical configuration of the Salish Sea model, in particular, determining appropriate forcing and boundary conditions for accurate reproduction of tidal amplitudes and phases as well as storm surge elevations.
+{\color{red} Although these features are not being used in the present work, NEMO's options for grid nesting and bio-geochemical coupling make it a useful tool for studying the complex physics and bio-geochemical interactions within the Strait of Georgia.} This work focuses on the physical configuration of the Salish Sea model, in particular, determining appropriate forcing and boundary conditions for accurate reproduction of tidal amplitudes and phases as well as storm surge elevations.
 Future work will include bio-geochemical coupling and data assimilation.
 NEMO version 3.4 was used in this study.
+{\color{red} NEMO has been successfully used in other coastal modelling studies in the Northeast Atlantic \citep{maraldi2013nemo} and Mediterranean (CITATION)!}
 
@@
 This restriction required choices in either widening or closing narrow straits.
-In particular, Active Pass was closed and Porlier Pass was widened.
+In particular, Active Pass was closed and Porlier Pass was widened (Figure \ref{fig:zoom}).
 Shallow regions were represented by constant depth of 4 m.
@@
 The model is relaxed to the forced temperature and salinity over the 10 grid points (about 5 km) closest to the open boundaries, using a flow relaxation scheme \citep{engedahl1995use}.
-%other references?
+{\color{red} At the north, seasonal climatologies from \citet{thomson1981oceanography} were used.} %other references?
 
@@
 {\color{red}Tidal predictions were generated with a tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year of the storm surge.
-In order to prevent over-predicted constituents due to high surge years, the time series is first filtered by applying a Doodson filter (NOAA citation) and removing periods with large non-tidal amplitudes.
+In order to prevent over-predicted constituents due to high surge years, the time series is first filtered by applying a Doodson filter \citep{parker2007NOAA} and removing periods with large non-tidal amplitudes.
 Long period constituents such as SA, SSA, etc are not used in the tidal predictions since these constituents are often contaminated by non-tidal energy from seasonal meteorological trends.
@@
 
-{\color{red} Although Tofino is not the closest tide guage to the western boundary, we have found that the surge conditions on the western coast of Vancouver Island and near the mouth of the Strait of Juan de Fuca are fairly uniform.
+{\color{red} Although Tofino is not the closest tide guage to the western boundary, we have found that the surge conditions on the western coast of Vancouver Island and near the mouth of the Strait of Juan de Fuca are fairly uniform as summarized in Table \ref{tab:pacificsurge}.
+Generally, surge amplitudes at Tofino are slightly larger and earlier than Neah Bay and Bamfield.
 }
@@
 \centering 
-\tbl{Average surge amplitudes and delay in peak timing from Tofino at several tide gauges on the Pacific coast.
+\tbl{{\color{red}Average surge amplitudes and delay in peak timing from Tofino at several tide gauges on the Pacific coast.
 Averages obtained over the surge events discussed on this article.
-A negative delay indicates that the peak surge occurred later than at Tofino.}
-{\begin{tabular}{|l |c c} 
+A negative delay indicates that the peak surge occurred later than at Tofino.}}
+{\begin{tabular}{|l |c | c|}
 \hline 
 Location & Mean surge amplitude (m) & Mean delay from Tofino (hrs) \\
+\hline
 Tofino  & 0.707 & -- \\
@@
 Winter Harbour & 0.572 & 0.333 \\
-Port Hardy & 0.491 & -2.67
+Port Hardy & 0.491 & -2.67\\
 \hline
 \end{tabular}}
+\label{tab:pacificsurge}
 \end{table}
+
 The tidal forcing and remote forcing were applied using the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}, a radiation scheme that allows disturbances to propagate out of the domain without reflections off of the open boundary.
@@
 Zero gradient conditions were chosen because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain.
-At the north baroclinic velocities were set to zero.
+{\color{red} No boundary conditions were set on the baroclinic velocities at the north.}
 
@@
 
-The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2014new}.
-Forecasts from the period of 2002-2012 are available.
+{\color{red} The ocean surface is forced with momentum fluxes, solar and non-solar heat fluxes, and precipitation from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2014new}.} Forecasts from the period of 2002-2012 are available.
 One additional simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.
@@
 The monthly flow from each river was input as a point source from 0 to 3 metres at the model point closest to the mouth of each river.
-Incoming water was assumed to be fresh and at surface temperature.
+Incoming water was assumed to be fresh and at surface {\color{red}ocean temperature}.
 A total of 150 rivers were parametrized by this method and their locations are indicated by the green dots in Figure \ref{fig:domain}.
@@
 All storm surge runs were started at least three days prior to the event of interest with zero initial velocities and sea surface height and a stratification profile from model spin up.
-The modelled sea surface height adjusted to forcing in less than one day {\color{red} as determined by comparing a simulation with restart conditions to one with velocities and sea surface height initialized to zero}.
+{\color{red} The modelled sea surface height adjusted to forcing in less than one day as determined by comparing a simulation with restart conditions to one with velocities and sea surface height initialized to zero}.
 
@@
 The model's M$_2$ amphidrome, indicated by the drop in amplitude around station 5 (Victoria), is too strong resulting in lower than observed M$_2$ amplitudes in this region.
-Additionally, the M$_2$ phase errors in the Strait of Juan de Fuca (stations 1-7) are quite large, leading to large complex differences in this region.
-However, the match in the Strait of Georgia (stations 10-24) is very good, with complex differences less than 6 cm, in part due to tidal tuning for highly accurate tides in the main region of interest.
-Details on the tidal tuning procedure are provided in Appendix \ref{sec:appendix}.
+{\color{red} The model also represents the dip in M$_2$ amplitude around stations 25 and 26 reasonably well.} Additionally, the M$_2$ phase errors in the Strait of Juan de Fuca (stations 1-7) are quite large, leading to large complex differences in this region.
+The match in the Strait of Georgia (stations 10-24) is very good, with complex differences less than 6 cm, in part due to tidal tuning for highly accurate tides in the main region of interest.
+{\color{red}Because the tides were tuned for a good match in the Strait of Georgia, the match in the Strait of Juan de Fuca has suffered, particularly noticeable at the station closest to the open boundary (Port Renfrew, 1).
+} Details on the tidal tuning procedure are provided in Appendix \ref{sec:appendix}.
 
@@
 Stratification conditions from model spin up at a time close to the event (within five days) were chosen to initialize the temperature and salinity fields.
-Boundary conditions include atmospheric forcing and fresh water input from rivers at the surface, sea surface height anomaly from observations at the open boundary, and tidal forcing with eight tidal constituents.
+{\color{red} Boundary conditions include atmospheric forcing (momentum and heat fluxes and precipitation) and fresh water input from rivers at the surface, sea surface height anomaly from observations at the open boundary, and tidal forcing with eight tidal constituents.}
+
 Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height due to the omission of the next leading order tidal constituents such as J$_1$.
-Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can approach 40 cm during times of high tidal range, and in particular, during the storm surge season Nov-Feb.
-As such, storm surge model predictions were corrected by accounting for this error; for each storm surge location of interest, the difference between a full tidal prediction and a tidal prediction with eight constituents is added to the modelled sea surface height.
-This correction is only applied to the modelled water level and not the modelled residuals.
+Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can approach {\color{red} 20} cm during times of high tidal range.
+{\color{red} As such, storm surge model predictions were corrected by accounting for this error.
+However, the model will contain tidal energy from shallow water constituents due nonlinear interactions in the model questions.
+As such, the correction is calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (SA, SSA, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.
+The difference between these two predictions is added to the model values as a correction.
+Note that the harmonic analysis to generate the constituents is applied in the same manner as that used in the forcing.
+That is, the time series from the year of the storm surge is filtered to remove periods of large non-tidal energy and then the harmonic analysis is performed.}  This correction is only applied to the modelled water level and not the modelled residuals.
 Additionally, since the modelled sea surface height is calculated about a zero reference level, the modelled water level is determined by adding a long term mean sea level at the specified location to the modelled sea surface height.
@@
 Water level observations are taken from the Canadian Tides and Water Level Data Archive \citep{DFOObservations} and are used to calculate the observed residual, defined as the observed water level minus tidal predictions.
-Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series, in this case, from the year prior to the surge event of interest.
+Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series.
+{\color{red} The same method for calculating forcing residuals described in section \ref{sec:config} is used to calculate the observed residuals.}
 
@@
 The simulation ran from Feb 1-8, 2006.
-The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations to corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River.
+The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations {\color{red} with} corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River.
 First, the timing of maximum water level matches well between the observations and the model at all four locations, the most significant contribution to high water being from the tides which are generally well-represented by the model.
@@
 It is surprising that Victoria and Patricia Bay agree so closely with observations because the M$_2$ tidal amplitudes in the model are known to be too low in this area.
-When considering the over predicted water levels at Point Atkinson and Campbell River, it is likely that one of the forcing factors is overestimated in this example.
+{\color{red} Given that the water levels are over predicted at Point Atkinson and Campbell River, it is likely that one of the forcing factors is overestimated in this example.
+}
 
 In general, there is also very good agreement between the observed and modelled residuals (Figure \ref{fig:feb2006}).
-The most noticeable difference is a 4-6 hour delay in the model's timing of the maximum residual except at Victoria where the maximum residual is in phase with observations.
-The surge manifests as a broad peak in both the modelled and observed residual leading to a high probability that large residuals coincide with the high tide.
-As such, it is of minor consequence that the timing of the observed and modelled maximum residual do not match perfectly.
+The most noticeable difference is {\color{red} a broader peak in the model's surge curve, leading to a delay of a few ours in timing of the peak surge for most of the stations (Table \ref{tab:peak}).
+Note that the surge manifests as a broad peak in both the modelled and observed residual leading to a high probability that large residuals coincide with the high tide.} As such, it is of minor consequence that the timing of the observed and modelled maximum residual do not match perfectly.
 
@@
 \hline
-Dec 2012-CRGF &  &  &  &  &  & \\
+Dec 2012-CGRF &  &  &  &  &  & \\
 Point Atkinson  & 5.345 & 5.416 & 0 & 0.663 & 0.555 & -4\\
@@
 The best match in maximum surge occurred at Campbell River with a 6.2 cm underestimate by the model.
-The timing of the maximum surge at Campbell River was identical with observations and too early at the other the stations.
+{\color{red} The timing of the maximum surge was too early at all stations except Campbell River.}
 
@@
 Fourth, a simulation with only tides and local wind forcing, that is, excluding remote forcing and local atmospheric pressure forcing in the dashed blue curve.
-Comparisons between these simulations indicate the relative importance of the remote forcing, the local wind forcing, and the local atmospheric pressure forcing.
-Finally, a simulation without tidal forcing was performed to study the importance of the tide-surge interaction (Figure \ref{fig:tidesurge}).
+{\color{red} Comparisons between these simulations will determine the contributions from the remote forcing, the local wind forcing, and the local atmospheric pressure forcing.} Finally, a simulation without tidal forcing was performed to study the importance of the tide-surge interaction (Figure \ref{fig:tidesurge}).
 Contributions due to stratification and bathymetric resolution will be left to a future study.
@@
 Our results suggest that the local atmospheric pressure generates a sea level change that is smaller than the IB response.
-This is related to the small spatial scale of the model domain.
+{\color{red} It is the atmopsheric pressure gradient that drives changes in the sea surface height.
+Since our domain is relatively small, so is the atmospheric pressure gradient, and hence, we see only a small contribution from local atmospheric pressure.
+}
 
@@
 Storms travelling over the Pacific Ocean towards the west coast of the northern United States and Canada can induce elevated sea levels along the coastline which then enter the Salish Sea system through the Strait of Juan de Fuca.
-Previous simulations by \citet{murty1995storm} suggest that the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain, consistent with the results of the present study.
+Previous simulations by \citet{murty1995storm} suggest that the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.
+{\color{red} Our study is consistent with these findings.}
 
@@
 The surge enters the system from the Strait of Juan de Fuca, peaking between Dec 15 at 9:30 UTC and Dec 15 at 11:30 UTC and then begins to subside throughout the entire domain.
-Notably, the Strait of Georgia experiences the largest surge values, upwards of 60 cm, whereas the Juan de Fuca region experiences a maximum surge around 45 cm.
+Notably, the Strait of Georgia experiences the largest surge values, upwards of {\color{red}70} cm, whereas the Juan de Fuca region experiences a maximum surge around {\color{red}60} cm {\color{red} (Figure \ref{fig:spatial})}.
 Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges ($\sim$10 cm) than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast.
@@
 While the surge entering the domain from the Pacific Ocean is the leading order contribution, the winds may induce localized changes to the surge amplitude.
-Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland.
-{\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps the incoming surge wave is partially reflected from the mainland coast in the Strait of Georgia}.
+Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland {\color{red}(Figure \ref{fig:spatial})}.
+{\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps the incoming surge wave is partially reflected from the mainland coast in the Strait of Georgia.
+Note that since local winds are stronger in this case, thelocal wind contribution is more significant than in the Feb 2006 case (6 cm contribution at Point Atkinson).
+However, this is still a small percentage of the total surge amplitude.
+}
 
@@
 Two significant surge events are noted, the first occurring on Nov 19 with an observed amplitude of about 50 cm at Point Atkinson, and a second, larger surge of 62.5 cm on Nov 20.
-Surprisingly, the strong wind event of Nov 19 was associated with the smaller of the two surges.
-Neither of the surges occurred during a high tide and so no high water level events were recorded.
-Although not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and another surge entering the system at Tofino and Port Hardy.
+{\color{red} The strong wind event of Nov 19 was associated with the smaller of the two surges, supporting the fact that local winds do not contribute significantly to surge amplitudes.}  Neither of the surges occurred during a high tide and so no high water level events were recorded.
+Although {\color{red} the sea surface heigh plots are} not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and another surge entering the system at Tofino and Port Hardy.
 
@@
 Willmott skill scores for both modelled residuals are above 0.92 at these four locations.
-Further, the modelled residuals deviate only slightly from each other, most noticeable at Point Atkinson during the large wind event on Dec 15, 2012 where simulation with the low resolution atmospheric forcing has a residual that is 5 cm higher.
+Further, the modelled residuals deviate only slightly from each other, most noticeable at Point Atkinson during the {\color{red}strong} wind event on Dec 15, 2012 where simulation with the low resolution atmospheric forcing has a residual that is 5 cm higher.
 Although the wind speeds on Dec 15, 2012 are quite high, this event did not result in any flooding or residuals greater than 50 cm.
@@
 Next, the M$_2$ and K$_1$ fluxes at the northern boundary were adjusted because the cross-sectional area of transect used by \citet{thomson1980johnstone} to measure tidal currents in Johnstone Strait is smaller than the cross-sectional area of our northern boundary.
-Finally, the remaining constituents were added, starting with O$_1$, then S$_2$, and then in pairs P$_1$/N$_2$ and Q$_1$/K$_2$.
+{\color{red} The northern boundary tides were tuned for good matches at Yorke Island  and Kelsey Bay.} Finally, the remaining constituents were added, starting with O$_1$, then S$_2$, and then in pairs P$_1$/N$_2$ and Q$_1$/K$_2$.
 Final adjustments were made by slightly varying amplitude and phase, evaluating the changes to M$_2$, K$_1$ and the ratios S$_2$/M$_2$, N$_2$/M$_2$, K$_2$/M$_2$, O$_1$/K$_1$, P$_1$/K$_1$, Q$_1$/K$_1$ and then choosing new values to move closer to observations by linear interpolation.

### 2015-05-14 AO_guides/manuscriptAO.tex
addressed all referee comments

@@
 NEMO version 3.4 was used in this study.
-{\color{red} NEMO has been successfully used in other coastal modelling studies in the Northeast Atlantic \citep{maraldi2013nemo} and Mediterranean (CITATION)!}
+{\color{red} NEMO has been successfully used in other coastal modelling studies in the Northeast Atlantic \citep{maraldi2013nemo} and the Mediterranean Sea \citep{brossier2012sensitivity}.}
 
@@
 Bathymetry from the Cascadia physiography data set \citep{haugerud1999digital} was smoothed to limit large changes in depth across grid cells, so that $\Delta h/\bar{h} \leq 0.8$, where $\Delta h$ is the difference in depth between two adjacent grid cells and $\bar{h}$ is their average depth.
-For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells {\color{red} inward from the open boundary}.
-As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
+For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells {\color{red} inward from the open boundary.
+Depths between 0 m and 4 m were set to 4 m and a minimum depth of 4 m was by NEMO.}  As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
 {\color{red} A magnified image of the complex passages through the San Jaun and Gulf Islands is provided in Figure \ref{fig:zoom}.}
@@
 These values are the Tofino tide gauge values with the tides removed.
-{\color{red}Tidal predictions were generated with a tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year of the storm surge.
+{\color{red}Tidal predictions were generated with a tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year of the storm surge for 67 constituents.
 In order to prevent over-predicted constituents due to high surge years, the time series is first filtered by applying a Doodson filter \citep{parker2007NOAA} and removing periods with large non-tidal amplitudes.
@@
 Generally, surge amplitudes at Tofino are slightly larger and earlier than Neah Bay and Bamfield.
-}
+In the future, storm surge forecasts will make use of Neah Bay water level predictions from the National Oceanic and Atmospheric Administration (NOAA).}
 
@@
 Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can approach {\color{red} 20} cm during times of high tidal range.
-{\color{red} As such, storm surge model predictions were corrected by accounting for this error.
-However, the model will contain tidal energy from shallow water constituents due nonlinear interactions in the model questions.
+{\color{red} As such, model water level predictions were corrected by accounting for this error.Note that the model will contain tidal energy from shallow water constituents due nonlinear interactions in the model equations.
 As such, the correction is calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (SA, SSA, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.
-The difference between these two predictions is added to the model values as a correction.
+The difference between these two predictions is added to the model ses surface height as a correction.
 Note that the harmonic analysis to generate the constituents is applied in the same manner as that used in the forcing.
-That is, the time series from the year of the storm surge is filtered to remove periods of large non-tidal energy and then the harmonic analysis is performed.}  This correction is only applied to the modelled water level and not the modelled residuals.
-Additionally, since the modelled sea surface height is calculated about a zero reference level, the modelled water level is determined by adding a long term mean sea level at the specified location to the modelled sea surface height.
+That is, the time series from the year of the storm surge is filtered to remove periods of large non-tidal energy and then the harmonic analysis is performed.
+The total modelled water level is determined by adding the mean sea level from the tide gauge record over 1983-2001 to the corrected model sea surface height.} This correction is only applied to the modelled water level and not the modelled residuals.
 All times are reported in UTC.
@@
 In general, there is also very good agreement between the observed and modelled residuals (Figure \ref{fig:feb2006}).
-The most noticeable difference is {\color{red} a broader peak in the model's surge curve, leading to a delay of a few ours in timing of the peak surge for most of the stations (Table \ref{tab:peak}).
-Note that the surge manifests as a broad peak in both the modelled and observed residual leading to a high probability that large residuals coincide with the high tide.} As such, it is of minor consequence that the timing of the observed and modelled maximum residual do not match perfectly.
+{\color{red} The most noticeable difference is a broader peak in the model's surge curve, leading to a delay of a few hours in timing of the peak surge for most of the stations (Table \ref{tab:peak}).
+Note that the surge manifests as a broad peak in both the modelled and observed residual leading to a high probability that large residuals coincide with the high tide.} As such, it is of minor consequence that the timing of the observed and modelled maximum residuals do not match perfectly.
 
@@
 
+{\color{red} This case was also used to asses the importance of the northern boundary residual forcing and was used to brielfy examine stratified effects.
+When the northern boundary forcing is exlcuded the surge amplitude at Campbell River drops by about 10 cm.
+The other stations were not very much affected.
+This was also observed in another hindcast (Dec 2006).
+The effect of stratification was examined by considering a case with an initially horizontally uniform stratification and comparing to the fully stratified case.
+The largest differences were observed at Victoria, with an average 5 cm increase in sea surface height after an initial adjustment period.
+Other studies have seen an improvement in modelled surge amplitudes when using a baroclinic mdoel as opposed to a three-dimensional barotropic model \citep{ma2015oceanic}.
+We plan to investigate stratified effects in more detail in a future study.}
+
 \subsection{Factors Contributing to Storm Surges}\label{sec:factors}
@@
 {\color{red} Comparisons between these simulations will determine the contributions from the remote forcing, the local wind forcing, and the local atmospheric pressure forcing.} Finally, a simulation without tidal forcing was performed to study the importance of the tide-surge interaction (Figure \ref{fig:tidesurge}).
-Contributions due to stratification and bathymetric resolution will be left to a future study.
 
@@
 The observed residual at Tofino, BC (a point nearby but outside of the domain) is up to 10 cm  lower than the observed residual at Neah Bay, WA (a point nearby but inside of the domain).
-Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the cause.
+{\color{red} Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the cause, although it still does not account for all of the error.}
 
@@
 {\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps the incoming surge wave is partially reflected from the mainland coast in the Strait of Georgia.
-Note that since local winds are stronger in this case, thelocal wind contribution is more significant than in the Feb 2006 case (6 cm contribution at Point Atkinson).
-However, this is still a small percentage of the total surge amplitude.
+Note that since local winds are stronger in this case, the local wind contribution is more significant than in the Feb 2006 case (6 cm contribution at Point Atkinson).
+However, this is still les than 10 \% of the total surge amplitude.
 }
@@
 Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered.
-Although large wind speeds were observed, the surge associated with this event is rather small when compared to the previous cases (Figure \ref{fig:nov2009}).
+{\color{red} Wind speeds at Point Atkinson reached up to 18 m s$^{-1}$ which is approximately the same magnitude to that observed on Feb 4, 2006.}  Although large wind speeds were observed, the surge associated with this event is rather small when compared to the previous cases (Figure \ref{fig:nov2009}).
 Two significant surge events are noted, the first occurring on Nov 19 with an observed amplitude of about 50 cm at Point Atkinson, and a second, larger surge of 62.5 cm on Nov 20.
 {\color{red} The strong wind event of Nov 19 was associated with the smaller of the two surges, supporting the fact that local winds do not contribute significantly to surge amplitudes.}  Neither of the surges occurred during a high tide and so no high water level events were recorded.
-Although {\color{red} the sea surface heigh plots are} not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and another surge entering the system at Tofino and Port Hardy.
+Although {\color{red} the sea surface height plots are} not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and another surge entering the system at Tofino and Port Hardy.
 

### 2015-05-14 AO_guides/manuscriptAO.tex
Checked consistency up until Nov 2006

@@
 It is shown that the surge entering the domain from the Pacific Ocean contributes most significantly to the {\color{red} surge amplitude within the Strait of Georgia.
-Comparisons between simulations with high resolution and low resolution atmospheric forcing further emphasize tthat remote forcing is the dominant factor in surge amplitudes of this region.} In addition, local wind patterns cause a slight increase in surge amplitude on the mainland side of the Strait of Georgia when compared to the Vancouver Island coastal areas during a major wind storm on Dec 15, 2006.
+Comparisons between simulations with high resolution and low resolution atmospheric forcing further emphasize that remote forcing is the dominant factor in surge amplitudes of this region.} In addition, local wind patterns cause a slight increase in surge amplitude on the mainland side of the Strait of Georgia when compared to the Vancouver Island coastal areas during a major wind storm on Dec 15, 2006.
 Generally, surge amplitudes are found to be greater within the Strait of Georgia when compared to amplitudes in the Strait of Juan de Fuca.
@@
 For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells {\color{red} inward from the open boundary.
-Depths between 0 m and 4 m were set to 4 m and a minimum depth of 4 m was by NEMO.}  As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
-{\color{red} A magnified image of the complex passages through the San Jaun and Gulf Islands is provided in Figure \ref{fig:zoom}.}
+Depths between 0 m and 4 m were set to 4 m and a minimum depth of 4 m was imposed by NEMO.}  As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
+{\color{red} A magnified image of the complex passages through the San Juan and Gulf Islands is provided in Figure \ref{fig:zoom}.}
 
@@
 Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can approach {\color{red} 20} cm during times of high tidal range.
-{\color{red} As such, model water level predictions were corrected by accounting for this error.Note that the model will contain tidal energy from shallow water constituents due nonlinear interactions in the model equations.
+{\color{red} As such, model water level predictions were corrected by accounting for this error.
+Note that the model will contain tidal energy from shallow water constituents due nonlinear interactions in the model equations.
 As such, the correction is calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (SA, SSA, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.
@@
 These three statistics, along with the mean observed values and mean modelled values (which indicate the model's bias) are reported in Table \ref{tab:statistics} for each hindcast.
+{\color{red} Observed and modelled maximum water levels and surge values as well as differences in their timing are reported in Table \ref{tab:peak}.}
 
@@
 The statistics are calculated with hourly output over the three-day period displayed in Figure \ref{fig:feb2006}.
-The mean absolute error in the water level ranges from 7.6 cm at Victoria to 13.1 cm at Campbell River, whereas, the mean absolute error in the residuals is smaller at all stations; however, this measure does not account for the large variability in water levels due to the tides.
+The mean absolute error in the water level ranges from {\color{red}8.2 cm} at Victoria to {\color{red}12.2} cm at Campbell River, whereas, the mean absolute error in the residuals is smaller at all stations; however, this measure does not account for the large variability in water levels due to the tides.
 A statistic that does take into account this variability is $\gamma^2$ which suggests very good agreement in both the predicted water levels and residuals.
@@
 An additional storm surge hindcast for a Nov 16, 2006 storm surge is provided as another means of assessing model performance and is mentioned briefly here.
-The model performs reasonably well in this hindcast; Willmott skill scores are greater than 0.94 (Table \ref{tab:statistics}).
+The model performs reasonably well in this hindcast; Willmott skill scores are greater than {\color{red}0.95} (Table \ref{tab:statistics}).
 Further, the $\gamma^2$ statistic suggests that the standard deviation of the error in the modelled residuals is about 40\% ($\sqrt{0.16}$) of the standard deviation of the observed residuals.
-The best match in maximum surge occurred at Campbell River with a 6.2 cm underestimate by the model.
+The best match in maximum surge occurred at Campbell River with a {\color{red}4.3 cm} underestimate by the model.
 {\color{red} The timing of the maximum surge was too early at all stations except Campbell River.}

### 2015-05-14 AO_guides/manuscriptAO.tex
Finished constitency and checked for typos

@@
 These values are the Tofino tide gauge values with the tides removed.
-{\color{red}Tidal predictions were generated with a tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year of the storm surge for 67 constituents.
-In order to prevent over-predicted constituents due to high surge years, the time series is first filtered by applying a Doodson filter \citep{parker2007NOAA} and removing periods with large non-tidal amplitudes.
-Long period constituents such as SA, SSA, etc are not used in the tidal predictions since these constituents are often contaminated by non-tidal energy from seasonal meteorological trends.
+{\color{red}Tidal predictions were generated with a tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year of the storm surge.
+A harmonic analysis was performed for 67 constituents.
+In order to prevent over-predicted constituents due to high surge years, the time series was first filtered by applying a Doodson filter \citep{parker2007NOAA} and removing periods with large non-tidal amplitudes.
+Long period constituents such as SA, SSA, etc were not used in the tidal predictions since these constituents are often contaminated by non-tidal energy from seasonal meteorological trends.
 Further, constituents with a low signal to noise ratio were neglected.
@@
 As such, the correction is calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (SA, SSA, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.
-The difference between these two predictions is added to the model ses surface height as a correction.
-Note that the harmonic analysis to generate the constituents is applied in the same manner as that used in the forcing.
+The difference between these two predictions is added to the model sea surface height as a correction.
+Note that the harmonic analysis used to generate the constituents is applied in the same manner as that used in the forcing.
 That is, the time series from the year of the storm surge is filtered to remove periods of large non-tidal energy and then the harmonic analysis is performed.
@@
 The largest differences were observed at Victoria, with an average 5 cm increase in sea surface height after an initial adjustment period.
-Other studies have seen an improvement in modelled surge amplitudes when using a baroclinic mdoel as opposed to a three-dimensional barotropic model \citep{ma2015oceanic}.
+Other studies have seen an improvement in modelled surge amplitudes when using a baroclinic model as opposed to a three-dimensional barotropic model \citep{ma2015oceanic}.
 We plan to investigate stratified effects in more detail in a future study.}
@@
 First, it is clear that remote forcing at the open boundary contributes most significantly to the surge at each of these locations (Figure \ref{fig:factors}).
-When the remote forcing is not included as a forcing condition, the surge drops from a maximum 82 cm to 11 cm at Point Atkinson and its maximum occurs eight hours later.
+When the remote forcing is not included as a forcing condition, the surge drops from a maximum {\color{red}84 cm} to 11 cm at Point Atkinson and its maximum occurs eight hours later.
 A similar drop in the maximum surge and delay in timing of the maximum surge is observed at the other locations.
@@
 The contribution to the surge due to local atmospheric forcing is small and appears to be geographically dependent, most likely due to the effects of the local winds.
-Removal of the local atmospheric forcing leads to a drop of about 2.5 cm in the maximum surge at Point Atkinson and 2.3 cm at Campbell River.
-By contrast, the surge amplitude increases by 4.8 cm at Victoria and 3.6 cm at Patricia Bay.
+Removal of the local atmospheric forcing leads to a drop of about {\color{red}2 cm} in the maximum surge at Point Atkinson and {\color{red}2.5 cm} at Campbell River.
+By contrast, the surge amplitude increases by {\color{red} 4.5 cm} at Victoria and 3.6 cm at Patricia Bay.
 The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in a slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
@@
 Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow regions due to nonlinear effects and bottom friction \citep{bernier2007tide}, is presented for this hindcast (Figure \ref{fig:tidesurge}).
-The most significant differences occur at Point Atkinson where the maximum of the surge-only case is 8.7 cm higher than maximum modelled residual, about 10\% of the total surge amplitude.
+The most significant differences occur at Point Atkinson where the maximum of the surge-only case is {\color{red} 8.4 cm} higher than maximum modelled residual, about 10\% of the total surge amplitude.
 In contrast, the tide-surge interaction at Victoria is hardly noticeable.
@@
 A major wind storm on Dec 15, 2006 is hindcasted next.
-Although the storm was very strong (gusts over 120 km hr$^{-1}$) and the observed residual at Point Atkinson reached 84 cm, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area \citep{forseth2006adaptation}.
+Although the storm was very strong (gusts over 120 km hr$^{-1}$) and the observed residual at Point Atkinson reached {\color{red}94 cm}, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area \citep{forseth2006adaptation}.
 Because of the high winds, this event has been used to examine the propagation of the surge throughout the domain and spatial variations in the surge amplitude.
@@
 Modelled residuals are generally too low in this hindcast (Figure \ref{fig:dec2006}), however, the timing of the surge agrees well.
-All Willmott skill scores rank above 0.88 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}).
+All Willmott skill scores rank above {\color{red}0.89} and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}).
 The $\gamma^2$ measure suggests that the standard deviation of the error in the modelled residuals is roughly half of the standard deviation of the observed residuals.
@@
 The surge enters the system from the Strait of Juan de Fuca, peaking between Dec 15 at 9:30 UTC and Dec 15 at 11:30 UTC and then begins to subside throughout the entire domain.
-Notably, the Strait of Georgia experiences the largest surge values, upwards of {\color{red}70} cm, whereas the Juan de Fuca region experiences a maximum surge around {\color{red}60} cm {\color{red} (Figure \ref{fig:spatial})}.
+Notably, the Strait of Georgia experiences the largest surge values, upwards of {\color{red}70 cm}, whereas the Juan de Fuca region experiences a maximum surge around {\color{red}60 cm} {\color{red} (Figure \ref{fig:spatial})}.
 Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges ($\sim$10 cm) than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast.
@@
 {\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps the incoming surge wave is partially reflected from the mainland coast in the Strait of Georgia.
-Note that since local winds are stronger in this case, the local wind contribution is more significant than in the Feb 2006 case (6 cm contribution at Point Atkinson).
-However, this is still les than 10 \% of the total surge amplitude.
+Since local winds are stronger in this case, their contribution to the surge amplitude of about 6 cm at Point Atkionson is more significant than in the Feb 2006 case.
+However, this is still less than 10 \% of the total surge amplitude.
 }
@@
 {\color{red} Wind speeds at Point Atkinson reached up to 18 m s$^{-1}$ which is approximately the same magnitude to that observed on Feb 4, 2006.}  Although large wind speeds were observed, the surge associated with this event is rather small when compared to the previous cases (Figure \ref{fig:nov2009}).
-Two significant surge events are noted, the first occurring on Nov 19 with an observed amplitude of about 50 cm at Point Atkinson, and a second, larger surge of 62.5 cm on Nov 20.
+Two significant surge events are noted, the first occurring on Nov 19 with an observed amplitude of about {\color{red}55 cm} at Point Atkinson, and a second, larger surge of {\color{red} 70 cm} on Nov 20.
 {\color{red} The strong wind event of Nov 19 was associated with the smaller of the two surges, supporting the fact that local winds do not contribute significantly to surge amplitudes.}  Neither of the surges occurred during a high tide and so no high water level events were recorded.
@@
 During the high wind events on Nov 19 and Nov 20, the winds at Point Atkinson were directed to the west, explaining the lack of strong surge observed during this event since wind stresses would act to depress the water on the east side of the basin.
-In fact, the surge amplitude at Victoria on Nov 20 was 64.7 cm which is slightly higher than that observed at Point Atkinson (62.5 cm).
+In fact, the surge amplitude at Victoria on Nov 20 was {\color{red}71.8 cm} which is slightly higher than that observed at Point Atkinson {\color{red}(70 cm)}.
 Of the cases considered so far, this is the only example where the surge amplitude at Victoria was higher than that at Point Atkinson.
-The model results do not reflect a higher surge at Victoria (46.1 cm versus 50.5~cm at Point Atkinson), perhaps due to an underestimate in the wind speed and some discrepancies in the wind direction of the atmospheric forcing.
+The model results do not reflect a higher surge at Victoria {\color{red}(54 cm versus 58.7~cm at Point Atkinson)}, perhaps due to an underestimate in the wind speed and some discrepancies in the wind direction of the atmospheric forcing.
+{\color{red} There is a fairly significant delay in the timing of the maximum surge at most stations (Table \ref{tab:peak}).
+The model's surge curve is much broader and does not contain the narrow peak on Nov 20 as seen in the observations.}
 
@@
 
-The modelled water levels and residuals are too high for this hindcast, except at Victoria (Table \ref{tab:statistics}).
-Although Victoria has the smallest mean error in both the water level and residuals, its Willmott skill score and $\gamma^2$ values rank low when compared to the other stations.
-Generally, the skills scores are very good, scoring above 0.91 at all stations.
+The modelled water levels and residuals are, {\color{red} on average,} too high for this hindcast, except at Victoria (Table \ref{tab:statistics}).
+Generally, the skills scores are very good, scoring above {\color{red} 0.89} at all stations.
 
@@
 
-First, a comparison of the wind speed and direction for the two models and observations is provided in Figure \ref{fig:dec2012_weather} at Point Atkinson and at Sandheads, a lighthouse station in the southern Strait of Georgia.
-For each of the models, the wind speed and direction at the closest grid point to the longitude/latitude of Point Atkinson and Sandheads is used for comparison.
-Although the low resolution CGRF model has a grid spacing of 33 km, two grid points align closely with these locations.
+{\color{red} First, a comparison of the wind speed and direction for the two models and observations is provided in Figure \ref{fig:dec2012_weather} at Point Atkinson, Sandheads and a NOAA Neah Bay buoy.
+Sandheads is a lighthouse station in the southern Strait of Georgia.} For each of the models, the wind speed and direction {\color{red} the wind is heading} at the closest grid point to the longitude/latitude of Point Atkinson and Sandheads is used for comparison.
+Although the low resolution CGRF model has a grid spacing of 33 km, grid points align closely with these locations.
 At Point Atkinson, the high resolution model is in very good agreement with observations for both the wind speed and direction, however, the low resolution CGRF model significantly underestimates the wind speed during high wind periods on Dec 15, 2012 and Dec 17, 2012.
@@
 The low resolution model has a very good match at Sandheads likely because this grid cell happens to be mostly over water.
-Although not shown here, both models have a similar sea level atmospheric pressure at these two locations.
+Although not shown here, both models have a similar sea level atmospheric pressure at these locations.
 Note that data assimilation is used in the parent model of the nested HRDPS product.
@@
 Hindcasts with both atmospheric forcing products agree well with the observations (Figure \ref{fig:dec2012_surge}).
-Willmott skill scores for both modelled residuals are above 0.92 at these four locations.
-Further, the modelled residuals deviate only slightly from each other, most noticeable at Point Atkinson during the {\color{red}strong} wind event on Dec 15, 2012 where simulation with the low resolution atmospheric forcing has a residual that is 5 cm higher.
-Although the wind speeds on Dec 15, 2012 are quite high, this event did not result in any flooding or residuals greater than 50 cm.
-Significant flooding occurred later, on Dec 17, 2012, where the observed surge amplitude at Point Atkinson was 68.3 cm.
+Willmott skill scores for both modelled residuals are above {\color{red} 0.84} at these four locations.
+{\color{red} There is considerable noise in the observed Campbell River residual, which could explain its low skill score (Table \ref{tab:statistics}.} {\color{red} The modelled residuals in each simulation deviate only slightly from each other, most noticeable at Point Atkinson during the strong wind event on Dec 15, 2012 where simulation with the low resolution atmospheric forcing has a residual that is 5 cm higher.
+The maximum residuals on Dec 17, 2012 are almost identical at each station (Table \ref{tab:peak}).} Although the wind speeds on Dec 15, 2012 are quite high, this event did not result in any flooding or residuals greater than 50 cm.
+Significant flooding occurred later, on Dec 17, 2012, where the observed surge amplitude at Point Atkinson was {\color{red}66.3 cm}.
 Even though the low resolution model significantly under predicts the wind speed at some locations, it performs comparably to forcing with high resolution products.
@@
 Hindcasts of five significant events in February 2006, November 2006, December 2006, November 2009, and December 2012 were simulated.
-In general, the model performs well at reproducing the observed water level heights and residuals, with Willmott skill scores greater than 0.88 in all cases and generally higher skill scores for the water level heights.
+In general, the model performs well at reproducing the observed water level heights and residuals, with Willmott skill scores greater than {\color{red}0.84} in all cases and generally higher skill scores for the water level heights.
 Further, in a single simulation, the skill scores are fairly consistent between the four stations examined.
@@
 Finally, difficulties in reproducing accurate M$_2$ tidal harmonics in both the Strait of Georgia and the Strait of Juan de Fuca have been highlighted in this study among others \citep{stronach1993update, foreman2004m}.
-Future work will attempt to improve model accuracy by examining mixing parametrizations and energy dissipation due to bottom friction and their effect on tidal harmonics.
-Additionally, a biological-chemical modelling component will be added.
+Future work will attempt to improve model accuracy by examining mixing parametrizations and energy dissipation due to bottom friction and their effect on tidal harmonics {\color{red} and investigating stratified effects on surge amplitudes.} Additionally, a biological-chemical modelling component will be added.
 

### 2015-06-08 AO_guides/manuscriptAO.tex
more typos

@@
 The model is relaxed to the forced temperature and salinity over the 10 grid points (about 5 km) closest to the open boundaries, using a flow relaxation scheme \citep{engedahl1995use}.
-{\color{red} At the north, seasonal climatologies from \citet{thomson1981oceanography} were used.} %other references?
+{\color{red} At the north, seasonal climatologies from \citet{thomson1981oceanography} were used in the same manner.} %other references?
 
@@
 These values are the Tofino tide gauge values with the tides removed.
-{\color{red}Tidal predictions were generated with a tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year of the storm surge.
+{\color{red}Tidal predictions were generated with the tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year of the storm surge.
 A harmonic analysis was performed for 67 constituents.
@@
 The match in the Strait of Georgia (stations 10-24) is very good, with complex differences less than 6 cm, in part due to tidal tuning for highly accurate tides in the main region of interest.
-{\color{red}Because the tides were tuned for a good match in the Strait of Georgia, the match in the Strait of Juan de Fuca has suffered, particularly noticeable at the station closest to the open boundary (Port Renfrew, 1).
+{\color{red}Because the tides were tuned for a good match in the Strait of Georgia, the match in the Strait of Juan de Fuca was compromised, particularly noticeable at the station closest to the open boundary (Port Renfrew, 1).
 } Details on the tidal tuning procedure are provided in Appendix \ref{sec:appendix}.
@@
 {\color{red} As such, model water level predictions were corrected by accounting for this error.
-Note that the model will contain tidal energy from shallow water constituents due nonlinear interactions in the model equations.
+Note that the model will contain tidal energy from shallow water constituents due to nonlinear interactions in the model equations.
 As such, the correction is calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (SA, SSA, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.
@@
 
-The model's skill is reported through a few statistical measures of the the total water level $\eta$ and the residual $\delta$. First, the mean absolute error ($\bar{e}$), defined as
+The model's skill is reported through a few statistical measures of the total water level $\eta$ and the residual $\delta$. First, the mean absolute error ($\bar{e}$), defined as
 \begin{equation}
@@
 
-{\color{red} This case was also used to asses the importance of the northern boundary residual forcing and was used to brielfy examine stratified effects.
-When the northern boundary forcing is exlcuded the surge amplitude at Campbell River drops by about 10 cm.
-The other stations were not very much affected.
+{\color{red} This case was also used to assess the importance of the northern boundary residual forcing and was used to briefly examine stratified effects.
+When the northern boundary forcing is excluded the surge amplitude at Campbell River drops by about 10 cm.
+The other stations were not strongly affected.
 This was also observed in another hindcast (Dec 2006).
@@
 {\color{red} Comparisons between these simulations will determine the contributions from the remote forcing, the local wind forcing, and the local atmospheric pressure forcing.} Finally, a simulation without tidal forcing was performed to study the importance of the tide-surge interaction (Figure \ref{fig:tidesurge}).
+{\color{red} Although the winds and atmospheric pressure are influenced by large scale atmospheric features, we use the terms local and remote to differentiate between factors originating from inside and outside of the domain.}
 
@@
 Willmott skill scores for both modelled residuals are above {\color{red} 0.84} at these four locations.
-{\color{red} There is considerable noise in the observed Campbell River residual, which could explain its low skill score (Table \ref{tab:statistics}.} {\color{red} The modelled residuals in each simulation deviate only slightly from each other, most noticeable at Point Atkinson during the strong wind event on Dec 15, 2012 where simulation with the low resolution atmospheric forcing has a residual that is 5 cm higher.
+{\color{red} There is considerable noise in the observed Campbell River residual, which could explain its low skill score (Table \ref{tab:statistics}).} {\color{red} The modelled residuals in each simulation deviate only slightly from each other, most noticeable at Point Atkinson during the strong wind event on Dec 15, 2012 where simulation with the low resolution atmospheric forcing has a residual that is 5 cm higher.
 The maximum residuals on Dec 17, 2012 are almost identical at each station (Table \ref{tab:peak}).} Although the wind speeds on Dec 15, 2012 are quite high, this event did not result in any flooding or residuals greater than 50 cm.
@@
 Local winds have a significant impact on wave action which is a primary concern for flooding management.
-Further, the remote forcing is applied uniformly across the open boundary and so any spatial variability in the incoming surge are not accounted for in the model.
+Further, the remote forcing is applied uniformly across the open boundary and so any spatial variability in the incoming surge is not accounted for in the model.
 Since the remote forcing is an important contributing factor, any forecasting with this regional model must make use of results from a larger scale prediction system.

### 2015-06-08 AO_guides/manuscriptAO.tex
Susan's comments and some of Keith's wordy comments

@@
 This article presents the development of a three-dimensional numerical ocean model for the Strait of Georgia and its skill in simulating storm surge hindcasts.
-{\color{red} The main purposes are to identify the forcing factors that contribute most significantly to storm surges in the Strait of Georgia, to examine the spatial distribution of the storm surge, and to evaluate the performance of a new model of this region.}  To complement previous barotropic storm surge modelling in this region, this model employs a higher resolution and includes the effect of stratification and rivers.
+{\color{red} The main goals are to identify the forcing factors that contribute most significantly to storm surges in the Strait of Georgia, to examine the spatial distribution of the storm surges, and to evaluate the performance of a new model of this region.}  To complement previous barotropic storm surge modelling in this region, this model employs a higher resolution and includes the effect of stratification and rivers.
 However, wetting and drying is not represented.
@@
 For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells {\color{red} inward from the open boundary.
-Depths between 0 m and 4 m were set to 4 m and a minimum depth of 4 m was imposed by NEMO.}  As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
+Depths between 0 m and 4 m were set to 4 m.}  As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
 {\color{red} A magnified image of the complex passages through the San Juan and Gulf Islands is provided in Figure \ref{fig:zoom}.}
@@
 The model is relaxed to the forced temperature and salinity over the 10 grid points (about 5 km) closest to the open boundaries, using a flow relaxation scheme \citep{engedahl1995use}.
-{\color{red} At the north, seasonal climatologies from \citet{thomson1981oceanography} were used in the same manner.} %other references?
+{\color{red} In the north, the seasonal climatologies of \citet{thomson1981oceanography} were used in the same manner.} %other references?
 
@@
 A harmonic analysis was performed for 67 constituents.
-In order to prevent over-predicted constituents due to high surge years, the time series was first filtered by applying a Doodson filter \citep{parker2007NOAA} and removing periods with large non-tidal amplitudes.
-Long period constituents such as SA, SSA, etc were not used in the tidal predictions since these constituents are often contaminated by non-tidal energy from seasonal meteorological trends.
+In order to prevent over-predicted constituents due to high surge years, the time series was first filtered by applying a Doodson filter \citep{parker2007NOAA} and periods with large non-tidal amplitudes were removed.
+Long period constituents such as Sa, Ssa, etc were not used in the tidal predictions since these constituents are often contaminated by non-tidal energy from seasonal meteorological trends.
 Further, constituents with a low signal to noise ratio were neglected.
@@
 
-{\color{red} Although Tofino is not the closest tide guage to the western boundary, we have found that the surge conditions on the western coast of Vancouver Island and near the mouth of the Strait of Juan de Fuca are fairly uniform as summarized in Table \ref{tab:pacificsurge}.
+{\color{red} Although Tofino is not the closest tide gauge to the western boundary, we found that the surge conditions on the western coast of Vancouver Island and near the mouth of the Strait of Juan de Fuca are fairly uniform as summarized in Table \ref{tab:pacificsurge}.
 Generally, surge amplitudes at Tofino are slightly larger and earlier than Neah Bay and Bamfield.
@@
 \hline
+Port Hardy & 0.491 & -2.67\\
+Winter Harbour & 0.572 & 0.333 \\
 Tofino  & 0.707 & -- \\
+Bamfield & 0.658 & -0.5 \\
+Neah Bay & 0.658 & -1.83 \\
 Toke Point & 1.17 & -2.83 \\
-Neah Bay & 0.658 & -1.83 \\
-Bamfield & 0.658 & -0.5 \\
-Winter Harbour & 0.572 & 0.333 \\
-Port Hardy & 0.491 & -2.67\\
 \hline
@@
 \end{table}
+
+{\color{red} The remote forcing is applied uniformly across the open boundaries.
+Given the similarities between the surge amplitude and phase at Neah Bay and Bamfield, this seems to be a reasonable assumption.
+Further, a geostrophic slope calculation with velocity 20 cm s$^{-1}$ would imply a difference of less than 5 cm between the northern and southern sides of the Strait of Juan de Fuca.
+Non-uniform forcing conditions will be investigated in a future study.}
 

### 2015-06-08 AO_guides/manuscriptAO.tex
Finished Keith's wordy comments

@@
 \tbl{{\color{red}Average surge amplitudes and delay in peak timing from Tofino at several tide gauges on the Pacific coast.
-Averages obtained over the surge events discussed on this article.
+Averages obtained over the five surge events discussed on this article.
 A negative delay indicates that the peak surge occurred later than at Tofino.}}
@@
 Zero gradient conditions were chosen because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain.
-{\color{red} No boundary conditions were set on the baroclinic velocities at the north.}
+{\color{red} No boundary conditions were set on the baroclinic velocities at the northern boundary.}
 
-At coastal boundaries, the partial slip boundary condition, an approximation to no slip, is used.
+At coastal boundaries, a partial slip boundary condition, an approximation to no slip, is used.
 Partial slip allows one to include the frictional effects of lateral boundaries without the restrictive resolution required to represent the lateral boundary layer under no slip conditions.
@@
 The match in the Strait of Georgia (stations 10-24) is very good, with complex differences less than 6 cm, in part due to tidal tuning for highly accurate tides in the main region of interest.
-{\color{red}Because the tides were tuned for a good match in the Strait of Georgia, the match in the Strait of Juan de Fuca was compromised, particularly noticeable at the station closest to the open boundary (Port Renfrew, 1).
+{\color{red}Because the tides were tuned for a good match in the Strait of Georgia, the match in the Strait of Juan de Fuca was compromised; the effect is particularly noticeable at the station closest to the open boundary (Port Renfrew, 1).
 } Details on the tidal tuning procedure are provided in Appendix \ref{sec:appendix}.
@@
 {\color{red} As such, model water level predictions were corrected by accounting for this error.
-Note that the model will contain tidal energy from shallow water constituents due to nonlinear interactions in the model equations.
-As such, the correction is calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (SA, SSA, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.
+Note that the model will contain tidal energy from shallow water constituents due to nonlinear interactions in the model equations and so the correction as calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (Sa, Ssa, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.
 The difference between these two predictions is added to the model sea surface height as a correction.
-Note that the harmonic analysis used to generate the constituents is applied in the same manner as that used in the forcing.
-That is, the time series from the year of the storm surge is filtered to remove periods of large non-tidal energy and then the harmonic analysis is performed.
+Note that the harmonic analysis used to generate the constituents is applied in the same manner as that used in the forcing, i.e.\, the time series from the year of the storm surge is filtered to remove periods of large non-tidal energy before the harmonic analysis is performed.
 The total modelled water level is determined by adding the mean sea level from the tide gauge record over 1983-2001 to the corrected model sea surface height.} This correction is only applied to the modelled water level and not the modelled residuals.
@@
 
-The model's skill is reported through a few statistical measures of the total water level $\eta$ and the residual $\delta$. First, the mean absolute error ($\bar{e}$), defined as
+The model's skill is reported in terms of statistical measures of the total water level $\eta$ and the residual $\delta$. First, the mean absolute error ($\bar{e}$), defined as
 \begin{equation}
@@
 
-{\color{red} This case was also used to assess the importance of the northern boundary residual forcing and was used to briefly examine stratified effects.
+{\color{red} This case was also used to assess the importance of the northern boundary residual forcing and briefly examine the effects of stratification.
 When the northern boundary forcing is excluded the surge amplitude at Campbell River drops by about 10 cm.
@@
 where $C$ is the coefficient of friction, $\rho_{air}$ is the density of air, $g$ is the acceleration due to gravity, $U$ is the wind speed, and $\rho$ is the density of water.
-Given  $C=10^{-3}$, $\rho_{air}=1.23$ kg m$^{-3}$, $\rho= 1035$  kg m$^{-3}$, wind speed $U=20$ m s$^{-1}$, and depth $D=100$ m, the slope of the sea surface is approximately $5\times10^{-7}$.
+Assuming  $C=10^{-3}$, $\rho_{air}=1.23$ kg m$^{-3}$, $\rho= 1035$  kg m$^{-3}$, wind speed $U=20$ m s$^{-1}$, and depth $D=100$ m, the slope of the sea surface is approximately $5\times10^{-7}$.
 At Point Atkinson, a westerly wind can act over a small distance of approximately 50 km, giving a sea surface elevation on the order of centimetres in agreement with the model findings.
@@
 In this event, the pressure at the Vancouver International Airport dropped to about 99 000 Pa \citep{ECClimateArchive}, corresponding to a 20 cm rise in sea level.
-This magnitude is in line with the findings of \citet{murty1995storm} that the IB contributes up to $^1/_3$ of the total surge amplitude.
+This magnitude is in line with the findings of \citet{murty1995storm} that the IB contributes up to one third of the total surge amplitude.
 In our model, including the influence of remote atmospheric pressure forcing (in the sea level at the open boundary) produces realistic surge amplitudes compared with the runs only including local atmospheric forcing.
@@
 The observed residual at Tofino, BC (a point nearby but outside of the domain) is up to 10 cm  lower than the observed residual at Neah Bay, WA (a point nearby but inside of the domain).
-{\color{red} Since the storm surge is less sensitive to wind forcing at the surface than anomaly forcing at the open boundaries, the latter is likely the cause, although it still does not account for all of the error.}
+{\color{red} Since the storm surge is less sensitive to wind forcing than anomaly forcing at the open boundaries, the latter is likely the cause, although it still does not account for all of the error.}
 
@@
 Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland {\color{red}(Figure \ref{fig:spatial})}.
-{\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps the incoming surge wave is partially reflected from the mainland coast in the Strait of Georgia.
+{\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the incoming surge wave could be partially reflected from the mainland coast in the Strait of Georgia.
 Since local winds are stronger in this case, their contribution to the surge amplitude of about 6 cm at Point Atkionson is more significant than in the Feb 2006 case.
@@
 {\color{red} First, a comparison of the wind speed and direction for the two models and observations is provided in Figure \ref{fig:dec2012_weather} at Point Atkinson, Sandheads and a NOAA Neah Bay buoy.
-Sandheads is a lighthouse station in the southern Strait of Georgia.} For each of the models, the wind speed and direction {\color{red} the wind is heading} at the closest grid point to the longitude/latitude of Point Atkinson and Sandheads is used for comparison.
+Sandheads is a lighthouse station in the southern Strait of Georgia.} For each of the models, the wind speed and {\color{red} the direction towards which the wind is heading,} at the closest grid point to the longitude/latitude of Point Atkinson, Sandheads, and the NOAA buoy, is used for comparison.
 Although the low resolution CGRF model has a grid spacing of 33 km, grid points align closely with these locations.
@@
 Observations from Sandheads and Point Atkinson are from the Environment Canada Climate Data archives \citep{ECClimateArchive}.
-Observations at the NOAA Buoy are found at this url: \url{http://www.ndbc.noaa.gov/station_page.php?station=46087}.
+Observations at the NOAA Buoy were downloaded from: \url{http://www.ndbc.noaa.gov/station_page.php?station=46087}.
 Wind direction represents the direction that the wind is headed.} }

### 2015-06-09 AO_guides/manuscriptAO.tex
Added tide-killing

@@
 A harmonic analysis was performed for 67 constituents.
-In order to prevent over-predicted constituents due to high surge years, the time series was first filtered by applying a Doodson filter \citep{parker2007NOAA} and periods with large non-tidal amplitudes were removed.
+In order to prevent over-predicted constituents due to high surge years, the time series was first filtered by applying a tide-killing Doodson filter \citep{parker2007NOAA} and periods with large non-tidal amplitudes were removed.
 Long period constituents such as Sa, Ssa, etc were not used in the tidal predictions since these constituents are often contaminated by non-tidal energy from seasonal meteorological trends.

### 2015-06-10 AO_guides/manuscriptAO.tex
Finished Keith's comments

@@
 Generally, surge amplitudes at Tofino are slightly larger and earlier than Neah Bay and Bamfield.
-In the future, storm surge forecasts will make use of Neah Bay water level predictions from the National Oceanic and Atmospheric Administration (NOAA).}
+In the future, storm surge forecasts will make use of Neah Bay water level predictions from the National Oceanic and Atmospheric Administration (NOAA).
+The remote forcing is applied uniformly across the open boundaries.
+Given the similarities between the surge amplitude and phase at Neah Bay and Bamfield, this seems to be a reasonable assumption.
+Further, a geostrophic slope calculation with velocity 20 cm s$^{-1}$ would imply a difference of less than 5 cm between the northern and southern sides of the Strait of Juan de Fuca.}
 
@@
 \end{table}
-
-{\color{red} The remote forcing is applied uniformly across the open boundaries.
-Given the similarities between the surge amplitude and phase at Neah Bay and Bamfield, this seems to be a reasonable assumption.
-Further, a geostrophic slope calculation with velocity 20 cm s$^{-1}$ would imply a difference of less than 5 cm between the northern and southern sides of the Strait of Juan de Fuca.
-Non-uniform forcing conditions will be investigated in a future study.}
 
@@
 {\color{red} As such, model water level predictions were corrected by accounting for this error.
-Note that the model will contain tidal energy from shallow water constituents due to nonlinear interactions in the model equations and so the correction as calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (Sa, Ssa, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.
+Note that the model will contain tidal energy from shallow water constituents due to nonlinear interactions in the model equations and so the correction was calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (Sa, Ssa, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.
 The difference between these two predictions is added to the model sea surface height as a correction.
@@
 
-The model's skill is reported in terms of statistical measures of the total water level $\eta$ and the residual $\delta$. First, the mean absolute error ($\bar{e}$), defined as
+The model's skill is reported in terms of five statistical measures of the total water level $\eta$ and the residual $\delta$. First, the mean absolute error ($\bar{e}$), defined as
 \begin{equation}
@@
 Note that the surge manifests as a broad peak in both the modelled and observed residual leading to a high probability that large residuals coincide with the high tide.} As such, it is of minor consequence that the timing of the observed and modelled maximum residuals do not match perfectly.
+{\color{red}Note that the observed surge curve has a relatively fast rise at each of the stations, a feature that is not reproduced by the model where the surge curves are generally symmetric.
+This sharp rise in sea level is not present in the remote forcing.
+}
 
@@
 A negative number means the model's maximum was later than the observations.
-Values calucated over the same periods discusses in Table \ref{tab:statistics}.
+Values calculated over the same periods discussed in Table \ref{tab:statistics}.
 } }
@@
 Our results suggest that the local atmospheric pressure generates a sea level change that is smaller than the IB response.
-{\color{red} It is the atmopsheric pressure gradient that drives changes in the sea surface height.
+{\color{red} It is the atmospheric pressure gradient that drives changes in the sea surface height.
 Since our domain is relatively small, so is the atmospheric pressure gradient, and hence, we see only a small contribution from local atmospheric pressure.
@@
 {\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the incoming surge wave could be partially reflected from the mainland coast in the Strait of Georgia.
-Since local winds are stronger in this case, their contribution to the surge amplitude of about 6 cm at Point Atkionson is more significant than in the Feb 2006 case.
+Since local winds are stronger in this case, their contribution to the surge amplitude of about 6 cm at Point Atkinson is more significant than in the Feb 2006 case.
 However, this is still less than 10 \% of the total surge amplitude.
@@
 {\color{red} There is a fairly significant delay in the timing of the maximum surge at most stations (Table \ref{tab:peak}).
-The model's surge curve is much broader and does not contain the narrow peak on Nov 20 as seen in the observations.}
+The model's surge curve is much broader and does not contain the narrow peak on Nov 20 as seen in the observations.
+Both the observations and remote forcing contain a relatively fast rise in the surge curve that is not reproduced in the model, which might be explained by errors in the model wind speed and direction.
+}
 

### 2015-06-16 AO_guides/manuscriptAO.tex
Youyu/JP comments finished

@@
 In particular, Active Pass was closed and Porlier Pass was widened (Figure \ref{fig:zoom}).
-Shallow regions were represented by constant depth of 4 m.
 All of these compromises are expected to have significant effects locally but minimal effects on the large scale.}
@@
 Generally, surge amplitudes at Tofino are slightly larger and earlier than Neah Bay and Bamfield.
-In the future, storm surge forecasts will make use of Neah Bay water level predictions from the National Oceanic and Atmospheric Administration (NOAA).
 The remote forcing is applied uniformly across the open boundaries.
 Given the similarities between the surge amplitude and phase at Neah Bay and Bamfield, this seems to be a reasonable assumption.
-Further, a geostrophic slope calculation with velocity 20 cm s$^{-1}$ would imply a difference of less than 5 cm between the northern and southern sides of the Strait of Juan de Fuca.}
+Further, a geostrophic slope calculation with velocity 20 cm s$^{-1}$ would imply a difference of less than 5 cm between the northern and southern sides of the Strait of Juan de Fuca.
+In the future, this uniform specification of sea level at the western boundary will be relaxed by taking values from the solution of a large-scale forecasting model.}
 
@@
 
-The tidal forcing and remote forcing were applied using the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}, a radiation scheme that allows disturbances to propagate out of the domain without reflections off of the open boundary.
+The tidal forcing and {\color{red} non-tidal} remote forcing were applied using the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}, a radiation scheme that allows disturbances to propagate out of the domain without reflections off of the open boundary.
 The barotropic velocity component of the remote forcing was set to zero.
@@
 Zero gradient conditions were chosen because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain.
-{\color{red} No boundary conditions were set on the baroclinic velocities at the northern boundary.}
+{\color{red} At the northern boundary, we choose the `no boundary condition' of NEMO that sets up an estuarine circulation by relaxing the temperature and salinity over a width of 10 grids to the seasonal climatology.
+}
 
@@
 This magnitude is in line with the findings of \citet{murty1995storm} that the IB contributes up to one third of the total surge amplitude.
-In our model, including the influence of remote atmospheric pressure forcing (in the sea level at the open boundary) produces realistic surge amplitudes compared with the runs only including local atmospheric forcing.
-Our results suggest that the local atmospheric pressure generates a sea level change that is smaller than the IB response.
-{\color{red} It is the atmospheric pressure gradient that drives changes in the sea surface height.
-Since our domain is relatively small, so is the atmospheric pressure gradient, and hence, we see only a small contribution from local atmospheric pressure.
+{\color{red} In our model, realistic surge amplitudes are produced by including the influence of atmospheric pressure forcing in the sea level at the open boundary.
+This indicates that the IB response of sea level to atmospheric pressure variation is mostly accounted for.
+If the IB response at the boundary is set to zero (by excluding the remote forcing) and the local atmospheric pressure variation is included, the model obtains much a smaller response to atmospheric pressure forcing.
+This suggests that the local atmospheric pressure variation must be combined with the remote variation to generate the full IB response in the Salish Sea.
 }
@@
 Since both of these products provide the same level of accuracy for storm surge hindcasts, in the future we will use the high resolution atmospheric model to produce real-time storm surge predictions, even though most of our testing was performed with the low resolution model.
-Additionally, the sea surface height anomaly from Neah Bay, forecasted by the National Oceanic and Atmospheric Administration (NOAA), will be used in future real-time water level predictions during the storm surge season.
+{\color{red} Additionally, the sea surface height anomaly boundary condition of the present model will be taken from the solution of large-scale forecasting model.}
 

### 2015-06-17 AO_guides/manuscriptAO.tex
corrected mean error calculations
final look through of mansucript. Improved clairty, corrected typos

@@
 Given the similarities between the surge amplitude and phase at Neah Bay and Bamfield, this seems to be a reasonable assumption.
-Further, a geostrophic slope calculation with velocity 20 cm s$^{-1}$ would imply a difference of less than 5 cm between the northern and southern sides of the Strait of Juan de Fuca.
+In support of this assumption, a geostrophic slope calculation with an estuarine velocity of 20 cm s$^{-1}$ would imply a difference of less than 5 cm between the northern and southern sides of the Strait of Juan de Fuca.
 In the future, this uniform specification of sea level at the western boundary will be relaxed by taking values from the solution of a large-scale forecasting model.}
@@
 \centering 
-\tbl{{\color{red}Average surge amplitudes and delay in peak timing from Tofino at several tide gauges on the Pacific coast.
-Averages obtained over the five surge events discussed on this article.
+\tbl{{\color{red}Average peak surge amplitudes and delay in peak timing from Tofino at several tide gauges on the Pacific coast.
+Averages obtained over the five surge events hindcasted in this article.
 A negative delay indicates that the peak surge occurred later than at Tofino.}}
@@
 
-{\color{red} The ocean surface is forced with momentum fluxes, solar and non-solar heat fluxes, and precipitation from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2014new}.} Forecasts from the period of 2002-2012 are available.
+{\color{red} The ocean surface is forced with momentum fluxes, solar and non-solar heat fluxes, and precipitation from a 33-km global atmospheric reforecasting model {\color{red}called CGRF} suitable for use in ocean modelling \citep{smith2014new}.} Forecasts from the period of 2002-2012 are available.
 One additional simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.
@@
 Averages& & & & & & & & & & \\
-Point Atkinson & 11.6&	3.401&	3.483&	0.013&	0.995&	6.7&	0.302&	0.307	&0.192	&0.932\\
-Victoria       &10.8&	2.175	&2.162	&0.04	&0.989	&5.1	&0.29	&0.28&	0.167&	0.94\\
-Patricia Bay   &11.6&	2.573&	2.619&	0.028&	0.991&	6.0&	0.281&	0.279&	0.179&	0.932\\
-Campbell River & 11.6&	3.139&	3.177&	0.021&	0.994&	7.8&	0.246&	0.293&	0.188&	0.893\\
+Point Atkinson & 12.0 & 3.424 & 3.514 & 0.017 & 0.993 & 7.2 & 0.318 & 0.327 & 0.176 & 0.933\\
+Victoria       & 9.6 & 2.193 & 2.193 & 0.042 & 0.988 & 5.3 & 0.295 & 0.295 & 0.162 & 0.945\\
+Patricia Bay   & 11.0 & 2.587 & 2.648 & 0.030 & 0.989 & 6.6 & 0.285 & 0.293 & 0.175 & 0.932\\
+Campbell River & 11.4 & 3.184 & 3.219 & 0.026 & 0.993 & 7.4 & 0.280 & 0.311 & 0.172 & 0.920\\
 \hline
@@
 A negative number means the model's maximum was later than the observations.
-Values calculated over the same periods discussed in Table \ref{tab:statistics}.
+Values are calculated over the same periods discussed in Table \ref{tab:statistics}.
 } }
@@
 Averages  &  &  &  &  &  & \\
-Point Atkinson  & 5.196 & 5.277 & 0  & 0.752 & 0.634 & -3.125\\
-Victoria        & 3.147 & 3.125 & -0.625 & 0.684 & 0.563 & -2.125\\
-Patricia Bay    & 4.005 & 3.901 & -0.125 & 0.689 & 0.584 & -2.875\\
-Campbell River  & 4.756 & 4.835 & -0.250 & 0.673 & 0.590 & -3.875\\ 
+Point Atkinson & 5.106 & 5.194 & 0 & 0.806 & 0.681 & -2.6 \\
+Victoria       & 3.148 & 3.130 & -1.0 & 0.696 & 0.589 & -1.0\\
+Patricia Bay   & 3.939 & 3.856 & -0.2 & 0.724 & 0.614 & -2.2\\
+Campbell River & 4.726 & 4.796 & -0.4 & 0.730 & 0.640 & -2.6 \\
 \hline 
@@
 
-{\color{red} This case was also used to assess the importance of the northern boundary residual forcing and briefly examine the effects of stratification.
+{\color{red} This case was also used to assess the importance of the northern boundary residual forcing and to briefly examine the effects of stratification.
 When the northern boundary forcing is excluded the surge amplitude at Campbell River drops by about 10 cm.
@@
 Willmott skill scores for both modelled residuals are above {\color{red} 0.84} at these four locations.
-{\color{red} There is considerable noise in the observed Campbell River residual, which could explain its low skill score (Table \ref{tab:statistics}).} {\color{red} The modelled residuals in each simulation deviate only slightly from each other, most noticeable at Point Atkinson during the strong wind event on Dec 15, 2012 where simulation with the low resolution atmospheric forcing has a residual that is 5 cm higher.
+{\color{red} There is considerable noise in the observed Campbell River residual, which could explain its low skill score (Table \ref{tab:statistics}).} {\color{red} The modelled residuals in each simulation deviate only slightly from each other, most noticeable at Point Atkinson during the strong wind event on Dec 15, 2012 where the simulation with the low resolution atmospheric forcing has a residual that is 5 cm higher.
 The maximum residuals on Dec 17, 2012 are almost identical at each station (Table \ref{tab:peak}).} Although the wind speeds on Dec 15, 2012 are quite high, this event did not result in any flooding or residuals greater than 50 cm.
@@
 Comparisons between high resolution and low resolution atmospheric forcing further suggest that local winds are not the leading order factor in setting up large surge amplitudes.
-The low resolution model significantly underestimated wind speeds in certain areas of the domain, however, it performed just as accurately in reproducing surge amplitudes as the high resolution product.
+The low resolution model significantly underestimated wind speeds in certain areas of the domain; however, it performed just as accurately in reproducing surge amplitudes as the high resolution product.
 Although the high resolution model delivers fine scale detail and more accurate local wind patterns, it did not provide any improvement to the surge hindcast.

### 2015-08-06 AO_guides/manuscriptAO_final.tex
Removed red text from final version draft

--- 2015-06-17 AO_guides/manuscriptAO.tex
@@
 Such large domains typically use nested or unstructured grids to resolve the details of complicated coastlines.
-Appropriate treatment of rivers \citep{flather1994storm}, accurate bathymetry and the inclusion {\color{red} of} surface waves \citep{xu2010storm} have also been identified as important.
+Appropriate treatment of rivers \citep{flather1994storm}, accurate bathymetry and the inclusion  of surface waves \citep{xu2010storm} have also been identified as important.
 Most of these models neglect the effects of stratification by treating the ocean as a constant temperature and salinity.
@@
 %overview oF SoG/Salish Sea models
-The early modelling work of \citet{crean1988mathematical}, \citet{crean1988numerical} and \citet{stronach1993update} {\color{red} developed} a series of models covering the Strait of Juan de Fuca, the Strait of Georgia, and Puget Sound using a finite difference discretization of the equations of motion.
-%A barotropic version of their model was used to reproduce storm surge water level elevations with reasonable accuracy, however, this barotropic version did not include the effects of fresh water entering the system from rivers.
-Later, a baroclinic version of their model examined residual circulation and fresh water flushing rates \citep{stronach1993update, marinone1996model}.
+The early modelling work of \citet{crean1988mathematical}, \citet{crean1988numerical} and \citet{stronach1993update}  developed a series of models covering the Strait of Juan de Fuca, the Strait of Georgia, and Puget Sound using a finite difference discretization of the equations of motion.
 These models pointed to an important connection between the model's bottom friction parametrization and the location and amplitude of the M$_2$ amphidrome.
@@
 They noted that their semi-diurnal amplitudes were generally too low.
-{\color{red} An updated version of this model with added biogeochemical components and an extended domain produced higher than observed surface temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}.} Smaller domains have also been considered; notably, the passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.
+An updated version of this model with added biogeochemical components and an extended domain produced higher than observed surface temperatures in the Strait of Juan de Fuca, suggesting too little mixing within the Salish Sea \citep{giddings2014hindcasts}.
+Smaller domains have also been considered; notably, the passageways through the Discovery Islands north of the Strait of Georgia have been modelled baroclinically by \citet{foreman2012circulation} with the Finite Volume Coastal Ocean Model (FVCOM) pointing to difficulties in reproducing baroclinic currents on their grid.
 Accurate reproduction of observed tidal amplitudes and phases and appropriate treatment of mixing is a recurring difficulty in modelling this region.
@@
 This article presents the development of a three-dimensional numerical ocean model for the Strait of Georgia and its skill in simulating storm surge hindcasts.
-{\color{red} The main goals are to identify the forcing factors that contribute most significantly to storm surges in the Strait of Georgia, to examine the spatial distribution of the storm surges, and to evaluate the performance of a new model of this region.}  To complement previous barotropic storm surge modelling in this region, this model employs a higher resolution and includes the effect of stratification and rivers.
+The main goals are to identify the forcing factors that contribute most significantly to storm surges in the Strait of Georgia, to examine the spatial distribution of the storm surges, and to evaluate the performance of a new model of this region.
+To complement previous barotropic storm surge modelling in this region, this model employs a higher resolution and includes the effect of stratification and rivers.
 However, wetting and drying is not represented.
@@
 NEMO's ocean core solves the three-dimensional hydrostatic equations of motion for an incompressible fluid under the Boussinesq approximation on a structured computational grid.
-{\color{red} Although these features are not being used in the present work, NEMO's options for grid nesting and bio-geochemical coupling make it a useful tool for studying the complex physics and bio-geochemical interactions within the Strait of Georgia.} This work focuses on the physical configuration of the Salish Sea model, in particular, determining appropriate forcing and boundary conditions for accurate reproduction of tidal amplitudes and phases as well as storm surge elevations.
+Although these features are not being used in the present work, NEMO's options for grid nesting and bio-geochemical coupling make it a useful tool for studying the complex physics and bio-geochemical interactions within the Strait of Georgia.
+This work focuses on the physical configuration of the Salish Sea model, in particular, determining appropriate forcing and boundary conditions for accurate reproduction of tidal amplitudes and phases as well as storm surge elevations.
 Future work will include bio-geochemical coupling and data assimilation.
 NEMO version 3.4 was used in this study.
-{\color{red} NEMO has been successfully used in other coastal modelling studies in the Northeast Atlantic \citep{maraldi2013nemo} and the Mediterranean Sea \citep{brossier2012sensitivity}.}
+NEMO has been successfully used in other coastal modelling studies in the Northeast Atlantic \citep{maraldi2013nemo} and the Mediterranean Sea \citep{brossier2012sensitivity}.}
 
@@
 Bathymetry from the Cascadia physiography data set \citep{haugerud1999digital} was smoothed to limit large changes in depth across grid cells, so that $\Delta h/\bar{h} \leq 0.8$, where $\Delta h$ is the difference in depth between two adjacent grid cells and $\bar{h}$ is their average depth.
-For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells {\color{red} inward from the open boundary.
-Depths between 0 m and 4 m were set to 4 m.}  As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
-{\color{red} A magnified image of the complex passages through the San Juan and Gulf Islands is provided in Figure \ref{fig:zoom}.}
+For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells inward from the open boundary.
+Depths between 0 m and 4 m were set to 4 m.
+As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.
+A magnified image of the complex passages through the San Juan and Gulf Islands is provided in Figure \ref{fig:zoom}.
 
@@
 At the bottom boundary, partial $z$-levels were used in order to limit large changes in bathymetry across grid cells \citep{madec2012nemo}.
-{\color{red} Model constraints require that all passages be at least two grid points wide.
+Model constraints require that all passages be at least two grid points wide.
 This restriction required choices in either widening or closing narrow straits.
 In particular, Active Pass was closed and Porlier Pass was widened (Figure \ref{fig:zoom}).
-All of these compromises are expected to have significant effects locally but minimal effects on the large scale.}
+All of these compromises are expected to have significant effects locally but minimal effects on the large scale.
 
@@
 \includegraphics[scale=0.5]{Figures/bathy.pdf}
-\caption{{\color{red} Model domain (highlighted area) including bathymetry, rivers (green circles), and storm surge locations of interest  (red stars).
-The black rectangle represents the region displayed in Figure \ref{fig:zoom}.}}\label{fig:domain}
+\caption{ Model domain (highlighted area) including bathymetry, rivers (green circles), and storm surge locations of interest  (red stars).
+The black rectangle represents the region displayed in Figure \ref{fig:zoom}.}\label{fig:domain}
 \end{figure}
@@
 \includegraphics[scale=0.4]{Figures/bathy_zoom.pdf}
-\caption{{\color{red} Model bathymetry and coastlines in the passages between the San Juan and Gulf Islands.}}\label{fig:zoom}
+\caption{Model bathymetry and coastlines in the passages between the San Juan and Gulf Islands.}\label{fig:zoom}
 \end{figure}
@@
 The model is relaxed to the forced temperature and salinity over the 10 grid points (about 5 km) closest to the open boundaries, using a flow relaxation scheme \citep{engedahl1995use}.
-{\color{red} In the north, the seasonal climatologies of \citet{thomson1981oceanography} were used in the same manner.} %other references?
+In the north, the seasonal climatologies of \citet{thomson1981oceanography} were used in the same manner.
 
@@
 These values are the Tofino tide gauge values with the tides removed.
-{\color{red}Tidal predictions were generated with the tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year of the storm surge.
+Tidal predictions were generated with the tidal analysis package called t\_tide \citep{pawlowicz2002classical} using a time series from the year of the storm surge.
 A harmonic analysis was performed for 67 constituents.
@@
 Further, constituents with a low signal to noise ratio were neglected.
-} The storm surge simulations also used the sea surface height anomaly from Port Hardy, forced at the northern boundary in Johnstone Strait.
+The storm surge simulations also used the sea surface height anomaly from Port Hardy, forced at the northern boundary in Johnstone Strait.
 
-{\color{red} Although Tofino is not the closest tide gauge to the western boundary, we found that the surge conditions on the western coast of Vancouver Island and near the mouth of the Strait of Juan de Fuca are fairly uniform as summarized in Table \ref{tab:pacificsurge}.
+ Although Tofino is not the closest tide gauge to the western boundary, we found that the surge conditions on the western coast of Vancouver Island and near the mouth of the Strait of Juan de Fuca are fairly uniform as summarized in Table \ref{tab:pacificsurge}.
 Generally, surge amplitudes at Tofino are slightly larger and earlier than Neah Bay and Bamfield.
@@
 \centering 
-\tbl{{\color{red}Average peak surge amplitudes and delay in peak timing from Tofino at several tide gauges on the Pacific coast.
+\tbl{}Average peak surge amplitudes and delay in peak timing from Tofino at several tide gauges on the Pacific coast.
 Averages obtained over the five surge events hindcasted in this article.
-A negative delay indicates that the peak surge occurred later than at Tofino.}}
+A negative delay indicates that the peak surge occurred later than at Tofino.}
 {\begin{tabular}{|l |c | c|}
@@
 
-The tidal forcing and {\color{red} non-tidal} remote forcing were applied using the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}, a radiation scheme that allows disturbances to propagate out of the domain without reflections off of the open boundary.
+The tidal forcing and  non-tidal remote forcing were applied using the NEMO Flather scheme \citep{flather1994storm, madec2012nemo}, a radiation scheme that allows disturbances to propagate out of the domain without reflections off of the open boundary.
 The barotropic velocity component of the remote forcing was set to zero.
@@
 Zero gradient conditions were chosen because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain.
-{\color{red} At the northern boundary, we choose the `no boundary condition' of NEMO that sets up an estuarine circulation by relaxing the temperature and salinity over a width of 10 grids to the seasonal climatology.
-}
+At the northern boundary, we choose the `no boundary condition' of NEMO that sets up an estuarine circulation by relaxing the temperature and salinity over a width of 10 grids to the seasonal climatology.
 
@@
 
-{\color{red} The ocean surface is forced with momentum fluxes, solar and non-solar heat fluxes, and precipitation from a 33-km global atmospheric reforecasting model {\color{red}called CGRF} suitable for use in ocean modelling \citep{smith2014new}.} Forecasts from the period of 2002-2012 are available.
+ The ocean surface is forced with momentum fluxes, solar and non-solar heat fluxes, and precipitation from a 33-km global atmospheric reforecasting model called CGRF suitable for use in ocean modelling \citep{smith2014new}.
+Forecasts from the period of 2002-2012 are available.
 One additional simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.
@@
 The monthly flow from each river was input as a point source from 0 to 3 metres at the model point closest to the mouth of each river.
-Incoming water was assumed to be fresh and at surface {\color{red}ocean temperature}.
+Incoming water was assumed to be fresh and at surface ocean temperature.
 A total of 150 rivers were parametrized by this method and their locations are indicated by the green dots in Figure \ref{fig:domain}.
@@
 All storm surge runs were started at least three days prior to the event of interest with zero initial velocities and sea surface height and a stratification profile from model spin up.
-{\color{red} The modelled sea surface height adjusted to forcing in less than one day as determined by comparing a simulation with restart conditions to one with velocities and sea surface height initialized to zero}.
+The modelled sea surface height adjusted to forcing in less than one day as determined by comparing a simulation with restart conditions to one with velocities and sea surface height initialized to zero.
 
@@
 The model's M$_2$ amphidrome, indicated by the drop in amplitude around station 5 (Victoria), is too strong resulting in lower than observed M$_2$ amplitudes in this region.
-{\color{red} The model also represents the dip in M$_2$ amplitude around stations 25 and 26 reasonably well.} Additionally, the M$_2$ phase errors in the Strait of Juan de Fuca (stations 1-7) are quite large, leading to large complex differences in this region.
+The model also represents the dip in M$_2$ amplitude around stations 25 and 26 reasonably well.
+Additionally, the M$_2$ phase errors in the Strait of Juan de Fuca (stations 1-7) are quite large, leading to large complex differences in this region.
 The match in the Strait of Georgia (stations 10-24) is very good, with complex differences less than 6 cm, in part due to tidal tuning for highly accurate tides in the main region of interest.
-{\color{red}Because the tides were tuned for a good match in the Strait of Georgia, the match in the Strait of Juan de Fuca was compromised; the effect is particularly noticeable at the station closest to the open boundary (Port Renfrew, 1).
-} Details on the tidal tuning procedure are provided in Appendix \ref{sec:appendix}.
+Because the tides were tuned for a good match in the Strait of Georgia, the match in the Strait of Juan de Fuca was compromised; the effect is particularly noticeable at the station closest to the open boundary (Port Renfrew, 1).
+Details on the tidal tuning procedure are provided in Appendix \ref{sec:appendix}.
 
@@
 Stratification conditions from model spin up at a time close to the event (within five days) were chosen to initialize the temperature and salinity fields.
-{\color{red} Boundary conditions include atmospheric forcing (momentum and heat fluxes and precipitation) and fresh water input from rivers at the surface, sea surface height anomaly from observations at the open boundary, and tidal forcing with eight tidal constituents.}
+Boundary conditions include atmospheric forcing (momentum and heat fluxes and precipitation) and fresh water input from rivers at the surface, sea surface height anomaly from observations at the open boundary, and tidal forcing with eight tidal constituents.
 
 Forcing the model with only eight tidal constituents leads to an error in modelled sea surface height due to the omission of the next leading order tidal constituents such as J$_1$.
-Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can approach {\color{red} 20} cm during times of high tidal range.
-{\color{red} As such, model water level predictions were corrected by accounting for this error.
+Comparisons between full tidal predictions and tidal predictions using eight constituents suggest that this error can approach  20 cm during times of high tidal range.
+As such, model water level predictions were corrected by accounting for this error.
 Note that the model will contain tidal energy from shallow water constituents due to nonlinear interactions in the model equations and so the correction was calculated in the following way:  For each storm surge location of interest, two tidal predictions are calculated: 1) A tidal prediction with all constituents excluding long period constituents (Sa, Ssa, etc), shallow water constituents, and constituents with signal to noise ratio less than two, and 2) a tidal prediction with only the eight constituents used to force the model.
@@
 Note that the harmonic analysis used to generate the constituents is applied in the same manner as that used in the forcing, i.e.\, the time series from the year of the storm surge is filtered to remove periods of large non-tidal energy before the harmonic analysis is performed.
-The total modelled water level is determined by adding the mean sea level from the tide gauge record over 1983-2001 to the corrected model sea surface height.} This correction is only applied to the modelled water level and not the modelled residuals.
+The total modelled water level is determined by adding the mean sea level from the tide gauge record over 1983-2001 to the corrected model sea surface height.
+This correction is only applied to the modelled water level and not the modelled residuals.
 All times are reported in UTC.
@@
 Tidal predictions are calculated using t\_tide \citep{pawlowicz2002classical}, a MATLAB tool that calculates tidal harmonics from a given time series.
-{\color{red} The same method for calculating forcing residuals described in section \ref{sec:config} is used to calculate the observed residuals.}
+The same method for calculating forcing residuals described in section \ref{sec:config} is used to calculate the observed residuals.
 
@@
 These three statistics, along with the mean observed values and mean modelled values (which indicate the model's bias) are reported in Table \ref{tab:statistics} for each hindcast.
-{\color{red} Observed and modelled maximum water levels and surge values as well as differences in their timing are reported in Table \ref{tab:peak}.}
+Observed and modelled maximum water levels and surge values as well as differences in their timing are reported in Table \ref{tab:peak}.
 
@@
 The simulation ran from Feb 1-8, 2006.
-The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations {\color{red} with} corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River.
+The model's ability to reproduce this event is demonstrated in Figure \ref{fig:feb2006} which compares observations  with corrected model output at Point Atkinson, Victoria, Patricia Bay, and Campbell River.
 First, the timing of maximum water level matches well between the observations and the model at all four locations, the most significant contribution to high water being from the tides which are generally well-represented by the model.
@@
 It is surprising that Victoria and Patricia Bay agree so closely with observations because the M$_2$ tidal amplitudes in the model are known to be too low in this area.
-{\color{red} Given that the water levels are over predicted at Point Atkinson and Campbell River, it is likely that one of the forcing factors is overestimated in this example.
-}
+Given that the water levels are over predicted at Point Atkinson and Campbell River, it is likely that one of the forcing factors is overestimated in this example.
 
 In general, there is also very good agreement between the observed and modelled residuals (Figure \ref{fig:feb2006}).
-{\color{red} The most noticeable difference is a broader peak in the model's surge curve, leading to a delay of a few hours in timing of the peak surge for most of the stations (Table \ref{tab:peak}).
-Note that the surge manifests as a broad peak in both the modelled and observed residual leading to a high probability that large residuals coincide with the high tide.} As such, it is of minor consequence that the timing of the observed and modelled maximum residuals do not match perfectly.
-{\color{red}Note that the observed surge curve has a relatively fast rise at each of the stations, a feature that is not reproduced by the model where the surge curves are generally symmetric.
+The most noticeable difference is a broader peak in the model's surge curve, leading to a delay of a few hours in timing of the peak surge for most of the stations (Table \ref{tab:peak}).
+Note that the surge manifests as a broad peak in both the modelled and observed residual leading to a high probability that large residuals coincide with the high tide.
+As such, it is of minor consequence that the timing of the observed and modelled maximum residuals do not match perfectly.
+Note that the observed surge curve has a relatively fast rise at each of the stations, a feature that is not reproduced by the model where the surge curves are generally symmetric.
 This sharp rise in sea level is not present in the remote forcing.
-}
 
@@
 The statistics are calculated with hourly output over the three-day period displayed in Figure \ref{fig:feb2006}.
-The mean absolute error in the water level ranges from {\color{red}8.2 cm} at Victoria to {\color{red}12.2} cm at Campbell River, whereas, the mean absolute error in the residuals is smaller at all stations; however, this measure does not account for the large variability in water levels due to the tides.
+The mean absolute error in the water level ranges from 8.2 cm at Victoria to 12.2 cm at Campbell River, whereas, the mean absolute error in the residuals is smaller at all stations; however, this measure does not account for the large variability in water levels due to the tides.
 A statistic that does take into account this variability is $\gamma^2$ which suggests very good agreement in both the predicted water levels and residuals.
@@
 \end{table}
-%anything else to include? statistics calculated in Feb 2006 -paper production.ipynb and saved in statistics_feb2006.CV 
-%I also have standard deviations, correlations, ...
+
 \begin{table}[h]
 \centering 
-\tbl{{\color{red}Summary of maximum water level, maximum surge amplitude and model delay for all simulations.
+\tbl{Summary of maximum water level, maximum surge amplitude and model delay for all simulations.
 The model time delay is the number of hours between the maximum in the observations and the maximum in the model.
@@
 Values are calculated over the same periods discussed in Table \ref{tab:statistics}.
-} }
+}
 {\begin{tabular}{|l |c c c | c c c|} 
@@
 An additional storm surge hindcast for a Nov 16, 2006 storm surge is provided as another means of assessing model performance and is mentioned briefly here.
-The model performs reasonably well in this hindcast; Willmott skill scores are greater than {\color{red}0.95} (Table \ref{tab:statistics}).
+The model performs reasonably well in this hindcast; Willmott skill scores are greater than 0.95 (Table \ref{tab:statistics}).
 Further, the $\gamma^2$ statistic suggests that the standard deviation of the error in the modelled residuals is about 40\% ($\sqrt{0.16}$) of the standard deviation of the observed residuals.
-The best match in maximum surge occurred at Campbell River with a {\color{red}4.3 cm} underestimate by the model.
-{\color{red} The timing of the maximum surge was too early at all stations except Campbell River.}
+The best match in maximum surge occurred at Campbell River with a 4.3 cm underestimate by the model.
+The timing of the maximum surge was too early at all stations except Campbell River.
 
-{\color{red} This case was also used to assess the importance of the northern boundary residual forcing and to briefly examine the effects of stratification.
+This case was also used to assess the importance of the northern boundary residual forcing and to briefly examine the effects of stratification.
 When the northern boundary forcing is excluded the surge amplitude at Campbell River drops by about 10 cm.
@@
 Other studies have seen an improvement in modelled surge amplitudes when using a baroclinic model as opposed to a three-dimensional barotropic model \citep{ma2015oceanic}.
-We plan to investigate stratified effects in more detail in a future study.}
+We plan to investigate stratified effects in more detail in a future study.
 
@@
 Fourth, a simulation with only tides and local wind forcing, that is, excluding remote forcing and local atmospheric pressure forcing in the dashed blue curve.
-{\color{red} Comparisons between these simulations will determine the contributions from the remote forcing, the local wind forcing, and the local atmospheric pressure forcing.} Finally, a simulation without tidal forcing was performed to study the importance of the tide-surge interaction (Figure \ref{fig:tidesurge}).
-{\color{red} Although the winds and atmospheric pressure are influenced by large scale atmospheric features, we use the terms local and remote to differentiate between factors originating from inside and outside of the domain.}
+Comparisons between these simulations will determine the contributions from the remote forcing, the local wind forcing, and the local atmospheric pressure forcing.
+Finally, a simulation without tidal forcing was performed to study the importance of the tide-surge interaction (Figure \ref{fig:tidesurge}).
+Although the winds and atmospheric pressure are influenced by large scale atmospheric features, we use the terms local and remote to differentiate between factors originating from inside and outside of the domain.
 
 First, it is clear that remote forcing at the open boundary contributes most significantly to the surge at each of these locations (Figure \ref{fig:factors}).
-When the remote forcing is not included as a forcing condition, the surge drops from a maximum {\color{red}84 cm} to 11 cm at Point Atkinson and its maximum occurs eight hours later.
+When the remote forcing is not included as a forcing condition, the surge drops from a maximum 84 cm to 11 cm at Point Atkinson and its maximum occurs eight hours later.
 A similar drop in the maximum surge and delay in timing of the maximum surge is observed at the other locations.
@@
 The contribution to the surge due to local atmospheric forcing is small and appears to be geographically dependent, most likely due to the effects of the local winds.
-Removal of the local atmospheric forcing leads to a drop of about {\color{red}2 cm} in the maximum surge at Point Atkinson and {\color{red}2.5 cm} at Campbell River.
-By contrast, the surge amplitude increases by {\color{red} 4.5 cm} at Victoria and 3.6 cm at Patricia Bay.
+Removal of the local atmospheric forcing leads to a drop of about 2 cm in the maximum surge at Point Atkinson and 2.5 cm at Campbell River.
+By contrast, the surge amplitude increases by  4.5 cm at Victoria and 3.6 cm at Patricia Bay.
 The southeasterly winds over the Strait of Georgia could be acting to push water away from the southern tip of Vancouver Island, resulting in a slight decrease in the surge amplitude at Victoria and Patricia Bay when wind effects are included in the model.
@@
 This magnitude is in line with the findings of \citet{murty1995storm} that the IB contributes up to one third of the total surge amplitude.
-{\color{red} In our model, realistic surge amplitudes are produced by including the influence of atmospheric pressure forcing in the sea level at the open boundary.
+In our model, realistic surge amplitudes are produced by including the influence of atmospheric pressure forcing in the sea level at the open boundary.
 This indicates that the IB response of sea level to atmospheric pressure variation is mostly accounted for.
@@
 This suggests that the local atmospheric pressure variation must be combined with the remote variation to generate the full IB response in the Salish Sea.
-}
 
@@
 Previous simulations by \citet{murty1995storm} suggest that the surge entering the system through the Strait of Juan de Fuca contribute most significantly to storm surges in this domain.
-{\color{red} Our study is consistent with these findings.}
+Our study is consistent with these findings.
 
@@
 Next, an assessment of the importance of including the tide-surge interaction, which is important in shallow regions due to nonlinear effects and bottom friction \citep{bernier2007tide}, is presented for this hindcast (Figure \ref{fig:tidesurge}).
-The most significant differences occur at Point Atkinson where the maximum of the surge-only case is {\color{red} 8.4 cm} higher than maximum modelled residual, about 10\% of the total surge amplitude.
+The most significant differences occur at Point Atkinson where the maximum of the surge-only case is 8.4 cm higher than maximum modelled residual, about 10\% of the total surge amplitude.
 In contrast, the tide-surge interaction at Victoria is hardly noticeable.
@@
 Modelled residuals are generally too low in this hindcast (Figure \ref{fig:dec2006}), however, the timing of the surge agrees well.
-All Willmott skill scores rank above {\color{red}0.89} and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}).
+All Willmott skill scores rank above 0.89 and the $\gamma^2$ measure also indicates good agreement between model and observations (Table \ref{tab:statistics}).
 The $\gamma^2$ measure suggests that the standard deviation of the error in the modelled residuals is roughly half of the standard deviation of the observed residuals.
@@
 The observed residual at Tofino, BC (a point nearby but outside of the domain) is up to 10 cm  lower than the observed residual at Neah Bay, WA (a point nearby but inside of the domain).
-{\color{red} Since the storm surge is less sensitive to wind forcing than anomaly forcing at the open boundaries, the latter is likely the cause, although it still does not account for all of the error.}
+Since the storm surge is less sensitive to wind forcing than anomaly forcing at the open boundaries, the latter is likely the cause, although it still does not account for all of the error.
 
@@
 
-The evolution and spatial distribution of the surge is highlighted in Figure \ref{fig:spatial} where the modelled residual is plotted every hour beginning on Dec 15 at 05:30 UTC through to Dec 15 at 14:30 UTC.
+The evolution and spatial distribution of the surge is highlighted in Figure \ref{fig:spatial} where the modelled residual is plotted every hour beginning on Dec 15 at 02:30 UTC through to Dec 15 at 11:30 UTC.
 The surge enters the system from the Strait of Juan de Fuca, peaking between Dec 15 at 9:30 UTC and Dec 15 at 11:30 UTC and then begins to subside throughout the entire domain.
-Notably, the Strait of Georgia experiences the largest surge values, upwards of {\color{red}70 cm}, whereas the Juan de Fuca region experiences a maximum surge around {\color{red}60 cm} {\color{red} (Figure \ref{fig:spatial})}.
+Notably, the Strait of Georgia experiences the largest surge values, upwards of 70 cm, whereas the Juan de Fuca region experiences a maximum surge around 60 cm (Figure \ref{fig:spatial}).
 Additionally, the coastal areas on the mainland side of the Strait of Georgia experience slightly higher surges ($\sim$10 cm) than the Vancouver Island side, likely due to the southerly wind in the atmospheric model pushing water against the mainland coast.
@@
 \caption{Propagation of the storm surge on Dec 15, 2006.
-Modelled residual every hour starting on Dec 15 at 05:30 UTC.
+Modelled residual every hour starting on Dec 15 at 02:30 UTC.
 Time advances from top left to bottom right.
@@
 While the surge entering the domain from the Pacific Ocean is the leading order contribution, the winds may induce localized changes to the surge amplitude.
-Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland {\color{red}(Figure \ref{fig:spatial})}.
-{\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the incoming surge wave could be partially reflected from the mainland coast in the Strait of Georgia.
+Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland (Figure \ref{fig:spatial}).
+Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the incoming surge wave could be partially reflected from the mainland coast in the Strait of Georgia.
 Since local winds are stronger in this case, their contribution to the surge amplitude of about 6 cm at Point Atkinson is more significant than in the Feb 2006 case.
 However, this is still less than 10 \% of the total surge amplitude.
-}
 
@@
 Next, a relatively strong wind event recorded at Point Atkinson on Nov 19, 2009 is considered.
-{\color{red} Wind speeds at Point Atkinson reached up to 18 m s$^{-1}$ which is approximately the same magnitude to that observed on Feb 4, 2006.}  Although large wind speeds were observed, the surge associated with this event is rather small when compared to the previous cases (Figure \ref{fig:nov2009}).
-Two significant surge events are noted, the first occurring on Nov 19 with an observed amplitude of about {\color{red}55 cm} at Point Atkinson, and a second, larger surge of {\color{red} 70 cm} on Nov 20.
-{\color{red} The strong wind event of Nov 19 was associated with the smaller of the two surges, supporting the fact that local winds do not contribute significantly to surge amplitudes.}  Neither of the surges occurred during a high tide and so no high water level events were recorded.
-Although {\color{red} the sea surface height plots are} not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and another surge entering the system at Tofino and Port Hardy.
+Wind speeds at Point Atkinson reached up to 18 m s$^{-1}$ which is approximately the same magnitude to that observed on Feb 4, 2006.
+Although large wind speeds were observed, the surge associated with this event is rather small when compared to the previous cases (Figure \ref{fig:nov2009}).
+Two significant surge events are noted, the first occurring on Nov 19 with an observed amplitude of about 55 cm at Point Atkinson, and a second, larger surge of 70 cm on Nov 20.
+The strong wind event of Nov 19 was associated with the smaller of the two surges, supporting the fact that local winds do not contribute significantly to surge amplitudes.
+Neither of the surges occurred during a high tide and so no high water level events were recorded.
+Although the sea surface height plots are not shown here, the other three locations also observed a maximum surge on Nov 20, which appears to be linked to a secondary high wind event and another surge entering the system at Tofino and Port Hardy.
 
@@
 During the high wind events on Nov 19 and Nov 20, the winds at Point Atkinson were directed to the west, explaining the lack of strong surge observed during this event since wind stresses would act to depress the water on the east side of the basin.
-In fact, the surge amplitude at Victoria on Nov 20 was {\color{red}71.8 cm} which is slightly higher than that observed at Point Atkinson {\color{red}(70 cm)}.
+In fact, the surge amplitude at Victoria on Nov 20 was 71.8 cm which is slightly higher than that observed at Point Atkinson (70 cm).
 Of the cases considered so far, this is the only example where the surge amplitude at Victoria was higher than that at Point Atkinson.
-The model results do not reflect a higher surge at Victoria {\color{red}(54 cm versus 58.7~cm at Point Atkinson)}, perhaps due to an underestimate in the wind speed and some discrepancies in the wind direction of the atmospheric forcing.
-{\color{red} There is a fairly significant delay in the timing of the maximum surge at most stations (Table \ref{tab:peak}).
+The model results do not reflect a higher surge at Victoria (54 cm versus 58.7~cm at Point Atkinson), perhaps due to an underestimate in the wind speed and some discrepancies in the wind direction of the atmospheric forcing.
+There is a fairly significant delay in the timing of the maximum surge at most stations (Table \ref{tab:peak}).
 The model's surge curve is much broader and does not contain the narrow peak on Nov 20 as seen in the observations.
 Both the observations and remote forcing contain a relatively fast rise in the surge curve that is not reproduced in the model, which might be explained by errors in the model wind speed and direction.
-}
 
@@
 
-The modelled water levels and residuals are, {\color{red} on average,} too high for this hindcast, except at Victoria (Table \ref{tab:statistics}).
-Generally, the skills scores are very good, scoring above {\color{red} 0.89} at all stations.
+The modelled water levels and residuals are,  on average, too high for this hindcast, except at Victoria (Table \ref{tab:statistics}).
+Generally, the skills scores are very good, scoring above 0.89 at all stations.
 
@@
 
-{\color{red} First, a comparison of the wind speed and direction for the two models and observations is provided in Figure \ref{fig:dec2012_weather} at Point Atkinson, Sandheads and a NOAA Neah Bay buoy.
-Sandheads is a lighthouse station in the southern Strait of Georgia.} For each of the models, the wind speed and {\color{red} the direction towards which the wind is heading,} at the closest grid point to the longitude/latitude of Point Atkinson, Sandheads, and the NOAA buoy, is used for comparison.
+First, a comparison of the wind speed and direction for the two models and observations is provided in Figure \ref{fig:dec2012_weather} at Point Atkinson, Sandheads and a NOAA Neah Bay buoy.
+Sandheads is a lighthouse station in the southern Strait of Georgia.
+For each of the models, the wind speed and  the direction towards which the wind is heading, at the closest grid point to the longitude/latitude of Point Atkinson, Sandheads, and the NOAA buoy, is used for comparison.
 Although the low resolution CGRF model has a grid spacing of 33 km, grid points align closely with these locations.
@@
 Hindcasts with both atmospheric forcing products agree well with the observations (Figure \ref{fig:dec2012_surge}).
-Willmott skill scores for both modelled residuals are above {\color{red} 0.84} at these four locations.
-{\color{red} There is considerable noise in the observed Campbell River residual, which could explain its low skill score (Table \ref{tab:statistics}).} {\color{red} The modelled residuals in each simulation deviate only slightly from each other, most noticeable at Point Atkinson during the strong wind event on Dec 15, 2012 where the simulation with the low resolution atmospheric forcing has a residual that is 5 cm higher.
-The maximum residuals on Dec 17, 2012 are almost identical at each station (Table \ref{tab:peak}).} Although the wind speeds on Dec 15, 2012 are quite high, this event did not result in any flooding or residuals greater than 50 cm.
-Significant flooding occurred later, on Dec 17, 2012, where the observed surge amplitude at Point Atkinson was {\color{red}66.3 cm}.
+Willmott skill scores for both modelled residuals are above 0.84 at these four locations.
+There is considerable noise in the observed Campbell River residual, which could explain its low skill score (Table \ref{tab:statistics}).
+The modelled residuals in each simulation deviate only slightly from each other, most noticeable at Point Atkinson during the strong wind event on Dec 15, 2012 where the simulation with the low resolution atmospheric forcing has a residual that is 5 cm higher.
+The maximum residuals on Dec 17, 2012 are almost identical at each station (Table \ref{tab:peak}).
+Although the wind speeds on Dec 15, 2012 are quite high, this event did not result in any flooding or residuals greater than 50 cm.
+Significant flooding occurred later, on Dec 17, 2012, where the observed surge amplitude at Point Atkinson was 66.3 cm.
 Even though the low resolution model significantly under predicts the wind speed at some locations, it performs comparably to forcing with high resolution products.
@@
 Hindcasts of five significant events in February 2006, November 2006, December 2006, November 2009, and December 2012 were simulated.
-In general, the model performs well at reproducing the observed water level heights and residuals, with Willmott skill scores greater than {\color{red}0.84} in all cases and generally higher skill scores for the water level heights.
+In general, the model performs well at reproducing the observed water level heights and residuals, with Willmott skill scores greater than 0.84 in all cases and generally higher skill scores for the water level heights.
 Further, in a single simulation, the skill scores are fairly consistent between the four stations examined.
@@
 However, the local wind patterns can set up small spatial differences in the surge, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season.
-{\color{red}Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due a partial reflection of the surge from the mainland coast.} Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important north of the San Juan and Gulf Islands.
+Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due a partial reflection of the surge from the mainland coast.
+Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important north of the San Juan and Gulf Islands.
 The modelled surge amplitude increased by about 10\% at Point Atkinson when the tide-surge interaction was neglected in one example.
@@
 Since both of these products provide the same level of accuracy for storm surge hindcasts, in the future we will use the high resolution atmospheric model to produce real-time storm surge predictions, even though most of our testing was performed with the low resolution model.
-{\color{red} Additionally, the sea surface height anomaly boundary condition of the present model will be taken from the solution of large-scale forecasting model.}
+Additionally, the sea surface height anomaly boundary condition of the present model will be taken from the solution of large-scale forecasting model.
 
@@
 Finally, difficulties in reproducing accurate M$_2$ tidal harmonics in both the Strait of Georgia and the Strait of Juan de Fuca have been highlighted in this study among others \citep{stronach1993update, foreman2004m}.
-Future work will attempt to improve model accuracy by examining mixing parametrizations and energy dissipation due to bottom friction and their effect on tidal harmonics {\color{red} and investigating stratified effects on surge amplitudes.} Additionally, a biological-chemical modelling component will be added.
-
-%future directions: running real time, biology and chemistry, model improvements?
-%summary of skill scores
+Future work will attempt to improve model accuracy by examining mixing parametrizations and energy dissipation due to bottom friction and their effect on tidal harmonics and investigating stratified effects on surge amplitudes.
+Additionally, a biological-chemical modelling component will be added.
 
@@
 Next, the M$_2$ and K$_1$ fluxes at the northern boundary were adjusted because the cross-sectional area of transect used by \citet{thomson1980johnstone} to measure tidal currents in Johnstone Strait is smaller than the cross-sectional area of our northern boundary.
-{\color{red} The northern boundary tides were tuned for good matches at Yorke Island  and Kelsey Bay.} Finally, the remaining constituents were added, starting with O$_1$, then S$_2$, and then in pairs P$_1$/N$_2$ and Q$_1$/K$_2$.
+The northern boundary tides were tuned for good matches at Yorke Island  and Kelsey Bay.
+Finally, the remaining constituents were added, starting with O$_1$, then S$_2$, and then in pairs P$_1$/N$_2$ and Q$_1$/K$_2$.
 Final adjustments were made by slightly varying amplitude and phase, evaluating the changes to M$_2$, K$_1$ and the ratios S$_2$/M$_2$, N$_2$/M$_2$, K$_2$/M$_2$, O$_1$/K$_1$, P$_1$/K$_1$, Q$_1$/K$_1$ and then choosing new values to move closer to observations by linear interpolation.

### 2015-08-06 AO_guides/manuscriptAO_final.tex
removed path to figures
Fixed up mistakes from removing color and changing paths to figures

@@
 The numerical model employs the Nucleus for European Modelling of the Ocean (NEMO) architecture in a regional configuration.
-The model is evaluated through {\color{red} the} comparisons of tidal elevation harmonics and storm surge with observations.
+The model is evaluated through the comparisons of tidal elevation harmonics and storm surge with observations.
 Important forcing factors contributing to storm surges have been assessed.
-It is shown that the surge entering the domain from the Pacific Ocean contributes most significantly to the {\color{red} surge amplitude within the Strait of Georgia.
-Comparisons between simulations with high resolution and low resolution atmospheric forcing further emphasize that remote forcing is the dominant factor in surge amplitudes of this region.} In addition, local wind patterns cause a slight increase in surge amplitude on the mainland side of the Strait of Georgia when compared to the Vancouver Island coastal areas during a major wind storm on Dec 15, 2006.
+It is shown that the surge entering the domain from the Pacific Ocean contributes most significantly to the surge amplitude within the Strait of Georgia.
+Comparisons between simulations with high resolution and low resolution atmospheric forcing further emphasize that remote forcing is the dominant factor in surge amplitudes of this region.
+In addition, local wind patterns cause a slight increase in surge amplitude on the mainland side of the Strait of Georgia when compared to the Vancouver Island coastal areas during a major wind storm on Dec 15, 2006.
 Generally, surge amplitudes are found to be greater within the Strait of Georgia when compared to amplitudes in the Strait of Juan de Fuca.
@@
 
-The Strait of Georgia is a strongly stratified, semi-enclosed body of water located between Vancouver Island and the mainland of British Columbia {\color{red}\citep{thomson1981oceanography}}.
+The Strait of Georgia is a strongly stratified, semi-enclosed body of water located between Vancouver Island and the mainland of British Columbia \citep{thomson1981oceanography}.
 It is part of a larger system of waterways collectively known as the Salish Sea and is connected to the Pacific Ocean via the Strait of Juan de Fuca and Johnstone Strait (Figure \ref{fig:domain}).
 At its southern end, the Strait of Georgia is connected to the Strait of Juan de Fuca through several passages, the largest being Haro Strait.
-The San Juan and Gulf Islands are scattered throughout these passages where complex bathymetry and steep topographic slopes induce vigorous tidal mixing {\color{red}\citep{thomson1981oceanography}}.
+The San Juan and Gulf Islands are scattered throughout these passages where complex bathymetry and steep topographic slopes induce vigorous tidal mixing \citep{thomson1981oceanography}.
 Mixed diurnal and semi-diurnal tides are found throughout most of the Salish Sea but a semi-diurnal amphidrome is present nearby Victoria, where the tides are mainly diurnal \citep{thomson1981oceanography}.
-This system is {\color{red} a} home to several large coastal communities, such as Vancouver and Victoria, serves as a waterway for commercial traffic and recreational use, and supports a large and active fishing industry.
+This system is a home to several large coastal communities, such as Vancouver and Victoria, serves as a waterway for commercial traffic and recreational use, and supports a large and active fishing industry.
 
@@
 NEMO version 3.4 was used in this study.
-NEMO has been successfully used in other coastal modelling studies in the Northeast Atlantic \citep{maraldi2013nemo} and the Mediterranean Sea \citep{brossier2012sensitivity}.}
+NEMO has been successfully used in other coastal modelling studies in the Northeast Atlantic \citep{maraldi2013nemo} and the Mediterranean Sea \citep{brossier2012sensitivity}.
 
@@
 \centering
-\includegraphics[scale=0.5]{Figures/bathy.pdf}
+\includegraphics[scale=0.5]{bathy.pdf}
 \caption{ Model domain (highlighted area) including bathymetry, rivers (green circles), and storm surge locations of interest  (red stars).
@@
 \centering
-\includegraphics[scale=0.4]{Figures/bathy_zoom.pdf}
+\includegraphics[scale=0.4]{bathy_zoom.pdf}
 \caption{Model bathymetry and coastlines in the passages between the San Juan and Gulf Islands.}\label{fig:zoom}
@@
 In support of this assumption, a geostrophic slope calculation with an estuarine velocity of 20 cm s$^{-1}$ would imply a difference of less than 5 cm between the northern and southern sides of the Strait of Juan de Fuca.
-In the future, this uniform specification of sea level at the western boundary will be relaxed by taking values from the solution of a large-scale forecasting model.}
+In the future, this uniform specification of sea level at the western boundary will be relaxed by taking values from the solution of a large-scale forecasting model.
 
@@
 \centering 
-\tbl{}Average peak surge amplitudes and delay in peak timing from Tofino at several tide gauges on the Pacific coast.
+\tbl{Average peak surge amplitudes and delay in peak timing from Tofino at several tide gauges on the Pacific coast.
 Averages obtained over the five surge events hindcasted in this article.
@@
 
-%perhaps a table of complex differences at tidal stations similar to Table 1 of Foreman et al (1995)  (such as the one produced by tidetools.calc_diffs_meas_mod) here?
-%(would be cool to include the complex differences calculated at the VENUS nodes too)
-%perhaps a nice contour map of M2 and K1 amps and phases goes here?
-%if it's favourable, we could compare our complex differences to Foreman et al (1995), who got an average of D=3cm for M2 and D=2.5cm for K1
-
 \begin{figure}
 \centering
-\includegraphics[scale=0.6]{Figures/tides.pdf}
+\includegraphics[scale=0.6]{tides.pdf}
 \caption{Evaluation of model tidal amplitude and phases for the M$_2$ and K$_1$ constituents.
@@
 \centering
-\includegraphics[scale=0.6]{Figures/feb2006.pdf}
+\includegraphics[scale=0.6]{feb2006.pdf}
 \caption{Comparison of observations and model output for the Feb 4, 2006 storm surge.
@@
 \centering
-\includegraphics[scale=0.6]{Figures/feb2006_factors.pdf}
+\includegraphics[scale=0.6]{feb2006_factors.pdf}
 \caption{Comparison of the modelled residual for four simulations in Feb 2006: A simulation with all forcing (solid black), a simulation without local atmospheric forcing (dashed green), a simulation without remote forcing (solid red), and a simulation with only tides and local wind forcing, i.e.\ no remote forcing or local atmospheric pressure (dashed blue).
@@
 \centering
-\includegraphics[scale=0.6]{Figures/feb2006_tidesurge.pdf}
+\includegraphics[scale=0.6]{feb2006_tidesurge.pdf}
 \caption{Modelled residuals (solid) compared with a simulation of the surge propagation without tidal forcing (dashed) for the Feb 2006 hindcast. }
@@
 \centering
-\includegraphics[scale=0.6]{Figures/feb2006_tidesurge_spatial.pdf}
+\includegraphics[scale=0.6]{feb2006_tidesurge_spatial.pdf}
 \caption{Spatial dependency of tide-surge interaction in Feb 2006 hindcast.
@@
 A major wind storm on Dec 15, 2006 is hindcasted next.
-Although the storm was very strong (gusts over 120 km hr$^{-1}$) and the observed residual at Point Atkinson reached {\color{red}94 cm}, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area \citep{forseth2006adaptation}.
+Although the storm was very strong (gusts over 120 km hr$^{-1}$) and the observed residual at Point Atkinson reached 94 cm, this event did not coincide with high tide and so no significant or reported flooding occurred in the Vancouver area \citep{forseth2006adaptation}.
 Because of the high winds, this event has been used to examine the propagation of the surge throughout the domain and spatial variations in the surge amplitude.
@@
 \centering
-\includegraphics[scale=0.6]{Figures/dec2006.pdf}
+\includegraphics[scale=0.6]{dec2006.pdf}
 \caption{Comparison of observed and modelled water level and residuals during the Dec 15, 2006 storm surge simulation. }
@@
 \centering
-\includegraphics[scale=0.6]{Figures/dec2006_spatial.pdf}
+\includegraphics[scale=0.6]{dec2006_spatial.pdf}
 \caption{Propagation of the storm surge on Dec 15, 2006.
@@
 \centering
-\includegraphics[scale=0.6]{Figures/dec2006_spatial_sshonly.pdf}
+\includegraphics[scale=0.6]{dec2006_spatial_sshonly.pdf}
 \caption{As in Figure \ref{fig:spatial} but for a simulation that does not include atmospheric forcing. }
@@
 \centering
-\includegraphics[scale=0.6]{Figures/nov2009.pdf}
+\includegraphics[scale=0.6]{nov2009.pdf}
 \caption{Wind and storm surge conditions at Point Atkinson for a strong wind event on Nov 19, 2009.
@@
 \centering
-\includegraphics[scale=0.6]{Figures/dec2012_weather.pdf}
-\caption{{\color{red}Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson (top), Sandheads (middle) and NOAA Buoy 46087 (bottom).
+\includegraphics[scale=0.6]{dec2012_weather.pdf}
+\caption{Comparison of observed and modelled winds during Dec 2012 storm surge simulation at Point Atkinson (top), Sandheads (middle) and NOAA Buoy 46087 (bottom).
 Observed winds shown in solid black, low resolution CGRF model in dashed black and high resolution model in grey.
@@
 Observations at the NOAA Buoy were downloaded from: \url{http://www.ndbc.noaa.gov/station_page.php?station=46087}.
-Wind direction represents the direction that the wind is headed.} }
+Wind direction represents the direction that the wind is headed.}
 \label{fig:dec2012_weather}
@@
 \centering
-\includegraphics[scale=0.6]{Figures/dec2012_surge.pdf}
+\includegraphics[scale=0.6]{dec2012_surge.pdf}
 \caption{Comparison of observed (solid black) and modelled residuals for the Dec 2012 storm surge simulation.
